### 2014-05-29 manuscript/styron_hetland_wenchuan_topo_manuscript.tex
working manuscript tex (agu format), before hetland's final mods

\documentclass[twocolumn,jgrga]{AGUTeX}
\usepackage{mathptmx}
\usepackage{amssymb,amsmath}

%\usepackage[unicode=true]{hyperref}
%\hypersetup{breaklinks=true,
%            bookmarks=true,
%            pdfauthor={},
%            pdftitle={Topographic stresses and the Wenchuan earthquake},
%            colorlinks=true,
%            citecolor=blue,
%            urlcolor=blue,
%            linkcolor=magenta,
%            pdfborder={0 0 0}}
%\urlstyle{same}

%\usepackage{lineno}
%\linenumbers*[1]

\usepackage[pdftex]{graphicx}
\setkeys{Gin}{draft=false}

% Document metadata stuff
\authorrunninghead{STYRON AND HETLAND}
\titlerunninghead{WENCHUAN TOPOGRAPHIC STRESS}
\authoraddr{Corresponding author: Richard H. Styron, Department of Earth and
Environmental Sciences, University of Michigan, 2534 CC Little Bldg.,
1100 N. University Ave., Ann Arbor, MI 48104, USA
(richard.h.styron@gmail.com)}

\begin{document}
\title{Topographic stresses and the Wenchuan earthquake}

\authors{Richard H. Styron \altaffilmark{1} and Eric A. Hetland \altaffilmark{1}}

\altaffiltext{1}{Department of Earth and Environmental Sciences, University of Michigan,
Ann Arbor, Michigan, USA.}

% document
\begin{abstract}
Though it is widely recognized that large mountain ranges may produce
significant stresses in the earth's crust, these stresses are not
commonly quantified. Nonetheless, it is possible that active faults near
high topography may be affected by topographic stresses. In this work,
we calculate the topographic stress field in the Longmen Shan region of
eastern Tibet, where faulting occurs at the base of \textasciitilde{} 4
km relief.
\end{abstract}

\begin{article}
\section{Introduction}\label{introduction}

\begin{quote}
``And He has set up on the earth mountains standing firm, lest it should
shake with you'' -- \emph{The Holy Qur'an}, 15:16
\end{quote}

Stress is of fundamental importance to many processes in the earth. Both
the isotropic (pressure) and anisotropic (shear) components of stress
exert control on the deformation state of the earth at any point in the
brittle regime. However, unlike other fundamental quantities such as
temperature, stress is typically difficult to measure \emph{in situ},
without drilling-based techniques. Therefore, stress is often treated in
a semi-quantitative manner, with an emphasis on directions and relative
magnitudes of the principal stresses, either locally or regionally
\citep[e.g.,][]{angelier1994}. These estimates of stress are commonly
derived from strain, for example from studies of earthquake focal
mechanisms \citep[e.g.,][]{michael1987} or of geological fault slip data
\citep[e.g.,][]{reches1987}.

In areas of substantial relief, high terrain and steep slopes generate
large stresses in the crust beneath and adjacent to the high topography
\citep{jeffreys1924, coblentz1996}. Because of the irregularity of
topography in mountainous regions, the stresses produced by topography
are also heterogeneous, and may play a prominent role in local or
regional deformation, particularly if the region is tectonically active.
For example, shear and normal stresses on a fault from topographic
loading may push a particular fault closer to or (as suggested by the
Qur'an) farther from failure, or reorient the net shear stress direction
on a fault. These effects may affect the localization or deformational
style in a region. Furthermore, heterogeneous topographic stresses on a
particular fault may affect the way earthquake ruptures propagate across
the fault plane. Despite this, the degree to which topographic stresses
affect faulting has received little direct study.

In this work, we investigate the effects of topographic stresses on the
faults that ruptured in the 2008 \emph{M}7.9 Wenchuan, China earthquake.
This earthquake is an ideal candidate for study because it occurred at
the base of the Longmen Shan, one of the largest and steepest
escarpments on Earth (Figure~\ref{fig:lms_map}), and has a well-studied
coseismic slip distribution characterized by significant along-strike
variations in coseismic slip. Additionally, because the earthquake
occurred after \textasciitilde{}2000 years of seismic quiescence,
postseismic stresses in the lithosphere are likely to be negligible,
suggesting that the stress state on the faults at the time of rupture
can be approximated as the sum of topographic, lithostatic, and
accumulated tectonic stresses. We then use the Mohr-Coulomb failure
criterion to bracket the tectonic stress field, pore fluid pressure, and
static friction of the fault.

\subsection{Previous work on topographic
stresses}\label{previous-work-on-topographic-stresses}

Aspects of topographic stresses and their relevance to tectonics have
been studied for some time. \citet{jeffreys1924} noted that the
presence of high mountains is evidence that the Earth's crust can
support significant heterogeneous differential stresses over long time
periods. \citet{dalmayrac1981} and \citet{molnar1988} discussed how 
extensional deformation in
the high parts of orogens temporally coincident with contractional
deformation at the low-elevation margins of the orogen may be explained
by a spatially invariant, depth-integrated horizontal compressive stress
with spatially varying vertical stresses caused by topography.
\citet{richardson1994} exploited this
relationship in the central Andes to estimate horizontal tectonic
stresses through finite element modeling, and more recently 
\citet{copley2009} 
performed similar work in Albania. \citet{bollinger2004} showed how 
increased normal stress on the Main
Himalayan thrust due to loading of the high Himalayan massifs locally
suppressed microseismicity and increased fault locking. Meade and Conrad
\citet{meade2008} demonstrated that the increased weight of the
uplifting Andes influenced the Nazca-South America convergence rate.

Additionally, topographic stresses play a role in common models of
orogenic dynamics. For example, topographic loading is central to thin
viscous sheet models of lithospheric deformation \citep[e.g.,][]{birdpiper80,
flesch2010gpe}. Critical taper models for thrust or extensional
wedges \citep[e.g.,][]{dahlen1990, xiao1991} incorporate both variation in
vertical stress due to changing elevation as well as a shear stress
contributed by the slope of the wedge surface. Critical taper models
also incorporate the idea that in a growing wedge, progressively
increasing topographic stresses may eventually prevent continued slip on
a given fault plane, and strain will instead be transferred to the toe
of the thrust wedge where the stress state is more favorable. Of
particular relevance to this work are `fixed boundary' models of
gravitational collapse and spreading \citep{rey2001}, where an excess of
gravitational potential energy associated with the high topography and
thickened crust of eastern Tibet causes a transfer of rock to the
foreland through horizontal contraction at the rangefront \citep[e.g.,][]
{dewey1988, liuyang2003, copleymckenzie2007}. This process may be
aided by a weak (sub)horizontal structure at depth (such as a shear zone
or weak crustal channel) that is capable of transferring the vertical
and radial stresses from topography throughout an orogen to its margins,
leading to contraction there \citep[e.g.,][]{clark2005, burchfiel2008,
fleschbendick2012}.

The contributions of variable topography to the full stress field in the
elastic upper crust has been studied on smaller spatial scales. 
\citet{mctiguemei1981} and
\citet{savageswolfs1986} investigated the stress components from long,
symmetric ridges and showed how horizontal tension is induced under
ridge crests and horizontal compression is induced under valleys, mostly
due to shear stresses generated by slopes. Work in this vein was
continued by \citet{millerdunne1996} and \citet{martel2006}, who have 
focused on shallow rock fracturing
resulting from topographic stresses. \citet{liuzoback1992} investigated 
whether the topographic stresses
generated by the mountains around Cajon Pass (California, USA)
contributed to the observed left-lateral shear stress on a shallow
portion of the right-lateral San Andreas fault. In their study, they
developed methods for calculating the three dimensional elastic stress
tensor field arbitrary topography, where previous solutions were limited
to two dimensions and required topography to be mathematically-defined
(e.g., sinusoidal).

\citet{luttrell2011} inferred the coseismic shear stress
changes during the 2010 \emph{Mw}8.8 Maule, Chile earthquake, and using
the topographic stresses due to the overlying fore arc, constrained the
the stresses that led to this earthquake. In their study, they
calculated topographic stresses following a similar procedure as
\citet{liuzoback1992} (which we describe below), although
they only considered the component of the topographic load that can be
described by convolving a Boussinesq solution with the topography.
\citet{luttrell2011} also considered the contribution to
stresses due to buoyancy.

\subsection{The 2008 Wenchuan, China
earthquake}\label{the-2008-wenchuan-china-earthquake}

The 2008 \emph{M}7.9 Wenchuan, China earthquake is one of the most
devastating earthquakes in recent history (for an in-depth review and
discussion of the earthquake, see \citet{zhang2010}).
Surface rupture occurred along a 240 km segment of the Beichuan fault
and a parallel 72 km segment of the Pengguan fault \citep{xu2009} 
(Figure~\ref{fig:lms_map}). These faults lie at the base of the central and
northeastern Longmen Shan, a mountain range that forms the eastern
margin of the Tibetan plateau. Total relief across the central Longmen
Shan is around 4 km, though relief subsides somewhat to the northeast.
The central and southwestern Longmen Shan is the steepest margin of the
Tibetan plateau \citep{clarkroyden2000} and one of the highest and
steepest escarpments on earth. This is most apparent in the southwestern
portion of the earthquake rupture, where elevations \textgreater{} 4000
m over the Pengguan massif (a Precambrian crystalline massif in the
hanging wall of the Beichuan thrust) drop to \textasciitilde{}1200 m in
as little as 6 km map distance.

\begin{figure*}[ht!]
\includegraphics[width=40pc]{../figures/lms_map.pdf}
\caption{Map of eastern Tibet and the Sichuan basin, showing active
structures from \citet{styron2010}. Faults that ruptured
in the 2008 Wenchuan earthquake are shown in pink. GPS velocities are
relative to the mean velocity of sites within Sichuan basin, with
1$\sigma$ uncertainty, from the dataset of Liang et al.
\citet{liang2013}. Beachball is from the Global CMT focal mechanism
solution for the 2008 Wenchuan earthquake. BF = Beichuan fault. PF =
Pengguan fault. P = Pengguan massif. Grey box shows the extent of
Figures 3 and 9.}
\label{fig:lms_map}
\end{figure*}

Surface ruptures during the Wenchuan earthquake are highly variable and
show vertical (reverse-sense) displacements up to 9 m and horizontal
(right-lateral sense) displacements up to 5 m \citep{lin2009, liu2009,
xu2009}. In general, vertical displacements are higher in the
southwestern to central portions of the Beichuan rupture and decrease in
the northeast, whereas horizontal offsets are higher in the central to
northeast, though considerable variation exists. Coseismic slip models
constrained by seismic and geodetic data reveal a complicated pattern of
coseismic slip in the Wenchaun earthquake, with several high-slip
patches that dominate the seismic moment release and significant
variation coseismic slip rake along strike \citep[e.g.,][]{nakamura2010,
shen2009, tong2010, feng2010, zhang2011, qi2011, fielding2013}.
The variation in rake is such that the southwest portions of the fault
slipped largely in a reverse sense, while the northeast portions sliped
largely in a right-lateral sense. This change in rake is associated with
a change in inferred fault dip. Sections of faults that ruptured in the
Wenchuan earthquake (which we simply refer to as the ``Wenchuan
earthquake faults'') with shallow to moderate dips largerly ruptured as
thrust, and sections with steeper dips largely ruptured as strike-slip.
\citet{medinaluna2013} concluded that this
relationship is consistent with a uniform orientation of principal
stresses, where the variation of the dip of the fault leads to a change
in the direction of maximum fault shear stress, which they assumed to be
parallel to the coseismic slip rake.

\subsection{This study}\label{this-study}

We seek to quantify the topographic stress field in the Longmen Shan
region, and on the Wenchuan earthquake faults themselves. Specifically,
we evaluate the extent to which topographic stresses promote or inhibit
slip in the 2008 Wenchuan earthquake. If topographically induced shear
stresses on the fault are in roughly in the same direction as the
coseismic slip, then topographic loading general promotes coseismic slip
(Figure \ref{fig:topo_fault_scenarios} a). On the other hand, if the
topographic shear stresses are roughly in the opposite direction of
coseismic slip, then topographic loading inhibits coseismic slip (Figure
\ref{fig:topo_fault_scenarios} b). If topographic loading resisted slip
across the Beichuan faults, then tectonic stresses would need to
counteract the topographic fault stresses for the coseismic slip to
result. On a smaller scale, the heterogeneity of coseismic slip in the
earthquake may be influenced by shorter wavelength variations in
topography and topographic stresses.

\begin{figure*}[t!]
\centering
\includegraphics[width=40pc]{../figures/topo_stress_possibilities.pdf}
\caption{Scenarios for topographic effects on rangefront thrust
faulting. (a) Topographic stresses promote thrust faulting. (b)
Topographic stresses inhibit thrust faulting.}
\label{fig:topo_fault_scenarios}
\end{figure*}

Topographic stresses are only one component of the total stress field in
the crust \citep{molnar1988}. Coseismic slip in the Wenchaun earthquake
is due to the total accumulated stress on the faults, which also
includes components from lithostatic and tectonic stresses. (In the
present study, we do not consider stresses due to flexure \citep[e.g.,][]
{luttrell2007} or buoyancy \citep[e.g.,][]{luttrell2011}.) By quantifying
both the topographic and lithostatic stresses, we can use coseismic slip
models to solve for the tectonic stress assuming that (1) the total
shear stress on the fault is in the direction of coseismic slip \citep[e.g.,]
[]{angelier1994} and (2) the fault is at Mohr-Coulomb failure everywhere
that is slipped. If topographic stresses are significant and produce
shear in the direction of fault slip, then for given values of static
friction and pore fluid pressure, we can calculate the amount of
tectonic stress that can be added to the ambient stress field before the
faults should rupture; given limited acceptable ranges for friction and
fluid pressure, we are essentially able to place maximum constraints on
tectonic stress. Alternately, if topographic stresses work against
coseismic slip, for given friction and fluid pressures we can estimate
the minimum magnitudes of tectonic stresses necessary to overcome shear
and frictional resistance to slip. In a scenario with complex faulting
and topography, it may be possible to put tight bounds on minimum and
maximum magnitudes and directions of tectonic stresses. To account for
the non-uniqueness of the solution, we use a Bayesian Monte-Carlo
methodology to estimate posterior probability density functions (PDFs)
of tectonic stresses, static fault friction, and pore fluid pressure.

\section{Topographic stresses on the Longmen Shan
faults}\label{topographic-stresses-on-the-longmen-shan-faults}

To quantify tectonic stresses on the Wenchuan earthquake faults, we
first calculate the topographic stress field in the upper crust
throughout eastern Tibet, then interpolate those stresses onto three
dimensional models of the faults taken from coseismic slip models.
Finally, we can calculate topographic shear and normal stresses on the
faults and compare those to the coseismic slip patterns.

\subsection{Topographic stress tensor field
calculations}\label{topographic-stress-tensor-field-calculations}

\subsection{Analytical description}\label{analytical-description}

We calculate the stress tensor field induced by topography throughout
eastern Tibet using methods developed by 
\citet{liuzoback1992}. They show that the topographic stress tensor
field can be determined by a convolution of a topographic loading
function with Green's functions describing the stresses in an elastic
halfspace due to a point load at the surface. In our notation, the
stress tensor field due to topography, $M(x,y,z)$, is given by

\begin{equation}
M(x, y, z) = G(x, y, z) * F(x, y) \; ,
\end{equation}

where $G(x,y,z)$ is a set of Green's functions for the six stress tensor
elements, and $F(x,y)$ is a topographic loading function, described
below. We assume compressive stresses are positive, and $x>0$ is east,
$y>$ is north, and $z>0$ is depth. \citet{liuzoback1992}
show that $M(x,y,z)$ can be decomposed into two components as

\begin{equation}
M(x,y,z) = M^B(x,y,z) + M^C(x,y,z) \; .
\end{equation}

$M^B(x,y,z)$ is the component of the stress field due to the vertical
loading of the topography, and is

\begin{equation}
M^B(x, y, z) = G^B(x,y,z) * F_v(x, y) \; ,
\label{eqn:bous}
\end{equation}

where $G^B(x,y,z)$ are the Boussinesq solutions for stresses in a
halfspace due to a vertical point load on the surface (see Appendix A1)
and $F_v(x,y) = \rho g h(x,y)$. Note that $h(x,y)<0$ since $z<0$ is
depth. $M^C(x,y,z)$ is the component of the stress field due to the
mechanical coupling of the topography to the half-space, i.e.,
describing the lateral spreading forces in the rock above the halfspace,
and is given by

\begin{equation}
M^C(x, y, z) = G_x^C(x,y,z) * F_{h, x}(x, y) + G_y^C(x,y,z) * F_{h, y}(x, y) \; ,
\end{equation}

where $G_i^C(x,y,z)$ are the Cerruti solutions for a horizontal point
source load in the $i$ direction on the halfspace surface (see Appendix
A1). The the horizontal loading functions are given by

\begin{equation}
\begin{split}
F_{h, \, x}(x,y) = ( \rho g h(x,y) + M_{xx}^B(x,y,0) + T_{xx} )\, \frac{\partial h}{ \partial x} \\
\quad + (M_{xy}^{B}(x,y,0) + T_{xy}) \frac{\partial h}{ \partial y}
\end{split}
\label{eqn:f_hor_xz}
\end{equation}

and

\begin{equation}
\begin{split}
F_{hor, \, yz}(x,y) = ( \rho g h(x,y) + \sigma_{yy}^B(x,y,0) + T_{yy} )\, \frac{\partial h}{ \partial y} \\
\quad + (M_{xy}^{B}(x,y,0) + T_{xy}) \frac{\partial h}{ \partial x}\; . 
\end{split}
\label{eqn:f_hor_yz}
\end{equation}

$M_{ij}^B(x,y,0)$ is the stress from the vertical (Boussinesq) load
evaluated at $z=0$, and $T_{ij}$ is the tectonic stress component, which
we neglect in the present calculations.

\subsection{Numerical implementation}\label{numerical-implementation}

Topography was taken from the CGIAR-CSI v.4 release
\citep{jarvis2008srtm} of the Shuttle Radar Topographic Mission
\citep{farr2007srtm} Digital Elevation Model (DEM) at 1 km nominal
resolution. The DEM was projected from native WGS84 geographic
coordinates to UTM zone 48N, decreasing the nominal horizontal
resolution to 851 m. We assume a Poisson halfspace, and Green's
functions for the Boussinesq and Cerruti point-source solutions were
calculated at regular points in large square, 2-D grid at each depth
considered, with the point-source centered in the grid (see Table
\ref{table:convo_params} for model parameters). A mask was applied to
each grid such that values outside the kernel radius (i.e.~the `corners'
of the array) were zero, so that the kernel is circular. The size of the
grid was chosen to be quite large to incorporate potential contributions
from the elevated topography throughout eastern Tibet. So that the
Green's functions and the topography were discretized on the same size
grid, we pad the Green's function array with zeros. Because of
singularities in the Green's functions at depth $z=0$, we use
$\sigma^B(x,y,z)$ with $z=851$ m, the shallowest level of our
calculations, in construction of the horizontal loading functions in
Equations \ref{eqn:f_hor_xz} and \ref{eqn:f_hor_yz}. Convolutions were
computed using a 2D fast Fourier transform. All calculations were
implemented in Python (v. 2.7.3) using IPython \citep{perez2007ipython},
NumPy (v. 1.7) \citep{oliphant2007numpy} and Pandas (v. 12)
\citep{mckinney2010}; additional statistical analysis was performed with
StatsModels \citep{seabold2010}. We created an open-source Python package
to calculate topographic stresses in a reasonably automated way, which
is available at \url{https://github.com/cossatot/halfspace}. The package
is being expanded to encompass a wide range of elastic stress and strain
solutions as time permits. All data and scripts for this particular
project are available at
\url{https://github.com/cossatot/wenchuan_topo_stress}.

\begin{table}
\centering
\begin{tabular}{l c l}
\hline
Parameter & Value & Unit \\
\hline
horizontal spacing & 851 & m \\ 
vertical spacing & 1000 & m \\ 
minimum depth & 851 & m (below sea level) \\ 
maximum depth & 35851 & m (below sea level) \\ 
density ($\rho$) & 2700 & kg m$^{-3}$ \\ 
g & 9.81 & m s$^{-2}$ \\ 
Green's function radius & 9e5 & m \\ 
Poisson ratio & 0.25 & - \\ 
\hline
\end{tabular}
\caption{Parameters for numerical calculations of topographic stresses.}
\label{table:convo_params}
\end{table}

\subsection{Topographic fault stress
calculations}\label{topographic-fault-stress-calculations}

Topographic stresses on the Wenchuan faults are calculated on point sets
representing the faults taken from coseismic slip models. We use the
coseismic slip models of \citet{shen2009},
\citet{feng2010}, \citet{qi2011}, \citet{zhang2011},
and \citet{fielding2013}, and discard
points above 851 m below sea level, as this is above the depth at which
we compute $G^C(x,y,z)$. The six stress tensor components calculated at
the regular grid point are linearly interpolated to points describing
the faults. Because the fault points are completely surrounded by the
grid nodes at which topographic stresses were calculated and those nodes
are spaced \textless{}1 km apart, the fault points cannot be more than a
few hundred meters from the nearest grid node, so a higher order
interpolation is not necessary. We then project the topographic stress
tensor to fault normal stress, $\sigma_n^M$, down-dip shear stress,
$\tau_d^M$ and strike-slip shear stress, $\tau_s^M$ at each point in
describing the fault geometry.

\section{Results of topographic stress calculations on the Wenchuan
faults}\label{results-of-topographic-stress-calculations-on-the-wenchuan-faults}

Topographic stresses on the Wenchuan faults are in the 1--10s of MPa
range (Figure~\ref{fig:lms_topo_stresses_rot}, ~\ref{fig:fault_stress_3d}). 
Stresses are highest in the southwest,
beneath the Pengguan massif (the highest topography of the Longmen Shan
front), and decrease to the northeast. $M_{zz}$ is typically larger,
though not substantially, than $M_{xx}$ or $M_{yy}$. Maximum horizontal
stress is not typically aligned with either cardinal horizontal
component, and is typically larger than $M_{zz}$ above 10 km. Maximum
$M_{zz}$ is near 80 MPa, on the southwestern Beichuan fault below the
high peaks of the Pengguan massif, except for in slip models containing
near-horizontal fault segments in the mid-crust, where $M_{zz}$ reaches
100 MPa. Vertical shear stresses ($M_{xz}$ and $M_{yz}$) are on the
order of 1 MPa, and horizontal shear stress ($M_{xy}$) is on the order
of 0.1 MPa.

\begin{figure}[t]
\centering
\includegraphics[width=20pc]{../figures/lms_topo_stresses_rot.pdf}
\caption{Horizontal topographic stresses in the Longmen Shan region at 5
km depth: black and red lines signify most and least compressive
horizontal stresses. Other symbols are the same as in Figure
1. Stresses shown are down-sampled from the
discretization used in the calculations by a factor of 6.}
\label{fig:lms_topo_stresses_rot}
\end{figure}

Because the compressive stresses are near equal, $M$ contains a large
isotropic component and a smaller shear component. Consequently, $M$
resolves on the Wenchuan faults with a large $\sigma^M_n$ (median of
about 40-60 MPa for each slip model) and much smaller $\tau^M_d$. The
median $\tau^M_d$ is about -3 to -6 MPa in each slip model, where values
less than zero indicate normal-sense shear, and $\tau^M_s$ median values
range from about -2 to 1 MPa, where values less than zero indicate
sinistral shear; fault models with positive median $\tau^M_s$ have broad
thrust flats at depth, where little slip occurred. On the steeper fault
segments (where most of the moment release occurred), topographic shear
stresses are typically normal-sinistral, as opposed to the dominant mode
of coseismic slip, which is reverse-dextral.

\begin{figure*}[htbp]
\centering
\includegraphics[width=40pc]{../figures/fault_stress_3d.pdf}
\caption{Southwest-looking views of topographic stresses and coseismic
slip on selected slip models. All models share same lateral extent, but
the perspectives for B and C are more inclined than A. (A) $\sigma_n^M$
(colors), slip magnitude (contours, 1 m interval) and hanging-wall
topography on the \citet{feng2010} model of the Beichuan
fault. Note the suppression of fault slip where normal stress is
highest, such as below the Pengguan massif (P). Fault and topography
share the same scale. No vertical exaggeration. (B) $\tau_d^M$ (colors)
and dip slip (contours, 1 m interval) on the \citet{qi2011}
`rough' slip model. (C) $\tau_s^M$ (colors) and strike slip (contours, 1
m interval) on the \citet{qi2011} `rough' slip model.}
\label{fig:fault_stress_3d}
\end{figure*}

In contrast to the steeper fault segments, much of the shallowly-dipping
fault segments (the Pengguan fault and flats at the base of the Beichuan
fault, where present) have $\tau^M$ in the direction of coseismic slip.
The full stress tensor $M$ is not significantly different in these
locations, but because of the low dip angle, $M_{zz}$ contributes more
significantly to $\sigma^M_n$ than to $\tau^M_d$, which is then
dominated by horizontal compression, leading to reverse-sense shear.
Additionally, the stresses caused by the Pengguan massif locally resolve
as right-lateral on these segments as well. Coseismic slip on these
fault patches is much lower than on the steeper Beichuan fault, where
the majority of slip occurred and which is topographically loaded in the
opposite shear sense.

Compellingly, similar patterns exist in the spatial distributions of
$\sigma^M_n$ and coseismic slip. Most obvious is the coincidence of
locally high $\sigma^M_n$ and locally low slip magnitude on the
southwestern Beichuan fault below the culmination of the Pengguan
massif, in an area of otherwise high slip (Figure
\ref{fig:fault_stress_3d}). These correlations exist for other fault
patches, but they are not as clear (Figure
\ref{fig:feng_slip_sig_n_scatter}). This raises the possibility that
$\sigma^M_n$ is capable of limiting slip once failure has occurred, and
may have implications for estimations of dynamic friction and the
completeness of stress drop during the earthquake. Preliminary analysis
of this is currently being performed, and will be described in a
forthcoming manuscript.

\begin{figure}%[htbp]
\centering
\includegraphics[width=20pc]{../figures/feng_slip_sig_n_scatter.pdf}
\caption{Coseismic slip magnitude and $\sigma^M_n$ on the four segments
of the \citet{feng2010} coseismic slip model. Trendlines are
L1 regressions and do not include points with no slip. NE B =
Northeastern Beichuan fault. C B = central Beichuan fault. SW B =
Southwestern Beichuan fault. P = Pengguan fault.}
\label{fig:feng_slip_sig_n_scatter}
\end{figure}

\section{Calculations of tectonic stress, fault friction and pore fluid
pressure}\label{calculations-of-tectonic-stress-fault-friction-and-pore-fluid-pressure}

Faults fail in earthquakes when the shear stresses on the fault overcome
the frictional stresses resisting slip on the fault \citep[e.g.,][]
{scholz2002}. We assume that the entire fault was at at the point of
failure when the Wenchuan fault initiated, and use the Mohr-Coulomb
failure criterion

\begin{equation} 
\tau = \mu ( \sigma_n - \sigma_p ) \; ,
\label{eqn:amonton_raw} 
\end{equation}

where where $\mu$ is the coefficient of static friction on the fault and
$\sigma_p$ is the pore fluid pressure \citep[e.g.,][]{sibson1985}. We
describe the pore fluid pressure using a scalar, $0 \leq \phi \leq 1$,
which is the pore fluid pressure as a fraction of total pressure, and so
the failure criterion is

\begin{equation} 
\tau = \mu (1 - \phi) \sigma_n \; 
\label{eqn:amonton} 
\end{equation}

\citep[e.g.,][]{sibson1985}. We assume that both $\mu$ and $\phi$ are
constant across the Wenchuan faults.

We estimate the tectonic stress tensor field, $\mu$, and $\phi$
consistent with published coseismic slip models of the Wenchuan
earthquake using a Bayesian estimation. Bayesian estimation yields
posterior probability density functions of the model parameters \citep[e.g.,]
[]{mosegaard1995}. We first estimate posteriors of $T$ cosistent with
the coseismic slip models, and then estimate $\mu$ and $\phi$ consistent
with Mohr-Coulomb failure. The nature of Bayesian estimation allows us
to quantify both the relative likelihoods of model parameters and the
tradeoffs between them.

\subsection{Description of the stress
state}\label{description-of-the-stress-state}

We consider the complete stress tensor, $S$, at a point in the crust to
be

\begin{equation}
S = M + T + L \; ,
\end{equation}

where $M$ is described above, $T$ is the tectonic stress tensor, and $L$
is the lithostatic pressure tensor. $L$ is isotrapic, with diagonal
components equal to $\rho g z$. We assume that $T$ is latterally
homogeneous and only has horizontal stress components (i.e.,
$T_{xz} = T_{yz} = T_{zz} = 0$, with $T_{xx}$, $T_{yy}$ and $T_{xy}$
non-zero). We further assume that $T$ increases linearly with depth so
that the entire upper crust is near the critical failure envelope
\citep[e.g.,][]{townend2000}, are thus we parameterize the components of $T$
as scalars multiplied by lithostatic pressure.

\subsection{Bayesian inversion of tectonic
stresses}\label{bayesian-inversion-of-tectonic-stresses}

We invert topographic stresses and coseismic slip models for tectonics
stresses using Bayesian methods, and making the common `Wallace-Bott'
assumption (named after \citet{wallace1951} and \citet{bott1959}) 
that slip on the fault occurs in the general direction
of the maximum resolved shear stress on the fault \citep[e.g.,][]
{mckenzie1969, angelier1994}. We estimate the tectonic stresses in
light of the topographic stresses and slip distributions through the
relation

\begin{equation} 
p(T|D) \propto p(T) \, p(D|T) \; , 
\label{eqn:bayes_rule} 
\end{equation}

where $p(T)$ is the prior PDF (or \emph{priors}) of $T$, $p(D|T)$ is the
likelihood of observing the coseismic slip distribution $D$ given the
tectonic stresses $T$, and $p(T|D)$ is the posterior PDF of $T$ given
$D$, which is the solution to the inversion \citep[e.g.,][]{mosegaard1995}.
Due to the unknown proportionality in equation (\ref{eqn:bayes_rule}),
our posterior only gives likelihood of $T$ relative to the most likely
estiamte (MLE) \citep{tarantola2005}. We follow a Monte-Carlo strategy,
where samples of the prior PDF are retained as samples of the posterior
in proportion to $p(D|T)$ \citep[e.g.,][]{mosegaard1995}.

We parameterize $T$ by the magnitudes and orientation of the maximum and
minimum principal tectonic stresses. We assume priors such that the
magintudes of principal tectonic stresses are equally likely within
bounds and that all orientations are equally likely. Because the
Wenchuan event was an oblique reverse faulting earthquake, we assume
that total horizontal stresses are greater than the vertical stress,
which is satisfied if the tectonic stresses are positive. Prior samples
of maximum principal stress are taken from a uniform distribution
between $\rho g z$ and 2.5 $\rho g z$. Samples of the minumum principal
stress are from a uniform distribution between 0 and the value for
maximum stress. We describe the orientation of the tectonic stress using
the azimuth of the maximum tectonic stress, which are sampled uniformly
from 0 to 360$^{\circ}$.

We test 100,000 unique samples drawn from the prior using a seeded
pseudorandom number generator. We test the same prior samples against
each of the coseismic slip models. $S$ is then constructed for each
point discretizing the fault geometries in the coseismic slip models.
The rake of the maximum shear stress $\lambda^S$ on each point of the
fault is calculated and compared to the coseismic slip rake $\lambda^D$
at that point. A weighted mean misfit is calculated by

\begin{equation}
\bar{\lambda}^m = \sum \nolimits_{i=1}^n \frac{(\lambda^S_i - \lambda^D_i) D_i} 
{\bar{ D}} \; ,
\label{eqn:rake_misfit}
\end{equation}

where $D$ is the coseismic slip and $\bar{D}$ is the average coseismic
slip in a given coseismic slip model. Finally, the relative likelihood
of each model is computed using a Von Mises distribution as

\begin{equation}
p (D | T) = \frac{ \exp ( \kappa \cos \bar{\lambda}^m )} 
{\exp (\kappa \cos \bar{\lambda}^m_{\min})} \;,
\label{eqn:rel_likelihood}
\end{equation}

where $\kappa$ = 8.529, which is calculated so that the 68.2\%
confidence interval the Von Mises distribution is within $\pi$/9 radians
(20$^{\circ}$), the estimated 1$\sigma$ uncertainty of the coseismic
slip models based on comparisons between rakes of high-slip fault
patches (note that for a planar fault, $\tau$ at $\pi/9$ radians from
$\lambda_{max}$ is still \textgreater{}90\% of $\tau_{\mathrm{\max}}$
\citep{lisle2013}). Prior samples are retained in propotion to $p(D|T)$,
and the retained samples are then samples of the posterior.

\subsection{Analysis of friction and pore fluid
pressure}\label{analysis-of-friction-and-pore-fluid-pressure}

Once the tectonic stress distributions consistent with the coseismic
slip models have been determined, we deduce the distributions of $\mu$
and $\phi$ assuming that the stress is at the failure criterion in
equation (\ref{eqn:amonton}). We do this in three steps: First, we draw
a random $\phi$ from a uniform distribution, assuming $0 \leq \phi < 1$.
We again use a seeded pseudorandom number generator, such that each
stress model has a uniquely assigned $\phi$ that is consistent priors
accross all coseismic slip models. Second, we calculate $\tau^S$ and
$\sigma_n^S(1-\phi)$ for each point on the fault. Third, we solve
Equation \ref{eqn:amonton} for $\mu$. Finally, we filter the results so
that only models with $0 \le \mu < 1$ are retained, as values outside of
that range have not been suggested for rocks.

After this analysis has been done for all coseismic slip model, we find
the joint posterior (i.e., the posterior consistent with all of the
coseismic slip models) by taking the samples that are common to all of
the individual posteriors. We denote the joint posterior as
$p_{\mathrm{J}}(P | D)$, where $P$ is the parameter of interest.

\section{T, $\mu$, $\phi$ results}\label{t-mu-phi-results}

\subsection{Individual slip models}\label{individual-slip-models}

Results for $T^\prime$, $\mu$ and $\phi$ are quite consistent across all
coseismic slip models (Figure \ref{fig:T_scatters}). Maximum compressive
stress $T^\prime_{\mathrm{max}}$ is broadly east-west for all models,
with a mode trending at $90^{\circ} -- 105^{\circ}$.
$p(T^\prime_{\mathrm{max}}|D)$ for each slip model increases from
$T^\prime = 0$ to 0.5 or 1 before essentially leveling off, though some
slip models, particularly the \citet{qi2011} model, show a
slight decrease in relative likelihood past the initial mode at
$T^\prime =$ 0.5--1 (recall that $T^\prime$ is relative to $\rho g z$).
The low likelihood below $T^\prime \approx 0.5$ indicates that lower
tectonic stresses are unlikely to overcome fault friction and
topographic shear stresses resisting reverse-dextral slip on the
Wenchuan faults. $p(T^\prime_{\mathrm{min}} | D)$ for each slip model
has a mode close to $T^\prime_{\mathrm{min}} = 0.2 $ and decreases
abruptly at higher values, though all slip models show values for
$T^\prime_{\mathrm{min}}$ up to 2.5. $T^\prime_{\mathrm{min}}$ is
typically 0--0.4 of $T^\prime_{\mathrm{max}}$, but rarely higher.

\begin{figure*}%[htbp]
\centering
\includegraphics[width=40pc]{../figures/T_scatters.pdf}
\caption{Samples of
$p(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$ associated with
each of the coseismic slip models we consider, along with marginals of
$T^\prime_{\mathrm{max}}$ and $T^\prime_{\mathrm{min}}$. Inset rose
diagrams are histograms of azimuth of
$T^\prime_{max}$.}
\label{fig:T_scatters}
\end{figure*}

All slip models show $p(\phi | D)$ to be uniformly high from $\phi = 0$
to 0.4--0.6 and to decrease linearly to $p(\phi) = 0$ at $\phi = 1$
(Figure \ref{fig:mu_phi_scatters}). $p(\mu | D)$ for all slip models has
a mode at $\mu =$ 0.1--0.4 and $p(\mu)$ decreases at higher values.
$T^\prime_{\mathrm{max}}$, $\phi$ and $\mu$ are highly correlated, where
higher values of $T^\prime_{\mathrm{max}}$ are associated with higher
$\mu$ and lower $\phi$. Combinations of high $\mu$ and low $\phi$
require much higher $T^\prime_{\mathrm{max}}$ to overcome fault friction
and cause slip, and so are not represented in the posteriors. Since our
maximum $T_{\mathrm{max}}$ of $2.5 \rho g z$ is quite high
($\approx 660$ MPa at 10 km), we view high $\mu$ and low $\phi$
combinations as unrealistic for the Wenchuan faults. Similarly,
combinations of very low $\mu$ and very high $\phi$ are associated with
very low $T^\prime_{\mathrm{max}}$, and have a low probability density,
as it is unlikely that very low $T^\prime_{\mathrm{max}}$ values can
overcome sinistral and normal-sense topographic shear stresses to cause
the observed coseismic slip kinematics.

\begin{figure*}[htbp]
\centering
\includegraphics[width=40pc]{../figures/mu_phi_fms.pdf}
\caption{Samples of $p(\mu,\phi | D)$ for each coseismic slip model.
Colors indicate magnitude of $T^\prime_{\mathrm{max}}$. Contour lines
indicate relative density (i.e., likelihood) of posteriors; darker lines
mean higher densities.}
\label{fig:mu_phi_scatters}
\end{figure*}

\subsection{Joint posteriors}\label{joint-posteriors}

We define a joint posterior, $p_{J}(T^\prime | D)$, by the samples of
the individual posteriors estimated from each slip model that common.
Unsuprisingly, given the broad similarity between the posteriors from
the various slip models, $p_{J}(T^\prime_{\mathrm{max}} | D)$ is not
substantially different from any of the constituent model posteriors.
$p_{J}(T^\prime_{\mathrm{max}} | D)$ has a somewhat more well-defined
mode at $T^\prime_{\mathrm{max}} \approx 0.6$.

In our estimation of $\phi$ and $\mu$ in the posteriors associated with
each slip model, we have used the same random combinations of $T$ and
$\phi$ for each slip model, and then solved for $\mu$ so that the fault
is at a critical stress state (Equation \ref{eqn:amonton}). Because of
differences in the location and slip among the coseismic slip models,
some variability exists in $\mu$ for each prior sample. We therefore
choose $p_{\mathrm{J}}(\mu | D)$ to be the median $\mu$ of each slip
model for each sample. $p_{\mathrm{J}}(\mu | D)$ has a mostly similar
distribution as $p(\mu | D)$ for any of the slip models. However,
$p_{\mathrm{J}}(\mu | D)$ has a lower relative likelihood on the
high-$\mu$ tail compared to the constituent $p(\mu | D)$. This lower
likelihood of $\mu$ in the joint posterior is likely becuase it is the
average $\mu$ of all slip models. On the other hand, it is not similarly
sparse on the low-$\mu$ side, suggesting that low values for $\mu$ are
more robust.

\begin{figure*}[t]
\centering
\includegraphics[width=40pc]{../figures/joint_pdfs.pdf}
\caption{Joint model results. (A) Samples of
$p_{\mathrm{J}}(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$,
along with marginals of $T^\prime_{\mathrm{max}}$ and
$T^\prime_{\mathrm{min}}$. (B) Samples of
$p_{\mathrm{J}}(T^\prime_{\mathrm{max}}, T^\prime_{\mathrm{min}}/T^\prime_{\mathrm{max}} | D)$,
along with marginals of $T^\prime_{\mathrm{max}}$ and
$T^\prime_{\mathrm{min}}T^\prime_{\mathrm{max}}$. (C) Histogram of
azimuths of $T^\prime_{mathrm{max}}$. (D) Samples of
$p_{\mathrm{J}}(\mu, \phi | D)$, with marginals. Colors in the
scatterplot indicate magnitude of $T^\prime_{\mathrm{max}}$. Contour
lines indicate relative density (i.e., likelihood) of posteriors; darker
lines mean higher densities.}
\label{fig:joint_posteriors}
\end{figure*}

\section{Discussion}\label{discussion}

Few studies have performed similar quantification of static stress
fields on faults (see Section
\ref{previous-work-on-topographic-stresses} for some examples). This is
important to address because most studies of fault rupture dynamics
assume either a homogeneous or stochastic shear stress distribution
\citep[e.g.,][]{oglesbyday2002} and few assume any variation in normal
stress \citep[e.g.,][]{aagaard2001}, despite the importance that stress
variations likely have in earthquake dynamics \citep[e.g.,][]{day1982,
olsen1997}. Additionally, quantifying friction and pore fluid
pressure involved in faulting is a major challenge in studies of
faulting and orogenic dynamics \citep[e.g.,][]{meissner1982,
oglesbyday2002}.

Previous workers have demonstrated that by quantifying topographic
stress, other components in the Coulomb stress balance may be bracketed
\citep[e.g.,][]{cattin1997, lamb2006, luttrell2011}. Each of these studies
uses somewhat different approaches. Our approach is most similar to that
of \citet{luttrell2011}, although there are significant
differences: (1) We use the topographic stresss calculations proposed by
\citet{liuzoback1992}, whereas 
\citet{luttrell2011} only uses the vertical loading from topography,
equivalent to $M^B$ in equation (\ref{eqn:bous}). (2) We do not consider
bouancy forces due to lateral variations in density, due for instance to
Moho variation, as done by \citet{luttrell2011}. In the
Longmen Shan region, Moho variation is small compared to the change in
Moho depths at the Andean plate boundary, and so the buoyancy terms
should be relatively small. (3) We consider a full range of tectonic
stresses, instead of simply calculating the minimum principal tectonic
stress and its orientation. (4) We consider the stress tensor at each
point due to topographic loading, lithostatic stress, and horizontal
tectonic stress, and use inferred coseismic slip rake as a constraint on
the allowable stresses, rather than inferring the earthquake stress drop
from the coseismic slip models, as done by 
\citet{luttrell2011}. (5) We use both normal and shear stresses to
constrain pore fluid pressure and friction. (6) We use a Bayesian
estimation, resulting in a PDF of tectonic stress, as well as $\mu$ and
$\phi$, rather than solving for just the minimum tectonic stress
required for faulting, as done by \citet{luttrell2011}.

\subsection{Topographic stresses on the Wenchuan
faults}\label{topographic-stresses-on-the-wenchuan-faults}

Topographic stresses on the main Wenchuan faults are of considerable
magnitude: $\tau^M$ ranges from about -20 to 10 MPa, which is on the
order of inferred stress drop in earthquakes \citep[e.g.,][]{kanamori1975,
allmann2009}. Topographic stresses are
generally opposed to the tectonic slip direction, and therefore have to
be overcome by tectonic stresses in order to produce the observed
rupture patterns. The topographic shear stresses are likely persistent
over the lifespan of the topography (i.e.~on the order of millions to
tens of millions of years), otherwise the Wenchaun faults may fail in a
normal sense simply due to the weight of the Longmen Shan. This would
argue that tectonic stress drop in the Wenchuan earthquake was not
complete, as some residual tectonic shear stress must remain on the
fault to cancel out $\tau^M$.

The spatial variation in topographic stresses decreases in frequency and
amplitude with depth. This is not surprising, because with increasing
depth, the stress field at any point is more sensitive to surface loads
averaged over a greater region, and is less dominated by smaller scale
topographic features (i.e., individual mountains). The spatial
variablity of the topography stress with depth is similar to the spatial
variablity of coseismic slip in the models considered \citep[e.g.,][]
{zhang2011}, which are similarly more smooth at depth. Some of the
estimated slip variability is likely partially due to the more limited
detectability of displacements at depth versus near the surface.
However, the negative spatial correlations of slip versus stress
(especially $\sigma^M_n$) (Figures \ref{fig:fault_stress_3d},
\ref{fig:feng_slip_sig_n_scatter}) suggest this may be a real signal.

\subsection{Tectonic stresses in eastern
Tibet}\label{tectonic-stresses-in-eastern-tibet}

The maximum tectonic stress, $T_{\mathrm{max}}$, is consistently
oriented roughly E-W in our results. This orientation is oblique to the
Longmen Shan, which produces oblique (right-lateral and reverse sense)
shear on the Beichuan fault. $T_{\mathrm{min}}$ is \textasciitilde{}N-S
oriented, and is only slightly larger than lithostatic pressure. This
stress configuration is compatible with the observed kinematics of the
Wenchuan faults, and in close agreement pre-earthquake stress
measurements near the rupture zone (Figure \ref{fig:lms_stress_map}),
mostly from borehole breakout data from 2-5 km depth
\citep{heidbach2009}, which is the zone of maximum slip in the coseismic
slip models. It is somewhat discrepant with stress orientations
estimated at \textasciitilde{}800 m depth adjacent to the Beichuan fault
in the WFSD-1 borehole of the Wenchuan Earthquake Fault Scientific
Drilling Project several years after the 2008 earthquake \citep{cui2014},
which show $\sigma_H {\mathrm{\max}}$ to be more orthogonal to the fault
trace, suggesting that much of the right-lateral component of shear
stress was released during the earthquake. These results are similar to
the orientations of total stress obtained by 
\citet{medinaluna2013}, although they were unable to constrain the
magnitudes of stresses.

\begin{figure}[t]
\centering
\includegraphics[width=20pc]{../figures/lms_map_stresses_rot.pdf}
\caption{Topographic and tectonic horizontal stresses (taken from the
most likely estimates of $p(T|D)$ in the Wenchuan rupture region (black
and red crosses) with horizontal maximum stress orientation data taken
from before the 2008 Wenchuan event from the World Stress Map
\citep{heidbach2009} (purple arrows), and horizontal maximum stress
orientation data from after the earthquake at the WFSD-1 drill hole
\citep{cui2014} (blue arrows). Other symbology is the same as in Figure
\ref{fig:lms_map}. Stresses shown are downsampled from grid resolution
by a factor of 9.}
\label{fig:lms_stress_map}
\end{figure}

The magnitudes of $T_{\mathrm{max}}$ and $T_{\mathrm{min}}$ are
dominantly constrained on the low end by our analysis, which is apparent
by the sharp decrease in the frequency of $p(T_{\mathrm{max}}|D)$ below
about $T^\prime = 0.5$. This indicates that $T_{\mathrm{max}}$ of at
least \textasciitilde{}13.25 MPa km$^{-1}$ is necessary to overcome
topographic stresses resisting reverse and right-lateral slip on the
faults. We find that the highest likely ratio of strike slip to dip slip
shear along the Wenchuan earthquake faults is close to 1. This is
similar to the inferences of strain accumulation rates inferred from
squishy block modeling by Loveless and Meade \citet{loveless2011}, who
concluded that the rate of slip deficit accumulation in the thrust and
dextral sense was approximately equal. It should be noted that the
tectonic stresses we estimate here represent the accumulated stresses
prior to the Wenchuan earthquake, and the relation of these stresses to
accumulated slip deficit needs to be through a model of strain
accumulation.

The orientation of both the tectonic and total stresses near the
Wenchuan faults shows a larger difference with patterns of strain from
elsewhere in the orogen. For example, the presence of N-S contraction
and E-W extension throughout the high Tibetan plateau and much of the
Himalaya \citep[e.g.,][]{armijo1986, molnar1988, taylor2003} indicates a
roughly N-S $T_{\mathrm{max}}$ and E-W $T_{\mathrm{min}}$. Because
$T_{\mathrm{min}}$ is only slighly above lithostatic pressure on the
Wenchuan faults, it is quite possible that the N-S compression in the
Himalaya and Tibet, which is almost certainly due to Indo-Asian plate
collision, has almost completely decayed at the Longmen Shan, some 850
km northeast of the easternmost Himalaya. Therefore, contraction across
the Longmen Shan cannot easily be interpreted to directly reflect
stresses due to the Indo-Asian collision alone, unless some additional
mechanism of redirecting crustal stresses is incorporated.

\subsection{Slip on the Beichuan fault vs.~optimally-oriented
faults}\label{slip-on-the-beichuan-fault-vs.optimally-oriented-faults}

Our highest likelihood estimates of $\mu$ are in the range of 0.2--0.3.
These values are slightly lower than $\mu \approx 0.4$ inferred in
laboratory experiments on samples recoverd from the WFSD-1 drill hole
into the Beichuan fault \citep{kuo2014}. However, it should be noted that
$\mu \approx 0.4$ is likely relatively high likelihood in our posteriors
(Figure \ref{fig:joint_posteriors}). These values of friction, however,
are lower than typical values derived from laboratory experiments on
intact rock \citep[e.g.,][]{byerlee1978}, suggesting that slip occurred on
preexisting faults because they are weaker than optimally-oriented new
faults.

The obliquity of slip on the Wenchuan earthquake faults also suggests
that these faults may not be optimally oriented for slip given the total
stress state in the region of these faults. However, the Longmen Shan
fault zone dates back to the Indosinian orogeny, locally late Triassic
(226-206 Ma) \citep{yong2003} and has had multiple episodes of
reactivation since \citep[e.g.,][]{burchfiel1995, wang2012}, accumulating
tens of kilometers of shortening \citep[e.g.,][]{hubbard2010}. Such a mature
fault system may be expected to have low coefficients of friction due to
processes such as gouge development \citep[e.g.,][]{kuo2014}, and therefore
may slip in non-optimal orientations, with high $\phi$ values also
potentially contributing to this \citep[e.g.,][]{sibson1985}.

Our posterior estimates for $T$, $\phi$ and $\mu$ let us quantitatively
evaluate to what extent slip on the Wenchuan faults is more favorable
than on optimally oriented faults with more typical friction
coefficients. We use the same failure conditions as in determing the
posteriors above, and that optimally oriented faults exist.
Additionally, we evaluate the relative contributions of $T$, $\phi$ and
$\mu$ on potential fault weakening and reactivation. To explore these
relationships, we perform some preliminary analysis on a single fault
model (from \citet{zhang2011}) using a subset of 1000
samples drawn randomly from the joint posteriors. Given the similarity
of the fault models and of the posteriors for each model, we do not
expect that an analysis of all results on all fault models will yield
different conclusions.

First, we establish a metric with which to evaluate the favorability of
slip on a given fault plane, which we call the Coulomb failure ratio, or
CFR:

\begin{equation}
\mathrm{CFR} = \tau / \mu (1 - \phi) \sigma_n \; .
\label{eqn:cfr}
\end{equation}

CFR indicates whether a fault should fail under a given stress state:
CFR \textgreater{} 1 indicates failure, while a CFR \textless{} 1
indicates fault stability. We then calculate the CFR on each point in
the model of the Beichuan fault (594 points describe the fault model of
\citet{zhang2011}) based on the full stress field $S$ at
each point on the fault, for each of the 1000 samples of $T$, $\phi$ and
$\mu$ drawn randomly from the posteriors. We call this CFR$_f$. Then,
using the same $S$ and $\phi$, we calculate the CFR on an optimally
oriented fault with $\mu=0.6$ and no cohesion, which we call CFR$_o$.
Note this $\mu$ is typical for crustal rocks with fault normal stresses
above 200MPa, but lower than $\mu = 0.85$ for smaller $\sigma_n$
\citep{byerlee1978}, but may be appropriate for an immature crustal
fault.

Figure \ref{fig:cfr_ratios} shows
$log (\mathrm{CFR}_o / \mathrm{CFR}_f)$ plotted against $\mu$, $\phi$
and $T^\prime_{xx}$ on the Beichuan fault for all samples. Though
considerable scatter exists, it is clear that in most instances, slip on
the Beichuan fault is preferred over slip on an optimal fault. The
exceptions are at high values of $\mu$, $\phi$, or $T^\prime$, where
slip on an optimal plane is preferred. Because $T$, $\phi$ and $\mu$ can
all affect fault reactivation \citep[e.g.,][]{sibson1985}, we compare the
relative contributions of each with a simple mutliple linear regression,
using $T^\prime_{xx}$ normalized to [0,1) (the same range $\phi$ and
$\mu$) as a proxy for $T$ ($T^\prime_{xx}$ is essentially
$T^\prime_{\mathrm{max}}$ in most of the posteriors). The results are
shown in Table \ref{table:cfr_regress}. It is clear that $\mu$ is most
strongly correlated with $\mathrm{CFR}_o/\mathrm{CFR}_f$, followed
closely by $T^\prime_{xx}$ and then $\phi$; nonetheless, all
significantly affect the relative ease of faulting on the Beichuan fault
versus optimal faults.

\begin{figure}%[tb]
\centering
\includegraphics{../figures/cfr_plots.pdf}
\caption{Comparison of Coulomb failure ratio (CFR) on the Beichuan fault
from the \citet{zhang2011} model to CFR on an
optimally-oriented fault with $\mu = 0.6$, versus estimated $\mu$ on the
Beichuan fault. Values less than 0 (1 in linear space) indicate that
slip is favored on the Beichuan fault, even if it is not optimally
oriented. Values are calculated for each point in the slip model for
1000 randomly-drawn samples from $p(T,\mu,\phi |D)$.}
\label{fig:cfr_ratios}
\end{figure}

\begin{table}[htb]
\centering
\begin{tabular}{l c c c c}
\hline
Parameter & Coef. & Std. Err. & t statistic & $P>|t|$ \\
\hline
Intercept & -0.2566 & 0.0020 & -128.8 & \textless{}0.0001 \\ 
$\mu$ & 1.1168 & 0.0086 & 130.6 & \textless{}0.0001 \\ 
$T^\prime_{xx}$ & 0.9138 & 0.0048 & 191.2 & \textless{}0.0001 \\ 
$\phi$ & 0.5762 & 0.0055 & 105.7 & \textless{}0.0001 \\ 
\hline
\end{tabular}
\caption{Sensitivity of CFR ratios to relevant stress state parameters:
Results of multivariate linear regression of
$\mathrm{CFR}_o/ \mathrm{CFR}_f)$ against $\mu$, $\phi$ and
$T^\prime_{xx}$.}
\label{table:cfr_regress}
\end{table}

\subsection{The role of topography in orogenic development and strain
localization}\label{the-role-of-topography-in-orogenic-development-and-strain-localization}

Because the rise of broad elevated regions creates significant stresses
in the crust \citep[e.g.,][]{jeffreys1924}, these stresses have the
potential to influence orogenesis, from changing the deformation state
in the interior of an orogen \citep[e.g.,][]{dewey1988, molnar1988}, the
convergence rates of the plates surrounding the orogen \citep[e.g.][]
{meade2008}, or the location and style of deformation at the orogen's
margins \citep[e.g.,][]{beaumont2001, decelles2009}. These studies show that
as an orogen rises, horizontal contraction via crustal thickening will
occur at the thin margins of the orogen, and if elevations reach a
threshold or tectonic compression decreases, extension will take place
in the orogen's high interior, as is well described in Tibet. Therefore,
Tibet is commonly suggested to be undergoing some manner of
gravitational collapse \citep[e.g.,][]{england1989}; where the plateau abuts
rigid cratons, the deformational patterns resemble `fixed boundary'
collapse \citep{rey2001}, with concentrated crustal thickening at the
orogen's margins \citep[e.g.,][]{cook2008}.

Though much work has been done exploring the feedback mechanisms between
topography and deformation, topography is typically greatly smoothed,
stresses from it are folded into gravitational potential energy
estimates of the entire lithosphere, and the whole lithospheric column
is often considered viscous \citep{copleymckenzie2007}. Therefore many
models seeking to predict deformation from gravitational forces yield
continuous deformation and do not consider how the gravitational
stresses resolve on heterogeneities embedded in the crust. However, the
presence of structures such as weak faults \citep{bird1994} or
low-viscosity channels or shear zones \citep[e.g.,][]{clark2005} can localize
deformation \citep[e.g.,][]{bird1994, fleschbendick2012}.

Our results indicating that topographic stresses on the Wenchuan
earthquake faults are largely loaded in the opposite sense to the
coseismic slip direction shows that gravitational collapse is probably
not the driver of reverse faulting in the Longmen Shan. However, as
noted previously, thrust flats in the middle crust present in some
coseismic slip models, particularly the Qi model, are loaded in a thrust
and dextral sense. Given the very low dips of the thrust flats
($0^{\circ}$--$5^{\circ}$), they receive very little loading from
horizontal tectonic stress, so any slip on them may be in response to
topographic loading and therefore to gravitational collapse of the high
Longmen Shan and its hinterland. $M_{H{\mathrm{\max}}$ also changes
orientation from fault (or rangefront) normal to closer to more oblique
near the front of the range (Figure \ref{fig:lms_topo_stresses_rot}),
suggesting that any transfer of mass from the highlands to the lowlands
due to topographic stresses should terminate near the rangefront.

However, because $M_{zz}$ is greater than $M_{H{\mathrm{max}}$
underneath the higher topography, topographic stresses can only lead to
reverse faulting on the Wenchuan faults if subhorizontal shear zones or
channels are present, and are very weak so shear can occur at relatively
low $\tau$ relative to $\sigma_n$. This conclusion has been reached in
studies of generalized orogens \citep[e.g.,][]{fleschbendick2012} and
gravitational collapse of volcanic edifices \citep[e.g.,][]{byrne2013},
indicating that is a general requirement of gravitational spreading. Our
calculations of $\mu$ and $\phi$ are based on $T$, which is weighted
towards the high slip patches in the upper crust (Equation
\ref{eqn:rake_misfit}), yield low $\mu$ and low to moderate $\phi$,
which is probably insufficient for slip at high pressures at depth,
though $\phi$ may be higher there. Similarly, a low viscosity horizontal
channel that underlies eastern Tibet, likely much deeper than the
seismogenic zone we consider here, would be able to flow in response to
$\tau$ regardless of $\sigma_n$ and may be able to facilitate
gravitational spreading of the orogen \citep[e.g.,][]{clark2005, cook2008,
fleschbendick2012}.

This is all quite speculative, and we wish only to describe the
conditions under which gravitational collapse can occur given the
observations of deformation and the topographic stress field. It is not
at all clear that a suitably large decollement exists at the base of the
Beichuan fault (it is only present in one of the slip models
considered). Nonetheless, this topic has implications not only for
orogenic development, but for the recurrence interval on the Wenchuan
earthquake faults. If topographic stress contributes in some fashion to
the observed displacements on the fault, then those stresses were likely
barely diminished by coseismic stress change, and earthquake recurrence
of on the Wenchuan faults may be governed by different processes than
elastic rebound due to tectonic strain accumulation.

\section{Conclusions}\label{conclusions}

We have calculated shear and normal stresses due to topographic loading
on the Wenchuan earthquake faults, and used those stresses to constrain
tectonic stresses, fault friction and pore fluid pressure. Topographic
stresses on the main Wenchuan faults are large, with $\tau^M$ on the
faults up to \textbar{}20\textbar{} MPa, and $\sigma^M_n$ up to 80 MPa.
$\sigma^M_n$ reaches up to 100 MPa on mid-crustal thrust flats present
in some coseismic slip models. These topographic stresses generally
resist the slip inferred during the 2008 Wenchuan earthquake, with
$\tau^M$ opposed to the slip direction over much of the fault,
indicating that weight of the topography resist coseismic slip. High
values of $\sigma^M_n$ increase the frictional resistance to slip,
potentially limiting slip magnitude in locations such as below the
Pengguan massif.

Assuming that the Beichuan faults were at a Mohr-Coulomb fault criterion
immediately prior to the Wenchuan earthquake, we estimate the tectonic
stresses required for the faults to fail. We use a Bayesian estimation,
resulting in posteriors representing likelihood of tectonic stress,
static friction, and a pore pressure parameter. The results of our
Bayesian inversion indicate that the maximum tectonic stress is oriented
\textasciitilde{}E-W and has a likely minimum of 13.25 MPa km$^{-1}$.
The minimum tectonic stress is oriented \textasciitilde{}N-S and is
fairly low, with highest likelihood one half of the E-W tectonic stress.
The highest likelihood coefficient of static friction on the fault is
estimated at about 0.2--0.3, although values up to 0.5--0.6 are
permissable. Fluid pressures are likely 0--0.5 of the total pressure.

These results suggest that local stress conditions caused by topographic
loading are unfavorable to slip on the Wenchuan earthquake faults, but
because of the low friction coefficients and moderate fluid pressures,
slip occurred on these faults instead of more favorably-oriented faults
elsewhere in the region.

\appendix

\section{Green's functions for point-source
loads}\label{greens-functions-for-point-source-loads}

Note that in these solutions, $\lambda$ and $\mu$ are re-defined, and
are the first and second Lame's parameters, respectively, instead of
rake and fault friction as in the body of the manuscript.

\subsection{Boussinesq's solutions for vertical point-source
loads}\label{boussinesqs-solutions-for-vertical-point-source-loads}

\begin{equation}
\begin{split}
G_{xx}^B & = \frac{ F_v }{ 2\pi } \left[ \frac{ 3x^2 }{ r^5 } \right.
+ \frac{\mu (y^2 + z^2)}{(\lambda + \mu) r^3 (z + r)}
- \frac{\mu z}{(\lambda + \mu) r^3} \\
&\qquad \quad \left. - \frac{\mu x^2}{ (\lambda + \mu) r^2 (z + r)^2 }\right ]
\end{split}
\end{equation}

\begin{equation}
\begin{split}
G_{yy}^B & = \frac{F_v}{2\pi } \left [ \frac{3y^2}{r^5} \right.
+ \frac{\mu (x^{2} + z^{2})}{(\lambda + \mu) r^{3}(z + r)}
- \frac{\mu z}{(\lambda + \mu) r ^{3}} \\
& \qquad \quad \left. - \frac{\mu y^{2}}{(\lambda + \mu ) r^2 (z +r)^2} \right]
\end{split}
\end{equation}

\begin{equation}
G_{xy}^{B} = \frac{F _{v}}{2\pi} \left[ \frac{3xyz}{r^{5}}
- \frac{\mu x y (z + 2r)}{(\lambda + \mu) r^{3} (z + r)^{2}} \right]
\end{equation}

\begin{equation}
G_{zz}^{B} = 3 F _{v} z^{3} / 2 \pi r^{5}
\end{equation}

\begin{equation}
G_{xz}^{B} = 3 F _{v} xz^{2} / 2 \pi r^{5}
\end{equation}

\begin{equation}
G_{yz}^{B} = 3 F _{v} yz^{2} / 2 \pi r^{5}
\end{equation}

\subsection{Cerruti's solutions for horizontal point-source
loads}\label{cerrutis-solutions-for-horizontal-point-source-loads}

\begin{equation}
\begin{split}
G_{xx}^{C_x} &= \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3x^2}{r^2} \right.
- \frac{\mu}{(\lambda + \mu)(z+r)^2} 
\\
&\qquad \quad  \left. (r^2 - y^2 - \frac{2ry^2}{r+z}) \right]
\end{split}
\end{equation}

\begin{equation}
\begin{split}
G_{yy}^{C_x} & = \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3y^2}{r^2} \right.
- \frac{\mu}{(\lambda + \mu)(z+r)^2} \\
& \qquad \quad \left. (3r^2 - x^2 - \frac{2rx^2}{r+z}) \right]
\end{split}
\end{equation}

\begin{equation}
\begin{split}
G_{xy}^{C_x} & = \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3x^2}{r^2} \right.
- \frac{\mu}{(\lambda + \mu)(z+r)^2} 
\\
&\qquad \quad \left. \cdot (r^2 - x^2 - \frac{2rx^2}{r+z}) \right]
\end{split}
\end{equation}

\begin{equation}
    G^{C_x}_{zz} = \frac{ 3 F_{h,x} x z^2 }{2 \pi r^5}
\end{equation}

\begin{equation}
    G^{C_x}_{xz} = \frac{ 3 F_{h,x} z x^2 }{2 \pi r^5}
\end{equation}

\begin{equation}
    G^{C_x}_{yz} = \frac{ 3 F_{h,x} x y z }{2 \pi r^5}
\end{equation}

\begin{acknowledgements}
Joe Kington wrote the code to plot Figure \ref{fig:joint_posteriors}c.
\end{acknowledgements}
%\section{References}\label{references}

%\bibliography{wench_jgr_bib}{}
%\bibliographystyle{agufull08}

\begin{thebibliography}{85}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Aagaard et~al.}(2001)\textit{Aagaard, Heaton, and
  Hall}}]{aagaard2001}
Aagaard, B.~T., T.~H. Heaton, and J.~F. Hall (2001), Dynamic earthquake
  ruptures in the presence of lithostatic normal stresses: {I}mplications for
  friction models and heat production, \textit{Bulletin of the Seismological
  Society of America}, \textit{91}(6), 1765--1796.

\bibitem[{\textit{Allmann and Shearer}(2009)}]{allmann2009}
Allmann, B.~P., and P.~M. Shearer (2009), Global variations of stress drop for
  moderate to large earthquakes, \textit{Journal of Geophysical Research: Solid
  Earth (1978--2012)}, \textit{114}(B1).

\bibitem[{\textit{Angelier}(1994)}]{angelier1994}
Angelier, J. (1994), Fault slip analysis and paleostress reconstruction,
  \textit{Continental deformation}, \textit{4}.

\bibitem[{\textit{Armijo et~al.}(1986)\textit{Armijo, Tapponnier, Mercier, and
  Han}}]{armijo1986}
Armijo, R., P.~Tapponnier, J.~Mercier, and T.-L. Han (1986), Quaternary
  extension in southern {T}ibet: {F}ield observations and tectonic
  implications, \textit{Journal of Geophysical Research: Solid Earth
  (1978--2012)}, \textit{91}(B14), 13,803--13,872.

\bibitem[{\textit{Beaumont et~al.}(2001)\textit{Beaumont, Jamieson, Nguyen, and
  Lee}}]{beaumont2001}
Beaumont, C., R.~A. Jamieson, M.~Nguyen, and B.~Lee (2001), Himalayan tectonics
  explained by extrusion of a low-viscosity crustal channel coupled to focused
  surface denudation, \textit{Nature}, \textit{414}(6865), 738--742.

\bibitem[{\textit{Bird and Kong}(1994)}]{bird1994}
Bird, P., and X.~Kong (1994), Computer simulations of {C}alifornia tectonics
  confirm very low strength of major faults, \textit{Geological Society of
  America Bulletin}, \textit{106}(2), 159--174.

\bibitem[{\textit{Bird and Piper}(1980)}]{birdpiper80}
Bird, P., and K.~Piper (1980), Plane-stress finite-element models of tectonic
  flow in southern {C}alifornia, \textit{Physics of the earth and planetary
  interiors}, \textit{21}(2), 158--175.

\bibitem[{\textit{Bollinger et~al.}(2004)\textit{Bollinger, Avouac, Cattin, and
  Pandey}}]{bollinger2004}
Bollinger, L., J.~Avouac, R.~Cattin, and M.~Pandey (2004), Stress buildup in
  the {H}imalaya, \textit{Journal of Geophysical Research: Solid Earth},
  \textit{109}(B11).

\bibitem[{\textit{Bott}(1959)}]{bott1959}
Bott, M. H.~P. (1959), The mechanics of oblique slip faulting,
  \textit{Geological Magazine}, \textit{96}(02), 109--117.

\bibitem[{\textit{Burchfiel et~al.}(1995)\textit{Burchfiel, Zhiliang, Yupinc,
  and Royden}}]{burchfiel1995}
Burchfiel, B., C.~Zhiliang, L.~Yupinc, and L.~Royden (1995), Tectonics of the
  {L}ongmen {S}han and adjacent regions, central {}china, \textit{International
  Geology Review}, \textit{37}(8), 661--735.

\bibitem[{\textit{Burchfiel et~al.}(2008)\textit{Burchfiel, Royden, van~der
  Hilst, Hager, Chen, King, Li, L{\"u}, Yao, and Kirby}}]{burchfiel2008}
Burchfiel, B., L.~Royden, R.~van~der Hilst, B.~Hager, Z.~Chen, R.~King, C.~Li,
  J.~L{\"u}, H.~Yao, and E.~Kirby (2008), A geological and geophysical context
  for the {W}enchuan earthquake of 12 {M}ay 2008, {S}ichuan, {P}eople's
  {R}epublic of {C}hina, \textit{GSA Today}, \textit{18}(7), 4--11.

\bibitem[{\textit{Byerlee}(1978)}]{byerlee1978}
Byerlee, J. (1978), Friction of rocks, \textit{Pure and applied Geophysics},
  \textit{116}(4-5), 615--626.

\bibitem[{\textit{Byrne et~al.}(2013)\textit{Byrne, Holohan, Kervyn, de~Vries,
  Troll, and Murray}}]{byrne2013}
Byrne, P., E.~Holohan, M.~Kervyn, B.~v.~W. de~Vries, V.~R. Troll, and J.~Murray
  (2013), A sagging-spreading continuum of large volcano structure,
  \textit{Geology}, \textit{41}(3), 339--342.

\bibitem[{\textit{Cattin et~al.}(1997)\textit{Cattin, Lyon-Caen, and
  Ch{\'e}ry}}]{cattin1997}
Cattin, R., H.~Lyon-Caen, and J.~Ch{\'e}ry (1997), Quantification of interplate
  coupling in subduction zones and forearc topography, \textit{Geophysical
  Research Letters}, \textit{24}(13), 1563--1566.

\bibitem[{\textit{Clark and Royden}(2000)}]{clarkroyden2000}
Clark, M.~K., and L.~H. Royden (2000), Topographic ooze: {B}uilding the eastern
  margin of {T}ibet by lower crustal flow, \textit{Geology}, \textit{28}(8),
  703--706.

\bibitem[{\textit{Clark et~al.}(2005)\textit{Clark, Bush, and
  Royden}}]{clark2005}
Clark, M.~K., J.~W. Bush, and L.~H. Royden (2005), Dynamic topography produced
  by lower crustal flow against rheological strength heterogeneities bordering
  the {T}ibetan {P}lateau, \textit{Geophysical Journal International},
  \textit{162}(2), 575--590.

\bibitem[{\textit{Coblentz and Richardson}(1996)}]{coblentz1996}
Coblentz, D.~D., and R.~M. Richardson (1996), Analysis of the {S}outh
  {A}merican intraplate stress field, \textit{Journal of Geophysical Research:
  Solid Earth (1978--2012)}, \textit{101}(B4), 8643--8657.

\bibitem[{\textit{Cook and Royden}(2008)}]{cook2008}
Cook, K.~L., and L.~H. Royden (2008), The role of crustal strength variations
  in shaping orogenic plateaus, with application to {T}ibet, \textit{Journal of
  Geophysical Research: Solid Earth (1978--2012)}, \textit{113}(B8).

\bibitem[{\textit{Copley and McKenzie}(2007)}]{copleymckenzie2007}
Copley, A., and D.~McKenzie (2007), Models of crustal flow in the
  {I}ndia--{A}sia collision zone, \textit{Geophysical Journal International},
  \textit{169}(2), 683--698.

\bibitem[{\textit{Copley et~al.}(2009)\textit{Copley, Boait, Hollingsworth,
  Jackson, and McKenzie}}]{copley2009}
Copley, A., F.~Boait, J.~Hollingsworth, J.~Jackson, and D.~McKenzie (2009),
  Subparallel thrust and normal faulting in {A}lbania and the roles of
  gravitational potential energy and rheology contrasts in mountain belts,
  \textit{Journal of Geophysical Research: Solid Earth}, \textit{114}(B5).

\bibitem[{\textit{Cui et~al.}(2014)\textit{Cui, Lin, Wang, Gao, Huang, Wang,
  Sun, Li, Zhou, Qian et~al.}}]{cui2014}
Cui, J., W.~Lin, L.~Wang, L.~Gao, Y.~Huang, W.~Wang, D.~Sun, Z.~Li, C.~Zhou,
  H.~Qian, et~al. (2014), Determination of three-dimensional in situ stresses
  by anelastic strain recovery in {W}enchuan {E}arthquake {F}ault {S}cientific
  {D}rilling {P}roject {H}ole-1 ({W}{F}{S}{D}-1), \textit{Tectonophysics},
  \textit{619–620}, 123 -- 132.

\bibitem[{\textit{Dahlen}(1990)}]{dahlen1990}
Dahlen, F. (1990), Critical taper model of fold-and-thrust belts and
  accretionary wedges, \textit{Annual Review of Earth and Planetary Sciences},
  \textit{18}, 55.

\bibitem[{\textit{Dalmayrac and Molnar}(1981)}]{dalmayrac1981}
Dalmayrac, B., and P.~Molnar (1981), Parallel thrust and normal faulting in
  {P}eru and constraints on the state of stress, \textit{Earth and Planetary
  Science Letters}, \textit{55}(3), 473--481.

\bibitem[{\textit{Day}(1982)}]{day1982}
Day, S.~M. (1982), Three-dimensional simulation of spontaneous rupture: {T}he
  effect of nonuniform prestress, \textit{Bulletin of the Seismological Society
  of America}, \textit{72}(6A), 1881--1902.

\bibitem[{\textit{DeCelles et~al.}(2009)\textit{DeCelles, Ducea, Kapp, and
  Zandt}}]{decelles2009}
DeCelles, P.~G., M.~N. Ducea, P.~Kapp, and G.~Zandt (2009), Cyclicity in
  cordilleran orogenic systems, \textit{Nature Geoscience}, \textit{2}(4),
  251--257.

\bibitem[{\textit{Dewey}(1988)}]{dewey1988}
Dewey, J.~F. (1988), Extensional collapse of orogens, \textit{Tectonics},
  \textit{7}(6), 1123--1139.

\bibitem[{\textit{England and Houseman}(1989)}]{england1989}
England, P., and G.~Houseman (1989), Extension during continental convergence,
  with application to the tibetan plateau, \textit{Journal of Geophysical
  Research: Solid Earth (1978--2012)}, \textit{94}(B12), 17,561--17,579.

\bibitem[{\textit{Farr et~al.}(2007)\textit{Farr, Rosen, Caro, Crippen, Duren,
  Hensley, Kobrick, Paller, Rodriguez, Roth et~al.}}]{farr2007srtm}
Farr, T.~G., P.~A. Rosen, E.~Caro, R.~Crippen, R.~Duren, S.~Hensley,
  M.~Kobrick, M.~Paller, E.~Rodriguez, L.~Roth, et~al. (2007), The shuttle
  radar topography mission, \textit{Reviews of geophysics}, \textit{45}(2).

\bibitem[{\textit{Feng et~al.}(2010)\textit{Feng, Hetland, Ding, Li, and
  Zhang}}]{feng2010}
Feng, G., E.~A. Hetland, X.~Ding, Z.~Li, and L.~Zhang (2010), Coseismic fault
  slip of the 2008 {M}w 7.9 {W}enchuan earthquake estimated from {I}n{S}{A}{R}
  and {G}{P}{S} measurements, \textit{Geophysical research letters},
  \textit{37}(1).

\bibitem[{\textit{Fielding et~al.}(2013)\textit{Fielding, Sladen, Li, Avouac,
  B{\"u}rgmann, and Ryder}}]{fielding2013}
Fielding, E.~J., A.~Sladen, Z.~Li, J.-P. Avouac, R.~B{\"u}rgmann, and I.~Ryder
  (2013), Kinematic fault slip evolution source models of the 2008 m7.9
  wenchuan earthquake in china from sar interferometry, gps and teleseismic
  analysis and implications for longmen shan tectonics, \textit{Geophysical
  Journal International}, \textit{194}(2), 1138--1166.

\bibitem[{\textit{Flesch and Bendick}(2012)}]{fleschbendick2012}
Flesch, L., and R.~Bendick (2012), The relationship between surface kinematics
  and deformation of the whole lithosphere, \textit{Geology}, \textit{40}(8),
  711--714.

\bibitem[{\textit{Flesch and Kreemer}(2010)}]{flesch2010gpe}
Flesch, L.~M., and C.~Kreemer (2010), Gravitational potential energy and
  regional stress and strain rate fields for continental plateaus: {E}xamples
  from the central {A}ndes and {C}olorado {P}lateau, \textit{Tectonophysics},
  \textit{482}(1), 182--192.

\bibitem[{\textit{Heidbach et~al.}(2009)\textit{Heidbach, Tingay, and
  Barth}}]{heidbach2009}
Heidbach, O., M.~Tingay, and A.~Barth (2009), \textit{The World Stress Map
  based on the database release 2008, equatorial scale 1:46,000,000},
  Commission for the Geological Map of the World, Paris.

\bibitem[{\textit{Hubbard et~al.}(2010)\textit{Hubbard, Shaw, and
  Klinger}}]{hubbard2010}
Hubbard, J., J.~H. Shaw, and Y.~Klinger (2010), Structural setting of the 2008
  {M}w 7.9 {W}enchuan, {C}hina, earthquake, \textit{Bulletin of the
  Seismological Society of America}, \textit{100}(5B), 2713--2735.

\bibitem[{\textit{Jarvis et~al.}(2008)\textit{Jarvis, Reuter, Nelson, and
  Guevara}}]{jarvis2008srtm}
Jarvis, A., H.~Reuter, A.~Nelson, and E.~Guevara (2008), Hole-filled srtm for
  the globe version 3, available from the cgiar-csi srtm 90m database,
  \textit{Accessed: March}, \textit{12}, 2012.

\bibitem[{\textit{Jeffreys}(1924)}]{jeffreys1924}
Jeffreys, H. (1924), \textit{The earth: its origin, history and physical
  constitution}, Cambridge University Press.

\bibitem[{\textit{Kanamori and Anderson}(1975)}]{kanamori1975}
Kanamori, H., and D.~L. Anderson (1975), Theoretical basis of some empirical
  relations in seismology, \textit{Bulletin of the Seismological Society of
  America}, \textit{65}(5), 1073--1095.

\bibitem[{\textit{Kuo et~al.}(2014)\textit{Kuo, Li, Smith, Di~Toro, Suppe,
  Song, Nielsen, Sheu, and Si}}]{kuo2014}
Kuo, L.-W., H.~Li, S.~A. Smith, G.~Di~Toro, J.~Suppe, S.-R. Song, S.~Nielsen,
  H.-S. Sheu, and J.~Si (2014), Gouge graphitization and dynamic fault
  weakening during the 2008 {M}w 7.9 {W}enchuan earthquake, \textit{Geology},
  \textit{42}(1), 47--50.

\bibitem[{\textit{Lamb}(2006)}]{lamb2006}
Lamb, S. (2006), Shear stresses on megathrusts: {I}mplications for mountain
  building behind subduction zones, \textit{Journal of Geophysical Research:
  Solid Earth}, \textit{111}(B7).

\bibitem[{\textit{Liang et~al.}(2013)\textit{Liang, Gan, Shen, Xiao, Liu, Chen,
  Ding, and Zhou}}]{liang2013}
Liang, S., W.~Gan, C.~Shen, G.~Xiao, J.~Liu, W.~Chen, X.~Ding, and D.~Zhou
  (2013), Three-dimensional velocity field of present-day crustal motion of the
  {T}ibetan {P}lateau derived from {G}{P}{S} measurements, \textit{Journal of
  Geophysical Research: Solid Earth}, \textit{118}(10), 5722--5732.

\bibitem[{\textit{Lin et~al.}(2009)\textit{Lin, Ren, Jia, and Wu}}]{lin2009}
Lin, A., Z.~Ren, D.~Jia, and X.~Wu (2009), Co-seismic thrusting rupture and
  slip distribution produced by the 2008< i> m</i>< sub> w</sub> 7.9 {W}enchuan
  earthquake, china, \textit{Tectonophysics}, \textit{471}(3), 203--215.

\bibitem[{\textit{Lisle}(2013)}]{lisle2013}
Lisle, R.~J. (2013), A critical look at the {W}allace-{B}ott hypothesis in
  fault-slip analysis, \textit{Bulletin de la Societe Geologique de France},
  \textit{184}(4-5), 299--306.

\bibitem[{\textit{Liu and Zoback}(1992)}]{liuzoback1992}
Liu, L., and M.~D. Zoback (1992), The effect of topography on the state of
  stress in the crust: {A}pplication to the site of the {C}ajon {P}ass
  {S}cientific {D}rilling {P}roject, \textit{Journal of Geophysical Research:
  Solid Earth (1978--2012)}, \textit{97}(B4), 5095--5108.

\bibitem[{\textit{Liu and Yang}(2003)}]{liuyang2003}
Liu, M., and Y.~Yang (2003), Extensional collapse of the {T}ibetan {P}lateau:
  {R}esults of three-dimensional finite element modeling, \textit{Journal of
  Geophysical Research: Solid Earth (1978--2012)}, \textit{108}(B8).

\bibitem[{\textit{Liu-Zeng et~al.}(2009)\textit{Liu-Zeng, Zhang, Wen,
  Tapponnier, Sun, Xing, Hu, Xu, Zeng, Ding et~al.}}]{liu2009}
Liu-Zeng, J., Z.~Zhang, L.~Wen, P.~Tapponnier, J.~Sun, X.~Xing, G.~Hu, Q.~Xu,
  L.~Zeng, L.~Ding, et~al. (2009), Co-seismic ruptures of the 12 may 2008,< i>
  m</i>< sub> s</sub> 8.0 wenchuan earthquake, sichuan: East--west crustal
  shortening on oblique, parallel thrusts along the eastern edge of tibet,
  \textit{Earth and Planetary Science Letters}, \textit{286}(3), 355--370.

\bibitem[{\textit{Loveless and Meade}(2011)}]{loveless2011}
Loveless, J., and B.~Meade (2011), Partitioning of localized and diffuse
  deformation in the {T}ibetan {P}lateau from joint inversions of geologic and
  geodetic observations, \textit{Earth and Planetary Science Letters},
  \textit{303}(1), 11--24.

\bibitem[{\textit{Luttrell et~al.}(2007)\textit{Luttrell, Sandwell,
  Smith-Konter, Bills, and Bock}}]{luttrell2007}
Luttrell, K., D.~Sandwell, B.~Smith-Konter, B.~Bills, and Y.~Bock (2007),
  Modulation of the earthquake cycle at the southern {S}an {A}ndreas fault by
  lake loading, \textit{Journal of Geophysical Research: Solid Earth
  (1978--2012)}, \textit{112}(B8).

\bibitem[{\textit{Luttrell et~al.}(2011)\textit{Luttrell, Tong, Sandwell,
  Brooks, and Bevis}}]{luttrell2011}
Luttrell, K.~M., X.~Tong, D.~T. Sandwell, B.~A. Brooks, and M.~G. Bevis (2011),
  Estimates of stress drop and crustal tectonic stress from the 27 {F}ebruary
  2010 {M}aule, {C}hile, earthquake: Implications for fault strength,
  \textit{Journal of Geophysical Research: Solid Earth}, \textit{116}(B11).

\bibitem[{\textit{Martel}(2006)}]{martel2006}
Martel, S.~J. (2006), Effect of topographic curvature on near-surface stresses
  and application to sheeting joints, \textit{Geophysical Research Letters},
  \textit{33}(1).

\bibitem[{\textit{McKenzie}(1969)}]{mckenzie1969}
McKenzie, D.~P. (1969), The relation between fault plane solutions for
  earthquakes and the directions of the principal stresses, \textit{Bulletin of
  the Seismological Society of America}, \textit{59}(2), 591--601.

\bibitem[{\textit{McKinney}(2010)}]{mckinney2010}
McKinney, W. (2010), Data structures for statistical computing in python, in
  \textit{Proceedings of the 9th Python in Science Conference}, edited by
  S.~van~der Walt and J.~Millman, pp. 51 -- 56, SciPy.

\bibitem[{\textit{McTigue and Mei}(1981)}]{mctiguemei1981}
McTigue, D.~F., and C.~C. Mei (1981), Gravity-induced stresses near topography
  of small slope, \textit{Journal of Geophysical Research: Solid Earth},
  \textit{86}(B10), 9268--9278.

\bibitem[{\textit{Meade and Conrad}(2008)}]{meade2008}
Meade, B.~J., and C.~P. Conrad (2008), Andean growth and the deceleration of
  {S}outh {A}merican subduction: {T}ime evolution of a coupled
  orogen-subduction system, \textit{Earth and Planetary Science Letters},
  \textit{275}(1), 93--101.

\bibitem[{\textit{Medina~Luna and Hetland}(2013)}]{medinaluna2013}
Medina~Luna, L., and E.~A. Hetland (2013), Regional stresses inferred from
  coseismic slip models of the 2008 {M}w 7.9 {W}enchuan, {C}hina, earthquake,
  \textit{Tectonophysics}, \textit{584}, 43--53.

\bibitem[{\textit{Meissner and Strehlau}(1982)}]{meissner1982}
Meissner, R., and J.~Strehlau (1982), Limits of stresses in continental crusts
  and their relation to the depth-frequency distribution of shallow
  earthquakes, \textit{Tectonics}, \textit{1}(1), 73--89.

\bibitem[{\textit{Michael}(1987)}]{michael1987}
Michael, A.~J. (1987), Use of focal mechanisms to determine stress: a control
  study, \textit{Journal of Geophysical Research: Solid Earth (1978--2012)},
  \textit{92}(B1), 357--368.

\bibitem[{\textit{Miller and Dunne}(1996)}]{millerdunne1996}
Miller, D.~J., and T.~Dunne (1996), Topographic perturbations of regional
  stresses and consequent bedrock fracturing, \textit{Journal of Geophysical
  Research: Solid Earth (1978--2012)}, \textit{101}(B11), 25,523--25,536.

\bibitem[{\textit{Molnar and Lyon-Caen}(1988)}]{molnar1988}
Molnar, P., and H.~Lyon-Caen (1988), Some simple physical aspects of the
  support, structure, and evolution of mountain belts, \textit{Spec. Pap. Geol.
  Soc. Am}, \textit{218}, 179--207.

\bibitem[{\textit{Mosegaard and Tarantola}(1995)}]{mosegaard1995}
Mosegaard, K., and A.~Tarantola (1995), Monte {C}arlo sampling of solutions to
  inverse problems, \textit{Journal of Geophysical Research: Solid Earth
  (1978--2012)}, \textit{100}(B7), 12,431--12,447.

\bibitem[{\textit{Nakamura et~al.}(2010)\textit{Nakamura, Tsuboi, Kaneda, and
  Yamanaka}}]{nakamura2010}
Nakamura, T., S.~Tsuboi, Y.~Kaneda, and Y.~Yamanaka (2010), Rupture process of
  the 2008 {W}enchuan, {C}hina earthquake inferred from teleseismic waveform
  inversion and forward modeling of broadband seismic waves,
  \textit{Tectonophysics}, \textit{491}(1), 72--84.

\bibitem[{\textit{Oglesby and Day}(2002)}]{oglesbyday2002}
Oglesby, D.~D., and S.~M. Day (2002), Stochastic fault stress: {I}mplications
  for fault dynamics and ground motion, \textit{Bulletin of the Seismological
  Society of America}, \textit{92}(8), 3006--3021.

\bibitem[{\textit{Oliphant}(2007)}]{oliphant2007numpy}
Oliphant, T.~E. (2007), Python for scientific computing, \textit{Computing in
  Science \& Engineering}, \textit{9}(3), 10--20.

\bibitem[{\textit{Olsen et~al.}(1997)\textit{Olsen, Madariaga, and
  Archuleta}}]{olsen1997}
Olsen, K., R.~Madariaga, and R.~Archuleta (1997), Three-dimensional dynamic
  simulation of the 1992 {L}anders earthquake, \textit{Science},
  \textit{278}(5339), 834--838.

\bibitem[{\textit{P\'erez and Granger}(2007)}]{perez2007ipython}
P\'erez, F., and B.~E. Granger (2007), {IP}ython: a {S}ystem for {I}nteractive
  {S}cientific {C}omputing, \textit{{C}omput. {S}ci. {E}ng.}, \textit{9}(3),
  21--29.

\bibitem[{\textit{Qi et~al.}(2011)\textit{Qi, Xuejun, Qigui, Freymueller,
  Shaomin, Caijun, Yonglin, Xinzhao, Kai, and Gang}}]{qi2011}
Qi, W., Q.~Xuejun, L.~Qigui, J.~Freymueller, Y.~Shaomin, X.~Caijun, Y.~Yonglin,
  Y.~Xinzhao, T.~Kai, and C.~Gang (2011), Rupture of deep faults in the 2008
  {W}enchuan earthquake and uplift of the {L}ongmen {S}han, \textit{Nature
  Geoscience}, \textit{4}(9), 634--640.

\bibitem[{\textit{Reches}(1987)}]{reches1987}
Reches, Z. (1987), Determination of the tectonic stress tensor from slip along
  faults that obey the coulomb yield condition, \textit{Tectonics},
  \textit{6}(6), 849--861.

\bibitem[{\textit{Rey et~al.}(2001)\textit{Rey, Vanderhaeghe, and
  Teyssier}}]{rey2001}
Rey, P., O.~Vanderhaeghe, and C.~Teyssier (2001), Gravitational collapse of the
  continental crust: definition, regimes and modes, \textit{Tectonophysics},
  \textit{342}(3), 435--449.

\bibitem[{\textit{Richardson and Coblentz}(1994)}]{richardson1994}
Richardson, R.~M., and D.~D. Coblentz (1994), Stress modeling in the {A}ndes:
  Constraints on the {S}outh {A}merican intraplate stress magnitudes,
  \textit{Journal of Geophysical Research: Solid Earth}, \textit{99}(B11),
  22,015--22,025.

\bibitem[{\textit{Savage and Swolfs}(1986)}]{savageswolfs1986}
Savage, W.~Z., and H.~S. Swolfs (1986), Tectonic and gravitational stress in
  long symmetric ridges and valleys, \textit{Journal of Geophysical Research:
  Solid Earth}, \textit{91}(B3), 3677--3685.

\bibitem[{\textit{Scholz}(2002)}]{scholz2002}
Scholz, C.~H. (2002), \textit{The mechanics of earthquakes and faulting},
  Cambridge university press.

\bibitem[{\textit{Seabold and Perktold}(2010)}]{seabold2010}
Seabold, S., and J.~Perktold (2010), Statsmodels: econometric and statistical
  modeling with python, in \textit{Proceedings of the 9th Python in Science
  Conference}, pp. 57--61, SciPy.

\bibitem[{\textit{Shen et~al.}(2009)\textit{Shen, Sun, Zhang, Wan, Wang,
  B{\"u}rgmann, Zeng, Gan, Liao, and Wang}}]{shen2009}
Shen, Z.-K., J.~Sun, P.~Zhang, Y.~Wan, M.~Wang, R.~B{\"u}rgmann, Y.~Zeng,
  W.~Gan, H.~Liao, and Q.~Wang (2009), Slip maxima at fault junctions and
  rupturing of barriers during the 2008 {W}enchuan earthquake, \textit{Nature
  Geoscience}, \textit{2}(10), 718--724.

\bibitem[{\textit{Sibson}(1985)}]{sibson1985}
Sibson, R.~H. (1985), A note on fault reactivation, \textit{Journal of
  Structural Geology}, \textit{7}(6), 751--754.

\bibitem[{\textit{Styron et~al.}(2010)\textit{Styron, Taylor, and
  Okoronkwo}}]{styron2010}
Styron, R., M.~Taylor, and K.~Okoronkwo (2010), Database of active structures
  from the {I}ndo-{A}sian collision, \textit{Eos, Transactions American
  Geophysical Union}, \textit{91}(20), 181--182.

\bibitem[{\textit{Tarantola}(2005)}]{tarantola2005}
Tarantola, A. (2005), \textit{Inverse problem theory and methods for model
  parameter estimation}, SIAM.

\bibitem[{\textit{Taylor et~al.}(2003)\textit{Taylor, Yin, Ryerson, Kapp, and
  Ding}}]{taylor2003}
Taylor, M., A.~Yin, F.~J. Ryerson, P.~Kapp, and L.~Ding (2003), Conjugate
  strike-slip faulting along the {B}angong-{N}ujiang suture zone accommodates
  coeval east-west extension and north-south shortening in the interior of the
  {T}ibetan {P}lateau, \textit{Tectonics}, \textit{22}(4).

\bibitem[{\textit{Tong et~al.}(2010)\textit{Tong, Sandwell, and
  Fialko}}]{tong2010}
Tong, X., D.~T. Sandwell, and Y.~Fialko (2010), Coseismic slip model of the
  2008 {W}enchuan earthquake derived from joint inversion of interferometric
  synthetic aperture radar, {G}{P}{S}, and field data, \textit{Journal of
  Geophysical Research: Solid Earth}, \textit{115}(B4).

\bibitem[{\textit{Townend and Zoback}(2000)}]{townend2000}
Townend, J., and M.~D. Zoback (2000), How faulting keeps the crust strong,
  \textit{Geology}, \textit{28}(5), 399--402.

\bibitem[{\textit{Wallace}(1951)}]{wallace1951}
Wallace, R.~E. (1951), Geometry of shearing stress and relation to faulting,
  \textit{The Journal of Geology}, pp. 118--130.

\bibitem[{\textit{Wang et~al.}(2012)\textit{Wang, Kirby, Furlong, Van~Soest,
  Xu, Shi, Kamp, and Hodges}}]{wang2012}
Wang, E., E.~Kirby, K.~P. Furlong, M.~Van~Soest, G.~Xu, X.~Shi, P.~J. Kamp, and
  K.~Hodges (2012), Two-phase growth of high topography in eastern {T}ibet
  during the {C}enozoic, \textit{Nature Geoscience}, \textit{5}(9), 640--645.

\bibitem[{\textit{Xiao et~al.}(1991)\textit{Xiao, Dahlen, and
  Suppe}}]{xiao1991}
Xiao, H.-B., F.~Dahlen, and J.~Suppe (1991), Mechanics of extensional wedges,
  \textit{Journal of Geophysical Research: Solid Earth}, \textit{96}(B6),
  10,301--10,318.

\bibitem[{\textit{Xu et~al.}(2009)\textit{Xu, Wen, Yu, Chen, Klinger, Hubbard,
  and Shaw}}]{xu2009}
Xu, X., X.~Wen, G.~Yu, G.~Chen, Y.~Klinger, J.~Hubbard, and J.~Shaw (2009),
  Coseismic reverse-and oblique-slip surface faulting generated by the 2008 mw
  7.9 wenchuan earthquake, china, \textit{Geology}, \textit{37}(6), 515--518.

\bibitem[{\textit{Yong et~al.}(2003)\textit{Yong, Allen, Densmore, and
  Qiang}}]{yong2003}
Yong, L., P.~A. Allen, A.~L. Densmore, and X.~Qiang (2003), Evolution of the
  {L}ongmen {S}han foreland basin (western {S}ichuan, {C}hina) during the
  {L}ate {T}riassic {I}ndosinian orogeny, \textit{Basin Research},
  \textit{15}(1), 117--138.

\bibitem[{\textit{Zhang et~al.}(2011)\textit{Zhang, Qu, Shan, Song, Zhang,
  Wang, Hu, and Wang}}]{zhang2011}
Zhang, G., C.~Qu, X.~Shan, X.~Song, G.~Zhang, C.~Wang, J.-C. Hu, and R.~Wang
  (2011), Slip distribution of the 2008 {W}enchuan {M}s 7.9 earthquake by joint
  inversion from {G}{P}{S} and {I}n{S}{A}{R} measurements: a resolution test
  study, \textit{Geophysical Journal International}, \textit{186}(1), 207--220.

\bibitem[{\textit{Zhang et~al.}(2010)\textit{Zhang, Wen, Shen, and
  Chen}}]{zhang2010}
Zhang, P.-Z., X.-z. Wen, Z.-K. Shen, and J.-h. Chen (2010), Oblique,
  high-angle, listric-reverse faulting and associated development of strain:
  The {W}enchuan earthquake of {M}ay 12, 2008, {S}ichuan, {C}hina,
  \textit{Annual Review of Earth and Planetary Sciences}, \textit{38},
  353--382.

\end{thebibliography}

\end{article}

\end{document}

### 2014-05-29 manuscript/styron_hetland_wenchuan_topo_manuscript.tex
added hetland's revisions to manuscript tex

@@
 \usepackage{amssymb,amsmath}
-
-%\usepackage[unicode=true]{hyperref}
-%\hypersetup{breaklinks=true,
-%            bookmarks=true,
-%            pdfauthor={},
-%            pdftitle={Topographic stresses and the Wenchuan earthquake},
-%            colorlinks=true,
-%            citecolor=blue,
-%            urlcolor=blue,
-%            linkcolor=magenta,
-%            pdfborder={0 0 0}}
-%\urlstyle{same}
 
@@
 Stress is of fundamental importance to many processes in the earth. Both
-the isotropic (pressure) and anisotropic (shear) components of stress
+the isotropic and deviatioric components of stress
 exert control on the deformation state of the earth at any point in the
@@
 derived from strain, for example from studies of earthquake focal
-mechanisms \citep[e.g.,][]{michael1987} or of geological fault slip data
-\citep[e.g.,][]{reches1987}.
+mechanisms \citep[e.g.,][]{michael1987} or of fault slip data
+\citep[e.g.,][]{reches1987, medinaluna2013}.
 
@@
 faults that ruptured in the 2008 \emph{M}7.9 Wenchuan, China earthquake.
-This earthquake is an ideal candidate for study because it occurred at
+This earthquake is an ideal candidate for this study because it occurred at
 the base of the Longmen Shan, one of the largest and steepest
@@
 extensional deformation in
-the high parts of orogens temporally coincident with contractional
+the high parts of orogens that is temporally coincident with contractional
 deformation at the low-elevation margins of the orogen may be explained
@@
 of the thrust wedge where the stress state is more favorable. Of
-particular relevance to this work are `fixed boundary' models of
+particular relevance to our work are `fixed boundary' models of
 gravitational collapse and spreading \citep{rey2001}, where an excess of
@@
 developed methods for calculating the three dimensional elastic stress
-tensor field arbitrary topography, where previous solutions were limited
+tensor field arbitrary topography, whereas previous solutions were limited
 to two dimensions and required topography to be mathematically-defined
@@
 []{angelier1994} and (2) the fault is at Mohr-Coulomb failure everywhere
-that is slipped. If topographic stresses are significant and produce
+that it slipped. If topographic stresses are significant and produce
 shear in the direction of fault slip, then for given values of static
@@
 dimensional models of the faults taken from coseismic slip models.
-Finally, we can calculate topographic shear and normal stresses on the
+Finally, we calculate topographic shear and normal stresses on the
 faults and compare those to the coseismic slip patterns.
@@
 calculations}\label{topographic-stress-tensor-field-calculations}
-
-\subsection{Analytical description}\label{analytical-description}
 
@@
 \citet{liuzoback1992}. They show that the topographic stress tensor
-field can be determined by a convolution of a topographic loading
-function with Green's functions describing the stresses in an elastic
+field can be determined by a convolution of topographic loading
+functions with Green's functions describing the stresses in an elastic
 halfspace due to a point load at the surface. In our notation, the
-stress tensor field due to topography, $M(x,y,z)$, is given by
+stress tensor field due to topography is given by
 
@@
 where $G^B(x,y,z)$ are the Boussinesq solutions for stresses in a
-halfspace due to a vertical point load on the surface (see Appendix A1)
-and $F_v(x,y) = \rho g h(x,y)$. Note that $h(x,y)<0$ since $z<0$ is
+halfspace due to a vertical point load on the surface (see Appendix A1), 
+$F_v(x,y) = \rho g h(x,y)$, and $h(x,y)$ is topography. 
+Note that $h(x,y)<0$ since $z<0$ is
 depth. $M^C(x,y,z)$ is the component of the stress field due to the
@@
 source load in the $i$ direction on the halfspace surface (see Appendix
-A1). The the horizontal loading functions are given by
+A1). The horizontal loading functions are given by
 
@@
 \begin{split}
-F_{hor, \, yz}(x,y) = ( \rho g h(x,y) + \sigma_{yy}^B(x,y,0) + T_{yy} )\, \frac{\partial h}{ \partial y} \\
+F_{h, \, yz}(x,y) = ( \rho g h(x,y) + \sigma_{yy}^B(x,y,0) + T_{yy} )\, \frac{\partial h}{ \partial y} \\
 \quad + (M_{xy}^{B}(x,y,0) + T_{xy}) \frac{\partial h}{ \partial x}\; . 
@@
 functions for the Boussinesq and Cerruti point-source solutions were
-calculated at regular points in large square, 2-D grid at each depth
+calculated at regular points in large 2-D grid at each depth
 considered, with the point-source centered in the grid (see Table
 \ref{table:convo_params} for model parameters). A mask was applied to
-each grid such that values outside the kernel radius (i.e.~the `corners'
-of the array) were zero, so that the kernel is circular. The size of the
+each of the discretized Green's functions such that values outside a 
+radius (i.e. the `corners' of the array) were set to zero, yielding a
+circular array. The size of the
 grid was chosen to be quite large to incorporate potential contributions
@@
 grid, we pad the Green's function array with zeros. Because of
-singularities in the Green's functions at depth $z=0$, we use
+singularities in the Green's functions at $z=0$, we use
 $\sigma^B(x,y,z)$ with $z=851$ m, the shallowest level of our
@@
 we compute $G^C(x,y,z)$. The six stress tensor components calculated at
-the regular grid point are linearly interpolated to points describing
+the regular grid points are linearly interpolated to points describing
 the faults. Because the fault points are completely surrounded by the
@@
 
-Topographic stresses on the Wenchuan faults are in the 1--10s of MPa
-range (Figure~\ref{fig:lms_topo_stresses_rot}, ~\ref{fig:fault_stress_3d}). 
+Topographic stresses on the Wenchuan faults are on the order 1--10s MPa
+(Figure~\ref{fig:lms_topo_stresses_rot}, ~\ref{fig:fault_stress_3d}). 
 Stresses are highest in the southwest,
@@
 stress is not typically aligned with either cardinal horizontal
-component, and is typically larger than $M_{zz}$ above 10 km. Maximum
+direction, and is typically larger than $M_{zz}$ above 10 km. Maximum
 $M_{zz}$ is near 80 MPa, on the southwestern Beichuan fault below the
@@
 km depth: black and red lines signify most and least compressive
-horizontal stresses. Other symbols are the same as in Figure
+horizontal stresses, respectively. Other symbols are the same as in Figure
 1. Stresses shown are down-sampled from the
-discretization used in the calculations by a factor of 6.}
+discretization used in the calculations by a factor of six.}
 \label{fig:lms_topo_stresses_rot}
@@
 Because the compressive stresses are near equal, $M$ contains a large
-isotropic component and a smaller shear component. Consequently, $M$
+isotropic component and a smaller deviatoric component. Consequently, $M$
 resolves on the Wenchuan faults with a large $\sigma^M_n$ (median of
@@
 highest, such as below the Pengguan massif (P). Fault and topography
-share the same scale. No vertical exaggeration. (B) $\tau_d^M$ (colors)
+share the same scale, with no vertical exaggeration. (B) $\tau_d^M$ (colors)
 and dip slip (contours, 1 m interval) on the \citet{qi2011}
 `rough' slip model. (C) $\tau_s^M$ (colors) and strike slip (contours, 1
-m interval) on the \citet{qi2011} `rough' slip model.}
+m interval) on the \citet{qi2011} slip model.}
 \label{fig:fault_stress_3d}
@@
 fault, where present) have $\tau^M$ in the direction of coseismic slip.
-The full stress tensor $M$ is not significantly different in these
+$M$ is not significantly different in these
 locations, but because of the low dip angle, $M_{zz}$ contributes more
@@
 dominated by horizontal compression, leading to reverse-sense shear.
-Additionally, the stresses caused by the Pengguan massif locally resolve
+The stresses caused by the Pengguan massif locally resolve
 as right-lateral on these segments as well. Coseismic slip on these
@@
 \ref{fig:feng_slip_sig_n_scatter}). This raises the possibility that
-$\sigma^M_n$ is capable of limiting slip once failure has occurred, and
+topographic loading of these faults contributes to limiting coseismic
+slip once failure has occurred, and
 may have implications for estimations of dynamic friction and the
@@
 of the \citet{feng2010} coseismic slip model. Trendlines are
-L1 regressions and do not include points with no slip. NE B =
-Northeastern Beichuan fault. C B = central Beichuan fault. SW B =
-Southwestern Beichuan fault. P = Pengguan fault.}
+L1 regressions and do not include points with no slip. ``NE B'' =
+Northeastern Beichuan fault. ``C B'' = central Beichuan fault. ``SW B'' =
+Southwestern Beichuan fault. ``P'' = Pengguan fault.}
 \label{fig:feng_slip_sig_n_scatter}
@@
 consistent with published coseismic slip models of the Wenchuan
-earthquake using a Bayesian estimation. Bayesian estimation yields
-posterior probability density functions of the model parameters \citep[e.g.,]
-[]{mosegaard1995}. We first estimate posteriors of $T$ cosistent with
+earthquake using a Bayesian estimation, resulting in posterior probability 
+density functions of the model parameters. 
+We first estimate posteriors of $T$ cosistent with
 the coseismic slip models, and then estimate $\mu$ and $\phi$ consistent
@@
 where $M$ is described above, $T$ is the tectonic stress tensor, and $L$
-is the lithostatic pressure tensor. $L$ is isotrapic, with diagonal
+is the lithostatic stress tensor. $L$ is isotropic, with diagonal
 components equal to $\rho g z$. We assume that $T$ is latterally
@@
 \citep[e.g.,][]{townend2000}, are thus we parameterize the components of $T$
-as scalars multiplied by lithostatic pressure.
+as scalars multiplied by lithostatic pressure, denoted as $T^\prime$
 
@@
 our posterior only gives likelihood of $T$ relative to the most likely
-estiamte (MLE) \citep{tarantola2005}. We follow a Monte-Carlo strategy,
+estimate (MLE) \citep{tarantola2005}. We follow a Monte-Carlo strategy,
 where samples of the prior PDF are retained as samples of the posterior
@@
 of maximum principal stress are taken from a uniform distribution
-between $\rho g z$ and 2.5 $\rho g z$. Samples of the minumum principal
+between $\rho g z$ and 2.5 $\rho g z$. Samples of the minimum principal
 stress are from a uniform distribution between 0 and the value for
@@
 stress $T^\prime_{\mathrm{max}}$ is broadly east-west for all models,
-with a mode trending at $90^{\circ} -- 105^{\circ}$.
+with a mode trending at $90^{\circ}$--$105^{\circ}$.
 $p(T^\prime_{\mathrm{max}}|D)$ for each slip model increases from
@@
 slight decrease in relative likelihood past the initial mode at
-$T^\prime =$ 0.5--1 (recall that $T^\prime$ is relative to $\rho g z$).
+$T^\prime =$ 0.5--1.
 The low likelihood below $T^\prime \approx 0.5$ indicates that lower
@@
 very low $T^\prime_{\mathrm{max}}$, and have a low probability density,
-as it is unlikely that very low $T^\prime_{\mathrm{max}}$ values can
+as it is unlikely that tectonic stress with very low $T^\prime_{\mathrm{max}}$
+values can
 overcome sinistral and normal-sense topographic shear stresses to cause
@@
 Colors indicate magnitude of $T^\prime_{\mathrm{max}}$. Contour lines
-indicate relative density (i.e., likelihood) of posteriors; darker lines
-mean higher densities.}
+indicate relative density (i.e., likelihood) of posteriors (darker lines
+signify higher densities).}
 \label{fig:mu_phi_scatters}
@@
 
-We define a joint posterior, $p_{J}(T^\prime | D)$, by the samples of
-the individual posteriors estimated from each slip model that common.
+We define a joint posterior, $p_{J}(T^\prime | D)$, by the samples that 
+are common to the individual posteriors estimated from each slip model.
 Unsuprisingly, given the broad similarity between the posteriors from
@@
 \includegraphics[width=40pc]{../figures/joint_pdfs.pdf}
-\caption{Joint model results. (A) Samples of
+\caption{(a) Samples of
 $p_{\mathrm{J}}(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$,
 along with marginals of $T^\prime_{\mathrm{max}}$ and
-$T^\prime_{\mathrm{min}}$. (B) Samples of
-$p_{\mathrm{J}}(T^\prime_{\mathrm{max}}, T^\prime_{\mathrm{min}}/T^\prime_{\mathrm{max}} | D)$,
+$T^\prime_{\mathrm{min}}$. (b) Samples of
+$p_{\mathrm{J}}(T^\prime_{\mathrm{\max}}, T^\prime_{\mathrm{\min}}/T^\prime_{\mathrm{max}} | D)$,
 along with marginals of $T^\prime_{\mathrm{max}}$ and
-$T^\prime_{\mathrm{min}}T^\prime_{\mathrm{max}}$. (C) Histogram of
-azimuths of $T^\prime_{mathrm{max}}$. (D) Samples of
-$p_{\mathrm{J}}(\mu, \phi | D)$, with marginals. Colors in the
-scatterplot indicate magnitude of $T^\prime_{\mathrm{max}}$. Contour
-lines indicate relative density (i.e., likelihood) of posteriors; darker
-lines mean higher densities.}
+$T^\prime_{\mathrm{min}} / T^\prime_{\mathrm{max}}$. (c) Histogram of
+azimuths of $T^\prime_{\mathrm{\max}}$. (d) Samples of
+$p_{\mathrm{J}}(\mu, \phi | D)$, with marginals, where color of the samples
+indicate magnitudes of $T^\prime_{\mathrm{\max}}$ and contour
+lines indicate relative density (i.e., likelihood) of posteriors (darker
+lines signify higher densities).}
 \label{fig:joint_posteriors}
@@
 fields on faults (see Section
-\ref{previous-work-on-topographic-stresses} for some examples). This is
-important to address because most studies of fault rupture dynamics
+\ref{previous-work-on-topographic-stresses} for some examples), even
+though it may have important ramifications to the earthquake process.
+Most studies of fault rupture dynamics
 assume either a homogeneous or stochastic shear stress distribution
@@
 estimation, resulting in a PDF of tectonic stress, as well as $\mu$ and
-$\phi$, rather than solving for just the minimum tectonic stress
+$\phi$, rather than just solving for the minimum tectonic stress
 required for faulting, as done by \citet{luttrell2011}.
@@
 
-The spatial variation in topographic stresses decreases in frequency and
+The spatial variation of topographic stresses decreases in frequency and
 amplitude with depth. This is not surprising, because with increasing
@@
 topographic features (i.e., individual mountains). The spatial
-variablity of the topography stress with depth is similar to the spatial
+variablity of the topographic stress with depth is similar to the spatial
 variablity of coseismic slip in the models considered \citep[e.g.,][]
-{zhang2011}, which are similarly more smooth at depth. Some of the
+{zhang2011}, which are both smoother at depth. Some of the
 estimated slip variability is likely partially due to the more limited
-detectability of displacements at depth versus near the surface.
+resolution of coseismic slip at depth using geodetic data.
 However, the negative spatial correlations of slip versus stress
@@
 Drilling Project several years after the 2008 earthquake \citep{cui2014},
-which show $\sigma_H {\mathrm{\max}}$ to be more orthogonal to the fault
+which show $\sigma_{H{\mathrm{\max}}}$ to be more orthogonal to the fault
 trace, suggesting that much of the right-lateral component of shear
@@
 orientation data from after the earthquake at the WFSD-1 drill hole
-\citep{cui2014} (blue arrows). Other symbology is the same as in Figure
-\ref{fig:lms_map}. Stresses shown are downsampled from grid resolution
-by a factor of 9.}
+\citep{cui2014} (blue arrows). Other symbols are as in Figure
+\ref{fig:lms_map}. Stresses shown are downsampled from our computational 
+grid resolution by a factor of nine.}
 \label{fig:lms_stress_map}
@@
 topographic stresses resisting reverse and right-lateral slip on the
-faults. We find that the highest likely ratio of strike slip to dip slip
+faults. We find that the highest likely ratio of strike-slip to dip-slip
 shear along the Wenchuan earthquake faults is close to 1. This is
@@
 Himalaya and Tibet, which is almost certainly due to Indo-Asian plate
-collision, has almost completely decayed at the Longmen Shan, some 850
+collision, has significantly decayed at the Longmen Shan, some 850
 km northeast of the easternmost Himalaya. Therefore, contraction across
@@
 into the Beichuan fault \citep{kuo2014}. However, it should be noted that
-$\mu \approx 0.4$ is likely relatively high likelihood in our posteriors
-(Figure \ref{fig:joint_posteriors}). These values of friction, however,
+$\mu \approx 0.4$ is relatively high likelihood in our posteriors
+(Figure \ref{fig:joint_posteriors}). These values of friction 
 are lower than typical values derived from laboratory experiments on
@@
 coefficients. We use the same failure conditions as in determing the
-posteriors above, and that optimally oriented faults exist.
+posteriors above, and assume optimally oriented faults exist
+(i.e., we do not consider the generation of new faults in relatively 
+intact rock).
 Additionally, we evaluate the relative contributions of $T$, $\phi$ and
@@
 Figure \ref{fig:cfr_ratios} shows
-$log (\mathrm{CFR}_o / \mathrm{CFR}_f)$ plotted against $\mu$, $\phi$
+$\log (\mathrm{CFR}_o / \mathrm{CFR}_f)$ plotted against $\mu$, $\phi$
 and $T^\prime_{xx}$ on the Beichuan fault for all samples. Though
@@
 \caption{Comparison of Coulomb failure ratio (CFR) on the Beichuan fault
-from the \citet{zhang2011} model to CFR on an
+from the \citet{zhang2011} coseismic slip model to CFR on an
 optimally-oriented fault with $\mu = 0.6$, versus estimated $\mu$ on the
@@
 in the crust \citep[e.g.,][]{jeffreys1924}, these stresses have the
-potential to influence orogenesis, from changing the deformation state
-in the interior of an orogen \citep[e.g.,][]{dewey1988, molnar1988}, the
+potential to influence orogenesis. The result may be to change the deformation
+state in the interior of the orogen \citep[e.g.,][]{dewey1988, molnar1988}, the
 convergence rates of the plates surrounding the orogen \citep[e.g.][]
@@
 Though much work has been done exploring the feedback mechanisms between
-topography and deformation, topography is typically greatly smoothed,
-stresses from it are folded into gravitational potential energy
+topography and deformation; however, topography is typically greatly smoothed,
+and topographic stresses are folded into gravitational potential energy
 estimates of the entire lithosphere, and the whole lithospheric column
@@
 
-Our results indicating that topographic stresses on the Wenchuan
+Our results that topographic stresses on the Wenchuan
 earthquake faults are largely loaded in the opposite sense to the
-coseismic slip direction shows that gravitational collapse is probably
+coseismic slip direction indicates that gravitational collapse is probably
 not the driver of reverse faulting in the Longmen Shan. However, as
@@
 horizontal tectonic stress, so any slip on them may be in response to
-topographic loading and therefore to gravitational collapse of the high
+topographic loading and thus to gravitational collapse of the high
 Longmen Shan and its hinterland. $M_{H{\mathrm{\max}}$ also changes
-orientation from fault (or rangefront) normal to closer to more oblique
+orientation from fault rangefront normal to closer to more oblique
 near the front of the range (Figure \ref{fig:lms_topo_stresses_rot}),
@@
 reverse faulting on the Wenchuan faults if subhorizontal shear zones or
-channels are present, and are very weak so shear can occur at relatively
+channels are present, and are very weak so shear can occur at 
 low $\tau$ relative to $\sigma_n$. This conclusion has been reached in
@@
 indicating that is a general requirement of gravitational spreading. Our
-calculations of $\mu$ and $\phi$ are based on $T$, which is weighted
+calculations of $\mu$ and $\phi$ are based on $T$, which is biased
 towards the high slip patches in the upper crust (Equation
 \ref{eqn:rake_misfit}), yield low $\mu$ and low to moderate $\phi$,
-which is probably insufficient for slip at high pressures at depth,
-though $\phi$ may be higher there. Similarly, a low viscosity horizontal
+which is probably insufficient for slip at high pressures at depth.
+Similarly, a low viscosity horizontal
 channel that underlies eastern Tibet, likely much deeper than the
@@
 
-This is all quite speculative, and we wish only to describe the
+These arguments are all quite speculative, and we wish only to describe the
 conditions under which gravitational collapse can occur given the
@@
 $\sigma^M_n$ reaches up to 100 MPa on mid-crustal thrust flats present
-in some coseismic slip models. These topographic stresses generally
-resist the slip inferred during the 2008 Wenchuan earthquake, with
-$\tau^M$ opposed to the slip direction over much of the fault,
-indicating that weight of the topography resist coseismic slip. High
+in some coseismic slip models. The direction of $\tau_M$ is generally
+opposed to coseismic slip inferred during the 2008 Wenchuan earthquake,
+indicating that weight of the topography resists coseismic slip. High
 values of $\sigma^M_n$ increase the frictional resistance to slip,
@@
 resulting in posteriors representing likelihood of tectonic stress,
-static friction, and a pore pressure parameter. The results of our
-Bayesian inversion indicate that the maximum tectonic stress is oriented
+static friction, and a pore pressure parameter. The posteriors
+indicate that the maximum tectonic stress is oriented
 \textasciitilde{}E-W and has a likely minimum of 13.25 MPa km$^{-1}$.
@@
 
-These results suggest that local stress conditions caused by topographic
-loading are unfavorable to slip on the Wenchuan earthquake faults, but
-because of the low friction coefficients and moderate fluid pressures,
-slip occurred on these faults instead of more favorably-oriented faults
-elsewhere in the region.
+Slip occurred on these faults instead of more favorably-oriented faults
+elsewhere in the region, due to the inferred low coefficient of friction
+and moderate fluid pressures.
 
@@
 loads}\label{greens-functions-for-point-source-loads}
-
+For completeness, we reproduce the Boussinesq \citep[e.g.,][]{jeffreys1970}
+and Cerruti \citep[e.g.,][]{love1927} solutions here.
 Note that in these solutions, $\lambda$ and $\mu$ are re-defined, and
@@
 
-\begin{thebibliography}{85}
+\begin{thebibliography}{87}
 \providecommand{\natexlab}[1]{#1}
@@
 \bibitem[{\textit{Jeffreys}(1924)}]{jeffreys1924}
-Jeffreys, H. (1924), \textit{The earth: its origin, history and physical
+Jeffreys, H. (1924), \textit{The Earth: its origin, history and physical
   constitution}, Cambridge University Press.
+
+\bibitem[{\textit{Jeffreys}(1970)}]{jeffreys1970}
+Jeffreys, H. (1970), \textit{The Earth}, Cambridge University Press.
 
@@
 
+\bibitem[{\textit{Love}(1927)}]{love1927}
+Love, E.~A.~H. (1927), \textit{The mathematical theory of elasticity.},
+  Cambridge University Press.
+
 \bibitem[{\textit{Loveless and Meade}(2011)}]{loveless2011}

### 2014-09-03 scripts/e_asia_bouss.py
modified bouss (tested) and cerruti (not finished running) for poisson ratio of 0.35

@@ -5,13 +5,16 @@ import scipy.fftpack as sf
 import time
 import h5py
 
-stress_dir = '/cmld/data7/styron/wenchuan_eq/wench_output/'
+#stress_dir = '/cmld/data7/styron/wenchuan_eq/wench_output/'
+stress_dir = '../../stress_files/'
 stress_file = stress_dir + 'e_asia_bouss.h5'
 
 print 'setting up problem...'
 rho = 2700  # density in kg m^-3
 g = 9.81    # gravitational force in m s^-2
 Fv =  -rho * g
+lame_1 = 70e9
+shear_mod = 30e9
 study_res = 851 # resolution for topography, filters, etc.
 z_res = 1000
 b_conv_mode = 'valid'
@@ -67,6 +70,7 @@ for comp in comp_list:
 
     for i, z in enumerate(z_vec):
        b_dict[comp][:,:,i] = hs.do_b_convo( component = comp,  z = z, Fv = Fv,
+                                        lamb = lame_1, mu = shear_mod,
                                         load = topo, load_mode = 'topo',
                                         conv_mode = b_conv_mode,
                                         kernel_radius = kernel_rad,

### 2014-09-03 scripts/e_asia_cerruti.py

@@ -5,7 +5,8 @@ import h5py
 
 t0 = time.time()
 
-stress_dir = '/cmld/data7/styron/wenchuan_eq/wench_output/'
+#stress_dir = '/cmld/data7/styron/wenchuan_eq/wench_output/'
+stress_dir = '../../stress_files/'
 b_stress_file = stress_dir + 'e_asia_bouss.h5'
 c_stress_file = stress_dir + 'e_asia_cerruti_topo.h5'
 stress_file = stress_dir + 'e_asia_topo_stress.h5'
@@ -21,6 +22,8 @@ conv_mode = 'same'
 
 z_min = 851.
 z_max = 35851.
+lame_1 = 70e9
+shear_mod = 30e9
 z_len = (z_max - z_min) / z_res + 1
 z_vec = np.linspace(z_min, z_max, num=z_len)
 
@@ -75,11 +78,13 @@ for comp in comp_list:
 
     for i, z in enumerate(z_vec):
         cerr_x[comp][:,:,i] = hs.do_c_convo( component=comp, f_dir = 'x', z=z,
+                                       lamb = lame_1, mu = shear_mod,
                                        load = Fh_x, kernel_res = study_res,
                                        kernel_radius = kernel_rad,
                                        conv_mode = conv_mode) * 1e-6
 
         cerr_y[comp][:,:,i] = hs.do_c_convo( component=comp, f_dir = 'y', z=z,
+                                       lamb = lame_1, mu = shear_mod,
                                        load = Fh_y, kernel_res = study_res,
                                        kernel_radius = kernel_rad,
                                        conv_mode = conv_mode) * 1e-6

### 2014-09-03 tect_stress/scripts/feng_bayes_fail.py
updated all run scripts for negative lower bound on Tmin priors, except for qi smooth rake (waiting for testing).  Also running 200k (instead of 100k) trials.

@@ -29,7 +29,7 @@ rho = 2700
 g = 9.81
 
 # priors for pore fluid pressure (phi[da])
-phi_priors = np.random.random(1e5)
+phi_priors = np.random.random(2e5)
 phi_priors = phi_priors[t_prior_df.index]
 
 # make dataframe

### 2014-09-03 tect_stress/scripts/feng_bayes_rake.py

@@ -17,7 +17,7 @@ lms = pd.concat((fb, fp), axis=0)
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700
@@ -27,7 +27,7 @@ g = 9.81
 # These are functions of lithostatic pressure (rho g depth)
 # Priors for each are uniform [-2, 2).
 s1s = np.random.uniform(0,2.5, n_trials)
-s3s = np.random.uniform(0, 1, n_trials) * s1s
+s3s = np.random.uniform(-1, 1, n_trials) * s1s
 thetas = np.random.uniform(0, 2 * np.pi, n_trials)
 
 xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)

### 2014-09-03 tect_stress/scripts/field_bayes_fail.py

@@ -25,7 +25,7 @@ rho = 2700
 g = 9.81
 
 # priors for pore fluid pressure (phi[da])
-phi_priors = np.random.random(1e5)
+phi_priors = np.random.random(2e5)
 phi_priors = phi_priors[t_prior_df.index]
 
 # make dataframe

### 2014-09-03 tect_stress/scripts/field_bayes_rake.py

@@ -15,7 +15,7 @@ lms = pd.read_csv('../../slip_models/fielding/field_topo_stress.csv',
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700
@@ -25,7 +25,7 @@ g = 9.81
 # These are functions of lithostatic pressure (rho g depth)
 # Priors for each are uniform [-2, 2).
 s1s = np.random.uniform(0,2.5, n_trials)
-s3s = np.random.uniform(0, 1, n_trials) * s1s
+s3s = np.random.uniform(-1, 1, n_trials) * s1s
 thetas = np.random.uniform(0, 2 * np.pi, n_trials)
 
 xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)

### 2014-09-03 tect_stress/scripts/qi_rough_bayes_rake.py

@@ -24,7 +24,7 @@ lms = lms[lms.slip_m>=0.001]
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700
@@ -33,21 +33,21 @@ g = 9.81
 # Priors for tectonic stresses (txx, tyy, txy).
 # These are functions of lithostatic pressure (rho g depth)
 # Priors for each are uniform [-2, 2).
-#s1s = np.random.uniform(1,2.5, n_trials)
-#s3s = np.random.uniform(0, 1, n_trials) * s1s
-#thetas = np.random.uniform(0, 2 * np.pi, n_trials)
-#
-#xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)
-#yys = scv.yy_stress_from_s1_s3_theta(s1s, s3s, thetas)
-#xys = scv.xy_stress_from_s1_s3_theta(s1s, s3s, thetas)
-#
-#del s1s, s3s, thetas
-#
-#xxs = xxs.reshape([n_trials, 1])
-#yys = yys.reshape([n_trials, 1])
-#xys = xys.reshape([n_trials, 1])
-#
-#t_priors = np.concatenate((xxs, yys, xys), axis=1)
+s1s = np.random.uniform(1,2.5, n_trials)
+s3s = np.random.uniform(-1, 1, n_trials) * s1s
+thetas = np.random.uniform(0, 2 * np.pi, n_trials)
+
+xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)
+yys = scv.yy_stress_from_s1_s3_theta(s1s, s3s, thetas)
+xys = scv.xy_stress_from_s1_s3_theta(s1s, s3s, thetas)
+
+del s1s, s3s, thetas
+
+xxs = xxs.reshape([n_trials, 1])
+yys = yys.reshape([n_trials, 1])
+xys = xys.reshape([n_trials, 1])
+
+t_priors = np.concatenate((xxs, yys, xys), axis=1)
 
 # Columns for search dataframe
 search_df_cols = ['iter','txx', 'tyy', 'txy', 'pt_index', 'depth', 'strike',
@@ -55,27 +55,27 @@ search_df_cols = ['iter','txx', 'tyy', 'txy', 'pt_index', 'depth', 'strike',
 
 
 ## making index list
-#iter_range = np.arange(n_trials, dtype='float')
[…]

### 2014-09-03 tect_stress/scripts/shen_bayes_fail.py

@@ -27,7 +27,7 @@ rho = 2700
 g = 9.81
 
 # priors for pore fluid pressure (phi[da])
-phi_priors = np.random.random(1e5)
+phi_priors = np.random.random(2e5)
 phi_priors = phi_priors[t_prior_df.index]
 
 # make dataframe

### 2014-09-03 tect_stress/scripts/shen_bayes_rake.py

@@ -14,7 +14,7 @@ lms = pd.read_csv('../../slip_models/shen/shen_topo_stress.csv')
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700
@@ -24,7 +24,7 @@ g = 9.81
 # These are functions of lithostatic pressure (rho g depth)
 # Priors for each are uniform [-2, 2).
 s1s = np.random.uniform(0,2.5, n_trials)
-s3s = np.random.uniform(0, 1, n_trials) * s1s
+s3s = np.random.uniform(-1, 1, n_trials) * s1s
 thetas = np.random.uniform(0, 2 * np.pi, n_trials)
 
 xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)

### 2014-09-03 tect_stress/scripts/zhang_bayes_rake.py

@@ -15,7 +15,7 @@ lms.rake_deg = hsp.get_rake_from_shear_components(strike_shear=lms.slp_strk_m,
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700
@@ -25,7 +25,7 @@ g = 9.81
 # These are functions of lithostatic pressure (rho g depth)
 # Priors for each are uniform [-2, 2).
 s1s = np.random.uniform(0,2.5, n_trials)
-s3s = np.random.uniform(0, 1, n_trials) * s1s
+s3s = np.random.uniform(-1, 1, n_trials) * s1s
 thetas = np.random.uniform(0, 2 * np.pi, n_trials)
 
 xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)

### 2014-09-04 tect_stress/scripts/feng_bayes_rake.py
removed old results files.  scripts updated (except for qi smooth?). am going to restart w/ larger RAM and do 200k runs.

@@ -17,7 +17,7 @@ lms = pd.concat((fb, fp), axis=0)
 np.random.seed(69)
 
 # some inital constants
-n_trials = 2e5
+n_trials = 1.5e5
 
 n_points = len(lms.index)
 rho = 2700

### 2014-09-04 tect_stress/scripts/qi_rough_bayes_rake.py

@@ -24,7 +24,7 @@ lms = lms[lms.slip_m>=0.001]
 np.random.seed(69)
 
 # some inital constants
-n_trials = 2e5
+n_trials = 1.5e5
 
 n_points = len(lms.index)
 rho = 2700

### 2014-09-04 slip_models/fielding/calc_fielding_topo_stress.py
calculated new values for zhang, shen and fielding (others are missing scripts)

@@ -9,7 +9,8 @@ import time
 t0 = time.time()
 print 'setting up and scaling stress arrays...'
 
-stress_f_dir = '/Volumes/cmld/data7/styron/wenchuan_eq/wench_output/'
+#stress_f_dir = '/Volumes/cmld/data7/styron/wenchuan_eq/wench_output/'
+stress_f_dir = '../../../stress_files/'
 
 stress_file = stress_f_dir + 'e_asia_topo_stress.h5'
 
@@ -102,9 +103,9 @@ for i in lms.index:
 
 print 'done!  now making new dataframes.'
 
-lms.to_csv('field_topo_stess.csv')
-lms[lms.segment == 'beichuan'].to_csv('field_beich_topo_stress.csv')
-lms[lms.segment == 'pengguan'].to_csv('field_peng_topo_stress.csv')
+lms.to_csv('field_topo_stress_035_pr.csv')
+#lms[lms.segment == 'beichuan'].to_csv('field_beich_topo_stress.csv')
+#lms[lms.segment == 'pengguan'].to_csv('field_peng_topo_stress.csv')
 
 
 

### 2014-09-04 slip_models/shen/calc_shen_fault_stresses.py

@@ -9,7 +9,8 @@ import time
 t0 = time.time()
 print 'setting up and scaling stress arrays...'
 
-stress_f_dir = '/Volumes/cmld/data7/styron/wenchuan_eq/wench_output/'
+#stress_f_dir = '/Volumes/cmld/data7/styron/wenchuan_eq/wench_output/'
+stress_f_dir = '../../../stress_files/'
 
 stress_file = stress_f_dir + 'e_asia_topo_stress.h5'
 
@@ -101,8 +102,8 @@ for i in lms.index:
 
 print 'done!  now making new dataframes.'
 
-lms.to_csv('shen_topo_stess.csv')
-lms[lms.segment == 'beichuan'].to_csv('shen_beich_topo_stress.csv')
-lms[lms.segment == 'pengguan'].to_csv('shen_peng_topo_stress.csv')
+lms.to_csv('shen_topo_stress_035_pr.csv')
+#lms[lms.segment == 'beichuan'].to_csv('shen_beich_topo_stress.csv')
+#lms[lms.segment == 'pengguan'].to_csv('shen_peng_topo_stress.csv')
 
 

### 2014-09-04 slip_models/zhang/calc_gz_fault_stresses.py

@@ -13,8 +13,8 @@ print 'setting up and scaling stress arrays...'
 
 
 wench_slip_dir = 'multi/'
-stress_f_dir = '/cmld/data7/styron/wenchuan_eq/wench_output/'
-
+stress_f_dir = '../../../stress_files/'
+#stress_f_dir = '/cmld/data7/styron/wenchuan_eq/wench_output/'
 stress_file = stress_f_dir + 'e_asia_topo_stress.h5'
 
 fs1 = wench_slip_dir + 's1_xyz.csv'
@@ -426,7 +426,7 @@ lms_slip = pd.read_csv('lms_slip.csv', index_col=0)
 
 lms = pd.concat([lms, lms_slip], axis=1)
 
-lms.to_csv('lms_stress_slip.csv')
+lms.to_csv('lms_stress_slip_035_pr.csv')
 
 t5 = time.time()
 print 'done with everything in', t5-t0,'s'

### 2014-09-04 revisions/rebuttal.md
first revisions commit

## R1
> The complete stress tensor is divided into contributions from topography,
> tectonic, and lithostatic stress (eq 9). The tectonic stress is then assumed
> to increase linearly with depth "so that the entire crust is near the critical
> failure envelope" (line 247), referencing Towend and Zoback 2000. This is
> essentially just a statement of Byerlee's law, in which brittle failure occurs
> in the upper crust at a stress level approximately equal to the lithostatic
> pressure applied to the rock scaled by ~0.6 (this study's parameter T'?).
> However, this means that the formulation expressed in equation 9 is counting
> the lithostatic contribution of stress twice. The Towend paper derives its
> results from deep drilling studies, so it would be more appropriate to assume
> that the total stress S increases linearly with depth in this manner.

(add whole tensor thing)

## R2
> The authors make the "Wallace-Bott" assumption (line 277) that fault slip
> direction aligns with maximum horizontal principal stress direction, and the
> subsequent analysis of tectonic stress is made with the criteria that
> acceptable solutions minimize the difference between these two (equation 11
> and lines 300-301). However, several of the results described in discussion
> section 6.2 interpret the alignment of the modeled tectonic stress
> orientation and coseismic slip direction as a meaningful result: e.g., line
> 401, line 407, line 417. For example, that the shear stress has non-trivial
> contributions of both strike and dip-slip was required since the stress must
> align with the oblique-slip earthquake.

In these comments, the reviewer basically asserts that we claim that is
significant that our model's predictions match the observations, and then gives
several example lines (401, 407, 417). Though this is not the interpretation of
the paragraph that we had intended, it is apparent how this interpretation
could be made. Nonetheless, the specific lines seem to be off the mark in this:

Line 401 is basically a quick description of the results (E-W maximum principal
tectonic stress produces oblique shear on the fault), and is not really a claim
about the significance of the results.

Line 407 is a comparison of our estimated maximum principal stress and stress
measurements from the WFSD drill hole at that location; the drill hole stress
measurements were not used in our inversion, and therefore this is an 
independent confirmation of our results, not a tautological statement.

Line 417 states that dip-slip and strike-slip shear are nearly equal in the
most-likely model results, and then compares that result to an orogen-scale
elastic block modeling study, not to the coseismic slip models used in the
inversion.

However, in one instance in this paragraph (lines 403-404) we are guilty as
charged, and make amendments. We change the lines:

> This stress configuration is compatible with the observed kinematics of the
> Wenchuan faults, and in close agreement pre-earthquake stress measurements
> near the rupture zone...

to

> This stress configuration is in close agreement with pre-earthquake stress
> measurements near the rupture zone...

## R3
> Line 188: In the Liu and Zoback formulation, tectonic stress is an integrated
> part of the horizontal loading functions (equations 5 and 6). Is it 
> appropriate to neglect tectonic stress in these equations but then add it 
> back in linearly during the analysis of section 4? I expect this has some
> ramifications.

This was something we puzzled over for a while. Ultimately, we chose our
approach because it is less than obvious how to incorporate a depth-dependent
tectonic stress into the calculations. The horizontal loading functions
specify horizontal surface loads that are then propagated into the subsurface using Cerruti's solutions [eq. xx]. The Liu and Zoback formulations of the loading
functions seems to require a fixed tectonic stress, or the tectonic stress at
zero depth. Because we assume that tectonic stress increases linearly from zero
with depth, the tectonic stress at zero depth (sea level) is zero.

## R4
> Line 193 assumes a "Poisson halfspace", presumably meaning an elastic
> halfspace made up of a Poisson solid with Poisson ratio of 0.25. While this
> is an oft-made assumption, in this situation it has more impact. Figure 2
> nicely illustrates how the topography loading on the fault planes can change
> sign depending on how effective the vertical load is at generating horizontal
> stresses, which is precisely what the Poisson ratio governs. It would be
> helpful, therefore, to know how robust these results are to choice of
> Poisson ratio.

This is a valid point, but it wasn't something that we had tested much because
of the computational effort required, both in terms of programmer time and CPU
time. However, in the intervening months since submitting this manuscript, we
have streamlined the analysis so that the programmer time required is much
less.

[test this].

## R5
> Line 292: do both the tectonic stresses have to be positive? Vertical 
> principal stress could still be the least compressive if Tmin were negative 
> and Tmax were positive, right? This would increase the available permitted 
> parameter space.

In general, neither of the tectonic stresses have to be positive. However,
vertical stress cannot be least compressive if $T_{min}$ were negative, unless
topographic stress produced a strong horizontal compression in the direction of
$T_{min}$ and less strong vertical compression.

To restate this mathematically, for clarity:
Let's consider $T_{min}$ to be aligned N-S, i.e. it is in the $xx$ direction.
The equation for total stress is $S = M + T + L$ (equation 9), and we consider
tectonic stress $T$ to be only horizontal stresses. So the vertical stress is
$M + L$. $L$ is isotropic, so we can ignore it here. The only way that 
$M_{xx} + T_{min} > $M_{zz}$, where $T_{min} < 0$, is to have 
$M_{xx} > M_{zz} - T_{min}$. Given that typically $M_{zz} > M_{xx}$ in our 
results (Section 3, lines 224-225 of the original manuscript), it is not
possible in this case to have a negative $T_{min}$ and still have vertical
stress ($S_{zz}$) as the least compressive stress (which is required by reverse
fault slip).

Furthermore,

## R6
> The emphasis on the Bayesian estimation providing a doubly bounded constraint 
> on tectonic stress (379) or on the on the ability of the study to identify 
> more than a minimum bound of the tectonic stress (line 373) is overstated. 
> The fundamental gist of this study is that topography provides a resistive
> force that must be overcome by tectonic stress (lines 140-142). This provides
> a physical minimum bound for tectonic stress, but there is no information
> considered that physically provides any upper bound. Unsurprisingly,
> therefore, the results in figure 6 show a pdf with a lower bound only, as
> noted on line 413. This is not a failure or triumph of the chosen statistical
> scheme, it's a physical limitation of the available data.

The reviewer is absolutely correct in his or her assessment of the ability of
the Bayesian inversion to capture the maximum tectonic stresses for this
particular earthquake, and we are up front about this (lines 323-326). (For
other events with less of a the thrust component, we can constrain maximum
tectonic stress).

The lines 373 and 379, chosen by the reviewer, are in a discussion of
methodological differences between this study and another that only attempts to
constrain the minimum tectonic stress in a similar environment; the differences
here are in approach more than results.

## R7
> I'm concerned that the Bayesian PDF estimation as used for T'min, e.g. in 
> figure 6, results in a biased estimate. Because there's a requirement that 
> Tmin < Tmax, half of parameter space is off limits a priori (lines 293-296).
> It therefore seems unsurprising that the T'min PDF is heavily weighted toward
> zero. Is there a way to re-scale the problem so that the parameters being
> estimated are independent?

This was something else that we spent some time on. The issue with picking
stresses independently is that they're not independent, because $T_{min}$ *has
to* < $T_{max}$. So they can't be sampled indendently, and there are two
approaches that can be taken. One of them is to take a random Nx2 array and
sort each row, so that one column is $T_{max}$ and the other is $T_{min}$.
This approach yields a uniform joint prior probability density over the
available half of the $T_{min}$ x $T_{max}$ parameter space, but gives strongly
nonuniform marginal priors for either $T_{max}$ (which linearly increases from
0 at T=0 to max) or $T_{min}$ (with a commensurate linear decrease).

The other is to do what we did, and take a random Nx1 array for T$_{max}$, and
then take $T_{min}$ as some random value from [0,1) of $T_{max}$. This gives
a uniform prior PDF for $T_{max}$, the same prior for $T_{min}$ as the first
option, and a higher joint probability density towards the low $T_{max}$ end.

So, although the joint probability density is higher towards the low $T_{max}$
end (visible in Figures 6 and 8a) in our approach, the marginal priors for
$T_{max}$ would be much more strongly biased in the first approach, but
$T_{min}$ is unaffected by the sampling strategy. Overall, we feel that our
approach is less biased.

On further reflection, it would be possible to pick a uniform $T_{min}$ and
then build $T_{max}$ around that, but then $T_{max}$ would be non-uniform. We
can think of no particular reason to favor this approach, since researchers are
generally more interested in estimating maximum principal tectonic stress than
minimum.

## R8
> The caption for Figure 4 is truncated on p 43, and is not present anywhere
> else in the text. The figure itself is not self-explanatory, and the details
> aren't sufficiently explained in the results section 3. This seems to be an
> unfortunate LaTEX mishap, but since this model is the primary finding of this
> study and the basis for the subsequent analysis, the reader unfortunately
> cannot just "read past" the omitted details. Some of the obvious gaps are
> noted below.

## R9
> It is not entirely clear which faults are shown in 4a, and not at all clear
> what faults or regions of space are represented in 4b-c (by the way, a b and
> c are not labeled either). What depths are represented in these models? The
> visible portion of the caption notes "same lateral extent, but different
> perspectives", but it is not clear what this means. And why is the maximum
> slip magnitude in "A" 7m, but the maximum component slips in "B" and "C" are
> 14m and 10m?

## R10
> Also, as section 3 represents the principal findings of this study, the 
> reader would benefit from some additional discussion here of the relevant
> features of the models. For example, the text on lines 237-239 and lines
> 244-247 seem to represent and important result worthy of more attention and
> discussion.

## R11
> The introduction of the study says the work intends to "investigate the
> effects of topographic stress on the faults" (line 48) and then "bracket the
> tectonic stress field, pore fluid pressure, and static friction of the fault"
> (line54). However, in beginning to describe the topographic stress resolved on
> faults, section 3 refers to an upcoming manuscript for analysis of how these
> stresses might affect faulting. This is indeed an interesting aspect of the
> study, but either that information should be included here or the focus on
> influencing rupture process should be removed from the abstract and
> introduction.

This study does indeed

## R12
> Overall, the manuscript, though well organized, seems to have an unbalanced emphasis. The specific details of the analytical methods are present, but overall the results should be more fully described and the discussion more fully developed.

## R13
> Line 24: Though I can see the appeal of the selected quotation, it would be more appropriate for an oral presentation or a monograph on the subject. In this context, it comes across as awkward and out of place.

## R14
> Section 1.1: additional relevant references to consider: Fialko, Rivera, and Kanamori (2005 GJI), Ghosh et al (2006 Geology), Flesch et al (2007 GJI)

## R15
> Lines 224-230 describe the magnitudes of stress tensor components, but this would be better explained with a figure. If not as its own figure, could Mzz at least be incorporated into figure 3?

## R16
> Section 6.0 contrasts the methods of this study to those of Luttrell et al. 2011 study of the Maule earthquake. This is an understandable comparison, but because the Wenchuan earthquake includes both strike-slip and dip-slip components, it is actually more similar to the Luttrell and Sandwell (2012 JGR) study of Mid Ocean Ridges that includes both strike-slip and normal earthquakes as constraints.

That study doesn't use coseismic slip models.

## R17
> The introduction lists 7 coseismic slip models of the Wenchuan earthquake (line 124), and then the analysis refers to 5 of them (line 213). As the authors noticed, all these models essentially draw from the same set of geodetic data, and all have essentially the same result, and thus all give this study essentially the same result (line 343). Is there a reason these particular models were chosen for analysis? Why not just choose one, or if they truly are different enough to influence these results, why not use them all? Some justification of this choice is warranted.

## R18
> --Typos 
> Lines 177 and 182: is z positive up or down? Clarify this 
> Line 258: earthquake, not fault, initiated 
> Line 274: "are" -> "and"? complete sentence with period 
> Line 509: typo, delete "to close" 
> Line 556: If you acknowledge some of your data sources, you should acknowledge all of your data sources (e.g., scientific drilling, WSM, etc) 
> Fig 3 and 9 need 33ºN label correction

### 2014-09-04 manuscript/styron_hetland_wenchuan_topo_manuscript.tex
first revisions commit

@@
 \begin{document}
-\title{Topographic stresses and the Wenchuan earthquake}
+\title{The weight of the mountains: Constraints on tectonic stress, friction,
+       and fluid pressure in the 2008 Wenchuan earthquake from estimates of 
+       topographic loading}
 
@@
 \begin{abstract}
-Though it is widely recognized that large mountain ranges may produce
-significant stresses in the earth's crust, these stresses are not
-commonly quantified. Nonetheless, it is possible that active faults near
-high topography may be affected by topographic stresses. In this work,
-we calculate the topographic stress field in the Longmen Shan region of
-eastern Tibet, where faulting occurs at the base of \textasciitilde{} 4
-km relief.
+Though it is widely recognized that large mountain ranges produce significant
+stresses in the Earth's crust, these stresses are not commonly quantified.
+Nonetheless, near large mountains topography may affect fault activity by
+changing the stress balance on the faults. In this work, we calculate the
+stress field from topography in the Longmen Shan (Sichuan, China) and resolve
+those stresses on several models of the faults that ruptured in the 2008
+M$_\mathrm{w}$ 7.9 Wenchuan earthquake. We find that the topography results in
+shear stresses up to 20 MPa and normal stresses up to 80 MPa on the faults,
+with significant variability across the faults. Topographic stresses generally
+load the fault in a normal and left-lateral shear sense, opposite to the
+inferred coseismic slip sense, and thus inhibit the coseismic slip. We estimate
+the accumulated tectonic stress needed to overcome topographic and lithostatic
+stresses so that the direction of maximum shear on the faults is roughly
+collinear with the inferred coseismic slip. We further estimate the static
+friction and pore fluid pressure assuming that the fault was at Mohr-Coulomb
+failure at the time of the Wenchuan earthquake. We use a Bayesian inversion
+strategy, yielding posterior probability distributions for the estimated
+parameters. We find most likely estimates of maximum tectonic compressive
+stress near 0.6 $\rho g z$ and oriented \textasciitilde{}E-W, and minimum
+tectonic stress near 0.2 $\rho g z$. Static friction on the fault is near 0.2,
+and pore fluid pressure is between 0 and 0.4 of the total lithostatic pressure.
 \end{abstract}
 
-\begin{article}
-\section{Introduction}\label{introduction}
-
-\begin{quote}
-``And He has set up on the earth mountains standing firm, lest it should
-shake with you'' -- \emph{The Holy Qur'an}, 15:16
-\end{quote}
-
-Stress is of fundamental importance to many processes in the earth. Both
-the isotropic and deviatioric components of stress
-exert control on the deformation state of the earth at any point in the
-brittle regime. However, unlike other fundamental quantities such as
-temperature, stress is typically difficult to measure \emph{in situ},
-without drilling-based techniques. Therefore, stress is often treated in
-a semi-quantitative manner, with an emphasis on directions and relative
-magnitudes of the principal stresses, either locally or regionally
-\citep[e.g.,][]{angelier1994}. These estimates of stress are commonly
-derived from strain, for example from studies of earthquake focal
-mechanisms \citep[e.g.,][]{michael1987} or of fault slip data
-\citep[e.g.,][]{reches1987, medinaluna2013}.
-
-In areas of substantial relief, high terrain and steep slopes generate
-large stresses in the crust beneath and adjacent to the high topography
-\citep{jeffreys1924, coblentz1996}. Because of the irregularity of
-topography in mountainous regions, the stresses produced by topography
-are also heterogeneous, and may play a prominent role in local or
-regional deformation, particularly if the region is tectonically active.
-For example, shear and normal stresses on a fault from topographic
-loading may push a particular fault closer to or (as suggested by the
-Qur'an) farther from failure, or reorient the net shear stress direction
-on a fault. These effects may affect the localization or deformational
-style in a region. Furthermore, heterogeneous topographic stresses on a
-particular fault may affect the way earthquake ruptures propagate across
-the fault plane. Despite this, the degree to which topographic stresses
-affect faulting has received little direct study.
-
-In this work, we investigate the effects of topographic stresses on the
-faults that ruptured in the 2008 \emph{M}7.9 Wenchuan, China earthquake.
-This earthquake is an ideal candidate for this study because it occurred at
-the base of the Longmen Shan, one of the largest and steepest
-escarpments on Earth (Figure~\ref{fig:lms_map}), and has a well-studied
-coseismic slip distribution characterized by significant along-strike
-variations in coseismic slip. Additionally, because the earthquake
-occurred after \textasciitilde{}2000 years of seismic quiescence,
-postseismic stresses in the lithosphere are likely to be negligible,
-suggesting that the stress state on the faults at the time of rupture
-can be approximated as the sum of topographic, lithostatic, and
-accumulated tectonic stresses. We then use the Mohr-Coulomb failure
-criterion to bracket the tectonic stress field, pore fluid pressure, and
-static friction of the fault.
+\begin{article} \section{Introduction}\label{introduction}
+
+\begin{quote} ``And He has set up on the earth mountains standing firm, lest it
+    should shake with you'' -- \emph{The Holy Qur'an}, 15:16 \end{quote}
+
+Stress is of fundamental importance to many processes in the earth. Both the
+isotropic and deviatioric components of stress exert control on the deformation
+state of the earth at any point in the brittle regime. However, unlike other
+fundamental quantities such as temperature, stress is typically difficult to
+measure \emph{in situ}, without drilling-based techniques. Therefore, stress is
+often treated in a semi-quantitative manner, with an emphasis on directions and
+relative magnitudes of the principal stresses, either locally or regionally
+\citep[e.g.,][]{angelier1994}. These estimates of stress are commonly derived
+from strain, for example from studies of earthquake focal mechanisms
+\citep[e.g.,][]{michael1987} or of fault slip data \citep[e.g.,][]{reches1987,
+medinaluna2013}.
+
+In areas of substantial relief, high terrain and steep slopes generate large
+stresses in the crust beneath and adjacent to the high topography
+\citep{jeffreys1924, coblentz1996}. Because of the irregularity of topography
+in mountainous regions, the stresses produced by topography are also
+heterogeneous, and may play a prominent role in local or regional deformation,
+particularly if the region is tectonically active.  For example, shear and
+normal stresses on a fault from topographic loading may push a particular fault
+closer to or (as suggested by the Qur'an) farther from failure, or reorient the
+net shear stress direction on a fault. These effects may affect the
+localization or deformational style in a region. Furthermore, heterogeneous
+topographic stresses on a particular fault may affect the way earthquake
+ruptures propagate across the fault plane. Despite this, the degree to which
+topographic stresses affect faulting has received little direct study.
+
+In this work, we investigate the effects of topographic stresses on the faults
+that ruptured in the 2008 \emph{M}7.9 Wenchuan, China earthquake.  This
+earthquake is an ideal candidate for this study because it occurred at the base
+of the Longmen Shan, one of the largest and steepest escarpments on Earth
+(Figure~\ref{fig:lms_map}), and has a well-studied coseismic slip distribution
+characterized by significant along-strike variations in coseismic slip.
+Additionally, because the earthquake occurred after \textasciitilde{}2000 years
+of seismic quiescence, postseismic stresses in the lithosphere are likely to be
+negligible, suggesting that the stress state on the faults at the time of
+rupture can be approximated as the sum of topographic, lithostatic, and
+accumulated tectonic stresses. We then use the Mohr-Coulomb failure criterion
+to bracket the tectonic stress field, pore fluid pressure, and static friction
+of the fault.
 
@@
 
-Aspects of topographic stresses and their relevance to tectonics have
-been studied for some time. \citet{jeffreys1924} noted that the
-presence of high mountains is evidence that the Earth's crust can
-support significant heterogeneous differential stresses over long time
-periods. \citet{dalmayrac1981} and \citet{molnar1988} discussed how 
-extensional deformation in
-the high parts of orogens that is temporally coincident with contractional
-deformation at the low-elevation margins of the orogen may be explained
-by a spatially invariant, depth-integrated horizontal compressive stress
-with spatially varying vertical stresses caused by topography.
-\citet{richardson1994} exploited this
-relationship in the central Andes to estimate horizontal tectonic
-stresses through finite element modeling, and more recently 
-\citet{copley2009} 
-performed similar work in Albania. \citet{bollinger2004} showed how 
-increased normal stress on the Main
-Himalayan thrust due to loading of the high Himalayan massifs locally
-suppressed microseismicity and increased fault locking. Meade and Conrad
-\citet{meade2008} demonstrated that the increased weight of the
-uplifting Andes influenced the Nazca-South America convergence rate.
-
-Additionally, topographic stresses play a role in common models of
-orogenic dynamics. For example, topographic loading is central to thin
-viscous sheet models of lithospheric deformation \citep[e.g.,][]{birdpiper80,
-flesch2010gpe}. Critical taper models for thrust or extensional
-wedges \citep[e.g.,][]{dahlen1990, xiao1991} incorporate both variation in
-vertical stress due to changing elevation as well as a shear stress
-contributed by the slope of the wedge surface. Critical taper models
-also incorporate the idea that in a growing wedge, progressively
-increasing topographic stresses may eventually prevent continued slip on
-a given fault plane, and strain will instead be transferred to the toe
-of the thrust wedge where the stress state is more favorable. Of
-particular relevance to our work are `fixed boundary' models of
-gravitational collapse and spreading \citep{rey2001}, where an excess of
+Aspects of topographic stresses and their relevance to tectonics have been
+studied for some time. \citet{jeffreys1924} noted that the presence of high
+mountains is evidence that the Earth's crust can support significant
+heterogeneous differential stresses over long time periods.
+\citet{dalmayrac1981} and \citet{molnar1988} discussed how extensional
+deformation in the high parts of orogens that is temporally coincident with
+contractional deformation at the low-elevation margins of the orogen may be
+explained by a spatially invariant, depth-integrated horizontal compressive
+stress with spatially varying vertical stresses caused by topography.
+\citet{richardson1994} exploited this relationship in the central Andes to
+estimate horizontal tectonic stresses through finite element modeling, and more
+recently \citet{copley2009} performed similar work in Albania.
+\citet{bollinger2004} showed how increased normal stress on the Main Himalayan
+thrust due to loading of the high Himalayan massifs locally suppressed
+microseismicity and increased fault locking. Meade and Conrad \citet{meade2008}
+demonstrated that the increased weight of the uplifting Andes influenced the
+Nazca-South America convergence rate.
+
+Additionally, topographic stresses play a role in common models of orogenic
+dynamics. For example, topographic loading is central to thin viscous sheet
+models of lithospheric deformation \citep[e.g.,][]{birdpiper80, flesch2010gpe}.
+Critical taper models for thrust or extensional wedges
+\citep[e.g.,][]{dahlen1990, xiao1991} incorporate both variation in vertical
+stress due to changing elevation as well as a shear stress contributed by the
+slope of the wedge surface. Critical taper models also incorporate the idea
+that in a growing wedge, progressively increasing topographic stresses may
+eventually prevent continued slip on a given fault plane, and strain will
+instead be transferred to the toe of the thrust wedge where the stress state is
+more favorable. Of particular relevance to our work are `fixed boundary' models
+of gravitational collapse and spreading \citep{rey2001}, where an excess of
 gravitational potential energy associated with the high topography and
-thickened crust of eastern Tibet causes a transfer of rock to the
-foreland through horizontal contraction at the rangefront \citep[e.g.,][]
-{dewey1988, liuyang2003, copleymckenzie2007}. This process may be
-aided by a weak (sub)horizontal structure at depth (such as a shear zone
-or weak crustal channel) that is capable of transferring the vertical
-and radial stresses from topography throughout an orogen to its margins,
-leading to contraction there \citep[e.g.,][]{clark2005, burchfiel2008,
-fleschbendick2012}.
+thickened crust of eastern Tibet causes a transfer of rock to the foreland
+through horizontal contraction at the rangefront \citep[e.g.,][] {dewey1988,
+liuyang2003, copleymckenzie2007}. This process may be aided by a weak
+(sub)horizontal structure at depth (such as a shear zone or weak crustal
+channel) that is capable of transferring the vertical and radial stresses from
+topography throughout an orogen to its margins, leading to contraction there
+\citep[e.g.,][]{clark2005, burchfiel2008, fleschbendick2012}.
 
@@
 elastic upper crust has been studied on smaller spatial scales.
-\citet{mctiguemei1981} and
-\citet{savageswolfs1986} investigated the stress components from long,
-symmetric ridges and showed how horizontal tension is induced under
-ridge crests and horizontal compression is induced under valleys, mostly
-due to shear stresses generated by slopes. Work in this vein was
-continued by \citet{millerdunne1996} and \citet{martel2006}, who have 
-focused on shallow rock fracturing
-resulting from topographic stresses. \citet{liuzoback1992} investigated 
-whether the topographic stresses
-generated by the mountains around Cajon Pass (California, USA)
-contributed to the observed left-lateral shear stress on a shallow
-portion of the right-lateral San Andreas fault. In their study, they
-developed methods for calculating the three dimensional elastic stress
-tensor field arbitrary topography, whereas previous solutions were limited
-to two dimensions and required topography to be mathematically-defined
-(e.g., sinusoidal).
-
-\citet{luttrell2011} inferred the coseismic shear stress
-changes during the 2010 \emph{Mw}8.8 Maule, Chile earthquake, and using
-the topographic stresses due to the overlying fore arc, constrained the
-the stresses that led to this earthquake. In their study, they
-calculated topographic stresses following a similar procedure as
-\citet{liuzoback1992} (which we describe below), although
+\citet{mctiguemei1981} and \citet{savageswolfs1986} investigated the stress
+components from long, symmetric ridges and showed how horizontal tension is
+induced under ridge crests and horizontal compression is induced under valleys,
+mostly due to shear stresses generated by slopes. Work in this vein was
+continued by \citet{millerdunne1996} and \citet{martel2006}, who have focused
+on shallow rock fracturing resulting from topographic stresses.
+\citet{liuzoback1992} investigated whether the topographic stresses generated
+by the mountains around Cajon Pass (California, USA) contributed to the
+observed left-lateral shear stress on a shallow portion of the right-lateral
+San Andreas fault. In their study, they developed methods for calculating the
+three dimensional elastic stress tensor field arbitrary topography, whereas
+previous solutions were limited to two dimensions and required topography to be
+mathematically-defined (e.g., sinusoidal).
+
+\citet{luttrell2011} inferred the coseismic shear stress changes during the
+2010 \emph{Mw}8.8 Maule, Chile earthquake, and using the topographic stresses
+due to the overlying fore arc, constrained the the stresses that led to this
+earthquake. In their study, they calculated topographic stresses following a
+similar procedure as \citet{liuzoback1992} (which we describe below), although
 they only considered the component of the topographic load that can be
 described by convolving a Boussinesq solution with the topography.
-\citet{luttrell2011} also considered the contribution to
-stresses due to buoyancy.
+\citet{luttrell2011} also considered the contribution to stresses due to
+buoyancy.
 
@@
 
-The 2008 \emph{M}7.9 Wenchuan, China earthquake is one of the most
-devastating earthquakes in recent history (for an in-depth review and
-discussion of the earthquake, see \citet{zhang2010}).
-Surface rupture occurred along a 240 km segment of the Beichuan fault
-and a parallel 72 km segment of the Pengguan fault \citep{xu2009} 
-(Figure~\ref{fig:lms_map}). These faults lie at the base of the central and
-northeastern Longmen Shan, a mountain range that forms the eastern
-margin of the Tibetan plateau. Total relief across the central Longmen
-Shan is around 4 km, though relief subsides somewhat to the northeast.
-The central and southwestern Longmen Shan is the steepest margin of the
-Tibetan plateau \citep{clarkroyden2000} and one of the highest and
-steepest escarpments on earth. This is most apparent in the southwestern
-portion of the earthquake rupture, where elevations \textgreater{} 4000
-m over the Pengguan massif (a Precambrian crystalline massif in the
-hanging wall of the Beichuan thrust) drop to \textasciitilde{}1200 m in
-as little as 6 km map distance.
-
-\begin{figure*}[ht!]
-\includegraphics[width=40pc]{../figures/lms_map.pdf}
+The 2008 \emph{M}7.9 Wenchuan, China earthquake is one of the most devastating
+earthquakes in recent history (for an in-depth review and discussion of the
+earthquake, see \citet{zhang2010}).  Surface rupture occurred along a 240 km
+segment of the Beichuan fault and a parallel 72 km segment of the Pengguan
+fault \citep{xu2009} (Figure~\ref{fig:lms_map}). These faults lie at the base
+of the central and northeastern Longmen Shan, a mountain range that forms the
+eastern margin of the Tibetan plateau. Total relief across the central Longmen
+Shan is around 4 km, though relief subsides somewhat to the northeast.  The
+central and southwestern Longmen Shan is the steepest margin of the Tibetan
+plateau \citep{clarkroyden2000} and one of the highest and steepest escarpments
+on earth. This is most apparent in the southwestern portion of the earthquake
+rupture, where elevations \textgreater{} 4000 m over the Pengguan massif (a
+Precambrian crystalline massif in the hanging wall of the Beichuan thrust) drop
+to \textasciitilde{}1200 m in as little as 6 km map distance.
+
+\begin{figure*}[ht!] \includegraphics[width=40pc]{../figures/lms_map.pdf}
     \caption{Map of eastern Tibet and the Sichuan basin, showing active
-structures from \citet{styron2010}. Faults that ruptured
-in the 2008 Wenchuan earthquake are shown in pink. GPS velocities are
-relative to the mean velocity of sites within Sichuan basin, with
-1$\sigma$ uncertainty, from the dataset of Liang et al.
-\citet{liang2013}. Beachball is from the Global CMT focal mechanism
-solution for the 2008 Wenchuan earthquake. BF = Beichuan fault. PF =
-Pengguan fault. P = Pengguan massif. Grey box shows the extent of
-Figures 3 and 9.}
-\label{fig:lms_map}
+        structures from \citet{styron2010}. Faults that ruptured in the 2008
+        Wenchuan earthquake are shown in pink. GPS velocities are relative to
+        the mean velocity of sites within Sichuan basin, with 1$\sigma$
+        uncertainty, from the dataset of Liang et al.  \citet{liang2013}.
+        Beachball is from the Global CMT focal mechanism solution for the 2008
+    Wenchuan earthquake. BF = Beichuan fault. PF = Pengguan fault. P = Pengguan
+massif. Grey box shows the extent of Figures 3 and 9.} \label{fig:lms_map}
 \end{figure*}
 
-Surface ruptures during the Wenchuan earthquake are highly variable and
-show vertical (reverse-sense) displacements up to 9 m and horizontal
-(right-lateral sense) displacements up to 5 m \citep{lin2009, liu2009,
-xu2009}. In general, vertical displacements are higher in the
-southwestern to central portions of the Beichuan rupture and decrease in
-the northeast, whereas horizontal offsets are higher in the central to
-northeast, though considerable variation exists. Coseismic slip models
-constrained by seismic and geodetic data reveal a complicated pattern of
-coseismic slip in the Wenchaun earthquake, with several high-slip
-patches that dominate the seismic moment release and significant
+Surface ruptures during the Wenchuan earthquake are highly variable and show
+vertical (reverse-sense) displacements up to 9 m and horizontal (right-lateral
+sense) displacements up to 5 m \citep{lin2009, liu2009, xu2009}. In general,
+vertical displacements are higher in the southwestern to central portions of
+the Beichuan rupture and decrease in the northeast, whereas horizontal offsets
+are higher in the central to northeast, though considerable variation exists.
+Coseismic slip models constrained by seismic and geodetic data reveal a
+complicated pattern of coseismic slip in the Wenchaun earthquake, with several
+high-slip patches that dominate the seismic moment release and significant
 variation coseismic slip rake along strike \citep[e.g.,][]{nakamura2010,
-shen2009, tong2010, feng2010, zhang2011, qi2011, fielding2013}.
-The variation in rake is such that the southwest portions of the fault
-slipped largely in a reverse sense, while the northeast portions sliped
-largely in a right-lateral sense. This change in rake is associated with
-a change in inferred fault dip. Sections of faults that ruptured in the
-Wenchuan earthquake (which we simply refer to as the ``Wenchuan
-earthquake faults'') with shallow to moderate dips largerly ruptured as
-thrust, and sections with steeper dips largely ruptured as strike-slip.
-\citet{medinaluna2013} concluded that this
-relationship is consistent with a uniform orientation of principal
-stresses, where the variation of the dip of the fault leads to a change
-in the direction of maximum fault shear stress, which they assumed to be
-parallel to the coseismic slip rake.
+shen2009, tong2010, feng2010, zhang2011, qi2011, fielding2013}.  The variation
+in rake is such that the southwest portions of the fault slipped largely in a
+reverse sense, while the northeast portions sliped largely in a right-lateral
+sense. This change in rake is associated with a change in inferred fault dip.
+Sections of faults that ruptured in the Wenchuan earthquake (which we simply
+refer to as the ``Wenchuan earthquake faults'') with shallow to moderate dips
+largerly ruptured as thrust, and sections with steeper dips largely ruptured as
+strike-slip.  \citet{medinaluna2013} concluded that this relationship is
+consistent with a uniform orientation of principal stresses, where the
+variation of the dip of the fault leads to a change in the direction of maximum
+fault shear stress, which they assumed to be parallel to the coseismic slip
+rake.
 
@@
 
-We seek to quantify the topographic stress field in the Longmen Shan
-region, and on the Wenchuan earthquake faults themselves. Specifically,
-we evaluate the extent to which topographic stresses promote or inhibit
-slip in the 2008 Wenchuan earthquake. If topographically induced shear
-stresses on the fault are in roughly in the same direction as the
-coseismic slip, then topographic loading general promotes coseismic slip
-(Figure \ref{fig:topo_fault_scenarios} a). On the other hand, if the
-topographic shear stresses are roughly in the opposite direction of
-coseismic slip, then topographic loading inhibits coseismic slip (Figure
-\ref{fig:topo_fault_scenarios} b). If topographic loading resisted slip
-across the Beichuan faults, then tectonic stresses would need to
-counteract the topographic fault stresses for the coseismic slip to
+We seek to quantify the topographic stress field in the Longmen Shan region,
+and on the Wenchuan earthquake faults themselves. Specifically, we evaluate the
+extent to which topographic stresses promote or inhibit slip in the 2008
+Wenchuan earthquake. If topographically induced shear stresses on the fault are
+in roughly in the same direction as the coseismic slip, then topographic
+loading general promotes coseismic slip (Figure \ref{fig:topo_fault_scenarios}
+a). On the other hand, if the topographic shear stresses are roughly in the
+opposite direction of coseismic slip, then topographic loading inhibits
+coseismic slip (Figure \ref{fig:topo_fault_scenarios} b). If topographic
+loading resisted slip across the Beichuan faults, then tectonic stresses would
+need to counteract the topographic fault stresses for the coseismic slip to
 result. On a smaller scale, the heterogeneity of coseismic slip in the
-earthquake may be influenced by shorter wavelength variations in
-topography and topographic stresses.
-
-\begin{figure*}[t!]
-\centering
+earthquake may be influenced by shorter wavelength variations in topography and
+topographic stresses.
+
+\begin{figure*}[t!] \centering
     \includegraphics[width=40pc]{../figures/topo_stress_possibilities.pdf}
-\caption{Scenarios for topographic effects on rangefront thrust
-faulting. (a) Topographic stresses promote thrust faulting. (b)
-Topographic stresses inhibit thrust faulting.}
-\label{fig:topo_fault_scenarios}
+    \caption{Scenarios for topographic effects on rangefront thrust faulting.
+        (a) Topographic stresses promote thrust faulting. (b) Topographic
+        stresses inhibit thrust faulting.} \label{fig:topo_fault_scenarios}
     \end{figure*}
 
-Topographic stresses are only one component of the total stress field in
-the crust \citep{molnar1988}. Coseismic slip in the Wenchaun earthquake
-is due to the total accumulated stress on the faults, which also
-includes components from lithostatic and tectonic stresses. (In the
-present study, we do not consider stresses due to flexure \citep[e.g.,][]
-{luttrell2007} or buoyancy \citep[e.g.,][]{luttrell2011}.) By quantifying
-both the topographic and lithostatic stresses, we can use coseismic slip
-models to solve for the tectonic stress assuming that (1) the total
-shear stress on the fault is in the direction of coseismic slip \citep[e.g.,]
-[]{angelier1994} and (2) the fault is at Mohr-Coulomb failure everywhere
-that it slipped. If topographic stresses are significant and produce
-shear in the direction of fault slip, then for given values of static
-friction and pore fluid pressure, we can calculate the amount of
-tectonic stress that can be added to the ambient stress field before the
-faults should rupture; given limited acceptable ranges for friction and
-fluid pressure, we are essentially able to place maximum constraints on
-tectonic stress. Alternately, if topographic stresses work against
-coseismic slip, for given friction and fluid pressures we can estimate
-the minimum magnitudes of tectonic stresses necessary to overcome shear
-and frictional resistance to slip. In a scenario with complex faulting
-and topography, it may be possible to put tight bounds on minimum and
-maximum magnitudes and directions of tectonic stresses. To account for
-the non-uniqueness of the solution, we use a Bayesian Monte-Carlo
-methodology to estimate posterior probability density functions (PDFs)
-of tectonic stresses, static fault friction, and pore fluid pressure.
+Topographic stresses are only one component of the total stress field in the
+crust \citep{molnar1988}. Coseismic slip in the Wenchaun earthquake is due to
+the total accumulated stress on the faults, which also includes components from
+lithostatic and tectonic stresses. (In the present study, we do not consider
+stresses due to flexure \citep[e.g.,][] {luttrell2007} or buoyancy
+\citep[e.g.,][]{luttrell2011}.) By quantifying both the topographic and
+lithostatic stresses, we can use coseismic slip models to solve for the
+tectonic stress assuming that (1) the total shear stress on the fault is in the
+direction of coseismic slip \citep[e.g.,] []{angelier1994} and (2) the fault is
+at Mohr-Coulomb failure everywhere that it slipped. If topographic stresses are
+significant and produce shear in the direction of fault slip, then for given
+values of static friction and pore fluid pressure, we can calculate the amount
+of tectonic stress that can be added to the ambient stress field before the
+faults should rupture; given limited acceptable ranges for friction and fluid
+pressure, we are essentially able to place maximum constraints on tectonic
+stress. Alternately, if topographic stresses work against coseismic slip, for
+given friction and fluid pressures we can estimate the minimum magnitudes of
+tectonic stresses necessary to overcome shear and frictional resistance to
+slip. In a scenario with complex faulting and topography, it may be possible to
+put tight bounds on minimum and maximum magnitudes and directions of tectonic
+stresses. To account for the non-uniqueness of the solution, we use a Bayesian
+Monte-Carlo methodology to estimate posterior probability density functions
+(PDFs) of tectonic stresses, static fault friction, and pore fluid pressure.
 

### 2014-09-04 tect_stress/scripts/feng_bayes_rake.py
done w/ new_T_priors

@@ -17,7 +17,7 @@ lms = pd.concat((fb, fp), axis=0)
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1.5e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700

### 2014-09-04 tect_stress/scripts/qi_rough_bayes_rake_ec2.py

@@ -22,7 +22,7 @@ lms = lms[lms.slip_m>=0.001]
 np.random.seed(69)
 
 # some inital constants
-n_trials = 1e5
+n_trials = 2e5
 
 n_points = len(lms.index)
 rho = 2700
@@ -32,7 +32,7 @@ g = 9.81
 # These are functions of lithostatic pressure (rho g depth)
 # Priors for each are uniform [-2, 2).
 s1s = np.random.uniform(0,2.5, n_trials)
-s3s = np.random.uniform(0, 1, n_trials) * s1s
+s3s = np.random.uniform(-1, 1, n_trials) * s1s
 thetas = np.random.uniform(0, 2 * np.pi, n_trials)
 
 xxs = scv.xx_stress_from_s1_s3_theta(s1s, s3s, thetas)

### 2014-09-15 scripts/get_stress_slices.py
added zz stress slices

@@ -84,26 +84,43 @@ xy_stress_vec_12 = pd.Series(xy_sparse_12km_la.x)
 
 del xy_la
 
+print 'doing zz stresses'
+zz_la = la.larry(stress_db['zz_MPa'][:,:,:], stress_labels)
+
+zz_sparse_5km_la = zz_la[::10, ::10, 4].flatten()
+zz_stress_vec_5 = pd.Series(zz_sparse_5km_la.x)
+
+zz_sparse_20km_la = zz_la[::10, ::10, 20].flatten()
+zz_stress_vec_20 = pd.Series(zz_sparse_20km_la.x)
+
+zz_sparse_12km_la = zz_la[::10, ::10, 12].flatten()
+zz_stress_vec_12 = pd.Series(zz_sparse_12km_la.x)
+
+del zz_la
+
 stress_db.close()
 
 print 'making dataframes'
 topo_stress_5km = pd.concat([east, north, xx_stress_vec_5, yy_stress_vec_5,
-                             xy_stress_vec_5], axis=1)
-topo_stress_5km.columns=['easting', 'northing', 'xx_MPa', 'yy_MPa', 'xy_MPa']
+                             xy_stress_vec_5, zz_stress_vec_5], axis=1)
+topo_stress_5km.columns=['easting', 'northing', 'xx_MPa', 'yy_MPa', 'xy_MPa',
+                         'zz_MPa']
 
 topo_stress_5km.to_csv('e_asia_horiz_stress_5km.csv', index=False)
 
 
 topo_stress_12km = pd.concat([east, north, xx_stress_vec_12, yy_stress_vec_12,
-                             xy_stress_vec_12], axis=1)
-topo_stress_12km.columns=['easting', 'northing', 'xx_MPa', 'yy_MPa', 'xy_MPa']
+                             xy_stress_vec_12, zz_stress_vec_12], axis=1)
+topo_stress_12km.columns=['easting', 'northing', 'xx_MPa', 'yy_MPa', 'xy_MPa',
+                          'zz_MPa']
 
 topo_stress_12km.to_csv('e_asia_horiz_stress_12km.csv', index=False)
 
 
 topo_stress_20km = pd.concat([east, north, xx_stress_vec_20, yy_stress_vec_20,
-                             xy_stress_vec_20], axis=1)
-topo_stress_20km.columns=['easting', 'northing', 'xx_MPa', 'yy_MPa', 'xy_MPa']
[…]

### 2014-09-17 revisions/rebuttal.md
revisions before sending to eric

@@
-## R1
+## R1a
 > The complete stress tensor is divided into contributions from topography,
@@
 > pressure applied to the rock scaled by ~0.6 (this study's parameter T'?).
+
+This is related to, but not really a restatement of, Byerlee's law. This is
+much more a statement of the Coulomb failure envelope. Byerlee's law is related
+to the failure of intact rock, which is often quoted (as in the review, and
+in our section 6.3) as 0.6. This coefficient is related to the coefficient of
+internal friction of intact rock, which is not our $T'$. In our study, we are
+primarily concerned with the failure of a specific, pre-existing fault, whose
+strength (coefficient of static friction) is unknown but likely less than
+unfractured rock. We scale tectonic stresses linearly with depth, and each
+component of tectonic stress ($T_{xx}$, $T_{xy}$, and $T_{yy}$, or $T_{max}$,
+$T_{min}$, and azimuth) increases with depth at a different rate.
+
+$T'$ is simply $T / \rho g z$ (line 275 of original manuscript).
+
+## R1b
 > However, this means that the formulation expressed in equation 9 is counting
-> the lithostatic contribution of stress twice. The Towend paper derives its
-> results from deep drilling studies, so it would be more appropriate to assume
-> that the total stress S increases linearly with depth in this manner.
+> the lithostatic contribution of stress twice.
 
-(add whole tensor thing)
+We have not double-counted lithostatic stress, we have simply been unclear in
+our description of the stress state. Lithostatic stress is an isotropic tensor,
+so (as described) it is only the diagonals of the tensor. Tectonic stress is
+only horizontal, $T'$ is scaled by $\rho g z$ (not by the tensor $L$, as we had
+previously, and erroneously stated). So $T$ and $L$ do not occupy the all same
+components of the stress tensor.
+
+We amend the submitted description:
+
+> We consider the complete stress tensor, $S$, at a point in the crust to
+> be
+> 
+> \begin{equation}
+> S = M + T + L \; ,
+> \end{equation}
+> 
+> where $M$ is described above, $T$ is the tectonic stress tensor, and $L$
+> is the lithostatic stress tensor. $L$ is isotropic, with diagonal
+> components equal to $\rho g z$. We assume that $T$ is laterally
+> homogeneous and only has horizontal stress components (i.e.,
+> $T_{xz} = T_{yz} = T_{zz} = 0$, with $T_{xx}$, $T_{yy}$ and $T_{xy}$
+> non-zero). We further assume that $T$ increases linearly with depth so
+> that the entire upper crust is near the critical failure envelope
+> [e.g., *Townend and Zoback, 2000*], and thus we parameterize the components of $T$
+> as scalars multiplied by lithostatic pressure, denoted as $T^\prime$
+
+to
+
+> We consider the complete stress tensor, $S$, at a point in the crust to
+> be
+
+> \begin{equation}
+> S = M + T + L \; ,
+> \end{equation}
+
+> where $M$ is described above, $T$ is the tectonic stress tensor, and $L$
+> is the lithostatic stress tensor. $L$ is isotropic, with diagonal
+> components equal to $\rho g z$. We assume that $T$ is laterally
+> homogeneous and only has horizontal stress components (i.e.,
+> $T_{xz} = T_{yz} = T_{zz} = 0$, with $T_{xx}$, $T_{yy}$ and $T_{xy}$
+> non-zero). The expanded stress tensor is:
+
+> \begin{equation}
+> S = \begin{bmatrix}
+>   M_{xx} + T_{xx} + L_{xx} & M_{xy} + T_{xy} &  M_{xz} \\
+> 	M_{xy} + T_{xy} &  M_{yy} + T_{yy} + L_{yy} & M_{yz} \\
+> 	M_{xz}     &  M_{yz}  &  M_{zz} + L_{zz}
+> 	\end{bmatrix} \; .
+> \label{eqn:stress_tensor}
+> \end{equation}
+
+> We further assume that $T$ increases linearly with depth so
+> that the entire upper crust is near the critical failure envelope
+> [e.g., *Townend and Zoback, 2000*], and thus we parameterize the components 
+> of $T$ as scalars multiplied by $\rho g z$, denoted as $T^\prime$. Therefore, 
+> if $T'_{xx} = 0.1$, at some point just below 1 km, $L=27$ MPa, so
+> $S_{xx} = M_{xx}$ + 27 MPa + 2.7 MPa.
+
+which hopefully clarifies this.
+
+## R1c
+> The Towend paper derives its results from deep drilling studies, so it
+> would be more appropriate to assume that the total stress S increases 
+> linearly with depth in this manner.
+
+We agree that $S$ increasing linearly with depth would be more analogous to
+*Townend and Zoback 2000*, but our model does not allow for this, because we
+would have to vary $T$ spatially to exactly counterbalance variations in $M$
+that deviate from linearity. Instead, we find it much simpler to approximate
+$T$ as increasing linearly with depth, because then we can quantify it with
+three numbers ($T'_{xx}$, $T'_{xy}$, and $T'_{yy}$, or equivalently $T'_{max}$,
+$T'_{min}$ and the azimuth of $T'_{max}$). Additionally, because $L$ increases
+rapidly with depth, it overwhelms the largest topographic stresses on the
+fault (~80 MPa) at about 3 km, so for most of the crust, $S$ does
+increase essentially linearly with depth.
 
@@
 
-In these comments, the reviewer basically asserts that we claim that is
-significant that our model's predictions match the observations, and then gives
-several example lines (401, 407, 417). Though this is not the interpretation of
-the paragraph that we had intended, it is apparent how this interpretation
-could be made. Nonetheless, the specific lines seem to be off the mark in this:
+In these comments, the reviewer basically asserts that we set up our model to
+replicate observations and then claim that is significant that our model's
+predictions match the observations, and then gives several example lines (401,
+407, 417). Though this is not the interpretation of the paragraph that we had
+intended, it is apparent how this interpretation could be made. Nonetheless,
+the specific lines seem to be off the mark in this:
 
 Line 401 is basically a quick description of the results (E-W maximum principal
-tectonic stress produces oblique shear on the fault), and is not really a claim
-about the significance of the results.
+tectonic stress produces oblique shear on the fault), and is not a claim about
+the significance of the results.
 
@@
 measurements were not used in our inversion, and therefore this is an
-independent confirmation of our results, not a tautological statement.
+independent confirmation of our results, not a direct consequence of the
+inversion.
 
@@
 approach because it is less than obvious how to incorporate a depth-dependent
-tectonic stress into the calculations. The horizontal loading functions
-specify horizontal surface loads that are then propagated into the subsurface using Cerruti's solutions [eq. xx]. The Liu and Zoback formulations of the loading
-functions seems to require a fixed tectonic stress, or the tectonic stress at
+tectonic stress into the calculations. The horizontal loading functions specify
+horizontal surface loads that are then propagated into the subsurface using
+Cerruti's solutions [Appendix A2]. The Liu and Zoback formulations of the
+loading functions require a fixed tectonic stress, or the tectonic stress at
 zero depth. Because we assume that tectonic stress increases linearly from zero
-with depth, the tectonic stress at zero depth (sea level) is zero.
+with depth, the tectonic stress at the surface is zero.
+
+An additional complication is that the calculation of topographic stresses in
+the region (not their interpolation onto the fault) takes about 9 hours, and
+so devising some iterative scheme where we could calculate topographic
+stresses, then find tectonic stress and then add that back into the topographic
+stress calculations and repeat until some criteria is reached, is really out
+of the question with the Bayesian inversion scheme we've chosen (there are
+about 18,000 models in the joint posterior).
 
@@
 
-This is a valid point, but it wasn't something that we had tested much because
-of the computational effort required, both in terms of programmer time and CPU
-time. However, in the intervening months since submitting this manuscript, we
-have streamlined the analysis so that the programmer time required is much
-less.
+We had chosen our the value of 0.25 for the Poisson ratio not only because of
+the calculational convenience (the shear modulus and Lame's parameters drop out
+of the equations) but because this is the most likely value for the rather
+felsic Tibetan crust based on seismic studies.  But in light of these comments,
+we have recalculated the fault stresses using a Poisson ratio of 0.28, as a
+test. We have found that the stresses are different by 1 or 2 MPa, which is
+a very small difference relative to the overall magnitudes. We have changed the
+relevant sentence in first paragraph of Section 2.2 from:
 
-[test this].
+> We assume a Poisson halfspace
+
+to
+
+> We assume a Poisson ratio of 0.25, following receiver function studies
+suggesting values of ca. 0.24-0.26 throughout central and western China [*Chen
+et al., 2010*] (we have tested a Poisson ratio of 0.28, which is on the higher
+end of values for intermediate rock compositions [*Zandt and Ammons, 1995*],
+and found the results to vary by a few percent).
 
@@
 In general, neither of the tectonic stresses have to be positive. However,
-vertical stress cannot be least compressive if $T_{min}$ were negative, unless
-topographic stress produced a strong horizontal compression in the direction of
-$T_{min}$ and less strong vertical compression.
+vertical stress cannot be least compressive if $T_{min}$ were negative
+(tensile), unless topographic stress produced a strong horizontal compression
+in the direction of $T_{min}$ and less strong vertical compression. Our results
+show that the topographic stresses are mostly stronger (more compressive) in
+the vertical direction than in the horizontal directions.
 
-To restate this mathematically, for clarity:
-Let's consider $T_{min}$ to be aligned N-S, i.e. it is in the $xx$ direction.
-The equation for total stress is $S = M + T + L$ (equation 9), and we consider
-tectonic stress $T$ to be only horizontal stresses. So the vertical stress is
-$M + L$. $L$ is isotropic, so we can ignore it here. The only way that 
-$M_{xx} + T_{min} > $M_{zz}$, where $T_{min} < 0$, is to have 
-$M_{xx} > M_{zz} - T_{min}$. Given that typically $M_{zz} > M_{xx}$ in our 
-results (Section 3, lines 224-225 of the original manuscript), it is not
-possible in this case to have a negative $T_{min}$ and still have vertical
-stress ($S_{zz}$) as the least compressive stress (which is required by reverse
-fault slip).
+Additionally, from a tectonic standpoint, there is no real reason to consider
+N-S tectonic tension (the direction of $T_{min}$ in our results) in the Sichuan Basin;
+this is a region surrounded by thrust and strike-slip faults of various
+orientations (e.g. Figure 1). If there was tectonic tension in the region, then
+the patterns of deformation would be different in the interior of the Sichuan
+Basin away from topography; currently faulting there is thrusting, which
+indicates a vertical $\sigma^3$. There is no evidence of N-S extension anywhere
+in the Himalayan-Tibetan orogen except for in the Miocene in the Himalaya,
+which is synchronous with parallel thrusting (the famed Himalayan extrusion).
 
-Furthermore,
+The purpose of Bayesian priors is to keep likely values in as much as to keep
+unlikely values out. Because we adopt a relatively simple scheme of uniform
+probabilities over some interval, we can't add unlikely values at the same
+rate as more likely values, without either strongly diluting the number of
+acceptable models in the posterior (and therefore needing to run many more
+models, and we are already at the limit of what can be done with 128 GB RAM),
+and/or introducing a lot of unlikely models into the posterior, depending on
+how sensitive the results are to the priors.
 
@@
 
-The reviewer is absolutely correct in his or her assessment of the ability of
-the Bayesian inversion to capture the maximum tectonic stresses for this
-particular earthquake, and we are up front about this (lines 323-326). (For
-other events with less of a the thrust component, we can constrain maximum
-tectonic stress).
+While the results are certainly more sensitive to the low end of allowable
+$T'_{max}$ values, they are somewhat sensitive to the high end. This
+sensitivity varies with slip model, as is apparent in Figure 6. Figure 8,
+the joint results, show that $T'_{max}$ decreases past about 1.0, and the
+likelihood of $T'_{max}=2.5$ is about 60% of the likelihood of $T'_{max}=0.8$.
+Additionally, this is assuming that $T'_{max}$ is *a priori* equal at 0.8 and
+2.5 (which is our simple model); personally, I find it rather unlikely that
+tectonic stresses would be so high, although we chose the model prior to be
+a bit different than our personal prior beliefs.
 
@@
 
-This was something else that we spent some time on. The issue with picking
-stresses independently is that they're not independent, because $T_{min}$ *has
-to* < $T_{max}$. So they can't be sampled indendently, and there are two
-approaches that can be taken. One of them is to take a random Nx2 array and
-sort each row, so that one column is $T_{max}$ and the other is $T_{min}$.
-This approach yields a uniform joint prior probability density over the
-available half of the $T_{min}$ x $T_{max}$ parameter space, but gives strongly
-nonuniform marginal priors for either $T_{max}$ (which linearly increases from
-0 at T=0 to max) or $T_{min}$ (with a commensurate linear decrease).
+The requirement that $T_{min} < T_{max}$ is not simply an attribute of our
+priors or Bayesian estimation scheme, but basic math and logic. Minimum is
+defined to be less than maximum, so the half of the parameter space where this
+would occur is off limits by definition.
 
-The other is to do what we did, and take a random Nx1 array for T$_{max}$, and
-then take $T_{min}$ as some random value from [0,1) of $T_{max}$. This gives
-a uniform prior PDF for $T_{max}$, the same prior for $T_{min}$ as the first
-option, and a higher joint probability density towards the low $T_{max}$ end.
+Constructing suitable priors, though, was something that we spent some time on.
+The issue with picking stresses independently is that they're not independent,
+because $T_{min} < T_{max}$, regardless of the value of $T_{max}$, so there is
+an inherent conditional probability. So they can't be sampled indendently, and
+there are two approaches that can be taken. One of them is to take a random
+(uniform) Nx2 array and sort each row, so that one column is $T_{max}$ and the
+other is $T_{min}$. This approach yields a uniform joint prior probability
+density over the available half of the $T_{min}$ x $T_{max}$ parameter space,
+but yields marginal priors for $T_{max}$ that are strongly biased towards high
+values, and priors for $T_{min}$ that are biased towards the low end.
+
+The other is to do what we did, and take a random (uniform) Nx1 array for
+T$_{max}$, and then take $T_{min}$ as some random value from [0,1) of
+$T_{max}$. This gives a uniform prior PDF for $T_{max}$, a similar prior for
+$T_{min}$ as the first option, and a higher joint probability density towards
+the low $T_{max}$ end of the $T_{max}$ vs $T_{min}$ space.
 
@@
 $T_{max}$ would be much more strongly biased in the first approach, but
-$T_{min}$ is unaffected by the sampling strategy. Overall, we feel that our
-approach is less biased.
+$T_{min}$ is less affected by the sampling strategy. Overall, we feel that our
+approach is less biased, because it results in a uniform $T_{max}$, which is
+the more important value.
 
@@
 minimum.
+
+However, it is clear in Figure 8b, where $T'_{min}$ is scaled to $T'_{max}$,
+that the *ratio* of $T_{min}$ to $T_{max}$ is constrained by the inversion;
+the most likely value for $T_{min} \approx 0.3 T_{max}$, regardless of the
+value for $T_{max}$. This point was mentioned in passing in section 5.1 (line
+330) but could be discussed in more detail, particularly with reference to 
+Figure 8.1. Therefore, we add to Section 5.2:
+
+>It is also apparent in Figure 8b that regardless of the magnitude of
+$T^\_{\mathrm{max}$, $p_{J}(T^\prime_{\mathrm{min}} | D)$ is
+consistently 0--0.6 $p_{J}(T^\prime_{\mathrm{max}} | D)$, with a mode
+of about 0.3; this indicates that the relative magnitude of the
+tectonic stresses has a substantial influence on the rake of the maximum
+shear stress resolved on the fault.
 
@@
 
+This is indeed true, and it is unfortunate. However, the unclipped figure
+caption was available to the reviewer in the 'article file' available to
+him/her (not the merged pdf) on the review website. More fortunately, the
+general reader will be reading a professionally typeset version of the paper,
+not a submitted LaTeX draft.
+
 ## R9
@@
 
+The caption states that
+
+The magnitude of slip in the models varies by a factor of about 2. The rakes
+and the spatial distribution of slip vary much less.
+
 ## R10
@@
 > discussion.
+
+We have added a paragraph and a half describing the characteristics of the slip
+models used, including the similarities and differences of their geometry and
+slip distributions, to Section 2.3:
+
+>We use six models,
+those of *Shen, et al.* [*2009*], *Feng, et al.* [*2010*], *Zhang, et al.*
+[*2011*], *Fielding, et al.,* [*2013*] and two from *Qi, et al.* [*2011*] (these two models share a
+fault geometry but use different smoothing functions for the slip
+distribution). These vary somewhat in the fault geometry and specifics of the
+slip distribution, but nonetheless are of high resolution, are properly
+georeferenced. By using a suite of models in our calculations instead of a
+single model, we evaluate results that are persistent in most or all of the
+models to be more robust, and other results can be more confidently linked to
+specifics of the model geometry or slip distribution.
+
+>In general, the models share a common geometry and pattern of slip distribution
+above 10 km or so; i.e. the locations of high and low slip patches and the slip
+rake are similar in all of the models. Though some of models use discrete fault
+segments for the Beichuan fault, whereas others use a single continuous fault,
+the fault planes are essentially collocated. However, there is significant
+variability in the magnitude of slip in the different models; for example, the
+maximum slip magnitude in the Qi 'rough' model is about twice that of the
+Feng model. Large differences also exist in the deeper geometries of the
+models: the Qi, Fielding and Shen models all have horizontal or subhorizontal
+thrust flats at or below 15 km, though only minor slip is assumed to have
+occurred on these fault segments. In contrast, the Feng and Zhang models are
+planar to their lower end at 25-30 km.
+
+The significance of the findings highlighted here by the reviewer are discussed
+in the discussion (Section 6.4, particularly lines 504-511). In general, our
+chosen expository style in this paper is more geared towards description of the
+results rather than (over-)interpretation. The results themselves are described,
+and the results files are made available for readers who are interested in
+probing them for themselves. Because our results cover both topographic and
+tectonic stresses, as well as fault strength parameters, there are a lot of
+implications for anything from earthquake processes to orogenic development,
+and we feel it is better to be reasonably brief in our discussion of 
+individual aspects of these implications.
 
@@
 
-This study does indeed
+There is nothing in the abstract that references a topographic influence on
+rupture processes. In one sentence in the introduction (lines 41-45), we write
+that topographic stresses *may* influence rupture propagation. We do not
+'focus' on this. Additionally, as promised in the two quoted passages (from
+lines 48 and 54), we do investigate the effects of topographic stress on the
+faults, and bracket the tectonic stresses, friction, and fluid pressure on the
+faults. This is the focus of the paper.
+
+The first-order 'effects of topographic stress on the faults' is whether those
+stresses promote or inhibit slip on the faults. We show that topographic
+stresses are overwhelmingly resistive/inhibiting to fault slip, both by adding
+shear stress to the faults in the opposite direction to slip, and by adding
+significant normal stress to the faults.
+
+The influence of topographic stresses on a rupture that does occur may be a
+second-order effect. We see that topographic stresses, particularly normal
+stress, is negatively correlated with slip magnitude, and we discuss this in
+lines 248-255. However, any real discussion on this other than the observation
+that it occurs is beyond what we can comfortably say at this time, without
+being more speculative than we would wish. However, we believe it is an
+important observation, and it is certainly one which we have already made, and
+so it would be inappropriate to sweep it under the rug because we do not yet
+understand it at publishable level. It may be that other researchers are
+interested, or have better means to investigate it (i.e. they are experts in
+dynamic rupture modeling), and we should not inhibit them.
 
 ## R12
-> Overall, the manuscript, though well organized, seems to have an unbalanced emphasis.
-The specific details of the analytical methods are present, but overall the results should be more fully described and the discussion more fully developed.
+> Overall, the manuscript, though well organized, seems to have an unbalanced
+> emphasis. The specific details of the analytical methods are present, but
+> overall the results should be more fully described and the discussion more
+> fully developed.
+
+This is definitely a 'methods' paper, and this work was performed with the
+broader goal of working to quantify tectonic stresses in a larger sense. We
+chose this earthquake to start with because it's well-documented and has a
+complex coseismic slip distribution, and because it lies at the base of a very
+large plateau that has been suggested to be expanding outward due to
+topographic stresses. We looked into these effects, framed the investigation in
+terms of questions (e.g., do topographic stresses promote or inhibit the
+observed slip on the faults), and answered them.
+
+Nonetheless, the results and the discussion make up the bulk of the paper. The
+methods sections (2 and 4) are 6.5 'manuscript' pages in sum, which includes
+many equations that take up a lot of space. The results sections (3 and 5) are
+about 2 pages, and 6 of the 10 figures. The discussion is 8 pages, by far the
+longest section.
 
 ## R13
-> Line 24: Though I can see the appeal of the selected quotation, it would be more appropriate for an oral presentation or a monograph on the subject.
-In this context, it comes across as awkward and out of place.
+> Line 24: Though I can see the appeal of the selected quotation, it would be
+more appropriate for an oral presentation or a monograph on the subject. In
+this context, it comes across as awkward and out of place.
+
+We have removed it.
 
 ## R14
-> Section 1.1: additional relevant references to consider: Fialko, Rivera, and Kanamori (2005 GJI), Ghosh et al (2006 Geology), Flesch et al (2007 GJI)
+> Section 1.1: additional relevant references to consider: Fialko, Rivera, and
+Kanamori (2005 GJI), Ghosh et al (2006 Geology), Flesch et al (2007 GJI)
+
+We have added a reference to the Fialko paper. The Flesch and Ghosh papers
+mentioned here are 'thin viscous sheet' type papers, similar to those already
+mentioned in the review (including a later paper by Flesch); those types of
+studies are of a much larger (orogen to global) scale than our study, and
+both a longer list of studies and a more detailed discussion of the studies
+is a bit out of the scope of the current paper.
 
 ## R15
-> Lines 224-230 describe the magnitudes of stress tensor components, but this would be better explained with a figure. If not as its own figure, could Mzz at least be incorporated into figure 3?
+> Lines 224-230 describe the magnitudes of stress tensor components, but this
+would be better explained with a figure. If not as its own figure, could Mzz at
+least be incorporated into figure 3?
+
+We have added vertical topographic stress to Figures 3 and 9.
 
@@
 > Section 6.0 contrasts the methods of this study to those of Luttrell et al.
-2011 study of the Maule earthquake.
-This is an understandable comparison, but because the Wenchuan earthquake includes both strike-slip and dip-slip components, it is actually more similar to the Luttrell and Sandwell (2012 JGR) study of Mid Ocean Ridges that includes both strike-slip and normal earthquakes as constraints.
+2011 study of the Maule earthquake. This is an understandable comparison, but
+because the Wenchuan earthquake includes both strike-slip and dip-slip
+components, it is actually more similar to the Luttrell and Sandwell (2012 JGR)
+study of Mid Ocean Ridges that includes both strike-slip and normal earthquakes
+as constraints.
 
-That study doesn't use coseismic slip models.
-
+Actually, we feel that our manuscript is more similar to the Luttrell et al.
+2011 paper because that paper uses a coseismic slip model in the stress
+inversion, and because it involves a single event. The Luttrell and Sandwell
+paper uses focal mechanisms, which are basically point sources, and considers
+multiple earthquakes. This is a bit more different with respect to our paper.
 ## R17
@@
 
+We used all of the models that we could find that were available (i.e. the slip
+model files themselves were published), had proper georeferencing, and seemed
+to be of sufficiently high quality. There are many models that did not attempt
+a good geometric fit to the available data--e.g. they consist of a single,
+planar fault segment--or models with very poor coseismic slip resolution. We
+did not feel it wise to include these (garbage in equals garbage out, as they
+say). Additionally, we did not feel it wise to list models that we thought
+were of too low quality.
+
+We have added some discussion of the models used, and the criteria for
+including them, to Section 2.3, as described in R10.
+
 ## R18
 > --Typos
+
 > Lines 177 and 182: is z positive up or down? Clarify this
+
 > Line 258: earthquake, not fault, initiated
+
 > Line 274: "are" -> "and"? complete sentence with period
+
 > Line 509: typo, delete "to close"
+
 > Line 556: If you acknowledge some of your data sources, you should acknowledge all of your data sources (e.g., scientific drilling, WSM, etc)
+
 > Fig 3 and 9 need 33ºN label correction
+
+Corrected.

### 2014-09-17 revisions/styron_hetland_wenchuan_topo_revised.tex
revisions before sending to eric

--- 2014-09-04 manuscript/styron_hetland_wenchuan_topo_manuscript.tex
@@
 \begin{article} \section{Introduction}\label{introduction}
-
-\begin{quote} ``And He has set up on the earth mountains standing firm, lest it
-    should shake with you'' -- \emph{The Holy Qur'an}, 15:16 \end{quote}
-
 Stress is of fundamental importance to many processes in the earth. Both the
@@
 normal stresses on a fault from topographic loading may push a particular fault
-closer to or (as suggested by the Qur'an) farther from failure, or reorient the
+closer to or farther from failure, or reorient the
 net shear stress direction on a fault. These effects may affect the
@@
 estimate horizontal tectonic stresses through finite element modeling, and more
-recently \citet{copley2009} performed similar work in Albania.
+recently \citet{copley2009} performed similar work in Albania, and
+\citet{fialko2005} investigated the role of vertical topographic stresses on
+fault rotation in Southern California.
 \citet{bollinger2004} showed how increased normal stress on the Main Himalayan
@@
 coordinates to UTM zone 48N, decreasing the nominal horizontal
-resolution to 851 m. We assume a Poisson halfspace, and Green's
+resolution to 851 m. We assume a Poisson ratio of 0.25, following
+receiver function studies suggesting values of ca. 0.24-0.26 throughout
+central and western China \citep{chen2010receiverfn} (we have tested a Poisson
+ratio of 0.28, which is on the higher end of values for intermediate rock
+compositions \citep{zandt_ammons1995}, and found the results to vary by a few
+percent). Green's
 functions for the Boussinesq and Cerruti point-source solutions were
@@
 Topographic stresses on the Wenchuan faults are calculated on point sets
-representing the faults taken from coseismic slip models. We use the
-coseismic slip models of \citet{shen2009},
-\citet{feng2010}, \citet{qi2011}, \citet{zhang2011},
-and \citet{fielding2013}, and discard
+representing the faults taken from coseismic slip models. We use six models,
+those of \citet{shen2009}, \citet{feng2010}, \citet{zhang2011},
+\citet{fielding2013} and two from \citet{qi2011} (these two models share a
+fault geometry but use different smoothing functions for the slip
+distribution). These vary somewhat in the fault geometry and specifics of the
+slip distribution, but nonetheless are of high resolution, are properly
+georeferenced. By using a suite of models in our calculations instead of a
+single model, we evaluate results that are persistent in most or all of the
+models to be more robust, and other results can be more confidently linked to
+specifics of the model geometry or slip distribution.
+
+In general, the models share a common geometry and pattern of slip distribution
+above 10 km or so; i.e.\ the locations of high and low slip patches and the slip
+rake are similar in all of the models. Though some of models use discrete fault
+segments for the Beichuan fault, whereas others use a single continuous fault,
+the fault planes are essentially collocated. However, there is significant
+variability in the magnitude of slip in the different models; for example, the
+maximum slip magnitude in the Qi `rough' model is about twice that of the
+Feng model. Large differences also exist in the deeper geometries of the
+models: the Qi, Fielding and Shen models all have horizontal or subhorizontal
+thrust flats at or below 15 km, though only minor slip is assumed to have
+occurred on these fault segments. In contrast, the Feng and Zhang models are
+planar to their lower end at 25-30 km.
+
+In our stress calculations, we discard
 points above 851 m below sea level, as this is above the depth at which
@@
 is the lithostatic stress tensor. $L$ is isotropic, with diagonal
-components equal to $\rho g z$. We assume that $T$ is latterally
+components equal to $\rho g z$. We assume that $T$ is laterally
 homogeneous and only has horizontal stress components (i.e.,
 $T_{xz} = T_{yz} = T_{zz} = 0$, with $T_{xx}$, $T_{yy}$ and $T_{xy}$
-non-zero). We further assume that $T$ increases linearly with depth so
+non-zero). The expanded stress tensor is:
+
+\begin{equation}
+S = \begin{bmatrix}
+  M_{xx} + T_{xx} + L_{xx} & M_{xy} + T_{xy} &  M_{xz} \\
+	M_{xy} + T_{xy} &  M_{yy} + T_{yy} + L_{yy} & M_{yz} \\
+	M_{xz}     &  M_{yz}  &  M_{zz} + L_{zz}
+	\end{bmatrix} \; .
+\label{eqn:stress_tensor}
+\end{equation}
+
+We further assume that $T$ increases linearly with depth so
 that the entire upper crust is near the critical failure envelope
-\citep[e.g.,][]{townend2000}, are thus we parameterize the components of $T$
-as scalars multiplied by lithostatic pressure, denoted as $T^\prime$
+\citep[e.g.,][]{townend2000}, and thus we parameterize the components of $T$
+as scalars multiplied by $\rho g z$, denoted as $T^\prime$. Therefore, if
+$T\prime_{xx} = 0.1$, at some point just below 1 km, $L=27$ MPa, so
+$S_{xx} = M_{xx}$+ 27 MPa + 2.7 MPa.
 
@@
 $p_{J}(T^\prime_{\mathrm{max}} | D)$ has a somewhat more well-defined
-mode at $T^\prime_{\mathrm{max}} \approx 0.6$.
+mode at $T^\prime_{\mathrm{max}} \approx 0.6$. It is also apparent in
+Figure \ref{fig:joint_posteriors}~b that regardless of the magnitude of
+$T^\_{\mathrm{max}$, $p_{J}(T^\prime_{\mathrm{min}} | D)$ is
+consistently 0--0.6 $p_{J}(T^\prime_{\mathrm{max}} | D)$, with a mode
+of about 0.3; this indicates that the relative magnitude of the
+tectonic stresses has a substantial influence on the rake of the maximum
+shear stress resolved on the fault.
 
@@
 oriented, and is only slightly larger than lithostatic pressure. This
-stress configuration is compatible with the observed kinematics of the
-Wenchuan faults, and in close agreement pre-earthquake stress
+stress configuration is in close agreement with pre-earthquake stress
 measurements near the rupture zone (Figure \ref{fig:lms_stress_map}),
@@
 
-\begin{thebibliography}{87}
+\begin{thebibliography}{89}
 \providecommand{\natexlab}[1]{#1}
@@
 
+\bibitem[{\textit{Chen et~al.}(2010)\textit{Chen, Niu, Liu, Huang,
+  Tkal{\v{c}}i{\'c}, Sun, and Chan}}]{chen2010receiverfn}
+Chen, Y., F. Niu, R. Liu, Z. Huang, H. Tkal{\v{c}}i{\'c}, L. Sun, and
+  W. Chan (2010), Crustal structure beneath {C}hina from receiver function
+  analysis, \textit{Journal of Geophysical Research: Solid Earth}, 
+  \textit{115}(B03307), 1--22, doi:10.1029/2009JB006386.
+
 \bibitem[{\textit{Clark and Royden}(2000)}]{clarkroyden2000}
@@
   {A}merican intraplate stress field, \textit{Journal of Geophysical Research:
-  Solid Earth (1978--2012)}, \textit{101}(B4), 8643--8657.
+  Solid Earth}, \textit{101}(B4), 8643--8657.
 
@@
   M.~Kobrick, M.~Paller, E.~Rodriguez, L.~Roth, et~al. (2007), The shuttle
-  radar topography mission, \textit{Reviews of geophysics}, \textit{45}(2).
+  radar topography mission, \textit{Reviews of Geophysics}, \textit{45}(2).
 
@@
   slip of the 2008 {M}w 7.9 {W}enchuan earthquake estimated from {I}n{S}{A}{R}
-  and {G}{P}{S} measurements, \textit{Geophysical research letters},
+  and {G}{P}{S} measurements, \textit{Geophysical Research Letters},
   \textit{37}(1).
+
+\bibitem[{\textit{Fialko et~al.}(2005)\textit{Fialko, Riveral, and Kanamori}}]
+    {fialko2005}
+Fialko, Y., L.~Rivera, and H.~Kanamori (2005), Estimate of differential stress
+  in the upper crust from variations in topography and strike along the San
+  Andreas fault, \textit{Geophysical Journal International}, \textit{160},
+  527--532.
 
@@
 Lin, A., Z.~Ren, D.~Jia, and X.~Wu (2009), Co-seismic thrusting rupture and
-  slip distribution produced by the 2008< i> m</i>< sub> w</sub> 7.9 {W}enchuan
-  earthquake, china, \textit{Tectonophysics}, \textit{471}(3), 203--215.
+slip distribution produced by the 2008 {M}_w 7.9 {W}enchuan
+earthquake, {C}hina, \textit{Tectonophysics}, \textit{471}(3), 203--215.
 
@@
 Liu-Zeng, J., Z.~Zhang, L.~Wen, P.~Tapponnier, J.~Sun, X.~Xing, G.~Hu, Q.~Xu,
-  L.~Zeng, L.~Ding, et~al. (2009), Co-seismic ruptures of the 12 may 2008,< i>
-  m</i>< sub> s</sub> 8.0 wenchuan earthquake, sichuan: East--west crustal
-  shortening on oblique, parallel thrusts along the eastern edge of tibet,
+  L.~Zeng, L.~Ding, et~al. (2009), Co-seismic ruptures of the 12 {M}ay 2008, 
+  {M}_w 8.0 {W}enchuan earthquake, {s}ichuan: {E}ast--{w}est crustal
+  shortening on oblique, parallel thrusts along the eastern edge of {T}ibet,
   \textit{Earth and Planetary Science Letters}, \textit{286}(3), 355--370.

### 2014-09-25 revisions/rebuttal.md
added eric's rebuttal revisions

@@
 our description of the stress state. Lithostatic stress is an isotropic tensor,
-so (as described) it is only the diagonals of the tensor. Tectonic stress is
+so (as described) it has non-zero terms only the diagonals of the tensor. 
+Tectonic stress is
 only horizontal, $T'$ is scaled by $\rho g z$ (not by the tensor $L$, as we had
-previously, and erroneously stated). So $T$ and $L$ do not occupy the all same
+previously, and erroneously stated). So non-zero values of $T$ and $L$ are not 
+the same
 components of the stress tensor.
@@
 > We further assume that $T$ increases linearly with depth so
-> that the entire upper crust is near the critical failure envelope
+> that the entire upper crust is near the critical failure envelope at an unspecified coefficient of friction
 > [e.g., *Townend and Zoback, 2000*], and thus we parameterize the components 
@@
 
-We agree that $S$ increasing linearly with depth would be more analogous to
-*Townend and Zoback 2000*, but our model does not allow for this, because we
-would have to vary $T$ spatially to exactly counterbalance variations in $M$
+To increase $S$ linearly with depth,
+we would need to vary $T$ spatially to exactly counterbalance variations in $M$
 that deviate from linearity. Instead, we find it much simpler to approximate
@@
 
-This was something we puzzled over for a while. Ultimately, we chose our
-approach because it is less than obvious how to incorporate a depth-dependent
-tectonic stress into the calculations. The horizontal loading functions specify
+This was something we puzzled over for a while.
+However, the horizontal loading functions specify
 horizontal surface loads that are then propagated into the subsurface using
 Cerruti's solutions [Appendix A2]. The Liu and Zoback formulations of the
-loading functions require a fixed tectonic stress, or the tectonic stress at
-zero depth. Because we assume that tectonic stress increases linearly from zero
-with depth, the tectonic stress at the surface is zero.
+loading functions require a fixed tectonic stress at
+depths above zero (i.e., in the topography that is imposed on the surface of 
+the elastic half-space). Because we assume that tectonic stress increases 
+linearly
+with depth, with zero contribution of tectonic stress at the surface of the 
+half-space, tectonic stress does not need to be included in the horizontal 
+loading function of the second order terms of the topographic stresses.
 
@@
 of the question with the Bayesian inversion scheme we've chosen (there are
-about 18,000 models in the joint posterior).
+about 18,000 models in the joint posterior). While the assumption that the 
+tectonic component of stress vanishes in the upper few kilometers of the the 
+Earth (i.e., above the reference depth which is taken to be the top of the 
+elastic half-space), we feel we adaquately described this choice in our 
+construction of the model and with this choice are horizontal loading 
+functions in the calculation of topographic stresses is entirely 
+self-consistent.
 
@@
 We had chosen our the value of 0.25 for the Poisson ratio not only because of
-the calculational convenience (the shear modulus and Lame's parameters drop out
+ convenience (the shear modulus and Lame's parameters drop out
 of the equations) but because this is the most likely value for the rather
@@
 Additionally, from a tectonic standpoint, there is no real reason to consider
-N-S tectonic tension (the direction of $T_{min}$ in our results) in the Sichuan Basin;
+N-S tectonic tension (the direction of $T_{min}$ in our results) at the western 
+margin of the Sichuan Basin;
 this is a region surrounded by thrust and strike-slip faults of various
@@
 the patterns of deformation would be different in the interior of the Sichuan
-Basin away from topography; currently faulting there is thrusting, which
+Basin away from topography; currently here is thrust faulting in that region,
+which
 indicates a vertical $\sigma^3$. There is no evidence of N-S extension anywhere
@@
 
-The purpose of Bayesian priors is to keep likely values in as much as to keep
-unlikely values out. Because we adopt a relatively simple scheme of uniform
-probabilities over some interval, we can't add unlikely values at the same
-rate as more likely values, without either strongly diluting the number of
-acceptable models in the posterior (and therefore needing to run many more
-models, and we are already at the limit of what can be done with 128 GB RAM),
-and/or introducing a lot of unlikely models into the posterior, depending on
-how sensitive the results are to the priors.
+The purpose of Bayesian priors is to not only keep potentially likely models 
+in, but also as importantly to keep
+unlikely models out. Because we adopt a relatively simple scheme of uniform
+probabilities over some interval, adding unlikely models to the prior at the same
+rate as more likely models, will decrease the acceptance rate of models tested
+that are retained in the posterior,
+and therefore we would need to test many more models to obtain the same level 
+of sampling of the posterior as we do
+(we note that we are already at the limit of what can be done with 128 GB RAM).
 
@@
 tectonic stresses would be so high, although we chose the model prior to be
-a bit different than our personal prior beliefs.
+a bit different than our personal, unquantified prior beliefs. So in short,
+while we do not find a hard upper-bound, the posterior does not indicate that
+the likelihood of models at the upper bound is equal to the likelihood of 
+models at lower $T'_{max}$.
 
@@
 The requirement that $T_{min} < T_{max}$ is not simply an attribute of our
-priors or Bayesian estimation scheme, but basic math and logic. Minimum is
+priors or Bayesian estimation scheme, but minimum is
 defined to be less than maximum, so the half of the parameter space where this
@@
 
-So, although the joint probability density is higher towards the low $T_{max}$
-end (visible in Figures 6 and 8a) in our approach, the marginal priors for
-$T_{max}$ would be much more strongly biased in the first approach, but
+The marginal priors for
+$T_{max}$ would be much more strongly biased in the first sampling approach,
+while the marginals for
 $T_{min}$ is less affected by the sampling strategy. Overall, we feel that our
@@
 those of *Shen, et al.* [*2009*], *Feng, et al.* [*2010*], *Zhang, et al.*
-[*2011*], *Fielding, et al.,* [*2013*] and two from *Qi, et al.* [*2011*] (these two models share a
-fault geometry but use different smoothing functions for the slip
-distribution). These vary somewhat in the fault geometry and specifics of the
-slip distribution, but nonetheless are of high resolution, are properly
-georeferenced. By using a suite of models in our calculations instead of a
-single model, we evaluate results that are persistent in most or all of the
-models to be more robust, and other results can be more confidently linked to
-specifics of the model geometry or slip distribution.
+[*2011*], *Fielding, et al.,* [*2013*] and two from *Qi, et al.* [*2011*]. 
+DELETED LINE
+All of these models rely on geodetic data to some degree, but they do not all use the same data in their inversion (e.g., *Feng, et al.* [*2010*] uses a different InSAR catalogue as the others), all use differnet inversion stratagies, and different fault geometries (the two models of *Qi et al.* [*2011*] share a common fault geometry but use different regularization in the inferrence of the slip distribution).
+DELETED LINE
+By using a suite of models in our calculations,
+we can infer that results that are persistent in most or all of the
+models are more robust, while other results specific to only one of the coseismic slip models can be more confidently linked to
+specifics of the model geometry or inferred slip distribution in those models.
 
->In general, the models share a common geometry and pattern of slip distribution
-above 10 km or so; i.e. the locations of high and low slip patches and the slip
-rake are similar in all of the models. Though some of models use discrete fault
-segments for the Beichuan fault, whereas others use a single continuous fault,
-the fault planes are essentially collocated. However, there is significant
+>In general, the fault geometries in all of the coseismic slip models are simmilar, as well as the inferred pattern of slip distribution
+above 10 km or so (i.e., the locations of high and low slip patches and the slip
+rake are similar in all of the models).  Although some of models use discrete, planar fault
+segments, whereas others use a single continuous, non-planar fault,
+the fault geometry models are essentially collocated. However, there is significant
 variability in the magnitude of slip in the different models; for example, the
@@
 
-There is nothing in the abstract that references a topographic influence on
+There is nothing in the abstract that references a topographic influence on the
 rupture processes. In one sentence in the introduction (lines 41-45), we write
@@
 
-This is definitely a 'methods' paper, and this work was performed with the
-broader goal of working to quantify tectonic stresses in a larger sense. We
-chose this earthquake to start with because it's well-documented and has a
+We feel that it is crucially important to fully describe the methods of our calculations and estimation stratagy,
+We
+chose this earthquake because it's well-documented and has a
 complex coseismic slip distribution, and because it lies at the base of a very
@@
 
-Nonetheless, the results and the discussion make up the bulk of the paper. The
+Nevertheless, the results and the discussion make up the bulk of the paper. The
 methods sections (2 and 4) are 6.5 'manuscript' pages in sum, which includes
@@
 mentioned in the review (including a later paper by Flesch); those types of
-studies are of a much larger (orogen to global) scale than our study, and
-both a longer list of studies and a more detailed discussion of the studies
-is a bit out of the scope of the current paper.
+studies are of a much larger (orogen to global) scale than our study.
 
@@
 inversion, and because it involves a single event. The Luttrell and Sandwell
-paper uses focal mechanisms, which are basically point sources, and considers
+paper uses focal mechanisms, which are considered to represent constant coseismic slip over a planar fault plane corresponding to one of the nodal planes of the focal mechanism, and considers
 multiple earthquakes. This is a bit more different with respect to our paper.
+
 ## R17
@@
 planar fault segment--or models with very poor coseismic slip resolution. We
-did not feel it wise to include these (garbage in equals garbage out, as they
-say). Additionally, we did not feel it wise to list models that we thought
-were of too low quality.
+did not feel it wise to include these.
 
 We have added some discussion of the models used, and the criteria for
-including them, to Section 2.3, as described in R10.
+including them, to Section 2.3, as described in R10. We also reiterate here
+that not all of these coseismic slip models relied on the same data,
+approximated the fault geometry the same, or used the same inversion strategy
+to infer the coseismic slip. As any of these coseismis slip models are just 
+*models*, we do feel that basing our conclusions on just one particular model
+makes for the most robust study.
 
@@
 
-Corrected.
+All listed typos corrected.

### 2015-02-12 revisions/rebuttal2.md
committing my revisions prior to eric's edits

## General criticisms
>1 The grammar and spelling of the manuscript needs to be carefully checked.

In addition to errors caught by reviewers, we have rechecked and corrected
other errors.

>2 The fault geometry needs to be identified early in the manuscript. Right
now, the manuscript does not describe the fault dip anywhere, and this will
cause readers unnecessary difficulty.  These difficulties are particularly
apparent in section 3.

As described below, we have added descriptions of the faults in the text,
and made some figure annotations that we believe will be helpful. We state
here, as below, that the fault geometry is variable within each slip model
and between slip models, such that we cannot state a fault geometry as a
simple strike and dip.

>3 The manuscript confuses (or fails to clearly distinguish between) tractions
and stresses in several places. My experience is that this is common in the
Earth science community. It might be useful in an introductory section to state
that tractions are vectors that act on a particular plane, where as the stress
at a point is a tensor quantity.

We believe that most or all of our readers will know the difference between
tractions and stresses.  However, it is conventional (if slightly improper)
to refer to tractions as stresses, as in 'shear stress' and 'normal stress',
instead of 'shear traction' and 'normal traction'. Whether we are referring
to tensors, vectors, or scalars should be apparent and cause no confusion.

> 4 In several places I had difficulty telling whether the manuscript was
referring to points on the faults or to the fault as a whole. The manuscript
should make the distinction clear.

We have looked over the manuscript for this sort of problem but are not
sure where the reviewer is having trouble, beyond specific comments that are
addressed below.

> 5 The notation for stresses is unconventional, and that makes the manuscript
unnecessarily difficult to read. A more standard notation would be helpful.

We do not agree that it is quite so unconventional, and we have our reasons
for using the notation we chose; we have explained this below, and added some
discussion of this in the manuscript.

> 6 I think it would be better to describe certain distributions or functions
as posterior distributions or posterior functions rather than as “posteriors”
to avoid unintended reactions from some readers.

The term 'posteriors' has been well established in the literature for decades
and is unlikely to cause offense.

> 7 The figures and the text need to be better coordinated so that readers can
easily verify points made in the text when the figures are referred to. Several
of the figures can be modified to make them more useful to the reader. These
modifications should not be time-consuming to make.

We have made some minor modifications to figures.

> 8 A key modeling assumption seems to be that fault slip is not affected by
stress changes accompanying the fault rupture, but this assumption is not
stated. I think it should be.

We clearly state that our results are predicated upon the Wallace-Bott
assumption that fault slip occurs in the direction of maximum shear stress.

> 9 The comparison of the Beichuan fault to an optimally oriented fault needs
to be clarified. This seems to be a key point, but the discussion of this point
is not sufficiently clear.

We have reworked some of this, based on the comments.  However, it does seem
that the reviewer missed the key point that friction on the Beichuan fault
is less than friction on the hypothetical optimally-oriented faults, though
this is quite clear in the text.

## Specific Comments

### Abstract

**CA.1**

> Do they mean the whole fault was verging on failure at the time of the
earthquake, or only part of it?

We have changed the relevant sentence to 'We further estimate the static
friction and pore fluid pressure assuming that the fault was*, on average,* at
Mohr-Coulomb failure at the time of the Wenchuan earthquake.

**CA.2**

> Do they account for the topographic perturbation of the tectonic stresses in
coming up with this estimate? See for example Savage and Swolfs [1986].

We do not, because the method of Savage and Swolfs (as well as Liu and Zoback,
from whom we take our methods for topographic stress calculations) assume
an *a priori* tectonic stress value that is a constant. However, we assume that
tectonic stress is a tensor field that increases with depth; and we do not know
it *a priori*. Therefore, it is difficult to determine what the perturbation
to tectonic stresses is.

We have explained this more as part of the previous revisions.

### 1. Introduction

**C1.1**

> The only omission in literature coverage I am aware of is the work by
McGovern on topographic stresses associated with volcanoes. I think some of
McGovern’s papers from the 1990’s should be cited.

We have looked over these references, which are largely about volcanic
deformation on other planets, and have not found a good place to cite them,
without making a separate paragraph for discussions of topographic stresses in
volcanoes. This is far from our topic (the interaction of topographic stresses
and tectonics), and we did not read these papers during the work presented in
this paper, so we don't feel it appropriate to cite them.

**C1.2**

> This is the section of the manuscript where both the strike and dip of the
key faults should be stated.

We have added a paragraph and a half of description of the faults to the
introduction in the previous round of revisions.

**C1.3**

> (40-41) Referring to the Qur’an (or any other religious text) as one might
refer to a peer- reviewed scientific article is highly inappropriate. This
reference must be eliminated.

This reference has been eliminated, with the note that it is unfortunate that
we put up these sorts of walls in scholarship. The history of ideas shouldn't
be segregated so, and we did not refer to it 'as one might a peer-reviewed
scientific article'.

### 2. Topographic stresses on the Longmen Shan faults

**C2.1**

> This section came across well. I do think a description of the mechanics
behind equations 5 and 6 would be helpful though.

**C2.2**

> (169-170) The authors should be a bit more clear about some of the
limitations of the method of Liu and Zoback [1992]. That method cannot be used
to accurately calculate the topographically-induced stresses at shallow depths.
I think the method of Liu and Zoback is fine to use for the problem at hand,
but the method is not exact, and it does have its limitations. Those
limitations should be pointed out.

We change the sentence introducing the methods to "They show that the
topographic stress tensor field *beneath (but not within) topography* can be
determined by a convolution of topographic loading functions with Green's
functions describing the stresses in an elastic halfspace due to a point load
at the surface." (changes in italics). Additionally, in Section 2.2 on the
numerical implementation of the methods, we mention that our calculations only
apply to depths below 851 m.

**C2.3**

> (172) Why is M used to designate stress rather than $\sigma$, as is
conventional?  Some people might think M is for Moment. I strongly recommend
using σ for stresses.

Capital letters are quite commonly used to denote matrices, tensors, and
tensor fields. *M* is clearly defined in the manuscript.

See the response to Comment 3.9b, where we add text to the manuscript
explaining our notational usage.

**C2.4**

> (174) The line immediately after all the equations is indented and shouldn't
be. The indentations are supposed to signify the start of a new paragraph.

These are fixed.

**C2.5**

> (206-208) The open-source package is nice to make available to the community.

Yes, hopefully this will be useful for others in the future.

### 3. Results of topographic stress calculations on the Wenchuan faults

**C3.1**

> The manuscript needs to present figures and refer the reader to them to
verify the statements made in this section about the stresses. Right now
I can’t tell where to look to do this in several places in this section. Has
the manuscript omitted a figure?

No, the manuscript isn't missing any figures. Our results for this section
include each of the six stress tensor components, as well as the resolved
tractions ($\tau^M$, $\tau_d^M$, $\tau^M_s$, $\sigma_n^M$) for each of the six
coseismic slip models we've used. This would be 60 subfigures, clearly too many
to include as figures in the manuscript, so we have selected the most
illustrative results. All of the results are publicly available as .csv files
in the GitHub repository for project, and readers are welcome to download and
analyze these datasets if they wish to verify our claims, or otherwise inspect
the results. We have added a sentence informing readers of this to the end of
the first paragraph in Section 3:

"(Results not shown in figures are available as .csv files at
http://github.com/cossatot/wenchuan_topo_stress/)."

Additionally, in response to comments in the earlier round of revisions, we
have added vertical stresses to Figures 3 and 9.

**C3.2**

> Also, the manuscript needs a figure showing the model geometry. Also, I have
difficulty with the non-standard notation.

As mentioned in the response to Comment 1.2, we have added a descriptions of
the coseismic slip models used, and on the general geometry of the relevant
faults, to Section 2.3, as part of the previous rounds of revisions.

Additionally, as is evident in Figure 4, there is considerable variation in the
fault geometries assumed or inverted for by different researchers, so it is not
possible to accurately show a figure with model geometry, because there is no
real consensus on it. (To be clear, shallower than about 8 km depth, all of the
slip models we've used are quite similar in geometry and location, but below
this depth they start to diverge, as some models assume planar geometry and
others assume listric geometry, and so forth).

We have added some annotation of the dips and depths of various parts of the
slip models shown on Figure 4, in order to help readers.

The issue with notation is addressed in the response to Comment 3.9b.

**C3.3**

> Much of this section needs to be reworked so it is clear to the reader.

Specific criticisms of clarity are addressed in the following
comments and responses.

**C3.4a**

> (220) What is the purpose of the superscript M here for the traction
components? If does not seem to serve a purpose. Is it to designate tractions
due to topographic loads? If so, I recommend the authors be more explicit about
that.

Yes, the superscript M denotes the topographic origin for stresses. See the
response to Comment 3.9b.

**C3.4b**

> (220) See my comment on line 243 regarding the notation.

See our response to Comment 3.9b.

**C3.5a**

> (222-230) I can’t tell where to look to verify the statements about
stresses. Which figure, for example, shows Myz, the vertical normal stress?
Which figure shows Myz, the horizontal shear stress on a vertical plane that
strikes north?

See our response to Comment 3.1.

**C3.5b**

> (222-230) At what depth or location is the first sentence of the paragraph
referring to?

This sentence ("Topographic stresses on the Wenchuan faults are on the order
1-10s MPa") refers to the entirety of the faults, although it is a general
sentence and there are a few points on the faults where the stresses are below
this range.

**C3.6**

> (225-226)  Which figure shows the vertical normal stress? I can’t see a plot
in Fig. 3 or Fig. 4 that shows the vertical normal stress.

See our response to Comment 3.1.

**C3.7**

> (240-241) Which figure should the reader look at to see the faults in the
model?

We have added a reference to Figure 4b and c.

**C3.8**

> (242) Where are the locations being referred to here? What are their x,y,z
coordinates (or latitude, longitude, and depth)?

The Pengguan fault is shown on the maps (Figures 1, 3, 9). The flats in the
Beichuan fault are at the base of the southwestern part of the fault, as is
evident in Figure 4, which shows some of the fault models. We think that giving
'x,y,z' or other spatial coordinates would be quite unhelpful (particularly for
curved surfaces tens of kilometers in breadth), so we will not do this.  If
readers need to know this information, they can look at the .csv files on
GitHub or consult the sources for the slip models.

**C3.9a**

> (243) What is the dip of the fault in question?

This varies in the different fault models considered. If the reader wants to
know these specifics, he or she is welcome to read the papers that have
produced the fault models used.

**C3.9b**

> (243) The notation here, where M (with
double subscripts) refers to a stress, $\sigma$ (with a single subscript) to a
stress (but it would be more appropriately called a traction), and $\tau$ (with
a single subscript) to a stress (but again it would be more appropriately
called a traction) is very confusing to me. I strongly recommend using a
convention like so:

>$\sigma_{ij}$ = stress

>$\tau_n$= normal traction on a fault

>$\tau_s$ = strike-slip traction on a fault

>$\tau_d$ = dip-slip traction on a fault

>This is much more in keeping with standard usage and would be easier for
(most) readers to follow I think.

I do not think I have ever seen $\tau_n$ to refer to normal stress/traction;
we use $\tau$ for shear stress (tractions), which is quite standard.  We also
use $\sigma_n$ as normal stress (traction) which is standard. We use *M* to
denote the topographic stress tensor or tensor field, and it is clearly defined
as such. Capital letters are extremely commonly used to denote arrays,
matrices, tensors and tensor fields; this is not only a structural geology
convention but spans the science, mathematics and engineering disciplines.

We considered adopting the reviewer's suggestions (outside of using $\tau_n$
for normal stress) but in the end we do not use $\sigma$ to represent a stress
tensor or tensor field because we would have to exponentiate it to, for
example, $\sigma^{M^B}$ or $\sigma^{M,B}$ when referring to the stress tensor
field resulting from the convolution of topography with the Boussinesq Green's
functions. This is unwieldy. It would also be quite awkward to refer to the
normalized (dimensionless) tectonic stress field as $\sigma^{T^{\prime}}$ even
though it does not have values in Pa, or we could keep it as $T^\prime$ which
could further confuse readers as it is then one more step removed from *T*.

We have added an explanation of our notation to the first paragraph of Section
2.1:  "Note that in this work, we use capital letters (e.g., $M$, $T$) to
denote tensors and tensor fields (depending on context, which should be clear)
and Greek letters ($\tau$, $\sigma$) to denote stress components projected or
resolved on planes, which are more properly called tractions. We use
superscripts on these symbols to denote the origin of the stresses, and
subscripts to denote components of these tensors or stresses or tractions."

**C3.10**

> (244-245) Which figure should the reader look at to verify the statement
here?

We disagree that a figure is necessary, or particularly helpful, to verify a
claim that vertical stresses resolve more as normal stress than shear stress
on a plane with low dip. This should be apparent to readers who have taken a
course in Structural Geology or who have a continuum mechanics background,
which surely includes the vast majority of our readers. Readers without this
background will unfortunately have a quite difficult time with the paper, but
they are not our intended audience.

**C3.11**

> (249-251) I don't see where to look to verify the nature of the slip and the
tractions on Figure 4.  The blue background color of the slip magnitude figure
(upper left) looks very uniform, so I can't tell where the slip is high; also
note that the contour lines are not labeled with magnitudes. In the normal
stress plot in the upper right the red background color is very uniform, so
I can't tell where the normal traction is high (and the contour lines are not
labeled). This means I don't know how to compare the plots.

The colors shown are not simple 'background colors' but correspond to the
magnitudes of the stress component in question. Where they are uniform,
stresses are uniform. Labeling the contours would add unneccesary clutter to
the figures, making them harder to read. Nonetheless, the relevant comparisons
(for example, between high normal stress and low slip) are still quite evident
to us, both on the screen and in print, and other viewers have not had problems
with this. Maybe the reviewer has a monitor or printer calibration issue.

**C3.12**

> Where are these other fault patches? How big are they? Can the authors point
the reader to a figure to show these patches?

The 'other' fault patches are those that are not the ones described (i.e. fault
segments that are not the southwestern portion of the Beichuan fault). Figure
4 (referenced immediately before the sentence in question) illustrates where
the observed correlation between slip and normal stress is strong and where
it is not.

### 4. Calculations of tectonic stress, fault friction and pore fluid pressure

**C4.1**

> This section reads fairly well, although a fuller description of the Bayesian
statistics would have helped me.

We believe that we have explained the mathematics involved in our calculations
(Bayes' rule, the misfit equation, and the likelihood equation, which are
equations 10-12 in the original submission and equations 11-13 in the revised
submission). These are uncomplicated expressions. Unfortunately, though
Bayesian statistics are intuitive in practice and a formalized version of our
thought processes in many instances (revising ideas or beliefs based on
evidence), there is some sort of barrier for many people in grasping it
initially. For the first author, this was due to jargon, so we have tried to
define and explain each of the terms. We have also cited two excellent texts
[Tarantola, 2005; Mosegaard and Tarantola, 1995] that explain Bayesian
statistics in more detail.

**C4.2**

> (256-257) The statement here should also account for cohesion, although the
frictional resistance could be much larger than any intrinsic cohesion

This is true; we have added "and assuming that rock strength (cohesion) is
negligible" to the sentence describing the Mohr-Coulomb failure envelope.

**C4.3**

> (258) I think the authors mean when the earthquake initiated, not the fault.
The fault is presumably millions of years old.

Yes, this is what we meant.  We have replaced 'fault' with 'earthquake' in this
sentence.

**C4.4a**

> (265) The phrase "posteriors of T" could have unanticipated (and undesirable)
reactions from some readers. I suggest using the alternative phrase "these
functions".

The use of the word 'posterior' in statistics is well-established, and surely
won't incite any more unintended reactions or schoolboy jokes than 'thrusting'.
We advise readers who blush at the term 'posterior' to avoid television and
the internet at all costs.

**C4.4b**

> (265) Also, what is "T"?

*T* has been defined in this sentence as part of the previous revisions.

**C4.5**

> (274-275) The grammar and punctuation here need to be corrected.

We have fixed these problems as part of the previous revisions.

**C4.6**

> (276-287) Owing to my poor background in Bayesian statistics, I cannot
comment knowledgeably on the technical aspects of this paragraph except for one
point.  The authors should clarify whether the are evaluating the direction of
slip based on the conditions before the fault has begun to slip or as the fault
slips and the stress field around the slipped patch is evolving.

The direction of slip is given by the coseismic slip models, which are the net
slip from the earthquake; this is described in Section 4.2.  A consideration
of rake variability during the rupture would be excellent but unfortunately
we do not have the data to do this.

**C4.7**

> (290) The authors need to clarify what orientations they are referring to
here.

We have clarified the orientations as stress orientations.

**C4.8**

> (291-292) The statement here is true, but only because the lithostatic
stresses are defined to have a horizontal components equal to the vertical
components. The <u>definition</u> of the tensor L thus plays a key role in the
results.

Yes it does, but we have been explicit in our definition of *L* (section 4.1).

**C4.9**

> (293-294) Is the maximum tectonic principal stress being referred to here? If
so, I think it would be a good idea to state it here, even though the topic
sentence of the paragraph deals with principal tectonic stresses. It is very
important that the critical thoughts in the manuscript at this point be crystal
clear to the reader.

We have added 'tectonic' to 'maximum principal *tectonic* stress' and 'minimum
principal *tectonic* stress' to clarify this in these sentences.

### 5. T, $\mu$, $\phi$ results

**C5.1**

> Owing to my poor background in Bayesian statistics, this section was
difficult for me to understand. Other readers might be in the same situation.
I think the section would be substantially improved if it concluded with a new
brief paragraph that was written “in plain English” that gave the key points
the reader should take away from this section.

We do not find our descriptions of the results to require much of a knowledge
of Bayesian statistics.  It might be that the notation (which is the standard
notation of conditional probabilities) is confusing, but this is defined in
the preceeding sections.

**C5.2**

> In spite of my shortcomings in Bayesian statistics, I am pretty sure the key
point that the reader should take away from this section is not provided by the
last sentence in the section (“On the other hand, it is not similarly sparse on
the low-μ side, suggesting that low values for μ are more robust.”).

The placement of the key points of a section to be in the last sentence of that
section is a stylistic preference, and not one that we share (why not put the
important information towards the top of the paragraphs rather than at the
end?). There is not really a 'key point' here per se; this is a description of
the results. The important results are discussed as such.

**C5.3**

> (328-330) After this sentence, say what the significance of the peak in
T'(min) at 0.2 is, as was done in lines 326-328 for T'(max).

We don't view the value of the minimum tectonic stress to be as significant
as the value of the maximum stress, and there is simple statement that can
summarize the importance of this value (i.e. it does not directly affect much).
We do however discuss the results more in the discussion.

**C5.4**

> (331-332) The described ‘linear decrease’ is not linear. The histograms have
an envelope that appears to be consistently concave up. In some work I've been
involved with the difference between a linear decay and an almost linear decay
is very significant. I don't know if that is the case here, but I still think
it is important not to call something "linear" when it isn't.

We have added a wiggle word ('somewhat') so that readers do not interpret these
descriptions too literally.

**C5.5**

> (336-340) The authors describe two sets of conditions that they consider
<u>unlikely</u>. I think it would be helpful for them to conclude here what
they do think is <u>likely</u>.

We think the most likely values in our results are the most likely. When we
stated the conditions that were unlikely, we were explaining why they
were not well represented in the posteriors.

**C5.6**

> (341-345) I can't see a peak in T'(Max) at ~0.6 in any of the plots on Fig.
8. Some additional explanation is needed, at least for me.

The highest values of the $T'_{\max}$ marginal histograms at around 0.6 are
evident. This 'peak' (not our wording) is not particularly sharp, but it is
clearly there; there is a steep decrease in probability to the left of the
mode and a more gradual decrease to the right of the mode. As stated in the
response letter to the previous revision, the probability at 2.5 (the upper
boundary of $T'$ is 60% of the mode at 0.6.  We have amended this sentence
to read '0.6-0.8' so it is clear that we do not mean a narrow range of high
likelihoods around a single value.

**C5.7**

> (346-355) I could not follow this paragraph. Although my background in
Bayesian statistics is poor, many other readers might be in the same boat.
I think I am able to follow the general drift of this paragraph up until line
351. At that point the paragraph could reference some figures to clarify
matters, but no figures aren’t referenced. To be specific, here is what I am
referring to:

>“However, pJ(μ|D) has a lower relative likelihood on the high-μ tail compared
to the constituent p(μ|D).”

> <u>By citing the specific subfigure on Fig. 8 that shows this, the reader
would be better positioned to verify this.</u>

We have changed the sentence to read: "However, $p_{\mathrm{J}}(\mu | D)$ has
a lower relative likelihood on the high-$\mu$ tail (Figure 8d) compared to
$p(\mu | D)$ of the constituent slip model results (Figure 7)."

> “This lower likelihood of m in the joint posterior is likely bec[au]se it is
the average μ of all slip models. On the other hand, it is not similarly sparse
on the low-μ side, suggesting that low values for μ are more robust.”

> <u>How low is "low"? Can the authors be quantitative here?</u>

We added a reference to the figure (mentioned in the last response).  We would
rather readers look at the distribution itself than a number we throw out.

### 6. Discussion

**C6.1**

> The comparison of the Beichuan fault to an optimally oriented fault needs to
be clarified. This seems to be a key point, but the reasoning is not
sufficiently clear.

We address this in the comments below.

**C6.2**

> (356-363) This paragraph seems fine, but I do note that the structural
geology literature does contain some papers that discuss the effect of stress
variations on fault slip (e.g., Cooke, 1997; Cooke and Underwood, 2001; Martel
and Shacat, 2007; Savage and Cooke, 2010; Kaven et al., 2013). These papers
contain quasi-static analyses, not dynamic analyses, but they appear as though
they might be relevant to the discussion of lines 356-363.

We have looked through these papers and do not see an easy connection to our
work.

**C6.3**

> This paragraph looks fine to me.

excellent.

**C6.7**

> (389-390) I recommend that this concluding point of the paragraph be
developed a bit more.

It is sufficiently self-explanatory, and unimportant, that we don't feel it
warrants further discussion.

**C6.8**

> (391) How does a spatial variation decrease with frequency? Space and time
seem to be being mixed here. Also, what is the amplitude that is referred to
here? Do the authors mean that the spatial variation of topographic stresses at
various depths depends on the amplitude and wavelength of the topography?

We have clarified this sentence:

"The spatial variation of topographic stresses decreases in frequency and
amplitude with depth" is now: "The spatial variation of topographic stresses
increases in wavelength and decreases in magnitude with depth."

**C6.9**

> (399) What does "this" refer to? I encourage the authors to review the topic
sentence of this paragraph as they address this question of mine.

We have replaced "this" with "the relationship between stress variation, slip
variation, and depth", which is the topic of the paragraph.

**C6.10**

> (402) Neither the dip nor the dip direction of the Beichuan fault has been
described, so the reader cannot verify the statement here. The strike and dip
of all the key faults need to be identified.

As stated in response to Comment 1.2, we have added a section in the
introduction describing the faults and coseismic slip models.

**C6.11**

> (403-404) The authors should be very careful with the wording here. They have
just described the magnitude and orientation of Tmin, and many readers might
interpret the phrase “This stress configuration ... is in close agreement
[with] pre-earthquake stress measurements...” to mean that the measurements
include magnitude and orientation. Fig. 9, however, shows only orientation
data, not orientation and magnitude data. To my knowledge, borehole breakout
data have not proven yet to be reliable indicators of stress magnitudes.

We have changed the sentence from "pre-earthquake stress measurements" to
"pre-earthquake stress orientation measurements".

**C6.12**

> (413-414) Cite a figure to buttress the statement here.

A reference to Figure 8 has been added.

**C6.13**

> (415) What does the word “This” refer to? Again, I think referring to
a figure would help here considerably.

'This' changed to 'These results'.

**C6.14**

> (421-422) Check the grammar here. A word seems to be missing.

We have added 'any' which makes the sentence a bit more parseable.

**C6.15**

> (423-424) I'm not quite sure what the authors are trying to say here. I think
the wording here should be revised to clarify the sentence.

We have modified this sentence, which now reads "The orientation of both the
tectonic and total stresses near the Wenchuan faults shows a larger difference
with patterns of strain from elsewhere in the orogen *than in the Longmen Shan
region*." (changes in italics) which hopefuly clarifies our meaning.

**C6.16**

> (436-437) Need to revise the grammar here.

Revised.

**C6.17**

> (451) Where would these more optimally oriented faults be located, and how
would they be oriented?

As we state in the paper, these calculations apply to each point in the
coseismic slip model, so the optimal faults would be collocated with the
points of the slip model. Because we treat each point independently and
the comparisons are of local stresses, it doesn't make a difference that
with different orientations, the optimal faults wouldn't align into a single
plane.

**C6.18**

> (434-479) The conclusions of this paragraph, that the Beichuan fault is more
likely to slip than an optimally-oriented fault, will strike most readers as
impossible. How could the Beichuan fault be more optimally oriented for slip
than a fault that has the optimal orientation for slip? The "best" (in terms of
slip probability) that could be hoped for is that the Beichuan fault has the
orientation (and position) of the optimally-oriented fault, and therefore is
the optimally-oriented fault. How is it possible that slip on the Beichuan
fault is favored by a factor of as much as 1000 (the anti-log of 3) over the
optimally-oriented fault? I think the answer lies in the definition of what
constitutes the optimally-oriented fault. This needs to be more clearly and
completely explained to avoid confusing readers.

We believe most readers will not view slip on the Beichuan fault as impossible,
because it is believed by all researchers we know of to have happened in the
Wenchuan event.

The reviewer may be missing a key point, which is that the 'optimally-oriented
fault' has a canonical value of friction of 0.6 (lines 466-469 of the
submitted manuscript). In contrast, our model results indicate that the faults
that are thought to have ruptured in this event have a much lower coefficient
of static friction.  As is shown in Table 2, the coefficient of friction on the
fault has a strong influence on the favorability of slip on the Wenchuan
faults vs. optimally oriented faults of the stress parameters studied.
This is stated repeatedly throughout this section, so we do not believe that
we are unclear about this point.

Additionally, when we did the log transformation of the y-axis of Figure 10,
we used the base-*e* log, not base-10; the anti-log of 3 in this case is ~20,
which is far more reasonable value than 1000. We have updated the text and
figure captions so that 'log' now reads 'ln' to avoid this confusion.

Nonetheless, we have reviewed our code and an error in how the optimal
slip plane is found; essentially, we the algorithm worked assuming that dip
slip faulting would be favored (where the optimal plane strikes perpendicular
to the trend of $\sigma_1$) but produces an incorrect orientation otherwise.
We have fixed the bug, which has changed these results slightly (the bug only
affected these calculations; that particular function was not used in the rest
of the study). The new Figure 10 shows that more of the data plot in the
favorability region for optimal faulting than previously; however, most of the
data (with low friction values) still plot in the favorability region for the
Wenchuan faults.

We have also added some discussion of how the calculated ratios relate to the
parameters:

"It is clear that both $T^\prime_{xx}$ and $\mu$ are strongly correlated with
$\mathrm{CFR}_o/\mathrm{CFR}_f$, and $\phi$ is to a lesser degree; nonetheless,
all significantly affect the relative ease of faulting on the Wenchuan faults
versus optimal faults. In particular, lower values for any of them favor slip
on the Wenchuan faults. The lowest $\mathrm{CFR}_o/\mathrm{CFR}_f$ value
($\sim$0.057, $\sim$ -2.9 in log space) corresponds to the lowest value of
$\mu$ ($\sim$0.038), and is approximately the ratio of $\mu$ in the model to
0.6."

**C6.19**

> (480) I find more and more careful writers are reserving the term
"significant" to discuss "statistically significant" matters. That is not quite
how the term is being used here, even though the manuscript deals quite a bit
with statistics. I think the term "substantial" would be a better choice here.

This is an excellent suggestion.

**C6.20**

> (488) Cite a reference on deformation in Tibet to support this point.

We have added three references here.

**C6.21**

> (493-496) This sentence needs to be revised. The punctuation and/or grammar
here needs some revision (the semicolon is what initially caught my attention),
but the sentence contains too many incomplete thoughts without saying what the
main thought is. Also, if “much work has been done”, then wouldn’t it be
appropriate to cite more than one reference at the end of the sentence? See my
suggestion for rewording.

Grammatical changes to the manuscript were made as part of the previous round
of revisions. We have added a few more references to the sentence.

**C6.22**

> (499) I think a new paragraph should be started after the word "crust" here.
The sentence that starts with the word "However" does not really constitute the
conclusion of this paragraph as well as it perhaps serves as the topic sentence
for the next paragraph.

We disagree with this stylistic point. That paragraph discusses how the
treatment of large topographic or orogenic features as smooth and homogeneous
is an oversimplification, and the 'however' statement serves as a
counter-argument. It is also meant to lead to the next paragraph.

**C6.23**

> (504) Cite a reference. Say who noted this point.

This sentence clearly discusses our results.  We have added a reference to
Section 3.

**C6.24**

> (509-510) I think this statement is a bit too general considering what Fig.
3 shows. Near the southeast end of the fault trace, the most compressive
horizontal stress trends nearly normal to the fault trace. Near the northeast
of the fault trace, the most compressive horizontal stress trends nearly 60
degrees counterclockwise from the fault trace. So the relative orientation of
the most compressive stress relative to the fault trace trend varies with
position along the fault, not just the position perpendicular to the fault
trace. The relative orientation changes more along a traverse perpendicular to
the fault at the SW end of the fault than the NE end (I see little variation in
the stress trajectories there along a traverse normal to the fault trace.
I think the authors should be a bit more nuanced with their statement here.

We disagree with the reviewer's interpretation of the results in Figure 3.  The
orientation of maximum topographic stress in the southwest (the fault strikes
SW-NE and has no 'southeast end') rotates nearly 90 degrees from under the high
Longmen Shan across the fault into the Sichuan Basin. Close to the fault, the
orientations change very little along strike.

Additionally, this part of the paragraph concerns the loading of thrust flats
in the southwestern Beichuan fault. The coseismic slip models in this study
do not have these kinds of structures in the northeast, so the situation there
is not relevant.

**C6.25**

> (510-511) The direction of the most compressive principal stress apparently
is being equated with the direction of the greatest displacement. That
equivalence is incorrect. If the fault exhibits thrust slip, a displacement
component will exist normal to the trace of the fault.

If the fault exhibits a thrust slip, it will also have a component of the
greatest principal stress normal to the fault trace...  We are not sure of what
this comment is really referencing or what an appropriate revision would be.

**C6.26**

> (514) I think the authors need to say just what a "channel" is.

We reference several studies discussing channel flow in Tibet in this paragraph
and in preceeding ones. This part of the discussion is sort of an ancillary
section relating our results to several popular ideas about orogenic dynamics
and development; readers who are interested may look at those, but suffice it
to say that no strict definition of what an orogenic channel is exists beyond
the basic definition of a channel as an area of increased velocity and/or
decreased viscosity in between more rigid walls.

**C6.27**

> (517-519) think the thoughts in this sentence might come across more clearly
if the sentence is broken into two pieces. This sentence constrains more
thoughts than easily fit into one sentence.

I think this sentence was actually too unconstrained by thoughts, because it
doesn't make much sense several months later. I have rewritten the sentence to:

Our calculations of $\mu$ and $\phi$ are based on $T$, which is biased towards
the high slip patches in the upper crust (Equation 11).  The resulting low
values of $\mu$ and low to moderate values of $\phi$, probably yield stress
conditions insufficient for slip at high pressures found at depths > 10 km.

**C6.28**

> (524) Good paragraph.

Thanks!

### 7. Conclusions

**C7.1**

> The conclusions are straightforward and follow from the preceding material.
Please note, however, my comments on lines 544-547.

**C7.2**

> (536) Why are the absolute value bars included here?

Our convention in this paper is that positive values for $\tau$ indicate
reverse to right-lateral shear while negative values indicate normal to
left-lateral shear (Section 3). The absolute values are there because we
see shear stresses with both signs up to that magnitude.

**C7.3a**

> I don't think the data support going to four significant figures here. Two
significant figures seem more appropriate.

Changed.

**C7.3b**

> Stress has dimensions of MPa, not MPa/km.

In this study, we parameterize tectonic stress as a function of depth.
We have changed 'MPa km$^{-1}$' to 'MPa per kilometer of depth' to make this
more clear.

**C7.4**

> (544-547) It seems odd to give the most compressive stress to four
significant figures, and then in the next sentence describe the least
compressive stress qualitatively (no significant figures). I recommend that the
authors give quantitative values for both.

We have reworded this sentence for clarity, and given the results with both
more quantitative and less precise values, as suggested.

**C7.5**

> (550-551) I find that avoiding one-sentence paragraphs generally is good
procedure. I suggest expanding this final paragraph to at least two sentences,
or else merging it with the preceding paragraph.

We have merged it with the preceding paragraph.

### Appendices

**CAp1**

> I can't read equation (A9) because it is overwritten by the date.

This is a LaTeX problem which we couldn't easily fix and won't be
an issue with the final version.

**CAp2**

> Equations (A4)-(A6) match the solutions in Mal and Singh (1991), but
equations (A1)-(A2) do not match and have errors. Owing to the different
elastic coefficients Mal and Singh use, I can’t tell if equation (A3) is
correct or not. Also note that G must have dimensions of force/area, which
equations (A4)-(A6) do, but the lead terms in square brackets in equations (A1)
and (A2) do not. Aside from those differences, the elastic coefficients need to
be defined. Similarly, equations (A10) and (A12) match the solutions in Mal and
Singh (1991), but I can’t tell whether equations (A7)-(A9) do because of the
different elastic coefficients. Since the authors did not derive these
equations, they could simply cite a reference and they would not need to
include an Appendix, although they should check to make sure the errors in
equations (A1) and (A2) did not spill over to their codes.

The equations have been amended (a *z* was missing in the numerator of the
first term in the brackets). The mistake is not found in the code.  This fixes
the unit problem mentioned.  Although we reproduce the equations from Liu and
Zoback [1992], we reproduce them here for reference, and because we actually
only reproduce the right-hand sides of the equations from that reference, but
we use different terms on the left-hand sides.  (These equations are
essentially definitions, so there is no algebra or other operators on the
left-hand sides).  The elastic coefficients

**CAp3**

> For equation (6), what is σ B? Should it be M B? Also, shouldn't the
subscripts for F be h,y yy yy instead of h,yz?

Yes, $\sigma^B$ should be $M^B$, and $F_h,y$ instead of $F_h,yz$.

**CAp4**

> In equation (5) why are all three terms on the right side of the equal sign
in the first row multiplied by dh/dx? I can see why the ρgh term is, because if
the surface has a slope, then the slope will impart a horizontal shear stress.
I don’t understand why Mxx and Txx are multiplied by dh/dx though, because Mxx
and Txx already act horizontally. Similarly, why are Mxy and Txy multiplied by
dh/dy in the second row of the equation? Analogous comments apply to equation
(6). An explanation of this would be helpful.

### References

**CR1**

> I have not checked all the references because of time constraints.
I recommend adding at least one reference by McGovern and some of the suggested
references mentioned regarding lines 356-363.

We have not found any of these references to be appropriate.

### Figures and Captions

**CF1**

> Is the “Grey box” mentioned in the last sentence of the caption referring to
the small box in the inset figure in the lower right corner?

No, it's the oblique hollow grey rectangle in the center of the figure, labeled
'Figs. 3,9' (which is also referenced in the caption).

**CF2**

> What do the red and blue areas represent? They are what immediately attract
the reader’s eye, not the traction arrows. The emphasis should be on the
traction arrows in my opinion.

They represent crystalline rocks (in pink) and sedimentary basins (in blue).
The colors were chosen to match some of the geologic maps of the region.
We have explained the colors in the caption.

**CF3a**

> The explanation should define all the symbols on the map, including the
traces of the thrust faults.

The thrust fault symbols are standard across geology and geophysics and have
been for decades.

**CF3b**

> Are the fault traces shown the traces of real faults or model faults?

These are real faults.  As described in the caption for Figure 1, they are
active structures taken from Styron et al., 2010.

**CF4**

> I can’t tell what the plots in this figure represent. Are they plots of the
slip and various stresses on the plane(s) of the model faults? The caption
needs to be very clear about what these figures represent.

The first sentence of the caption reads: "Southwest-looking views of
topographic stress and coseismic slip on selected slip models." This should
clearly answer the reviewer's question here affirmatively.

**CF4a**

> The sub-figures here need to be labeled, and the subfigure in the upper
right or “normal stress” needs to describe the orientation of this normal
stress component.

Normal stress is by definition oriented perpendicular to the fault. It's not
customary to describe it with other orientations, as far we we are aware.

**CF4b**

> Say what the dip of the model fault is here.

The dips of the faults are variable throughout these models. We have annotated
the dips on the figure.

**CF4c**

> The caption is cut off in the pdf file I received.

There should have been a version of the manuscript without figures that clearly
displays the caption available through GEMS; at least it was visible to us.

**CF4d**

> The figure caption should say what the dip of the model fault is.

The dips of the faults are variable throughout these models. We have added
some annotation with dips in various places on the figure.

**CF4e**

> Is the slip being modeled in this manuscript, or is this the slip that
was modeled by others? If the slip is being modeled here, the manuscript needs
to say what the boundary conditions are along the model faults.

No, the slip is not being modeled here, and we are quite clear that we use
published coseismic slip models.

**CF4f**

> Where to the “selected slip models” come from?

From the sources listed in the caption.

**CF5**

> Fine.

**CF6a**

> What do the subfigure labels (e.g., “Zhang”, “Feng”, etc.) mean? Are they
faults? Authors of publications? Types of statistics (e.g., Qi rough and Qi
smooth)?

They are the authors of the slip models used in this study. 'Qi rough' and
'Qi smooth' describe the slip models, using different smoothing techniques,
as described in the *Qi et al* paper cited repeatedly in this manuscript.

**CF6b**

> The meanings of the subplots need to be described better. Each of the
6 subplots (e.g., Zhang, Feng, etc.) has four sub-subplots. Neither the caption
nor the text describe what the upper left sub-subplot is or what the lower
right sub-subplot is. Both of these sub-subplots are solid blue. They look like
histograms, and perhaps they give the number of points in bins that represent
thin columns or rows that extend from the apparent histograms. In any event,
I’ve never seen a plot like this – and I might not be the only person in this
situation – so a more complete description of these sub- subplots would be most
helpful.

We have changed the term 'marginals' to 'histograms of the marginal
distributions' which will hopefully give the uninitiated reader some help,
or at least some more precise or formal terms which will be more conducive to
searching.  It is rather basic statistical terminology and presentation.

**CF7**

> I don’t really understand how the contour lines were drawn in this figure.
Some clarification on this would help. See also my comments on Fig. 6.

We have added that the contours 'are constructed through kernel density
estimation' to the caption for Figure 7.

**CF8a**

> What are “marginals”? I don’t know what they are, and they are not defined.

We have changed the term 'marginals' to 'marginal distributions'. This is a
common statistical term and doesn't need to be defined here, any more than
we would define 'standard deviation'.

**CF8b**

> I can’t match up the number 0.6 in line 345 with a peak in the apparent
histograms at 0.6 in this figure. Is 0.6 the correct number? If it is, could
the subfigure that shows the peak at 0.6 be labeled so the reader can see where
to look?

We have changed that line to '0.6-0.8' to make it more clear that it is not
a 'peak' but a wide area at the top of the histogram.

**CF8c**

> The histogram of azimuths” is called a rose diagram earlier in the
manuscript. Why not call Fig. 8c a rose diagram?

Would it smell sweeter if we did?  The two have the same meaning, and we are
free to use both names.

**CF9**

> The wording "Topographic and tectonic horizontal stresses" implies that the
figure shows (a) topographic stresses, and (b) tectonic stresses, yet the
figure does not show them. The blue arrows show total stresses (topographic
+ tectonic + gravity) as do the purple arrows. The red-and-black crosses show
topographic + tectonic stresses, not total stresses, and not just topographic 
or tectonic. The description of just what this figure shows needs to be more
carefully described – this might involve the text and the caption.

And' in the first sentence is more appropriate than either writing out 'plus'
or putting a plus sign in a sentence.  The blue and purple arrows are clearly
defined as stress orientations in the caption.  We don't see where this
causes any confusion, and it does not seem that the reviewer is confused
either; rather he or she has an overly restrictive view of the meaning of
'and'.

**CF10**

> The figure and its caption look fine to me.

No changes.

## Additional changes

Introduction:

We have added "We then use the Mohr-Coulomb failure criterion
to bracket the tectonic stress field, pore fluid pressure, and static friction
of the fault*, so that these parameters are consistent with the toopographic
stresses and coseismic slip on the Wenchuan earthquake faults*." to the last
sentence of the third paragraph of the Introduction, where we describe the work
(changes in italics).

Section 'Tectonic stresses in eastern Tibet'

We have changed a clause from "$T_{\mathrm{min}}$... is slightly larger than
lithostatic pressure" to $T_{\mathrm{min}}$... is a small fraction of
lithostatic pressure", as the previous statement was incorrect.

### 2015-02-12 revisions/styron_hetland_wenchuan_topo_revised.tex
committing my revisions prior to eric's edits

@@
-\documentclass[twocolumn,jgrga]{AGUTeX}
+\documentclass[draft,jgrga]{AGUTeX}
 \usepackage{mathptmx}
 \usepackage{amssymb,amsmath}
-
-%\usepackage{lineno}
-%\linenumbers*[1]
+%\usepackage{hyperref}
+\usepackage{url}
+
+\usepackage{lineno}
+\linenumbers*[1]
 
@@
 \titlerunninghead{WENCHUAN TOPOGRAPHIC STRESS}
-\authoraddr{Corresponding author: Richard H. Styron, Department of Earth and
-Environmental Sciences, University of Michigan, 2534 CC Little Bldg.,
-1100 N. University Ave., Ann Arbor, MI 48104, USA
+\authoraddr{Corresponding author: Richard H. Styron, Earth Analysis,
+	4044 Latona Ave NE, Seattle, WA 98105, USA
 	(richard.h.styron@gmail.com)}
@@
 
-\authors{Richard H. Styron \altaffilmark{1} and Eric A. Hetland \altaffilmark{1}}
+\authors{Richard H. Styron \altaffilmark{1, 2} and Eric A. Hetland \altaffilmark{1}}
 
@@
 Ann Arbor, Michigan, USA.}
+
+\altaffiltext{2}{Earth Analysis, Seattle, Washington, USA}
 
@@
 collinear with the inferred coseismic slip. We further estimate the static
-friction and pore fluid pressure assuming that the fault was at Mohr-Coulomb
-failure at the time of the Wenchuan earthquake. We use a Bayesian inversion
-strategy, yielding posterior probability distributions for the estimated
-parameters. We find most likely estimates of maximum tectonic compressive
-stress near 0.6 $\rho g z$ and oriented \textasciitilde{}E-W, and minimum
+friction and pore fluid pressure assuming that the fault was, on average, at
+Mohr-Coulomb failure at the time of the Wenchuan earthquake. We use a Bayesian
+inversion strategy, yielding posterior probability distributions for the
+estimated parameters. We find most likely estimates of maximum tectonic
+compressive stress near 0.6 $\rho g z$ and oriented $\sim$E-W, and minimum
 tectonic stress near 0.2 $\rho g z$. Static friction on the fault is near 0.2,
-and pore fluid pressure is between 0 and 0.4 of the total lithostatic pressure.
-\end{abstract}
+and pore fluid pressure is between 0 and 0.4 of the total lithostatic
+pressure.  \end{abstract}
 
@@
 characterized by significant along-strike variations in coseismic slip.
-Additionally, because the earthquake occurred after \textasciitilde{}2000 years
+Additionally, because the earthquake occurred after $\sim$2000 years
 of seismic quiescence, postseismic stresses in the lithosphere are likely to be
@@
 to bracket the tectonic stress field, pore fluid pressure, and static friction
-of the fault.
+of the fault, so that these parameters are consistent with the toopographic
+stresses and coseismic slip on the Wenchuan earthquake faults.
 
@@
 thrust due to loading of the high Himalayan massifs locally suppressed
-microseismicity and increased fault locking. Meade and Conrad \citet{meade2008}
+microseismicity and increased fault locking. \citet{meade2008}
 demonstrated that the increased weight of the uplifting Andes influenced the
@@
 San Andreas fault. In their study, they developed methods for calculating the
-three dimensional elastic stress tensor field arbitrary topography, whereas
+three dimensional elastic stress tensor field due to 
+arbitrary topography, whereas
 previous solutions were limited to two dimensions and required topography to be
@@
 Precambrian crystalline massif in the hanging wall of the Beichuan thrust) drop
-to \textasciitilde{}1200 m in as little as 6 km map distance.
-
-\begin{figure*}[ht!] \includegraphics[width=40pc]{../figures/lms_map.pdf}
-    \caption{Map of eastern Tibet and the Sichuan basin, showing active
-        structures from \citet{styron2010}. Faults that ruptured in the 2008
-        Wenchuan earthquake are shown in pink. GPS velocities are relative to
-        the mean velocity of sites within Sichuan basin, with 1$\sigma$
-        uncertainty, from the dataset of Liang et al.  \citet{liang2013}.
-        Beachball is from the Global CMT focal mechanism solution for the 2008
-    Wenchuan earthquake. BF = Beichuan fault. PF = Pengguan fault. P = Pengguan
-massif. Grey box shows the extent of Figures 3 and 9.} \label{fig:lms_map}
-\end{figure*}
+to $\sim$1200 m in as little as 6 km map distance.
+
+%\begin{figure*}[ht!] \includegraphics[width=40pc]{../figures/lms_map.pdf}
+%    \caption{Map of eastern Tibet and the Sichuan basin, showing active
+%        structures from \citet{styron2010}. Faults that ruptured in the 2008
+%        Wenchuan earthquake are shown in pink. GPS velocities are relative to
+%        the mean velocity of sites within Sichuan basin, with 1$\sigma$
+%        uncertainty, from the dataset of Liang et al.  \citet{liang2013}.
+%        Beachball is from the Global CMT focal mechanism solution for the 2008
+%    Wenchuan earthquake. BF = Beichuan fault. PF = Pengguan fault. P = Pengguan
+%massif. Grey box shows the extent of Figures 3 and 9.} \label{fig:lms_map}
+%\end{figure*}
 
@@
 complicated pattern of coseismic slip in the Wenchaun earthquake, with several
-high-slip patches that dominate the seismic moment release and significant
-variation coseismic slip rake along strike \citep[e.g.,][]{nakamura2010,
+high-slip patches that dominate the seismic moment release and substantial
+variation in fault geometry and coseismic slip rake along strike 
+\citep[e.g.,][]{nakamura2010,
 shen2009, tong2010, feng2010, zhang2011, qi2011, fielding2013}.  The variation
@@
 refer to as the ``Wenchuan earthquake faults'') with shallow to moderate dips
-largerly ruptured as thrust, and sections with steeper dips largely ruptured as
+largely ruptured as thrust, and sections with steeper dips largely ruptured as
 strike-slip.  \citet{medinaluna2013} concluded that this relationship is
@@
 
-\begin{figure*}[t!] \centering
-    \includegraphics[width=40pc]{../figures/topo_stress_possibilities.pdf}
-    \caption{Scenarios for topographic effects on rangefront thrust faulting.
-        (a) Topographic stresses promote thrust faulting. (b) Topographic
-        stresses inhibit thrust faulting.} \label{fig:topo_fault_scenarios}
-    \end{figure*}
+%\begin{figure*}[t!] \centering
+%    \includegraphics[width=40pc]{../figures/topo_stress_possibilities.pdf}
+%    \caption{Scenarios for topographic effects on rangefront thrust faulting.
+%        (a) Topographic stresses promote thrust faulting. (b) Topographic
+%        stresses inhibit thrust faulting.} \label{fig:topo_fault_scenarios}
+%    \end{figure*}
 
@@
 slip. In a scenario with complex faulting and topography, it may be possible to
-put tight bounds on minimum and maximum magnitudes and directions of tectonic
+put bounds on both minimum and maximum magnitudes, in addition to the 
+directions of tectonic
 stresses. To account for the non-uniqueness of the solution, we use a Bayesian
@@
 \citet{liuzoback1992}. They show that the topographic stress tensor
-field can be determined by a convolution of topographic loading
-functions with Green's functions describing the stresses in an elastic
-halfspace due to a point load at the surface. In our notation, the
-stress tensor field due to topography is given by
+field beneath (but not within) topography can be determined by a convolution of
+topographic loading functions with Green's functions describing the stresses in
+an elastic halfspace due to a point load at the surface. Note that in this
+work, we use capital letters (e.g., $M$, $T$) to denote tensors and tensor
+fields (depending on context, which should be clear) and Greek letters ($\tau$,
+$\sigma$) to denote stress components projected or resolved on planes, which
+are more properly called tractions. We use superscripts on these symbols to
+denote the origin of the stresses, and subscripts to denote components of these
+tensors or stresses or tractions.
+
+We denote the stress tensor field resulting from topography as $M(x,y,z)$. It
+is given by
 
@@
 \end{equation}
-
 where $G(x,y,z)$ is a set of Green's functions for the six stress tensor
@@
 \end{equation}
-
 where $G^B(x,y,z)$ are the Boussinesq solutions for stresses in a
@@
 \end{equation}
-
 where $G_i^C(x,y,z)$ are the Cerruti solutions for a horizontal point
@@
 \begin{split}
-F_{h, \, x}(x,y) = ( \rho g h(x,y) + M_{xx}^B(x,y,0) + T_{xx} )\, \frac{\partial h}{ \partial x} \\
-\quad + (M_{xy}^{B}(x,y,0) + T_{xy}) \frac{\partial h}{ \partial y}
+F_{h, \, x}(x,y) = ( \rho g h(x,y) + M_{xx}^B(x,y,0) + T^0_{xx} )\, \frac{\partial h}{ \partial x} \\
+\quad + (M_{xy}^{B}(x,y,0) + T^0_{xy}) \frac{\partial h}{ \partial y}
 \end{split}
@@
 \begin{split}
-F_{h, \, yz}(x,y) = ( \rho g h(x,y) + \sigma_{yy}^B(x,y,0) + T_{yy} )\, \frac{\partial h}{ \partial y} \\
-\quad + (M_{xy}^{B}(x,y,0) + T_{xy}) \frac{\partial h}{ \partial x}\; . 
+F_{h, \, y}(x,y) = ( \rho g h(x,y) + M_{yy}^B(x,y,0) + T^0_{yy} )\, \frac{\partial h}{ \partial y} \\
+\quad + (M_{xy}^{B}(x,y,0) + T^0_{xy}) \frac{\partial h}{ \partial x}\; . 
 \end{split}
@@
 \end{equation}
-
 $M_{ij}^B(x,y,0)$ is the stress from the vertical (Boussinesq) load
-evaluated at $z=0$, and $T_{ij}$ is the tectonic stress component, which
-we neglect in the present calculations.
+evaluated at $z=0$, and $T^0_{ij}$ is the tectonic stress component at the
+reference depth (the top of the model), which we assume in the present 
+calculations, as described below.
 
@@
 resolution to 851 m. We assume a Poisson ratio of 0.25, following
-receiver function studies suggesting values of ca. 0.24-0.26 throughout
+receiver function studies suggesting values of about 0.24--0.26 throughout
 central and western China \citep{chen2010receiverfn} (we have tested a Poisson
 ratio of 0.28, which is on the higher end of values for intermediate rock
-compositions \citep{zandt_ammons1995}, and found the results to vary by a few
+compositions \citep{zandtammon1995}, and found the results to vary by a few
 percent). Green's
 functions for the Boussinesq and Cerruti point-source solutions were
-calculated at regular points in large 2-D grid at each depth
+calculated at regular points in a large 2-D grid at each depth
 considered, with the point-source centered in the grid (see Table
@@
 project are available at
-\url{https://github.com/cossatot/wenchuan_topo_stress}.
+\url{https://github.com/cossatot/wenchuan\_topo\_stress}.
+
+%\begin{table}
+%\centering
+%\begin{tabular}{l c l}
+%\hline
+%Parameter & Value & Unit \\
+%\hline
+%horizontal spacing & 851 & m \\ 
+%vertical spacing & 1000 & m \\ 
+%minimum depth & 851 & m (below sea level) \\ 
+%maximum depth & 35851 & m (below sea level) \\ 
+%density ($\rho$) & 2700 & kg m$^{-3}$ \\ 
+%g & 9.81 & m s$^{-2}$ \\ 
+%Green's function radius & 9e5 & m \\ 
+%Poisson ratio & 0.25 & - \\ 
+%\hline
+%\end{tabular}
+%\caption{Parameters for numerical calculations of topographic stresses.}
+%\label{table:convo_params}
+%\end{table}
+
+\subsection{Topographic fault stress
+calculations}\label{topographic-fault-stress-calculations}
+
+Topographic stresses on the Wenchuan faults are calculated on point sets
+representing the faults taken from coseismic slip models. We use six models,
+those of \citet{shen2009}, \citet{feng2010}, \citet{zhang2011},
+\citet{fielding2013} and two from \citet{qi2011}.
+All of these models rely on geodetic data to some degree, but they do not all 
+use the same data in their inversion (e.g., \citep{feng2010} uses a different
+InSAR catalogue as the others). All use different inversion strategies, and
+different fault geometries (the two models of \citep{qi2011} share a common
+fault geometry but use different regularization in the inference of the slip
+distribution). By using a suite of models in our calculations, we can infer
+that results which are persistent in most or all of the models are more robust,
+while other results specific to only one of the coseismic slip models can be
+more confidently linked to specifics of the model geometry or inferred slip
+distribution in those models.
+
+In general, the fault geometries in all of the coseismic slip models are
+similar, as well as the inferred pattern of slip distribution above 10 km or
+so (i.e., the locations of high and low slip patches and the slip rake are
+similar in all of the models).  Although some of models use discrete, planar
+fault segments, whereas others use a single continuous, non-planar fault, the
+fault geometry models are essentially collocated. However, there is significant
+variability in the magnitude of slip in the different models; for example, the
+maximum slip magnitude in the Qi `rough' model is about twice that of the
+Feng model. Large differences also exist in the deeper geometries of the
+models: the Qi, Fielding and Shen models all have horizontal or subhorizontal
+thrust flats at or below 15 km, though only minor slip is assumed to have
+occurred on these fault segments. In contrast, the Feng and Zhang models are
+planar to their lower end at 25-30 km depth.
+
+In our stress calculations, we discard
+points above 851 m below sea level, as this is above the depth at which
+we compute $G^C(x,y,z)$. The six stress tensor components calculated at
+the regular grid points are linearly interpolated to points describing
+the faults. Because the fault points are completely surrounded by the
+grid nodes at which topographic stresses were calculated and those nodes
+are spaced \textless{}1 km apart, the fault points cannot be more than a
+few hundred meters from the nearest grid node, so a higher order
+interpolation is not necessary. We then project the topographic stress
+tensor to fault normal stress, $\sigma_n^M$, down-dip shear stress,
+$\tau_d^M$ and strike-slip shear stress, $\tau_s^M$ at each point in
+describing the fault geometry. In this work, the superscripts on stress
+symbols denote the origin of the stresses.
+
+\section{Results of topographic stress calculations on the Wenchuan
+faults}\label{results-of-topographic-stress-calculations-on-the-wenchuan-faults}
+
+Topographic stresses on the Wenchuan faults are on the order 1--10s MPa
+(Figure~\ref{fig:lms_topo_stresses_rot}, ~\ref{fig:fault_stress_3d}). 
+Stresses are highest in the southwest,
+beneath the Pengguan massif (the highest topography of the Longmen Shan
+front), and decrease to the northeast. $M_{zz}$ is typically larger,
+though not substantially, than $M_{xx}$ or $M_{yy}$. Maximum horizontal
+stress is not typically aligned with either cardinal horizontal
+direction, and is typically larger than $M_{zz}$ above 10 km. Maximum
+$M_{zz}$ is near 80 MPa, on the southwestern Beichuan fault below the
+high peaks of the Pengguan massif, except for in slip models containing
+near-horizontal fault segments in the mid-crust, where $M_{zz}$ reaches
+100 MPa. Vertical shear stresses ($M_{xz}$ and $M_{yz}$) are on the
+order of 1 MPa, and horizontal shear stress ($M_{xy}$) is on the order
+of 0.1 MPa. (Results not shown in figures are available as .csv files
+at \url{http://github.com/cossatot/wenchuan_topo_stress/}).
+
+%\begin{figure}[t]
+%\centering
+%\includegraphics[width=20pc]{../figures/lms_topo_stresses_rot.pdf}
+%\caption{Horizontal topographic stresses in the Longmen Shan region at 5
+%km depth: black and red lines signify most and least compressive
+%horizontal stresses, respectively. Other symbols are the same as in Figure
+%1. Stresses shown are down-sampled from the
+%discretization used in the calculations by a factor of six.}
+%\label{fig:lms_topo_stresses_rot}
+%\end{figure}
+
+Because the compressive stresses are near equal, $M$ contains a large
+isotropic component and a smaller deviatoric component. Consequently, $M$
+resolves on the Wenchuan faults with a large $\sigma^M_n$ (median of
+about 40-60 MPa for each slip model) and much smaller $\tau^M_d$. The
+median $\tau^M_d$ is about -3 to -6 MPa in each slip model, where values
+less than zero indicate normal-sense shear, and $\tau^M_s$ median values
+range from about -2 to 1 MPa, where values less than zero indicate
+sinistral shear; fault models with positive median $\tau^M_s$ have broad
+thrust flats at depth, where little slip occurred. On the steeper fault
+segments (where most of the moment release occurred), topographic shear
+stresses are typically normal-sinistral, as opposed to the dominant mode
+of coseismic slip, which is reverse-dextral.
+
+%\begin{figure*}[htbp]
+%\centering
+%\includegraphics[width=40pc]{../figures/fault_stress_3d.pdf}
+%\caption{Southwest-looking views of topographic stresses and coseismic
+%slip on selected slip models. All views share same lateral extent, but
+%the perspectives for B and C are more inclined than A. (A) $\sigma_n^M$
+%(colors), slip magnitude (contours, 1 m interval) and hanging-wall
+%topography on the \citet{feng2010} model of the Beichuan
+%fault. Note the suppression of fault slip where normal stress is
+%highest, such as below the Pengguan massif (P). Fault and topography
+%share the same scale, with no vertical exaggeration. (B) $\tau_d^M$ (colors)
+%and dip slip (contours, 1 m interval) \citet{qi2011}
+%`rough' slip model of the Beichuan fault. (C) $\tau_s^M$ (colors) and strike
+%slip (contours, 1
+%m interval) on the \citet{qi2011} `rough' slip model if the Beichuan fault.}
+%\label{fig:fault_stress_3d}
+%\end{figure*}
+
+In contrast to the steeper fault segments, much of the shallowly-dipping
+fault segments (the Pengguan fault and flats at the base of the Beichuan
+fault, where present) have $\tau^M$ in the direction of coseismic slip
+(Figure~\ref{fig:fault_stress_3d}b,c).
+$M$ is not significantly different in these
+locations, but because of the low dip angle, $M_{zz}$ contributes more
+significantly to $\sigma^M_n$ than to $\tau^M_d$, which is then
+dominated by horizontal compression, leading to reverse-sense shear.
+The stresses caused by the Pengguan massif locally resolve
+as right-lateral on these segments as well. Coseismic slip on these
+fault patches is much lower than on the steeper Beichuan fault, where
+the majority of slip occurred and which is topographically loaded in the
+opposite shear sense.
+
+Compellingly, similar patterns exist in the spatial distributions of
+$\sigma^M_n$ and coseismic slip. Most obvious is the coincidence of
+locally high $\sigma^M_n$ and locally low slip magnitude on the
+southwestern Beichuan fault below the culmination of the Pengguan
+massif, in an area of otherwise high slip (Figure
+\ref{fig:fault_stress_3d}). These correlations exist for other fault
+patches, but they are not as clear (Figure
+\ref{fig:feng_slip_sig_n_scatter}). This raises the possibility that
+topographic loading of these faults contributes to limiting coseismic
+slip once failure has occurred, and
+may have implications for estimations of dynamic friction and the
+completeness of stress drop during the earthquake. Preliminary analysis
+of this is currently being performed, and will be described in a
+forthcoming manuscript.
+
+%\begin{figure}%[htbp]
+%\centering
+%\includegraphics[width=20pc]{../figures/feng_slip_sig_n_scatter.pdf}
+%\caption{Coseismic slip magnitude and $\sigma^M_n$ on the four segments
+%of the \citet{feng2010} coseismic slip model. Trendlines are
+%L1 regressions and do not include points with no slip. ``NE B'' =
+%Northeastern Beichuan fault. ``C B'' = central Beichuan fault. ``SW B'' =
+%Southwestern Beichuan fault. ``P'' = Pengguan fault.}
+%\label{fig:feng_slip_sig_n_scatter}
+%\end{figure}
+
+\section{Calculations of tectonic stress, fault friction and pore fluid
+pressure}\label{calculations-of-tectonic-stress-fault-friction-and-pore-fluid-pressure}
+
+Faults fail in earthquakes when the shear stresses on the fault overcome
+the frictional stresses resisting slip on the fault \citep[e.g.,][]
+{scholz2002}. We assume that the entire fault was at at the point of
+failure when the Wenchuan earthquake initiated, and use the Mohr-Coulomb
+failure criterion
+
+\begin{equation} 
+\tau = \mu ( \sigma_n - \sigma_p ) \; ,
+\label{eqn:amonton_raw} 
+\end{equation}
+where $\mu$ is the coefficient of static friction on the fault and
+$\sigma_p$ is the pore fluid pressure \citep[e.g.,][]{sibson1985}, and
+assuming that rock strength (cohesion) is negligible. We
+describe the pore fluid pressure using a scalar, $0 \leq \phi \leq 1$,
+which is the pore fluid pressure as a fraction of total pressure, and so
+the failure criterion is
+
+\begin{equation} 
+\tau = \mu (1 - \phi) \sigma_n \; 
+\label{eqn:amonton} 
+\end{equation}
+\citep[e.g.,][]{sibson1985}. We assume that both $\mu$ and $\phi$ are
+constant across the Wenchuan faults.
+
+We estimate the tectonic stress tensor field, $\mu$, and $\phi$
+consistent with published coseismic slip models of the Wenchuan
+earthquake using a Bayesian estimation, resulting in posterior probability 
+density functions of the model parameters. 
+We first estimate posteriors of the tectonic stress tensor field 
+$T$ consistent with
+the coseismic slip models, and then estimate $\mu$ and $\phi$ consistent
+with Mohr-Coulomb failure. The nature of Bayesian estimation allows us
+to quantify both the relative likelihoods of model parameters and the
+tradeoffs between them.
+
+\subsection{Description of the stress
+state}\label{description-of-the-stress-state}
+
+We consider the complete stress tensor, $S$, at a point in the crust to
+be
+
+\begin{equation}
+S = M + T + L \; ,
+\end{equation}
+where $M$ is described above, $T$ is the tectonic stress tensor, and $L$
+is the lithostatic stress tensor. $L$ is isotropic, with diagonal
+components equal to $\rho g z$. We assume that $T$ is laterally
+homogeneous and only has horizontal stress components (i.e.,
+$T_{xz} = T_{yz} = T_{zz} = 0$, with $T_{xx}$, $T_{yy}$ and $T_{xy}$
+non-zero). The expanded stress tensor is:
+
+\begin{equation}
+S = \begin{bmatrix}
+  M_{xx} + T_{xx} + L_{xx} & M_{xy} + T_{xy} &  M_{xz} \\
+	M_{xy} + T_{xy} &  M_{yy} + T_{yy} + L_{yy} & M_{yz} \\
+	M_{xz}     &  M_{yz}  &  M_{zz} + L_{zz}
+	\end{bmatrix} \; .
+\label{eqn:stress_tensor}
+\end{equation}
+
+We further assume that $T$ increases linearly with depth so
+that the entire upper crust is near the critical failure envelope
+at an unspecified coefficient of friction
+\citep[e.g.,][]{townend2000}, and thus we parameterize the components of $T$
+as scalars multiplied by $\rho g z$, denoted as $T^\prime$. Therefore, if
+$T^\prime_{xx} = 0.1$, at some point just below 1 km, $L=27$ MPa, so
+$S_{xx} = M_{xx}$+ 27 MPa + 2.7 MPa.
+
+\subsection{Bayesian inversion of tectonic
+stresses}\label{bayesian-inversion-of-tectonic-stresses}
+
+We invert topographic stresses and coseismic slip models for tectonics
+stresses using Bayesian methods, and making the common `Wallace-Bott'
+assumption (named after \citet{wallace1951} and \citet{bott1959}) 
+that slip on the fault occurs in the general direction
+of the maximum resolved shear stress on the fault \citep[e.g.,][]
+{mckenzie1969, angelier1994}. We estimate the tectonic stresses in
+light of the topographic stresses and slip distributions through the
+relation
+
+\begin{equation} 
+p(T|D) \propto p(T) \, p(D|T) \; , 
+\label{eqn:bayes_rule} 
+\end{equation}
+where $p(T)$ is the prior PDF (or \emph{prior}) of $T$, $p(D|T)$ is the
+likelihood of observing the coseismic slip distribution $D$ given the
+tectonic stresses $T$, and $p(T|D)$ is the posterior PDF of $T$ given
+$D$, which is the solution to the inversion \citep[e.g.,][]{mosegaard1995}.
+Due to the unknown proportionality in equation (\ref{eqn:bayes_rule}),
+our posterior only gives likelihood of $T$ relative to the most likely
+estimate (MLE) \citep{tarantola2005}. We follow a Monte Carlo strategy,
+where samples of the prior PDF are retained as samples of the posterior
+in proportion to $p(D|T)$ \citep[e.g.,][]{mosegaard1995}.
+
+We parameterize $T$ by the magnitudes and orientation of the maximum and
+minimum principal tectonic stresses. We assume priors such that the
+magnitudes of principal tectonic stresses are equally likely within
+bounds and that all stress orientations are equally likely. Because the
+Wenchuan event was an oblique reverse faulting earthquake, we assume
+that total horizontal stresses are greater than the vertical stress,
+which is satisfied if the tectonic stresses are positive. Prior samples
+of maximum principal tectonic stress are taken from a uniform distribution
+between $\rho g z$ and 2.5 $\rho g z$. Samples of the minimum principal
+tectonic stress are from a uniform distribution between 0 and the value for
+maximum stress. We describe the orientation of the tectonic stress using
+the azimuth of the maximum tectonic stress, which are sampled uniformly
+from 0 to 360$^{\circ}$.
+
+We test 100,000 unique samples drawn from the prior using a seeded
+pseudorandom number generator. We test the same prior samples against
+each of the coseismic slip models. $S$ is then constructed for each
+point discretizing the fault geometries in the coseismic slip models.
+The rake of the maximum shear stress $\lambda^S$ on each point of the
+fault is calculated and compared to the coseismic slip rake $\lambda^D$
+at that point. A weighted mean misfit is calculated by
+
+\begin{equation}
+\bar{\lambda}^m = \sum \nolimits_{i=1}^n \frac{(\lambda^S_i - \lambda^D_i) D_i} 
+{\bar{ D}} \; ,
+\label{eqn:rake_misfit}
+\end{equation}
+where $D$ is the coseismic slip and $\bar{D}$ is the average coseismic
+slip in a given coseismic slip model. Finally, the relative likelihood
+of each model is computed using a Von Mises distribution as
+
+\begin{equation}
+p (D | T) = \frac{ \exp ( \kappa \cos \bar{\lambda}^m )} 
+{\exp (\kappa \cos \bar{\lambda}^m_{\min})} \;,
+\label{eqn:rel_likelihood}
+\end{equation}
+where $\kappa$ = 8.529, which is calculated so that the 68.2\%
+confidence interval of the Von Mises distribution is within $\pi$/9 radians
+(20$^{\circ}$), the estimated 1$\sigma$ uncertainty of the coseismic
+slip models based on comparisons between rakes of high-slip fault
+patches (note that for a planar fault, $\tau$ at $\pi/9$ radians from
+$\lambda_{max}$ is still \textgreater{}90\% of $\tau_{\mathrm{\max}}$
+\citep{lisle2013}). Prior samples are retained in propotion to $p(D|T)$,
+and the retained samples are then samples of the posterior.
+
+\subsection{Analysis of friction and pore fluid
+pressure}\label{analysis-of-friction-and-pore-fluid-pressure}
+
+Once the tectonic stress distributions consistent with the coseismic
+slip models have been determined, we deduce the distributions of $\mu$
+and $\phi$ assuming that the stress is at the failure criterion in
+equation (\ref{eqn:amonton}). We do this in three steps: First, we draw
+a random $\phi$ from a uniform distribution, assuming $0 \leq \phi < 1$.
+We again use a seeded pseudorandom number generator, such that each
+stress model has a uniquely assigned $\phi$ that is consistent 
+across all coseismic slip models. Second, we calculate $\tau^S$ and
+$\sigma_n^S(1-\phi)$ for each point on the fault. Third, we solve
+Equation \ref{eqn:amonton} for $\mu$. Finally, we filter the results so
+that only models with $0 \le \mu < 1$ are retained, as values outside of
+that range have not been suggested for rocks.
+
+After this analysis has been done for all coseismic slip models, we find
+the joint posterior (i.e., the posterior consistent with all of the
+coseismic slip models) by taking the samples that are common to all of
+the individual posteriors. We denote the joint posterior as
+$p_{\mathrm{J}}(P | D)$, where $P$ is the parameter of interest.
+
+\section{T, $\mu$, $\phi$ results}\label{t-mu-phi-results}
+
+\subsection{Individual slip models}\label{individual-slip-models}
+
+Results for $T^\prime$, $\mu$ and $\phi$ are quite consistent across all
+coseismic slip models (Figure \ref{fig:T_scatters}). Maximum compressive
+tectonic stress $T^\prime_{\mathrm{max}}$ is broadly east-west for all models,
+with a mode trending at $90^{\circ}$--$105^{\circ}$.
+$p(T^\prime_{\mathrm{max}}|D)$ for each slip model increases from
+$T^\prime_{\mathrm{max}} = 0$ to 0.5 or 1 before essentially leveling off, though some
+slip models, particularly the \citet{qi2011} model, show a
+slight decrease in relative likelihood past the initial mode at
+$T^\prime_{\mathrm{max}} =$ 0.5--1.
+The low likelihood below $T^\prime_{\mathrm{max}} \approx 0.5$ indicates that lower
+tectonic stresses are unlikely to overcome fault friction and
+topographic shear stresses resisting reverse-dextral slip on the
+Wenchuan faults. $p(T^\prime_{\mathrm{min}} | D)$ for each slip model
+has a mode close to $T^\prime_{\mathrm{min}} = 0.2 $ and decreases
+abruptly at higher values, though all slip models show values for
+$T^\prime_{\mathrm{min}}$ up to 2.5. $T^\prime_{\mathrm{min}}$ is
+typically 0--0.4 of $T^\prime_{\mathrm{max}}$, but rarely higher.
+
+%\begin{figure*}%[htbp]
+%\centering
+%\includegraphics[width=40pc]{../figures/T_scatters.pdf}
+%\caption{Samples of
+%$p(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$ associated with
+%each of the coseismic slip models we consider, along with marginals of
+%$T^\prime_{\mathrm{max}}$ and $T^\prime_{\mathrm{min}}$. Inset rose
+%diagrams are histograms of azimuth of
+%$T^\prime_{max}$.}
+%\label{fig:T_scatters}
+%\end{figure*}
+
+All slip models show $p(\phi | D)$ to be uniformly high from $\phi = 0$
+to 0.4--0.6 and to decrease somewhat linearly to $p(\phi) = 0$ at $\phi = 1$
+(Figure \ref{fig:mu_phi_scatters}). $p(\mu | D)$ for all slip models has
+a mode at $\mu =$ 0.1--0.4 and $p(\mu)$ decreases at higher values.
+$T^\prime_{\mathrm{max}}$, $\phi$ and $\mu$ are highly correlated, where
+higher values of $T^\prime_{\mathrm{max}}$ are associated with higher
+$\mu$ and lower $\phi$. Combinations of high $\mu$ and low $\phi$
+require much higher $T^\prime_{\mathrm{max}}$ to overcome fault friction
+and cause slip, and so are not represented in the posteriors. Since our
+maximum $T_{\mathrm{max}}$ of $2.5 \rho g z$ is quite high
+($\approx 660$ MPa at 10 km), we view high $\mu$ and low $\phi$
+combinations as unrealistic for the Wenchuan faults. Similarly,
+combinations of very low $\mu$ and very high $\phi$ are associated with
+very low $T^\prime_{\mathrm{max}}$, and have a low probability density,
+as it is unlikely that tectonic stress with very low $T^\prime_{\mathrm{max}}$
+values can
+overcome sinistral and normal-sense topographic shear stresses to cause
+the observed coseismic slip kinematics.
+
+%\begin{figure*}[htbp]
+%\centering
+%\includegraphics[width=40pc]{../figures/mu_phi_fms.pdf}
+%\caption{Samples of $p(\mu,\phi | D)$ for each coseismic slip model.
+%Colors indicate magnitude of $T^\prime_{\mathrm{max}}$. Contour lines
+%indicate relative density (i.e., likelihood) of posteriors (darker lines
+%signify higher densities).}
+%\label{fig:mu_phi_scatters}
+%\end{figure*}
+
+\subsection{Joint posteriors}\label{joint-posteriors}
+
+We define a joint posterior, $p_{J}(T^\prime | D)$, by the samples that 
+are common to the individual posteriors estimated from each slip model.
+Unsurprisingly, given the broad similarity between the posteriors from
+the various slip models, $p_{J}(T^\prime_{\mathrm{max}} | D)$ is not
+substantially different from any of the constituent model posteriors.
+$p_{J}(T^\prime_{\mathrm{max}} | D)$ has a somewhat more well-defined
+mode at $T^\prime_{\mathrm{max}} \approx$0.6--0.8. It is also apparent in
+Figure \ref{fig:joint_posteriors}~b that regardless of the magnitude of
+$T^\prime_{\mathrm{max}}$, $T^\prime_{\mathrm{min}}$ is
+consistently 0--0.6 $T^\prime_{\mathrm{max}}$, with a mode
+of about 0.3; this indicates that the relative magnitude of the
+tectonic stresses has a substantial influence on the rake of the maximum
+shear stress resolved on the fault.
+
+In our estimation of $\phi$ and $\mu$ in the posteriors associated with
+each slip model, we have used the same random combinations of $T$ and
+$\phi$ for each slip model, and then solved for $\mu$ so that the fault
+is at a critical stress state (Equation \ref{eqn:amonton}). Because of
+differences in the location and slip among the coseismic slip models,
+some variability exists in $\mu$ for each prior sample. We therefore
+choose $p_{\mathrm{J}}(\mu | D)$ to be the median $\mu$ of each slip
+model for each sample. $p_{\mathrm{J}}(\mu | D)$ has a mostly similar
+distribution as $p(\mu | D)$ for any of the slip models. However,
+$p_{\mathrm{J}}(\mu | D)$ has a lower relative likelihood on the
+high-$\mu$ tail (Figure~\ref{fig:joint_posteriors}d) compared to
+$p(\mu | D)$ of the constituent slip model results (Figure~\ref{fig:mu_phi_scatters}). This lower
+likelihood of $\mu$ in the joint posterior is probably because it is the
+average $\mu$ of all slip models. On the other hand, it is not similarly
+sparse on the low-$\mu$ side, suggesting that low values for $\mu$ are
+more robust.
+
+%\begin{figure*}[t]
+%\centering
+%\includegraphics[width=40pc]{../figures/joint_pdfs.pdf}
+%\caption{(a) Samples of
+%$p_{\mathrm{J}}(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$,
+%along with marginals of $T^\prime_{\mathrm{max}}$ and
+%$T^\prime_{\mathrm{min}}$. (b) Samples of
+%$p_{\mathrm{J}}(T^\prime_{\mathrm{\max}}, T^\prime_{\mathrm{\min}}/T^\prime_{\mathrm{max}} | D)$,
+%along with marginals of $T^\prime_{\mathrm{max}}$ and
+%$T^\prime_{\mathrm{min}} / T^\prime_{\mathrm{max}}$. (c) Histogram of
+%azimuths of $T^\prime_{\mathrm{\max}}$. (d) Samples of
+%$p_{\mathrm{J}}(\mu, \phi | D)$, with marginals, where color of the samples
+%indicate magnitudes of $T^\prime_{\mathrm{\max}}$ and contour
+%lines indicate relative density (i.e., likelihood) of posteriors (darker
+%lines signify higher densities).}
+%\label{fig:joint_posteriors}
+%\end{figure*}
+
+\section{Discussion}\label{discussion}
+
+Few studies have performed similar quantification of static stress
+fields on faults (see Section
+\ref{previous-work-on-topographic-stresses} for some examples), even
+though it may have important ramifications to the earthquake process.
+Most studies of fault rupture dynamics
+assume either a homogeneous or stochastic shear stress distribution
+\citep[e.g.,][]{oglesbyday2002} and few assume any variation in normal
+stress \citep[e.g.,][]{aagaard2001}, despite the importance that stress
+variations likely have in earthquake dynamics \citep[e.g.,][]{day1982,
+olsen1997}. Additionally, quantifying friction and pore fluid
+pressure involved in faulting is a major challenge in studies of
+faulting and orogenic dynamics \citep[e.g.,][]{meissner1982,
+oglesbyday2002}.
+
+Previous workers have demonstrated that by quantifying topographic
+stress, other components in the Coulomb stress balance may be bracketed
+\citep[e.g.,][]{cattin1997, lamb2006, luttrell2011}. Each of these studies
+uses somewhat different approaches. Our approach is most similar to that
+of \citet{luttrell2011}, although there are significant
+differences: (1) We use the topographic stresss calculations proposed by
+\citet{liuzoback1992}, whereas 
+\citet{luttrell2011} only uses the vertical loading from topography,
+equivalent to $M^B$ in equation (\ref{eqn:bous}). (2) We do not consider
+bouyancy forces due to lateral variations in density, due for instance to
+Moho variation, as done by \citet{luttrell2011}. In the
+Longmen Shan region, Moho variation is small compared to the change in
+Moho depths at the Andean plate boundary, and so the buoyancy terms
+should be relatively small. (3) We consider a full range of tectonic
+stresses, instead of simply calculating the minimum principal tectonic
+stress and its orientation. (4) We consider the stress tensor at each
+point due to topographic loading, lithostatic stress, and horizontal
+tectonic stress, and use inferred coseismic slip rake as a constraint on
+the allowable stresses, rather than inferring the earthquake stress drop
+from the coseismic slip models, as done by 
+\citet{luttrell2011}. (5) We use both normal and shear stresses to
+constrain pore fluid pressure and friction. (6) We use a Bayesian
+estimation, resulting in a PDF of tectonic stress, as well as $\mu$ and
+$\phi$, rather than just solving for the minimum tectonic stress
+required for faulting, as done by \citet{luttrell2011}.
+
+\subsection{Topographic stresses on the Wenchuan
+faults}\label{topographic-stresses-on-the-wenchuan-faults}
+
+Topographic stresses on the main Wenchuan faults are of considerable
+magnitude: $\tau^M$ ranges from about -20 to 10 MPa, which is on the
+order of inferred stress drop in earthquakes \citep[e.g.,][]{kanamori1975,
+allmann2009}. Topographic stresses are
+generally opposed to the tectonic slip direction, and therefore have to
+be overcome by tectonic stresses in order to produce the observed
+rupture patterns. The topographic shear stresses are likely persistent
+over the lifespan of the topography (i.e.~on the order of millions to
+tens of millions of years), otherwise the Wenchaun faults may fail in a
+normal sense simply due to the weight of the Longmen Shan. This would
+argue that tectonic stress drop in the Wenchuan earthquake was not
+complete, as some residual tectonic shear stress must remain on the
+fault to cancel out $\tau^M$.
+
+The spatial variation of topographic stresses increases in wavelength and
+decreases in magnitude with depth. This is not surprising, because with
+increasing
+depth, the stress field at any point is more sensitive to surface loads
+averaged over a greater region, and is less dominated by smaller scale
+topographic features (i.e., individual mountains). The spatial
+variablity of the topographic stress with depth is similar to the spatial
+variablity of coseismic slip in the models considered \citep[e.g.,][]
+{zhang2011}, which are both smoother at depth. Some of the
+estimated slip variability is likely partially due to the more limited
+resolution of coseismic slip at depth using geodetic data.
+However, the negative spatial correlations of slip versus stress
+(especially $\sigma^M_n$) (Figures \ref{fig:fault_stress_3d},
+\ref{fig:feng_slip_sig_n_scatter}) suggest the relationship between
+stress variation, slip variation and depth may be a real signal.
+
+\subsection{Tectonic stresses in eastern
+Tibet}\label{tectonic-stresses-in-eastern-tibet}
+
+The maximum tectonic stress, $T_{\mathrm{max}}$, is consistently
+oriented roughly E-W in our results. This orientation is oblique to the
+Longmen Shan, which produces oblique (right-lateral and reverse sense)
+shear on the Beichuan fault. $T_{\mathrm{min}}$ is $\sim$N-S
+oriented, and is a small fraction of lithostatic pressure. This
+stress configuration is in close agreement with pre-earthquake stress
+orientation
+measurements near the rupture zone (Figure \ref{fig:lms_stress_map}),
+mostly from borehole breakout data from 2-5 km depth
+\citep{heidbach2009}, which is the zone of maximum slip in the coseismic
+slip models. It is somewhat discrepant with stress orientations
+estimated at $\sim$800 m depth adjacent to the Beichuan fault
+in the WFSD-1 borehole of the Wenchuan Earthquake Fault Scientific
+Drilling Project several years after the 2008 earthquake \citep{cui2014},
+which show $\sigma_{H{\mathrm{\max}}}$ to be more orthogonal to the fault
+trace, suggesting that much of the right-lateral component of shear
+stress was released during the earthquake. These results are similar to
+the orientations of total stress obtained by 
+\citet{medinaluna2013}, although they were unable to constrain the
+magnitudes of stresses.
+
+%\begin{figure}[t]
+%\centering
+%\includegraphics[width=20pc]{../figures/lms_map_stresses_rot.pdf}
+%\caption{Topographic and tectonic horizontal stresses (taken from the
+%most likely estimates of $p(T|D)$ in the Wenchuan rupture region (black
+%and red crosses) with horizontal maximum stress orientation data taken
+%from before the 2008 Wenchuan event from the World Stress Map
+%\citep{heidbach2009} (purple arrows), and horizontal maximum stress
+%orientation data from after the earthquake at the WFSD-1 drill hole
+%\citep{cui2014} (blue arrows). Other symbols are as in Figure
+%\ref{fig:lms_map}. Stresses shown are downsampled from our computational 
+%grid resolution by a factor of nine.}
+%\label{fig:lms_stress_map}
+%\end{figure}
+
+The magnitudes of $T_{\mathrm{max}}$ and $T_{\mathrm{min}}$ are
+dominantly constrained on the low end by our analysis, which is apparent
+by the sharp decrease in the frequency of $p(T_{\mathrm{max}}|D)$ below
+about $T^\prime_{\mathrm{max}} = 0.5$ (Figure~\ref{fig:joint_posteriors}). 
+These results indicates that $T_{\mathrm{max}}$ of at
+least $\sim$13.25 MPa km$^{-1}$ is necessary to overcome
+topographic stresses resisting reverse and right-lateral slip on the
+faults. We find that the highest likely ratio of strike-slip to dip-slip
+shear along the Wenchuan earthquake faults is close to 1. This is
+similar to the inferences of strain accumulation rates inferred from
+squishy block modeling by \citet{loveless2011}, who
+concluded that the rate of slip deficit accumulation in the thrust and
+dextral senses were approximately equal. It should be noted that the
+tectonic stresses we estimate here represent the accumulated stresses
+prior to the Wenchuan earthquake, and the relation of these stresses to any
+accumulated slip deficit needs to be through a model of strain
+accumulation.
+
+The orientation of both the tectonic and total stresses near the
+Wenchuan faults shows a larger difference with patterns of strain from
+elsewhere in the orogen than in the Longmen Shan region. For example, the
+presence of N-S contraction and E-W extension throughout the high Tibetan
+plateau and much of the Himalaya \citep[e.g.,][]{armijo1986, molnar1988,
+taylor2003} indicates a roughly N-S $T_{\mathrm{max}}$ and E-W
+$T_{\mathrm{min}}$. Because $ L + T_{\mathrm{min}}$ is only slightly above
+lithostatic pressure on the Wenchuan faults, it is quite possible that the N-S
+compression in the Himalaya and Tibet, which is almost certainly due to
+Indo-Asian plate collision, has significantly decayed at the Longmen Shan, some
+850 km northeast of the easternmost Himalaya. Therefore, contraction across the
+Longmen Shan cannot easily be interpreted to directly reflect stresses due to
+the Indo-Asian collision alone, unless some additional mechanism of redirecting
+crustal stresses is incorporated.
+
+\subsection{Slip on the Beichuan fault vs.~optimally-oriented
+faults}\label{slip-on-the-beichuan-fault-vs.optimally-oriented-faults}
+
+Our highest likelihood estimates of $\mu$ are in the range of 0.2--0.3.
+These values are slightly lower than $\mu \approx 0.4$ inferred in
+laboratory experiments on samples recovered from the WFSD-1 drill hole
+into the Beichuan fault \citep{kuo2014}. However, it should be noted that
+$\mu \approx 0.4$ has a relatively high likelihood in our posteriors
+(Figure \ref{fig:joint_posteriors}). These values of friction 
+are lower than typical values derived from laboratory experiments on
+intact rock \citep[e.g.,][]{byerlee1978}, suggesting that slip occurred on
+preexisting faults because they are weaker, rather than on optimally oriented
+new faults.
+
+The obliquity of slip on the Wenchuan earthquake faults also suggests
+that these faults may not be optimally oriented for slip given the total
+stress state in the region of these faults. However, the Longmen Shan
+fault zone dates back to the Indosinian orogeny, locally late Triassic
+(226-206 Ma) \citep{yong2003} and has had multiple episodes of
+reactivation since \citep[e.g.,][]{burchfiel1995, wang2012}, accumulating
+tens of kilometers of shortening \citep[e.g.,][]{hubbard2010}. Such a mature
+fault system may be expected to have low coefficients of friction due to
+processes such as gouge development \citep[e.g.,][]{kuo2014}, and therefore
+may slip in non-optimal orientations, with high $\phi$ values also
+potentially contributing to this \citep[e.g.,][]{sibson1985}.
+
+Our posterior estimates for $T$, $\phi$ and $\mu$ let us quantitatively
+evaluate to what extent slip on the Wenchuan faults is more favorable
+than on optimally oriented faults with more typical friction
+coefficients. We use the same failure conditions as in determining the
+posteriors above, and assume optimally oriented faults exist
+(i.e., we do not consider the generation of new faults in relatively 
+intact rock).
+Additionally, we evaluate the relative contributions of $T$, $\phi$ and
+$\mu$ on potential fault weakening and reactivation. To explore these
+relationships, we perform some preliminary analysis on a single fault
+model (from \citet{zhang2011}) using a subset of 1000
+samples drawn randomly from the joint posteriors. Given the similarity
+of the fault models and of the posteriors for each model, we do not
+expect that an analysis of all results on all fault models will yield
+different conclusions.
+
+First, we establish a metric with which to evaluate the favorability of
+slip on a given fault plane, which we call the Coulomb failure ratio, or
+CFR:
+
+\begin{equation}
+\mathrm{CFR} = \tau / \mu (1 - \phi) \sigma_n \; .
+\label{eqn:cfr}
+\end{equation}
+
+CFR indicates whether a fault should fail under a given stress state:
+CFR \textgreater{} 1 indicates failure, while a CFR \textless{} 1
+indicates fault stability. We then calculate the CFR on each point in
+the model of the Beichuan fault (594 points describe the fault model of
+\citet{zhang2011}) based on the full stress field $S$ at
+each point on the fault, for each of the 1000 samples of $T$, $\phi$ and
+$\mu$ drawn randomly from the posteriors. We call this CFR$_f$. Then,
+using the same $S$ and $\phi$, we calculate the CFR on an optimally
+oriented fault with $\mu=0.6$ and no cohesion, which we call CFR$_o$.
+The orientation of the optimally oriented fault is determined as being
+the angle $\beta$ away from $\sigma^S_1$, where $\beta = (\tan^{-1} \mu) / 2$
+and is in the $\sigma_1$--$\sigma_3$ plane \citep[e.g.,][]{sibson1985}.
+Note this $\mu$ is typical for crustal rocks with fault normal stresses
+above 200 MPa, but lower than $\mu = 0.85$ for smaller $\sigma_n$
+\citep{byerlee1978}, but may be appropriate for an immature crustal
+fault.
+
+Figure \ref{fig:cfr_ratios} shows
+$\ln (\mathrm{CFR}_o / \mathrm{CFR}_f)$ plotted against $\mu$, $\phi$
+and $T^\prime_{xx}$ on the Beichuan fault for all samples. Though
+considerable scatter exists, it is clear that in most instances, slip on
+the Beichuan fault is preferred over slip on an optimal fault. The
+exceptions are at high values of $\mu$, $\phi$, or $T^\prime$, where
+slip on an optimal plane is preferred. Because $T$, $\phi$ and $\mu$ can
+all affect fault reactivation \citep[e.g.,][]{sibson1985}, we compare the
+relative contributions of each with a simple multiple linear regression,
+using $T^\prime_{xx}$ normalized to [0,1) (the same range $\phi$ and
+$\mu$) as a proxy for $T$ ($T^\prime_{xx}$ is essentially
+$T^\prime_{\mathrm{max}}$ in most of the posteriors). The results are
+shown in Table \ref{table:cfr_regress}. It is clear that both $T^\prime_{xx}$
+and $\mu$ are strongly correlated with $\mathrm{CFR}_o/\mathrm{CFR}_f$, and $\phi$ is to a lesser degree; nonetheless, all
+significantly affect the relative ease of faulting on the Wenchuan faults
+versus optimal faults. In particular, lower values for any of them favor slip
+on the Wenchuan faults. The lowest $\mathrm{CFR}_o/\mathrm{CFR}_f$ value
+($\sim$0.057, $\sim$ -2.9 in log space) corresponds to the lowest value of
+$\mu$ ($\sim$0.038), and is approximately the ratio of $\mu$ in the model
+to 0.6.
+
+%\begin{figure}%[tb]
+%\centering
+%\includegraphics{../figures/cfr_plots.pdf}
+%\caption{Comparison of Coulomb failure ratio (CFR) on the Beichuan fault
+%from the \citet{zhang2011} coseismic slip model to CFR on an
+%optimally-oriented fault with $\mu = 0.6$, versus estimated $\mu$ on the
+%Beichuan fault. Values less than 0 (1 in linear space) indicate that
+%slip is favored on the Beichuan fault, even if it is not optimally
+%oriented. Values are calculated for each point in the slip model for
+%1000 randomly-drawn samples from $p(T,\mu,\phi |D)$.}
+%\label{fig:cfr_ratios}
+%\end{figure}
+%
+%\begin{table}[htb]
+%\centering
+%\begin{tabular}{l c c c c}
+%\hline
+%Parameter & Coef. & Std. Err. & t statistic & $P>|t|$ \\
+%\hline
+%Intercept & -0.2566 & 0.0020 & -128.8 & \textless{}0.0001 \\ 
+%$\mu$ & 1.1168 & 0.0086 & 130.6 & \textless{}0.0001 \\ 
+%$T^\prime_{xx}$ & 0.9138 & 0.0048 & 191.2 & \textless{}0.0001 \\ 
+%$\phi$ & 0.5762 & 0.0055 & 105.7 & \textless{}0.0001 \\ 
+%\hline
+%\end{tabular}
+%\caption{Sensitivity of CFR ratios to relevant stress state parameters:
+%Results of multivariate linear regression of
+%$\mathrm{CFR}_o/ \mathrm{CFR}_f)$ against $\mu$, $\phi$ and
+%$T^\prime_{xx}$.}
+%\label{table:cfr_regress}
+%\end{table}
+
+\subsection{The role of topography in orogenic development and strain
+localization}\label{the-role-of-topography-in-orogenic-development-and-strain-localization}
+
+Because the rise of broad elevated regions creates substantial stresses
+in the crust \citep[e.g.,][]{jeffreys1924}, these stresses have the
+potential to influence orogenesis. The result may be to change the deformation
+state in the interior of the orogen \citep[e.g.,][]{dewey1988, molnar1988}, the
+convergence rates of the plates surrounding the orogen \citep[e.g.][]
+{meade2008}, or the location and style of deformation at the orogen's
+margins \citep[e.g.,][]{beaumont2001, decelles2009}. These studies show that
+as an orogen rises, horizontal contraction via crustal thickening will
+occur at the thin margins of the orogen, and if elevations reach a
+threshold or tectonic compression decreases, extension will take place
+in the orogen's high interior, as is well described in Tibet
+\citep[e.g.,][]{armijo1986, taylor2003, styron2015}. Therefore,
+Tibet is commonly suggested to be undergoing some manner of
+gravitational collapse \citep[e.g.,][]{england1989}; where the plateau abuts
+rigid cratons, the deformational patterns resemble `fixed boundary'
+collapse \citep{rey2001}, with concentrated crustal thickening at the
+orogen's margins \citep[e.g.,][]{cook2008}.
+
+Though much work has been done exploring the feedback mechanisms between
+topography and deformation, topography is typically greatly smoothed,
+topographic stresses are folded into gravitational potential energy
+estimates of the entire lithosphere, and the whole lithospheric column
+is often considered viscous \citep[e.g.,][]{birdpiper80, copleymckenzie2007,
+flesch2010gpe}. Therefore many
+models seeking to predict deformation from gravitational forces yield
+continuous deformation and do not consider how the gravitational
+stresses resolve on heterogeneities embedded in the crust. However, the
+presence of structures such as weak faults \citep{bird1994} or
+low-viscosity channels or shear zones \citep[e.g.,][]{clark2005} can localize
+deformation \citep[e.g.,][]{bird1994, fleschbendick2012}.
+
+Our results that topographic stresses on the Wenchuan
+earthquake faults are largely loaded in the opposite sense to the
+coseismic slip direction indicates that gravitational collapse is probably
+not the driver of reverse faulting in the Longmen Shan. However, as
+noted previously, thrust flats in the middle crust present in some
+coseismic slip models, particularly the Qi model, are loaded in a thrust
+and dextral sense (Section 3). Given the very low dips of the thrust flats
+($0^{\circ}$--$5^{\circ}$), they receive very little loading from
+horizontal tectonic stress, so any slip on them may be in response to
+topographic loading and thus to gravitational collapse of the high
+Longmen Shan and its hinterland. $M_{H{\mathrm{\max}}}$ (the maximum horizontal
+compressive stress from topography) also changes
+orientation from fault rangefront normal to closer to more oblique
+near the front of the range (Figure \ref{fig:lms_topo_stresses_rot}),
+suggesting that any transfer of mass from the highlands to the lowlands
+due to topographic stresses should terminate near the rangefront.
+
+However, because $M_{zz}$ is greater than $M_{H{\mathrm{max}}}$
+underneath the higher topography, topographic stresses can only lead to
+reverse faulting on the Wenchuan faults if subhorizontal shear zones or
+channels are present, and are very weak so shear can occur at 
+low $\tau$ relative to $\sigma_n$. This conclusion has been reached in
+studies of generalized orogens \citep[e.g.,][]{fleschbendick2012} and
+gravitational collapse of volcanic edifices \citep[e.g.,][]{byrne2013},
+indicating that is a general requirement of gravitational spreading. Our
+calculations of $\mu$ and $\phi$ are based on $T$, which is biased
+towards the high slip patches in the upper crust (Equation
+\ref{eqn:rake_misfit}).  The resulting low values of $\mu$ and low
+to moderate values of $\phi$, probably yield stress conditions insufficient
+for slip at high pressures found at depths > 10 km.
+Similarly, a low viscosity horizontal
+channel that underlies eastern Tibet, likely much deeper than the
+seismogenic zone we consider here, would be able to flow in response to
+$\tau$ regardless of $\sigma_n$ and may be able to facilitate
+gravitational spreading of the orogen \citep[e.g.,][]{clark2005, cook2008,
+fleschbendick2012}.
+
+These arguments are all quite speculative, and we wish only to describe the
+conditions under which gravitational collapse can occur given the
+observations of deformation and the topographic stress field. It is not
+at all clear that a suitably large decollement exists at the base of the
+Beichuan fault (it is only present in one of the slip models
+considered). Nonetheless, this topic has implications not only for
+orogenic development, but for the recurrence interval on the Wenchuan
+earthquake faults. If topographic stress contributes in some fashion to
+the observed displacements on the fault, then those stresses were likely
+barely diminished by coseismic stress change, and earthquake recurrence
+of on the Wenchuan faults may be governed by different processes than
+elastic rebound due to tectonic strain accumulation.
+
+\section{Conclusions}\label{conclusions}
+
+We have calculated shear and normal stresses due to topographic loading
+on the Wenchuan earthquake faults, and used those stresses to constrain
+tectonic stresses, fault friction and pore fluid pressure. Topographic
+stresses on the main Wenchuan faults are large, with $\tau^M$ on the
+faults up to \textbar{}20\textbar{} MPa, and $\sigma^M_n$ up to 80 MPa.
+$\sigma^M_n$ reaches up to 100 MPa on mid-crustal thrust flats present
+in some coseismic slip models. The direction of $\tau_M$ is generally
+opposed to coseismic slip inferred during the 2008 Wenchuan earthquake,
+indicating that weight of the topography resists coseismic slip. High
+values of $\sigma^M_n$ increase the frictional resistance to slip,
+potentially limiting slip magnitude in locations such as below the
+Pengguan massif.
+
+Assuming that the Beichuan faults were at a Mohr-Coulomb fault criterion
+immediately prior to the Wenchuan earthquake, we estimate the tectonic
+stresses required for the faults to fail. We use a Bayesian estimation,
+resulting in posteriors representing likelihood of tectonic stress,
+static friction, and a pore pressure parameter. The posteriors
+indicate that the maximum tectonic stress is oriented
+$\sim$E-W and has a likely minimum of about 10 MPa per kilometer of depth
+(i.e. $T^\prime_{\mathrm{max}}$ of at least 0.4).
+The minimum tectonic stress is oriented $\sim$N-S and is
+fairly low, with the most likely values lower than 10-12 MPa per kilometer of
+depth ($T^\prime_{\mathrm{min}} <$ 0.5).
+The highest likelihood coefficient of static friction on the fault is
+estimated at about 0.2--0.3, although values up to 0.5--0.6 are
+permissible. Fluid pressures are likely 0--0.5 of the total pressure.
+Slip occurred on these faults instead of more favorably-oriented faults
+elsewhere in the region, due to the inferred low coefficient of friction
+and moderate fluid pressures.
+
+\appendix
+
+\section{Green's functions for point-source
+loads}\label{greens-functions-for-point-source-loads}
+For completeness, we reproduce the Boussinesq \citep[e.g.,][]{jeffreys1970}
+and Cerruti \citep[e.g.,][]{love1927} solutions here.
+Note that in these solutions, $\lambda$ and $\mu$ are re-defined, and
+are the first and second Lame's parameters, respectively, instead of
+rake and fault friction as in the body of the manuscript.
+
+\subsection{Boussinesq's solutions for vertical point-source
+loads}\label{boussinesqs-solutions-for-vertical-point-source-loads}
+
+\begin{equation}
+\begin{split}
+G_{xx}^B & = \frac{ F_v }{ 2\pi } \left[ \frac{ 3x^2 z}{ r^5 } \right.
++ \frac{\mu (y^2 + z^2)}{(\lambda + \mu) r^3 (z + r)}
+- \frac{\mu z}{(\lambda + \mu) r^3} \\
+&\qquad \quad \left. - \frac{\mu x^2}{ (\lambda + \mu) r^2 (z + r)^2 }\right ]
+\end{split}
+\end{equation}
+
+\begin{equation}
+\begin{split}
+G_{yy}^B & = \frac{F_v}{2\pi } \left [ \frac{3y^2 z}{r^5} \right.
++ \frac{\mu (x^{2} + z^{2})}{(\lambda + \mu) r^{3}(z + r)}
+- \frac{\mu z}{(\lambda + \mu) r ^{3}} \\
+& \qquad \quad \left. - \frac{\mu y^{2}}{(\lambda + \mu ) r^2 (z +r)^2} \right]
+\end{split}
+\end{equation}
+
+\begin{equation}
+G_{xy}^{B} = \frac{F _{v}}{2\pi} \left[ \frac{3xyz}{r^{5}}
+- \frac{\mu x y (z + 2r)}{(\lambda + \mu) r^{3} (z + r)^{2}} \right]
+\end{equation}
+
+\begin{equation}
+G_{zz}^{B} = 3 F _{v} z^{3} / 2 \pi r^{5}
+\end{equation}
+
+\begin{equation}
+G_{xz}^{B} = 3 F _{v} xz^{2} / 2 \pi r^{5}
+\end{equation}
+
+\begin{equation}
+G_{yz}^{B} = 3 F _{v} yz^{2} / 2 \pi r^{5}
+\end{equation}
+
+\subsection{Cerruti's solutions for horizontal point-source
+loads}\label{cerrutis-solutions-for-horizontal-point-source-loads}
+
+\begin{equation}
+\begin{split}
+G_{xx}^{C_x} &= \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3x^2}{r^2} \right.
+- \frac{\mu}{(\lambda + \mu)(z+r)^2} 
+\\
+&\qquad \quad  \left. (r^2 - y^2 - \frac{2ry^2}{r+z}) \right]
+\end{split}
+\end{equation}
+
+\begin{equation}
+\begin{split}
+G_{yy}^{C_x} & = \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3y^2}{r^2} \right.
+- \frac{\mu}{(\lambda + \mu)(z+r)^2} \\
+& \qquad \quad \left. (3r^2 - x^2 - \frac{2rx^2}{r+z}) \right]
+\end{split}
+\end{equation}
+
+\begin{equation}
+\begin{split}
+G_{xy}^{C_x} & = \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3x^2}{r^2} \right.
+- \frac{\mu}{(\lambda + \mu)(z+r)^2} 
+\\
+&\qquad \quad \left. \cdot (r^2 - x^2 - \frac{2rx^2}{r+z}) \right]
+\end{split}
+\end{equation}
+
+\begin{equation}
+    G^{C_x}_{zz} = \frac{ 3 F_{h,x} x z^2 }{2 \pi r^5}
+\end{equation}
+
+\begin{equation}
+    G^{C_x}_{xz} = \frac{ 3 F_{h,x} z x^2 }{2 \pi r^5}
+\end{equation}
+
+\begin{equation}
+    G^{C_x}_{yz} = \frac{ 3 F_{h,x} x y z }{2 \pi r^5}
+\end{equation}
+
+\begin{acknowledgements}
+All data used in this work are from published sources: \citet{shen2009};
+\citet{feng2010}; \citet{qi2011}; \citet{zhang2011}; \citet{fielding2013}. SRTM
+data are from CGIAR-SRTM: srtm.csi.cgiar.org. All code used is available at
+\url{https://github.com/cossatot/halfspace} and
+\url{https://github.com/wenchuan_topo_stress}. We thank the editor, Paul
+Tregoning, and Steve Marshak and an anonymous reviewer for reviews. Funding for
+this study was provided by the University of Michigan. Joe Kington wrote the
+code to plot Figure \ref{fig:joint_posteriors}c.  
+\end{acknowledgements}
+
+%\section{References}\label{references}
+
+%\bibliography{wench_jgr_bib}{}
+%\bibliographystyle{agufull08}
+
+\begin{thebibliography}{90}
+\providecommand{\natexlab}[1]{#1}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+
+\bibitem[{\textit{Aagaard et~al.}(2001)\textit{Aagaard, Heaton, and
+  Hall}}]{aagaard2001}
+Aagaard, B.~T., T.~H. Heaton, and J.~F. Hall (2001), Dynamic earthquake
+  ruptures in the presence of lithostatic normal stresses: {I}mplications for
+  friction models and heat production, \textit{Bulletin of the Seismological
+  Society of America}, \textit{91}(6), 1765--1796.
+
+\bibitem[{\textit{Allmann and Shearer}(2009)}]{allmann2009}
+Allmann, B.~P., and P.~M. Shearer (2009), Global variations of stress drop for
+  moderate to large earthquakes, \textit{Journal of Geophysical Research: Solid
+  Earth (1978--2012)}, \textit{114}(B1).
+
+\bibitem[{\textit{Angelier}(1994)}]{angelier1994}
+Angelier, J. (1994), Fault slip analysis and paleostress reconstruction,
+  \textit{Continental deformation}, \textit{4}.
+
+\bibitem[{\textit{Armijo et~al.}(1986)\textit{Armijo, Tapponnier, Mercier, and
+  Han}}]{armijo1986}
+Armijo, R., P.~Tapponnier, J.~Mercier, and T.-L. Han (1986), Quaternary
+  extension in southern {T}ibet: {F}ield observations and tectonic
+  implications, \textit{Journal of Geophysical Research: Solid Earth
+  (1978--2012)}, \textit{91}(B14), 13,803--13,872.
+
+\bibitem[{\textit{Beaumont et~al.}(2001)\textit{Beaumont, Jamieson, Nguyen, and
+  Lee}}]{beaumont2001}
+Beaumont, C., R.~A. Jamieson, M.~Nguyen, and B.~Lee (2001), Himalayan tectonics
+  explained by extrusion of a low-viscosity crustal channel coupled to focused
+  surface denudation, \textit{Nature}, \textit{414}(6865), 738--742.
+
+\bibitem[{\textit{Bird and Kong}(1994)}]{bird1994}
+Bird, P., and X.~Kong (1994), Computer simulations of {C}alifornia tectonics
+  confirm very low strength of major faults, \textit{Geological Society of
+  America Bulletin}, \textit{106}(2), 159--174.
+
+\bibitem[{\textit{Bird and Piper}(1980)}]{birdpiper80}
+Bird, P., and K.~Piper (1980), Plane-stress finite-element models of tectonic
+  flow in southern {C}alifornia, \textit{Physics of the earth and planetary
+  interiors}, \textit{21}(2), 158--175.
+
+\bibitem[{\textit{Bollinger et~al.}(2004)\textit{Bollinger, Avouac, Cattin, and
+  Pandey}}]{bollinger2004}
+Bollinger, L., J.~Avouac, R.~Cattin, and M.~Pandey (2004), Stress buildup in
+  the {H}imalaya, \textit{Journal of Geophysical Research: Solid Earth},
+  \textit{109}(B11).
+
+\bibitem[{\textit{Bott}(1959)}]{bott1959}
+Bott, M. H.~P. (1959), The mechanics of oblique slip faulting,
+  \textit{Geological Magazine}, \textit{96}(02), 109--117.
+
+\bibitem[{\textit{Burchfiel et~al.}(1995)\textit{Burchfiel, Zhiliang, Yupinc,
+  and Royden}}]{burchfiel1995}
+Burchfiel, B., C.~Zhiliang, L.~Yupinc, and L.~Royden (1995), Tectonics of the
+  {L}ongmen {S}han and adjacent regions, central {}china, \textit{International
+  Geology Review}, \textit{37}(8), 661--735.
+
+\bibitem[{\textit{Burchfiel et~al.}(2008)\textit{Burchfiel, Royden, van~der
+  Hilst, Hager, Chen, King, Li, L{\"u}, Yao, and Kirby}}]{burchfiel2008}
+Burchfiel, B., L.~Royden, R.~van~der Hilst, B.~Hager, Z.~Chen, R.~King, C.~Li,
+  J.~L{\"u}, H.~Yao, and E.~Kirby (2008), A geological and geophysical context
+  for the {W}enchuan earthquake of 12 {M}ay 2008, {S}ichuan, {P}eople's
+  {R}epublic of {C}hina, \textit{GSA Today}, \textit{18}(7), 4--11.
+
+\bibitem[{\textit{Byerlee}(1978)}]{byerlee1978}
+Byerlee, J. (1978), Friction of rocks, \textit{Pure and applied Geophysics},
+  \textit{116}(4-5), 615--626.
+
+\bibitem[{\textit{Byrne et~al.}(2013)\textit{Byrne, Holohan, Kervyn, de~Vries,
+  Troll, and Murray}}]{byrne2013}
+Byrne, P., E.~Holohan, M.~Kervyn, B.~v.~W. de~Vries, V.~R. Troll, and J.~Murray
+  (2013), A sagging-spreading continuum of large volcano structure,
+  \textit{Geology}, \textit{41}(3), 339--342.
+
+\bibitem[{\textit{Cattin et~al.}(1997)\textit{Cattin, Lyon-Caen, and
+  Ch{\'e}ry}}]{cattin1997}
+Cattin, R., H.~Lyon-Caen, and J.~Ch{\'e}ry (1997), Quantification of interplate
+  coupling in subduction zones and forearc topography, \textit{Geophysical
+  Research Letters}, \textit{24}(13), 1563--1566.
+
+\bibitem[{\textit{Chen et~al.}(2010)\textit{Chen, Niu, Liu, Huang,
+  Tkal{\v{c}}i{\'c}, Sun, and Chan}}]{chen2010receiverfn}
+Chen, Y., F. Niu, R. Liu, Z. Huang, H. Tkal{\v{c}}i{\'c}, L. Sun, and
+  W. Chan (2010), Crustal structure beneath {C}hina from receiver function
+  analysis, \textit{Journal of Geophysical Research: Solid Earth}, 
+  \textit{115}(B03307), 1--22, doi:10.1029/2009JB006386.
+
+\bibitem[{\textit{Clark and Royden}(2000)}]{clarkroyden2000}
+Clark, M.~K., and L.~H. Royden (2000), Topographic ooze: {B}uilding the eastern
+  margin of {T}ibet by lower crustal flow, \textit{Geology}, \textit{28}(8),
+  703--706.
+
+\bibitem[{\textit{Clark et~al.}(2005)\textit{Clark, Bush, and
+  Royden}}]{clark2005}
+Clark, M.~K., J.~W. Bush, and L.~H. Royden (2005), Dynamic topography produced
+  by lower crustal flow against rheological strength heterogeneities bordering
+  the {T}ibetan {P}lateau, \textit{Geophysical Journal International},
+  \textit{162}(2), 575--590.
+
+\bibitem[{\textit{Coblentz and Richardson}(1996)}]{coblentz1996}
+Coblentz, D.~D., and R.~M. Richardson (1996), Analysis of the {S}outh
+  {A}merican intraplate stress field, \textit{Journal of Geophysical Research:
+  Solid Earth}, \textit{101}(B4), 8643--8657.
+
+\bibitem[{\textit{Cook and Royden}(2008)}]{cook2008}
+Cook, K.~L., and L.~H. Royden (2008), The role of crustal strength variations
+  in shaping orogenic plateaus, with application to {T}ibet, \textit{Journal of
+  Geophysical Research: Solid Earth (1978--2012)}, \textit{113}(B8).
+
+\bibitem[{\textit{Copley and McKenzie}(2007)}]{copleymckenzie2007}
+Copley, A., and D.~McKenzie (2007), Models of crustal flow in the
+  {I}ndia--{A}sia collision zone, \textit{Geophysical Journal International},
+  \textit{169}(2), 683--698.
+
+\bibitem[{\textit{Copley et~al.}(2009)\textit{Copley, Boait, Hollingsworth,
+  Jackson, and McKenzie}}]{copley2009}
+Copley, A., F.~Boait, J.~Hollingsworth, J.~Jackson, and D.~McKenzie (2009),
+  Subparallel thrust and normal faulting in {A}lbania and the roles of
+  gravitational potential energy and rheology contrasts in mountain belts,
+  \textit{Journal of Geophysical Research: Solid Earth}, \textit{114}(B5).
+
+\bibitem[{\textit{Cui et~al.}(2014)\textit{Cui, Lin, Wang, Gao, Huang, Wang,
+  Sun, Li, Zhou, Qian et~al.}}]{cui2014}
+Cui, J., W.~Lin, L.~Wang, L.~Gao, Y.~Huang, W.~Wang, D.~Sun, Z.~Li, C.~Zhou,
+  H.~Qian, et~al. (2014), Determination of three-dimensional in situ stresses
+  by anelastic strain recovery in {W}enchuan {E}arthquake {F}ault {S}cientific
+  {D}rilling {P}roject {H}ole-1 ({W}{F}{S}{D}-1), \textit{Tectonophysics},
+  \textit{619–620}, 123 -- 132.
+
+\bibitem[{\textit{Dahlen}(1990)}]{dahlen1990}
+Dahlen, F. (1990), Critical taper model of fold-and-thrust belts and
+  accretionary wedges, \textit{Annual Review of Earth and Planetary Sciences},
+  \textit{18}, 55.
+
+\bibitem[{\textit{Dalmayrac and Molnar}(1981)}]{dalmayrac1981}
+Dalmayrac, B., and P.~Molnar (1981), Parallel thrust and normal faulting in
+  {P}eru and constraints on the state of stress, \textit{Earth and Planetary
+  Science Letters}, \textit{55}(3), 473--481.
+
+\bibitem[{\textit{Day}(1982)}]{day1982}
+Day, S.~M. (1982), Three-dimensional simulation of spontaneous rupture: {T}he
+  effect of nonuniform prestress, \textit{Bulletin of the Seismological Society
+  of America}, \textit{72}(6A), 1881--1902.
+
+\bibitem[{\textit{DeCelles et~al.}(2009)\textit{DeCelles, Ducea, Kapp, and
+  Zandt}}]{decelles2009}
+DeCelles, P.~G., M.~N. Ducea, P.~Kapp, and G.~Zandt (2009), Cyclicity in
+  cordilleran orogenic systems, \textit{Nature Geoscience}, \textit{2}(4),
+  251--257.
+
+\bibitem[{\textit{Dewey}(1988)}]{dewey1988}
+Dewey, J.~F. (1988), Extensional collapse of orogens, \textit{Tectonics},
+  \textit{7}(6), 1123--1139.
+
+\bibitem[{\textit{England and Houseman}(1989)}]{england1989}
+England, P., and G.~Houseman (1989), Extension during continental convergence,
+  with application to the tibetan plateau, \textit{Journal of Geophysical
+  Research: Solid Earth (1978--2012)}, \textit{94}(B12), 17,561--17,579.
+
+\bibitem[{\textit{Farr et~al.}(2007)\textit{Farr, Rosen, Caro, Crippen, Duren,
+  Hensley, Kobrick, Paller, Rodriguez, Roth et~al.}}]{farr2007srtm}
+Farr, T.~G., P.~A. Rosen, E.~Caro, R.~Crippen, R.~Duren, S.~Hensley,
+  M.~Kobrick, M.~Paller, E.~Rodriguez, L.~Roth, et~al. (2007), The shuttle
+  radar topography mission, \textit{Reviews of Geophysics}, \textit{45}(2).
+
+\bibitem[{\textit{Feng et~al.}(2010)\textit{Feng, Hetland, Ding, Li, and
+  Zhang}}]{feng2010}
+Feng, G., E.~A. Hetland, X.~Ding, Z.~Li, and L.~Zhang (2010), Coseismic fault
+  slip of the 2008 {M}w 7.9 {W}enchuan earthquake estimated from {I}n{S}{A}{R}
+  and {G}{P}{S} measurements, \textit{Geophysical Research Letters},
+  \textit{37}(1).
+
+\bibitem[{\textit{Fialko et~al.}(2005)\textit{Fialko, Riveral, and Kanamori}}]
+    {fialko2005}
+Fialko, Y., L.~Rivera, and H.~Kanamori (2005), Estimate of differential stress
+  in the upper crust from variations in topography and strike along the San
+  Andreas fault, \textit{Geophysical Journal International}, \textit{160},
+  527--532.
+
+\bibitem[{\textit{Fielding et~al.}(2013)\textit{Fielding, Sladen, Li, Avouac,
+  B{\"u}rgmann, and Ryder}}]{fielding2013}
+Fielding, E.~J., A.~Sladen, Z.~Li, J.-P. Avouac, R.~B{\"u}rgmann, and I.~Ryder
+  (2013), Kinematic fault slip evolution source models of the 2008 m7.9
+  wenchuan earthquake in china from sar interferometry, gps and teleseismic
+  analysis and implications for longmen shan tectonics, \textit{Geophysical
+  Journal International}, \textit{194}(2), 1138--1166.
+
+\bibitem[{\textit{Flesch and Bendick}(2012)}]{fleschbendick2012}
+Flesch, L., and R.~Bendick (2012), The relationship between surface kinematics
+  and deformation of the whole lithosphere, \textit{Geology}, \textit{40}(8),
+  711--714.
+
+\bibitem[{\textit{Flesch and Kreemer}(2010)}]{flesch2010gpe}
+Flesch, L.~M., and C.~Kreemer (2010), Gravitational potential energy and
+  regional stress and strain rate fields for continental plateaus: {E}xamples
+  from the central {A}ndes and {C}olorado {P}lateau, \textit{Tectonophysics},
+  \textit{482}(1), 182--192.
+
+\bibitem[{\textit{Heidbach et~al.}(2009)\textit{Heidbach, Tingay, and
+  Barth}}]{heidbach2009}
+Heidbach, O., M.~Tingay, and A.~Barth (2009), \textit{The World Stress Map
+  based on the database release 2008, equatorial scale 1:46,000,000},
+  Commission for the Geological Map of the World, Paris.
+
+\bibitem[{\textit{Hubbard et~al.}(2010)\textit{Hubbard, Shaw, and
+  Klinger}}]{hubbard2010}
+Hubbard, J., J.~H. Shaw, and Y.~Klinger (2010), Structural setting of the 2008
+  {M}w 7.9 {W}enchuan, {C}hina, earthquake, \textit{Bulletin of the
+  Seismological Society of America}, \textit{100}(5B), 2713--2735.
+
+\bibitem[{\textit{Jarvis et~al.}(2008)\textit{Jarvis, Reuter, Nelson, and
+  Guevara}}]{jarvis2008srtm}
+Jarvis, A., H.~Reuter, A.~Nelson, and E.~Guevara (2008), Hole-filled srtm for
+  the globe version 3, available from the cgiar-csi srtm 90m database,
+  \textit{Accessed: March}, \textit{12}, 2012.
+
+\bibitem[{\textit{Jeffreys}(1924)}]{jeffreys1924}
+Jeffreys, H. (1924), \textit{The Earth: its origin, history and physical
+  constitution}, Cambridge University Press.
+
+\bibitem[{\textit{Jeffreys}(1970)}]{jeffreys1970}
+Jeffreys, H. (1970), \textit{The Earth}, Cambridge University Press.
+
+\bibitem[{\textit{Kanamori and Anderson}(1975)}]{kanamori1975}
+Kanamori, H., and D.~L. Anderson (1975), Theoretical basis of some empirical
+  relations in seismology, \textit{Bulletin of the Seismological Society of
+  America}, \textit{65}(5), 1073--1095.
+
+\bibitem[{\textit{Kuo et~al.}(2014)\textit{Kuo, Li, Smith, Di~Toro, Suppe,
+  Song, Nielsen, Sheu, and Si}}]{kuo2014}
+Kuo, L.-W., H.~Li, S.~A. Smith, G.~Di~Toro, J.~Suppe, S.-R. Song, S.~Nielsen,
+  H.-S. Sheu, and J.~Si (2014), Gouge graphitization and dynamic fault
+  weakening during the 2008 {M}w 7.9 {W}enchuan earthquake, \textit{Geology},
+  \textit{42}(1), 47--50.
+
+\bibitem[{\textit{Lamb}(2006)}]{lamb2006}
+Lamb, S. (2006), Shear stresses on megathrusts: {I}mplications for mountain
+  building behind subduction zones, \textit{Journal of Geophysical Research:
+  Solid Earth}, \textit{111}(B7).
+
+\bibitem[{\textit{Liang et~al.}(2013)\textit{Liang, Gan, Shen, Xiao, Liu, Chen,
+  Ding, and Zhou}}]{liang2013}
+Liang, S., W.~Gan, C.~Shen, G.~Xiao, J.~Liu, W.~Chen, X.~Ding, and D.~Zhou
+  (2013), Three-dimensional velocity field of present-day crustal motion of the
+  {T}ibetan {P}lateau derived from {G}{P}{S} measurements, \textit{Journal of
+  Geophysical Research: Solid Earth}, \textit{118}(10), 5722--5732.
+
+\bibitem[{\textit{Lin et~al.}(2009)\textit{Lin, Ren, Jia, and Wu}}]{lin2009}
+Lin, A., Z.~Ren, D.~Jia, and X.~Wu (2009), Co-seismic thrusting rupture and
+slip distribution produced by the 2008 {M}w 7.9 {W}enchuan
+earthquake, {C}hina, \textit{Tectonophysics}, \textit{471}(3), 203--215.
+
+\bibitem[{\textit{Lisle}(2013)}]{lisle2013}
+Lisle, R.~J. (2013), A critical look at the {W}allace-{B}ott hypothesis in
+  fault-slip analysis, \textit{Bulletin de la Societe Geologique de France},
+  \textit{184}(4-5), 299--306.
+
+\bibitem[{\textit{Liu and Zoback}(1992)}]{liuzoback1992}
+Liu, L., and M.~D. Zoback (1992), The effect of topography on the state of
+  stress in the crust: {A}pplication to the site of the {C}ajon {P}ass
+  {S}cientific {D}rilling {P}roject, \textit{Journal of Geophysical Research:
+  Solid Earth (1978--2012)}, \textit{97}(B4), 5095--5108.
+
+\bibitem[{\textit{Liu and Yang}(2003)}]{liuyang2003}
+Liu, M., and Y.~Yang (2003), Extensional collapse of the {T}ibetan {P}lateau:
+  {R}esults of three-dimensional finite element modeling, \textit{Journal of
+  Geophysical Research: Solid Earth (1978--2012)}, \textit{108}(B8).
+
+\bibitem[{\textit{Liu-Zeng et~al.}(2009)\textit{Liu-Zeng, Zhang, Wen,
+  Tapponnier, Sun, Xing, Hu, Xu, Zeng, Ding et~al.}}]{liu2009}
+Liu-Zeng, J., Z.~Zhang, L.~Wen, P.~Tapponnier, J.~Sun, X.~Xing, G.~Hu, Q.~Xu,
+  L.~Zeng, L.~Ding, et~al. (2009), Co-seismic ruptures of the 12 {M}ay 2008, 
+  {M}w 8.0 {W}enchuan earthquake, {S}ichuan: {E}ast--{W}est crustal
+  shortening on oblique, parallel thrusts along the eastern edge of {T}ibet,
+  \textit{Earth and Planetary Science Letters}, \textit{286}(3), 355--370.
+
+\bibitem[{\textit{Love}(1927)}]{love1927}
+Love, E.~A.~H. (1927), \textit{The mathematical theory of elasticity.},
+  Cambridge University Press.
+
+\bibitem[{\textit{Loveless and Meade}(2011)}]{loveless2011}
+Loveless, J., and B.~Meade (2011), Partitioning of localized and diffuse
+  deformation in the {T}ibetan {P}lateau from joint inversions of geologic and
+  geodetic observations, \textit{Earth and Planetary Science Letters},
+  \textit{303}(1), 11--24.
+
+\bibitem[{\textit{Luttrell et~al.}(2007)\textit{Luttrell, Sandwell,
+  Smith-Konter, Bills, and Bock}}]{luttrell2007}
+Luttrell, K., D.~Sandwell, B.~Smith-Konter, B.~Bills, and Y.~Bock (2007),
+  Modulation of the earthquake cycle at the southern {S}an {A}ndreas fault by
+  lake loading, \textit{Journal of Geophysical Research: Solid Earth
+  (1978--2012)}, \textit{112}(B8).
+
+\bibitem[{\textit{Luttrell et~al.}(2011)\textit{Luttrell, Tong, Sandwell,
+  Brooks, and Bevis}}]{luttrell2011}
+Luttrell, K.~M., X.~Tong, D.~T. Sandwell, B.~A. Brooks, and M.~G. Bevis (2011),
+  Estimates of stress drop and crustal tectonic stress from the 27 {F}ebruary
+  2010 {M}aule, {C}hile, earthquake: Implications for fault strength,
+  \textit{Journal of Geophysical Research: Solid Earth}, \textit{116}(B11).
+
+\bibitem[{\textit{Martel}(2006)}]{martel2006}
+Martel, S.~J. (2006), Effect of topographic curvature on near-surface stresses
+  and application to sheeting joints, \textit{Geophysical Research Letters},
+  \textit{33}(1).
+
+\bibitem[{\textit{McKenzie}(1969)}]{mckenzie1969}
+McKenzie, D.~P. (1969), The relation between fault plane solutions for
+  earthquakes and the directions of the principal stresses, \textit{Bulletin of
+  the Seismological Society of America}, \textit{59}(2), 591--601.
+
+\bibitem[{\textit{McKinney}(2010)}]{mckinney2010}
+McKinney, W. (2010), Data structures for statistical computing in python, in
+  \textit{Proceedings of the 9th Python in Science Conference}, edited by
+  S.~van~der Walt and J.~Millman, pp. 51 -- 56, SciPy.
+
+\bibitem[{\textit{McTigue and Mei}(1981)}]{mctiguemei1981}
+McTigue, D.~F., and C.~C. Mei (1981), Gravity-induced stresses near topography
+  of small slope, \textit{Journal of Geophysical Research: Solid Earth},
+  \textit{86}(B10), 9268--9278.
+
+\bibitem[{\textit{Meade and Conrad}(2008)}]{meade2008}
+Meade, B.~J., and C.~P. Conrad (2008), Andean growth and the deceleration of
+  {S}outh {A}merican subduction: {T}ime evolution of a coupled
+  orogen-subduction system, \textit{Earth and Planetary Science Letters},
+  \textit{275}(1), 93--101.
+
+\bibitem[{\textit{Medina~Luna and Hetland}(2013)}]{medinaluna2013}
+Medina~Luna, L., and E.~A. Hetland (2013), Regional stresses inferred from
+  coseismic slip models of the 2008 {M}w 7.9 {W}enchuan, {C}hina, earthquake,
+  \textit{Tectonophysics}, \textit{584}, 43--53.
+
+\bibitem[{\textit{Meissner and Strehlau}(1982)}]{meissner1982}
+Meissner, R., and J.~Strehlau (1982), Limits of stresses in continental crusts
+  and their relation to the depth-frequency distribution of shallow
+  earthquakes, \textit{Tectonics}, \textit{1}(1), 73--89.
+
+\bibitem[{\textit{Michael}(1987)}]{michael1987}
+Michael, A.~J. (1987), Use of focal mechanisms to determine stress: a control
+  study, \textit{Journal of Geophysical Research: Solid Earth (1978--2012)},
+  \textit{92}(B1), 357--368.
+
+\bibitem[{\textit{Miller and Dunne}(1996)}]{millerdunne1996}
+Miller, D.~J., and T.~Dunne (1996), Topographic perturbations of regional
+  stresses and consequent bedrock fracturing, \textit{Journal of Geophysical
+  Research: Solid Earth (1978--2012)}, \textit{101}(B11), 25,523--25,536.
+
+\bibitem[{\textit{Molnar and Lyon-Caen}(1988)}]{molnar1988}
+Molnar, P., and H.~Lyon-Caen (1988), Some simple physical aspects of the
+  support, structure, and evolution of mountain belts, \textit{Spec. Pap. Geol.
+  Soc. Am}, \textit{218}, 179--207.
+
+\bibitem[{\textit{Mosegaard and Tarantola}(1995)}]{mosegaard1995}
+Mosegaard, K., and A.~Tarantola (1995), Monte {C}arlo sampling of solutions to
+  inverse problems, \textit{Journal of Geophysical Research: Solid Earth
+  (1978--2012)}, \textit{100}(B7), 12,431--12,447.
+
+\bibitem[{\textit{Nakamura et~al.}(2010)\textit{Nakamura, Tsuboi, Kaneda, and
+  Yamanaka}}]{nakamura2010}
+Nakamura, T., S.~Tsuboi, Y.~Kaneda, and Y.~Yamanaka (2010), Rupture process of
+  the 2008 {W}enchuan, {C}hina earthquake inferred from teleseismic waveform
+  inversion and forward modeling of broadband seismic waves,
+  \textit{Tectonophysics}, \textit{491}(1), 72--84.
+
+\bibitem[{\textit{Oglesby and Day}(2002)}]{oglesbyday2002}
+Oglesby, D.~D., and S.~M. Day (2002), Stochastic fault stress: {I}mplications
+  for fault dynamics and ground motion, \textit{Bulletin of the Seismological
+  Society of America}, \textit{92}(8), 3006--3021.
+
+\bibitem[{\textit{Oliphant}(2007)}]{oliphant2007numpy}
+Oliphant, T.~E. (2007), Python for scientific computing, \textit{Computing in
+  Science \& Engineering}, \textit{9}(3), 10--20.
+
+\bibitem[{\textit{Olsen et~al.}(1997)\textit{Olsen, Madariaga, and
+  Archuleta}}]{olsen1997}
+Olsen, K., R.~Madariaga, and R.~Archuleta (1997), Three-dimensional dynamic
+  simulation of the 1992 {L}anders earthquake, \textit{Science},
+  \textit{278}(5339), 834--838.
+
+\bibitem[{\textit{P\'erez and Granger}(2007)}]{perez2007ipython}
+P\'erez, F., and B.~E. Granger (2007), {IP}ython: a {S}ystem for {I}nteractive
+  {S}cientific {C}omputing, \textit{{C}omput. {S}ci. {E}ng.}, \textit{9}(3),
+  21--29.
+
+\bibitem[{\textit{Qi et~al.}(2011)\textit{Qi, Xuejun, Qigui, Freymueller,
+  Shaomin, Caijun, Yonglin, Xinzhao, Kai, and Gang}}]{qi2011}
+Qi, W., Q.~Xuejun, L.~Qigui, J.~Freymueller, Y.~Shaomin, X.~Caijun, Y.~Yonglin,
+  Y.~Xinzhao, T.~Kai, and C.~Gang (2011), Rupture of deep faults in the 2008
+  {W}enchuan earthquake and uplift of the {L}ongmen {S}han, \textit{Nature
+  Geoscience}, \textit{4}(9), 634--640.
+
+\bibitem[{\textit{Reches}(1987)}]{reches1987}
+Reches, Z. (1987), Determination of the tectonic stress tensor from slip along
+  faults that obey the coulomb yield condition, \textit{Tectonics},
+  \textit{6}(6), 849--861.
+
+\bibitem[{\textit{Rey et~al.}(2001)\textit{Rey, Vanderhaeghe, and
+  Teyssier}}]{rey2001}
+Rey, P., O.~Vanderhaeghe, and C.~Teyssier (2001), Gravitational collapse of the
+  continental crust: definition, regimes and modes, \textit{Tectonophysics},
+  \textit{342}(3), 435--449.
+
+\bibitem[{\textit{Richardson and Coblentz}(1994)}]{richardson1994}
+Richardson, R.~M., and D.~D. Coblentz (1994), Stress modeling in the {A}ndes:
+  Constraints on the {S}outh {A}merican intraplate stress magnitudes,
+  \textit{Journal of Geophysical Research: Solid Earth}, \textit{99}(B11),
+  22,015--22,025.
+
+\bibitem[{\textit{Savage and Swolfs}(1986)}]{savageswolfs1986}
+Savage, W.~Z., and H.~S. Swolfs (1986), Tectonic and gravitational stress in
+  long symmetric ridges and valleys, \textit{Journal of Geophysical Research:
+  Solid Earth}, \textit{91}(B3), 3677--3685.
+
+\bibitem[{\textit{Scholz}(2002)}]{scholz2002}
+Scholz, C.~H. (2002), \textit{The mechanics of earthquakes and faulting},
+  Cambridge university press.
+
+\bibitem[{\textit{Seabold and Perktold}(2010)}]{seabold2010}
+Seabold, S., and J.~Perktold (2010), Statsmodels: econometric and statistical
+  modeling with python, in \textit{Proceedings of the 9th Python in Science
+  Conference}, pp. 57--61, SciPy.
+
+\bibitem[{\textit{Shen et~al.}(2009)\textit{Shen, Sun, Zhang, Wan, Wang,
+  B{\"u}rgmann, Zeng, Gan, Liao, and Wang}}]{shen2009}
+Shen, Z.-K., J.~Sun, P.~Zhang, Y.~Wan, M.~Wang, R.~B{\"u}rgmann, Y.~Zeng,
+  W.~Gan, H.~Liao, and Q.~Wang (2009), Slip maxima at fault junctions and
+  rupturing of barriers during the 2008 {W}enchuan earthquake, \textit{Nature
+  Geoscience}, \textit{2}(10), 718--724.
+
+\bibitem[{\textit{Sibson}(1985)}]{sibson1985}
+Sibson, R.~H. (1985), A note on fault reactivation, \textit{Journal of
+  Structural Geology}, \textit{7}(6), 751--754.
+
+\bibitem[{\textit{Styron et~al.}(2010)\textit{Styron, Taylor, and
+  Okoronkwo}}]{styron2010}
+Styron, R., M.~Taylor, and K.~Okoronkwo (2010), Database of active structures
+  from the {I}ndo-{A}sian collision, \textit{Eos, Transactions American
+  Geophysical Union}, \textit{91}(20), 181--182.
+
+\bibitem[{\textit{Styron et~al.}(2015)\textit{Styron, Taylor and
+  Sundell}}]{styron2015}
+Styron, R., M.~Taylor, and K.~Sundell (2015), Accelerated extension of {T}ibet
+  linked to the northward underthrusting of {I}ndian crust, \textit{Nature
+  Geoscience}, \textit{8}(2), 131--134.
+
+\bibitem[{\textit{Tarantola}(2005)}]{tarantola2005}
+Tarantola, A. (2005), \textit{Inverse problem theory and methods for model
+  parameter estimation}, SIAM.
+
+\bibitem[{\textit{Taylor et~al.}(2003)\textit{Taylor, Yin, Ryerson, Kapp, and
+  Ding}}]{taylor2003}
+Taylor, M., A.~Yin, F.~J. Ryerson, P.~Kapp, and L.~Ding (2003), Conjugate
+  strike-slip faulting along the {B}angong-{N}ujiang suture zone accommodates
+  coeval east-west extension and north-south shortening in the interior of the
+  {T}ibetan {P}lateau, \textit{Tectonics}, \textit{22}(4).
+
+\bibitem[{\textit{Tong et~al.}(2010)\textit{Tong, Sandwell, and
+  Fialko}}]{tong2010}
+Tong, X., D.~T. Sandwell, and Y.~Fialko (2010), Coseismic slip model of the
+  2008 {W}enchuan earthquake derived from joint inversion of interferometric
+  synthetic aperture radar, {G}{P}{S}, and field data, \textit{Journal of
+  Geophysical Research: Solid Earth}, \textit{115}(B4).
+
+\bibitem[{\textit{Townend and Zoback}(2000)}]{townend2000}
+Townend, J., and M.~D. Zoback (2000), How faulting keeps the crust strong,
+  \textit{Geology}, \textit{28}(5), 399--402.
+
+\bibitem[{\textit{Wallace}(1951)}]{wallace1951}
+Wallace, R.~E. (1951), Geometry of shearing stress and relation to faulting,
+  \textit{The Journal of Geology}, pp. 118--130.
+
+\bibitem[{\textit{Wang et~al.}(2012)\textit{Wang, Kirby, Furlong, Van~Soest,
+  Xu, Shi, Kamp, and Hodges}}]{wang2012}
+Wang, E., E.~Kirby, K.~P. Furlong, M.~Van~Soest, G.~Xu, X.~Shi, P.~J. Kamp, and
+  K.~Hodges (2012), Two-phase growth of high topography in eastern {T}ibet
+  during the {C}enozoic, \textit{Nature Geoscience}, \textit{5}(9), 640--645.
+
+\bibitem[{\textit{Xiao et~al.}(1991)\textit{Xiao, Dahlen, and
+  Suppe}}]{xiao1991}
+Xiao, H.-B., F.~Dahlen, and J.~Suppe (1991), Mechanics of extensional wedges,
+  \textit{Journal of Geophysical Research: Solid Earth}, \textit{96}(B6),
+  10,301--10,318.
+
+\bibitem[{\textit{Xu et~al.}(2009)\textit{Xu, Wen, Yu, Chen, Klinger, Hubbard,
+  and Shaw}}]{xu2009}
+Xu, X., X.~Wen, G.~Yu, G.~Chen, Y.~Klinger, J.~Hubbard, and J.~Shaw (2009),
+  Coseismic reverse-and oblique-slip surface faulting generated by the 2008 mw
+  7.9 wenchuan earthquake, china, \textit{Geology}, \textit{37}(6), 515--518.
+
+\bibitem[{\textit{Yong et~al.}(2003)\textit{Yong, Allen, Densmore, and
+  Qiang}}]{yong2003}
+Yong, L., P.~A. Allen, A.~L. Densmore, and X.~Qiang (2003), Evolution of the
+  {L}ongmen {S}han foreland basin (western {S}ichuan, {C}hina) during the
+  {L}ate {T}riassic {I}ndosinian orogeny, \textit{Basin Research},
+  \textit{15}(1), 117--138.
+
+\bibitem[{\textit{Zandt and Ammon}(1995)\textit{Zandt and Ammon}}]
+	{zandtammon1995}
+Zandt, G., and C.~Ammon (1995), Continental crust composition constrained by
+measurements of crustal {P}oisson's ratio, \textit{Nature}, \textit{374}(6518),
+  152--154.
+
+\bibitem[{\textit{Zhang et~al.}(2011)\textit{Zhang, Qu, Shan, Song, Zhang,
+  Wang, Hu, and Wang}}]{zhang2011}
+Zhang, G., C.~Qu, X.~Shan, X.~Song, G.~Zhang, C.~Wang, J.-C. Hu, and R.~Wang
+  (2011), Slip distribution of the 2008 {W}enchuan {M}s 7.9 earthquake by joint
+  inversion from {G}{P}{S} and {I}n{S}{A}{R} measurements: a resolution test
+  study, \textit{Geophysical Journal International}, \textit{186}(1), 207--220.
+
+\bibitem[{\textit{Zhang et~al.}(2010)\textit{Zhang, Wen, Shen, and
+  Chen}}]{zhang2010}
+Zhang, P.-Z., X.-z. Wen, Z.-K. Shen, and J.-h. Chen (2010), Oblique,
+  high-angle, listric-reverse faulting and associated development of strain:
+  The {W}enchuan earthquake of {M}ay 12, 2008, {S}ichuan, {C}hina,
+  \textit{Annual Review of Earth and Planetary Sciences}, \textit{38},
+  353--382.
+
+\end{thebibliography}
+
+\end{article}
+
+\begin{figure}[ht!] \includegraphics[width=40pc]{../figures/lms_map.pdf}
+    \caption{Map of eastern Tibet and the Sichuan basin, showing active
+        structures from \citet{styron2010}. Faults that ruptured in the 2008
+        Wenchuan earthquake are shown in pink. GPS velocities are relative to
+        the mean velocity of sites within Sichuan basin, with 1$\sigma$
+        uncertainty, from the dataset of Liang et al.  \citet{liang2013}.
+        Beachball is from the Global CMT focal mechanism solution for the 2008
+    Wenchuan earthquake. BF = Beichuan fault. PF = Pengguan fault. P = Pengguan
+massif. Grey box shows the extent of Figures 3 and 9.} \label{fig:lms_map}
+\end{figure}
+
+\begin{figure}[t!] \centering
+    \includegraphics[width=40pc]{../figures/topo_stress_possibilities.pdf}
+    \caption{Scenarios for topographic effects on rangefront thrust faulting.
+        (a) Topographic stresses promote thrust faulting. (b) Topographic
+        stresses inhibit thrust faulting. Red color represents bedrock, while
+        blue color represents sedimentary basins.} 
+\label{fig:topo_fault_scenarios}
+\end{figure}
 
@@
 
-\subsection{Topographic fault stress
-calculations}\label{topographic-fault-stress-calculations}
-
-Topographic stresses on the Wenchuan faults are calculated on point sets
-representing the faults taken from coseismic slip models. We use six models,
-those of \citet{shen2009}, \citet{feng2010}, \citet{zhang2011},
-\citet{fielding2013} and two from \citet{qi2011} (these two models share a
-fault geometry but use different smoothing functions for the slip
-distribution). These vary somewhat in the fault geometry and specifics of the
-slip distribution, but nonetheless are of high resolution, are properly
-georeferenced. By using a suite of models in our calculations instead of a
-single model, we evaluate results that are persistent in most or all of the
-models to be more robust, and other results can be more confidently linked to
-specifics of the model geometry or slip distribution.
-
-In general, the models share a common geometry and pattern of slip distribution
-above 10 km or so; i.e.\ the locations of high and low slip patches and the slip
-rake are similar in all of the models. Though some of models use discrete fault
-segments for the Beichuan fault, whereas others use a single continuous fault,
-the fault planes are essentially collocated. However, there is significant
-variability in the magnitude of slip in the different models; for example, the
-maximum slip magnitude in the Qi `rough' model is about twice that of the
-Feng model. Large differences also exist in the deeper geometries of the
-models: the Qi, Fielding and Shen models all have horizontal or subhorizontal
-thrust flats at or below 15 km, though only minor slip is assumed to have
-occurred on these fault segments. In contrast, the Feng and Zhang models are
-planar to their lower end at 25-30 km.
-
-In our stress calculations, we discard
-points above 851 m below sea level, as this is above the depth at which
-we compute $G^C(x,y,z)$. The six stress tensor components calculated at
-the regular grid points are linearly interpolated to points describing
-the faults. Because the fault points are completely surrounded by the
-grid nodes at which topographic stresses were calculated and those nodes
-are spaced \textless{}1 km apart, the fault points cannot be more than a
-few hundred meters from the nearest grid node, so a higher order
-interpolation is not necessary. We then project the topographic stress
-tensor to fault normal stress, $\sigma_n^M$, down-dip shear stress,
-$\tau_d^M$ and strike-slip shear stress, $\tau_s^M$ at each point in
-describing the fault geometry.
-
-\section{Results of topographic stress calculations on the Wenchuan
-faults}\label{results-of-topographic-stress-calculations-on-the-wenchuan-faults}
-
-Topographic stresses on the Wenchuan faults are on the order 1--10s MPa
-(Figure~\ref{fig:lms_topo_stresses_rot}, ~\ref{fig:fault_stress_3d}). 
-Stresses are highest in the southwest,
-beneath the Pengguan massif (the highest topography of the Longmen Shan
-front), and decrease to the northeast. $M_{zz}$ is typically larger,
-though not substantially, than $M_{xx}$ or $M_{yy}$. Maximum horizontal
-stress is not typically aligned with either cardinal horizontal
-direction, and is typically larger than $M_{zz}$ above 10 km. Maximum
-$M_{zz}$ is near 80 MPa, on the southwestern Beichuan fault below the
-high peaks of the Pengguan massif, except for in slip models containing
-near-horizontal fault segments in the mid-crust, where $M_{zz}$ reaches
-100 MPa. Vertical shear stresses ($M_{xz}$ and $M_{yz}$) are on the
-order of 1 MPa, and horizontal shear stress ($M_{xy}$) is on the order
-of 0.1 MPa.
-
 \begin{figure}[t]
@@
 
-Because the compressive stresses are near equal, $M$ contains a large
-isotropic component and a smaller deviatoric component. Consequently, $M$
-resolves on the Wenchuan faults with a large $\sigma^M_n$ (median of
-about 40-60 MPa for each slip model) and much smaller $\tau^M_d$. The
-median $\tau^M_d$ is about -3 to -6 MPa in each slip model, where values
-less than zero indicate normal-sense shear, and $\tau^M_s$ median values
-range from about -2 to 1 MPa, where values less than zero indicate
-sinistral shear; fault models with positive median $\tau^M_s$ have broad
-thrust flats at depth, where little slip occurred. On the steeper fault
-segments (where most of the moment release occurred), topographic shear
-stresses are typically normal-sinistral, as opposed to the dominant mode
-of coseismic slip, which is reverse-dextral.
-
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
 \centering
@@
 \caption{Southwest-looking views of topographic stresses and coseismic
-slip on selected slip models. All models share same lateral extent, but
+slip on selected slip models. All views share same lateral extent, but
 the perspectives for B and C are more inclined than A. (A) $\sigma_n^M$
@@
 share the same scale, with no vertical exaggeration. (B) $\tau_d^M$ (colors)
-and dip slip (contours, 1 m interval) on the \citet{qi2011}
-`rough' slip model. (C) $\tau_s^M$ (colors) and strike slip (contours, 1
-m interval) on the \citet{qi2011} slip model.}
+and dip slip (contours, 1 m interval) \citet{qi2011}
+`rough' slip model of the Beichuan fault. (C) $\tau_s^M$ (colors) and strike
+slip (contours, 1
+m interval) on the \citet{qi2011} `rough' slip model if the Beichuan fault.}
 \label{fig:fault_stress_3d}
-\end{figure*}
-
-In contrast to the steeper fault segments, much of the shallowly-dipping
-fault segments (the Pengguan fault and flats at the base of the Beichuan
-fault, where present) have $\tau^M$ in the direction of coseismic slip.
-$M$ is not significantly different in these
-locations, but because of the low dip angle, $M_{zz}$ contributes more
-significantly to $\sigma^M_n$ than to $\tau^M_d$, which is then
-dominated by horizontal compression, leading to reverse-sense shear.
-The stresses caused by the Pengguan massif locally resolve
-as right-lateral on these segments as well. Coseismic slip on these
-fault patches is much lower than on the steeper Beichuan fault, where
-the majority of slip occurred and which is topographically loaded in the
-opposite shear sense.
-
-Compellingly, similar patterns exist in the spatial distributions of
-$\sigma^M_n$ and coseismic slip. Most obvious is the coincidence of
-locally high $\sigma^M_n$ and locally low slip magnitude on the
-southwestern Beichuan fault below the culmination of the Pengguan
-massif, in an area of otherwise high slip (Figure
-\ref{fig:fault_stress_3d}). These correlations exist for other fault
-patches, but they are not as clear (Figure
-\ref{fig:feng_slip_sig_n_scatter}). This raises the possibility that
-topographic loading of these faults contributes to limiting coseismic
-slip once failure has occurred, and
-may have implications for estimations of dynamic friction and the
-completeness of stress drop during the earthquake. Preliminary analysis
-of this is currently being performed, and will be described in a
-forthcoming manuscript.
+\end{figure}
 
@@
 
-\section{Calculations of tectonic stress, fault friction and pore fluid
-pressure}\label{calculations-of-tectonic-stress-fault-friction-and-pore-fluid-pressure}
-
-Faults fail in earthquakes when the shear stresses on the fault overcome
-the frictional stresses resisting slip on the fault \citep[e.g.,][]
-{scholz2002}. We assume that the entire fault was at at the point of
-failure when the Wenchuan fault initiated, and use the Mohr-Coulomb
-failure criterion
-
-\begin{equation} 
-\tau = \mu ( \sigma_n - \sigma_p ) \; ,
-\label{eqn:amonton_raw} 
-\end{equation}
-
-where where $\mu$ is the coefficient of static friction on the fault and
-$\sigma_p$ is the pore fluid pressure \citep[e.g.,][]{sibson1985}. We
-describe the pore fluid pressure using a scalar, $0 \leq \phi \leq 1$,
-which is the pore fluid pressure as a fraction of total pressure, and so
-the failure criterion is
-
-\begin{equation} 
-\tau = \mu (1 - \phi) \sigma_n \; 
-\label{eqn:amonton} 
-\end{equation}
-
-\citep[e.g.,][]{sibson1985}. We assume that both $\mu$ and $\phi$ are
-constant across the Wenchuan faults.
-
-We estimate the tectonic stress tensor field, $\mu$, and $\phi$
-consistent with published coseismic slip models of the Wenchuan
-earthquake using a Bayesian estimation, resulting in posterior probability 
-density functions of the model parameters. 
-We first estimate posteriors of $T$ cosistent with
-the coseismic slip models, and then estimate $\mu$ and $\phi$ consistent
-with Mohr-Coulomb failure. The nature of Bayesian estimation allows us
-to quantify both the relative likelihoods of model parameters and the
-tradeoffs between them.
-
-\subsection{Description of the stress
-state}\label{description-of-the-stress-state}
-
-We consider the complete stress tensor, $S$, at a point in the crust to
-be
-
-\begin{equation}
-S = M + T + L \; ,
-\end{equation}
-
-where $M$ is described above, $T$ is the tectonic stress tensor, and $L$
-is the lithostatic stress tensor. $L$ is isotropic, with diagonal
-components equal to $\rho g z$. We assume that $T$ is laterally
-homogeneous and only has horizontal stress components (i.e.,
-$T_{xz} = T_{yz} = T_{zz} = 0$, with $T_{xx}$, $T_{yy}$ and $T_{xy}$
-non-zero). The expanded stress tensor is:
-
-\begin{equation}
-S = \begin{bmatrix}
-  M_{xx} + T_{xx} + L_{xx} & M_{xy} + T_{xy} &  M_{xz} \\
-	M_{xy} + T_{xy} &  M_{yy} + T_{yy} + L_{yy} & M_{yz} \\
-	M_{xz}     &  M_{yz}  &  M_{zz} + L_{zz}
-	\end{bmatrix} \; .
-\label{eqn:stress_tensor}
-\end{equation}
-
-We further assume that $T$ increases linearly with depth so
-that the entire upper crust is near the critical failure envelope
-\citep[e.g.,][]{townend2000}, and thus we parameterize the components of $T$
-as scalars multiplied by $\rho g z$, denoted as $T^\prime$. Therefore, if
-$T\prime_{xx} = 0.1$, at some point just below 1 km, $L=27$ MPa, so
-$S_{xx} = M_{xx}$+ 27 MPa + 2.7 MPa.
-
-\subsection{Bayesian inversion of tectonic
-stresses}\label{bayesian-inversion-of-tectonic-stresses}
-
-We invert topographic stresses and coseismic slip models for tectonics
-stresses using Bayesian methods, and making the common `Wallace-Bott'
-assumption (named after \citet{wallace1951} and \citet{bott1959}) 
-that slip on the fault occurs in the general direction
-of the maximum resolved shear stress on the fault \citep[e.g.,][]
-{mckenzie1969, angelier1994}. We estimate the tectonic stresses in
-light of the topographic stresses and slip distributions through the
-relation
-
-\begin{equation} 
-p(T|D) \propto p(T) \, p(D|T) \; , 
-\label{eqn:bayes_rule} 
-\end{equation}
-
-where $p(T)$ is the prior PDF (or \emph{priors}) of $T$, $p(D|T)$ is the
-likelihood of observing the coseismic slip distribution $D$ given the
-tectonic stresses $T$, and $p(T|D)$ is the posterior PDF of $T$ given
-$D$, which is the solution to the inversion \citep[e.g.,][]{mosegaard1995}.
-Due to the unknown proportionality in equation (\ref{eqn:bayes_rule}),
-our posterior only gives likelihood of $T$ relative to the most likely
-estimate (MLE) \citep{tarantola2005}. We follow a Monte-Carlo strategy,
-where samples of the prior PDF are retained as samples of the posterior
-in proportion to $p(D|T)$ \citep[e.g.,][]{mosegaard1995}.
-
-We parameterize $T$ by the magnitudes and orientation of the maximum and
-minimum principal tectonic stresses. We assume priors such that the
-magintudes of principal tectonic stresses are equally likely within
-bounds and that all orientations are equally likely. Because the
-Wenchuan event was an oblique reverse faulting earthquake, we assume
-that total horizontal stresses are greater than the vertical stress,
-which is satisfied if the tectonic stresses are positive. Prior samples
-of maximum principal stress are taken from a uniform distribution
-between $\rho g z$ and 2.5 $\rho g z$. Samples of the minimum principal
-stress are from a uniform distribution between 0 and the value for
-maximum stress. We describe the orientation of the tectonic stress using
-the azimuth of the maximum tectonic stress, which are sampled uniformly
-from 0 to 360$^{\circ}$.
-
-We test 100,000 unique samples drawn from the prior using a seeded
-pseudorandom number generator. We test the same prior samples against
-each of the coseismic slip models. $S$ is then constructed for each
-point discretizing the fault geometries in the coseismic slip models.
-The rake of the maximum shear stress $\lambda^S$ on each point of the
-fault is calculated and compared to the coseismic slip rake $\lambda^D$
-at that point. A weighted mean misfit is calculated by
-
-\begin{equation}
-\bar{\lambda}^m = \sum \nolimits_{i=1}^n \frac{(\lambda^S_i - \lambda^D_i) D_i} 
-{\bar{ D}} \; ,
-\label{eqn:rake_misfit}
-\end{equation}
-
-where $D$ is the coseismic slip and $\bar{D}$ is the average coseismic
-slip in a given coseismic slip model. Finally, the relative likelihood
-of each model is computed using a Von Mises distribution as
-
-\begin{equation}
-p (D | T) = \frac{ \exp ( \kappa \cos \bar{\lambda}^m )} 
-{\exp (\kappa \cos \bar{\lambda}^m_{\min})} \;,
-\label{eqn:rel_likelihood}
-\end{equation}
-
-where $\kappa$ = 8.529, which is calculated so that the 68.2\%
-confidence interval the Von Mises distribution is within $\pi$/9 radians
-(20$^{\circ}$), the estimated 1$\sigma$ uncertainty of the coseismic
-slip models based on comparisons between rakes of high-slip fault
-patches (note that for a planar fault, $\tau$ at $\pi/9$ radians from
-$\lambda_{max}$ is still \textgreater{}90\% of $\tau_{\mathrm{\max}}$
-\citep{lisle2013}). Prior samples are retained in propotion to $p(D|T)$,
-and the retained samples are then samples of the posterior.
-
-\subsection{Analysis of friction and pore fluid
-pressure}\label{analysis-of-friction-and-pore-fluid-pressure}
-
-Once the tectonic stress distributions consistent with the coseismic
-slip models have been determined, we deduce the distributions of $\mu$
-and $\phi$ assuming that the stress is at the failure criterion in
-equation (\ref{eqn:amonton}). We do this in three steps: First, we draw
-a random $\phi$ from a uniform distribution, assuming $0 \leq \phi < 1$.
-We again use a seeded pseudorandom number generator, such that each
-stress model has a uniquely assigned $\phi$ that is consistent priors
-accross all coseismic slip models. Second, we calculate $\tau^S$ and
-$\sigma_n^S(1-\phi)$ for each point on the fault. Third, we solve
-Equation \ref{eqn:amonton} for $\mu$. Finally, we filter the results so
-that only models with $0 \le \mu < 1$ are retained, as values outside of
-that range have not been suggested for rocks.
-
-After this analysis has been done for all coseismic slip model, we find
-the joint posterior (i.e., the posterior consistent with all of the
-coseismic slip models) by taking the samples that are common to all of
-the individual posteriors. We denote the joint posterior as
-$p_{\mathrm{J}}(P | D)$, where $P$ is the parameter of interest.
-
-\section{T, $\mu$, $\phi$ results}\label{t-mu-phi-results}
-
-\subsection{Individual slip models}\label{individual-slip-models}
-
-Results for $T^\prime$, $\mu$ and $\phi$ are quite consistent across all
-coseismic slip models (Figure \ref{fig:T_scatters}). Maximum compressive
-stress $T^\prime_{\mathrm{max}}$ is broadly east-west for all models,
-with a mode trending at $90^{\circ}$--$105^{\circ}$.
-$p(T^\prime_{\mathrm{max}}|D)$ for each slip model increases from
-$T^\prime = 0$ to 0.5 or 1 before essentially leveling off, though some
-slip models, particularly the \citet{qi2011} model, show a
-slight decrease in relative likelihood past the initial mode at
-$T^\prime =$ 0.5--1.
-The low likelihood below $T^\prime \approx 0.5$ indicates that lower
-tectonic stresses are unlikely to overcome fault friction and
-topographic shear stresses resisting reverse-dextral slip on the
-Wenchuan faults. $p(T^\prime_{\mathrm{min}} | D)$ for each slip model
-has a mode close to $T^\prime_{\mathrm{min}} = 0.2 $ and decreases
-abruptly at higher values, though all slip models show values for
-$T^\prime_{\mathrm{min}}$ up to 2.5. $T^\prime_{\mathrm{min}}$ is
-typically 0--0.4 of $T^\prime_{\mathrm{max}}$, but rarely higher.
-
-\begin{figure*}%[htbp]
+\begin{figure}%[htbp]
 \centering
 \includegraphics[width=40pc]{../figures/T_scatters.pdf}
-\caption{Samples of
+\caption{Scatterplots of samples drawn from
 $p(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$ associated with
-each of the coseismic slip models we consider, along with marginals of
+each of the coseismic slip models we consider, along with marginal 
+distributions of
 $T^\prime_{\mathrm{max}}$ and $T^\prime_{\mathrm{min}}$. Inset rose
@@
 \label{fig:T_scatters}
-\end{figure*}
-
-All slip models show $p(\phi | D)$ to be uniformly high from $\phi = 0$
-to 0.4--0.6 and to decrease linearly to $p(\phi) = 0$ at $\phi = 1$
-(Figure \ref{fig:mu_phi_scatters}). $p(\mu | D)$ for all slip models has
-a mode at $\mu =$ 0.1--0.4 and $p(\mu)$ decreases at higher values.
-$T^\prime_{\mathrm{max}}$, $\phi$ and $\mu$ are highly correlated, where
-higher values of $T^\prime_{\mathrm{max}}$ are associated with higher
-$\mu$ and lower $\phi$. Combinations of high $\mu$ and low $\phi$
-require much higher $T^\prime_{\mathrm{max}}$ to overcome fault friction
-and cause slip, and so are not represented in the posteriors. Since our
-maximum $T_{\mathrm{max}}$ of $2.5 \rho g z$ is quite high
-($\approx 660$ MPa at 10 km), we view high $\mu$ and low $\phi$
-combinations as unrealistic for the Wenchuan faults. Similarly,
-combinations of very low $\mu$ and very high $\phi$ are associated with
-very low $T^\prime_{\mathrm{max}}$, and have a low probability density,
-as it is unlikely that tectonic stress with very low $T^\prime_{\mathrm{max}}$
-values can
-overcome sinistral and normal-sense topographic shear stresses to cause
-the observed coseismic slip kinematics.
-
-\begin{figure*}[htbp]
+\end{figure}
+
+\begin{figure}[htbp]
 \centering
@@
 indicate relative density (i.e., likelihood) of posteriors (darker lines
-signify higher densities).}
+signify higher densities), and are constructed through kernel density
+estimation.}
 \label{fig:mu_phi_scatters}
-\end{figure*}
-
-\subsection{Joint posteriors}\label{joint-posteriors}
-
-We define a joint posterior, $p_{J}(T^\prime | D)$, by the samples that 
-are common to the individual posteriors estimated from each slip model.
-Unsuprisingly, given the broad similarity between the posteriors from
-the various slip models, $p_{J}(T^\prime_{\mathrm{max}} | D)$ is not
-substantially different from any of the constituent model posteriors.
-$p_{J}(T^\prime_{\mathrm{max}} | D)$ has a somewhat more well-defined
-mode at $T^\prime_{\mathrm{max}} \approx 0.6$. It is also apparent in
-Figure \ref{fig:joint_posteriors}~b that regardless of the magnitude of
-$T^\_{\mathrm{max}$, $p_{J}(T^\prime_{\mathrm{min}} | D)$ is
-consistently 0--0.6 $p_{J}(T^\prime_{\mathrm{max}} | D)$, with a mode
-of about 0.3; this indicates that the relative magnitude of the
-tectonic stresses has a substantial influence on the rake of the maximum
-shear stress resolved on the fault.
-
-In our estimation of $\phi$ and $\mu$ in the posteriors associated with
-each slip model, we have used the same random combinations of $T$ and
-$\phi$ for each slip model, and then solved for $\mu$ so that the fault
-is at a critical stress state (Equation \ref{eqn:amonton}). Because of
-differences in the location and slip among the coseismic slip models,
-some variability exists in $\mu$ for each prior sample. We therefore
-choose $p_{\mathrm{J}}(\mu | D)$ to be the median $\mu$ of each slip
-model for each sample. $p_{\mathrm{J}}(\mu | D)$ has a mostly similar
-distribution as $p(\mu | D)$ for any of the slip models. However,
-$p_{\mathrm{J}}(\mu | D)$ has a lower relative likelihood on the
-high-$\mu$ tail compared to the constituent $p(\mu | D)$. This lower
-likelihood of $\mu$ in the joint posterior is likely becuase it is the
-average $\mu$ of all slip models. On the other hand, it is not similarly
-sparse on the low-$\mu$ side, suggesting that low values for $\mu$ are
-more robust.
-
-\begin{figure*}[t]
+\end{figure}
+
+\begin{figure}[t]
 \centering
@@
 $p_{\mathrm{J}}(T^\prime_{\mathrm{\max}}, T^\prime_{\mathrm{\min}}/T^\prime_{\mathrm{max}} | D)$,
-along with marginals of $T^\prime_{\mathrm{max}}$ and
+along with marginal distributions of $T^\prime_{\mathrm{max}}$ and
 $T^\prime_{\mathrm{min}} / T^\prime_{\mathrm{max}}$. (c) Histogram of
 azimuths of $T^\prime_{\mathrm{\max}}$. (d) Samples of
-$p_{\mathrm{J}}(\mu, \phi | D)$, with marginals, where color of the samples
+$p_{\mathrm{J}}(\mu, \phi | D)$, with marginal distributions,
+where color of the samples
 indicate magnitudes of $T^\prime_{\mathrm{\max}}$ and contour
@@
 \label{fig:joint_posteriors}
-\end{figure*}
-
-\section{Discussion}\label{discussion}
-
-Few studies have performed similar quantification of static stress
-fields on faults (see Section
-\ref{previous-work-on-topographic-stresses} for some examples), even
-though it may have important ramifications to the earthquake process.
-Most studies of fault rupture dynamics
-assume either a homogeneous or stochastic shear stress distribution
-\citep[e.g.,][]{oglesbyday2002} and few assume any variation in normal
-stress \citep[e.g.,][]{aagaard2001}, despite the importance that stress
-variations likely have in earthquake dynamics \citep[e.g.,][]{day1982,
-olsen1997}. Additionally, quantifying friction and pore fluid
-pressure involved in faulting is a major challenge in studies of
-faulting and orogenic dynamics \citep[e.g.,][]{meissner1982,
-oglesbyday2002}.
-
-Previous workers have demonstrated that by quantifying topographic
-stress, other components in the Coulomb stress balance may be bracketed
-\citep[e.g.,][]{cattin1997, lamb2006, luttrell2011}. Each of these studies
-uses somewhat different approaches. Our approach is most similar to that
-of \citet{luttrell2011}, although there are significant
-differences: (1) We use the topographic stresss calculations proposed by
-\citet{liuzoback1992}, whereas 
-\citet{luttrell2011} only uses the vertical loading from topography,
-equivalent to $M^B$ in equation (\ref{eqn:bous}). (2) We do not consider
-bouancy forces due to lateral variations in density, due for instance to
-Moho variation, as done by \citet{luttrell2011}. In the
-Longmen Shan region, Moho variation is small compared to the change in
-Moho depths at the Andean plate boundary, and so the buoyancy terms
-should be relatively small. (3) We consider a full range of tectonic
-stresses, instead of simply calculating the minimum principal tectonic
-stress and its orientation. (4) We consider the stress tensor at each
-point due to topographic loading, lithostatic stress, and horizontal
-tectonic stress, and use inferred coseismic slip rake as a constraint on
-the allowable stresses, rather than inferring the earthquake stress drop
-from the coseismic slip models, as done by 
-\citet{luttrell2011}. (5) We use both normal and shear stresses to
-constrain pore fluid pressure and friction. (6) We use a Bayesian
-estimation, resulting in a PDF of tectonic stress, as well as $\mu$ and
-$\phi$, rather than just solving for the minimum tectonic stress
-required for faulting, as done by \citet{luttrell2011}.
-
-\subsection{Topographic stresses on the Wenchuan
-faults}\label{topographic-stresses-on-the-wenchuan-faults}
-
-Topographic stresses on the main Wenchuan faults are of considerable
-magnitude: $\tau^M$ ranges from about -20 to 10 MPa, which is on the
-order of inferred stress drop in earthquakes \citep[e.g.,][]{kanamori1975,
-allmann2009}. Topographic stresses are
-generally opposed to the tectonic slip direction, and therefore have to
-be overcome by tectonic stresses in order to produce the observed
-rupture patterns. The topographic shear stresses are likely persistent
-over the lifespan of the topography (i.e.~on the order of millions to
-tens of millions of years), otherwise the Wenchaun faults may fail in a
-normal sense simply due to the weight of the Longmen Shan. This would
-argue that tectonic stress drop in the Wenchuan earthquake was not
-complete, as some residual tectonic shear stress must remain on the
-fault to cancel out $\tau^M$.
-
-The spatial variation of topographic stresses decreases in frequency and
-amplitude with depth. This is not surprising, because with increasing
-depth, the stress field at any point is more sensitive to surface loads
-averaged over a greater region, and is less dominated by smaller scale
-topographic features (i.e., individual mountains). The spatial
-variablity of the topographic stress with depth is similar to the spatial
-variablity of coseismic slip in the models considered \citep[e.g.,][]
-{zhang2011}, which are both smoother at depth. Some of the
-estimated slip variability is likely partially due to the more limited
-resolution of coseismic slip at depth using geodetic data.
-However, the negative spatial correlations of slip versus stress
-(especially $\sigma^M_n$) (Figures \ref{fig:fault_stress_3d},
-\ref{fig:feng_slip_sig_n_scatter}) suggest this may be a real signal.
-
-\subsection{Tectonic stresses in eastern
-Tibet}\label{tectonic-stresses-in-eastern-tibet}
-
-The maximum tectonic stress, $T_{\mathrm{max}}$, is consistently
-oriented roughly E-W in our results. This orientation is oblique to the
-Longmen Shan, which produces oblique (right-lateral and reverse sense)
-shear on the Beichuan fault. $T_{\mathrm{min}}$ is \textasciitilde{}N-S
-oriented, and is only slightly larger than lithostatic pressure. This
-stress configuration is in close agreement with pre-earthquake stress
-measurements near the rupture zone (Figure \ref{fig:lms_stress_map}),
-mostly from borehole breakout data from 2-5 km depth
-\citep{heidbach2009}, which is the zone of maximum slip in the coseismic
-slip models. It is somewhat discrepant with stress orientations
-estimated at \textasciitilde{}800 m depth adjacent to the Beichuan fault
-in the WFSD-1 borehole of the Wenchuan Earthquake Fault Scientific
-Drilling Project several years after the 2008 earthquake \citep{cui2014},
-which show $\sigma_{H{\mathrm{\max}}}$ to be more orthogonal to the fault
-trace, suggesting that much of the right-lateral component of shear
-stress was released during the earthquake. These results are similar to
-the orientations of total stress obtained by 
-\citet{medinaluna2013}, although they were unable to constrain the
-magnitudes of stresses.
+\end{figure}
 
@@
 
-The magnitudes of $T_{\mathrm{max}}$ and $T_{\mathrm{min}}$ are
-dominantly constrained on the low end by our analysis, which is apparent
-by the sharp decrease in the frequency of $p(T_{\mathrm{max}}|D)$ below
-about $T^\prime = 0.5$. This indicates that $T_{\mathrm{max}}$ of at
-least \textasciitilde{}13.25 MPa km$^{-1}$ is necessary to overcome
-topographic stresses resisting reverse and right-lateral slip on the
-faults. We find that the highest likely ratio of strike-slip to dip-slip
-shear along the Wenchuan earthquake faults is close to 1. This is
-similar to the inferences of strain accumulation rates inferred from
-squishy block modeling by Loveless and Meade \citet{loveless2011}, who
-concluded that the rate of slip deficit accumulation in the thrust and
-dextral sense was approximately equal. It should be noted that the
-tectonic stresses we estimate here represent the accumulated stresses
-prior to the Wenchuan earthquake, and the relation of these stresses to
-accumulated slip deficit needs to be through a model of strain
-accumulation.
-
-The orientation of both the tectonic and total stresses near the
-Wenchuan faults shows a larger difference with patterns of strain from
-elsewhere in the orogen. For example, the presence of N-S contraction
-and E-W extension throughout the high Tibetan plateau and much of the
-Himalaya \citep[e.g.,][]{armijo1986, molnar1988, taylor2003} indicates a
-roughly N-S $T_{\mathrm{max}}$ and E-W $T_{\mathrm{min}}$. Because
-$T_{\mathrm{min}}$ is only slighly above lithostatic pressure on the
-Wenchuan faults, it is quite possible that the N-S compression in the
-Himalaya and Tibet, which is almost certainly due to Indo-Asian plate
-collision, has significantly decayed at the Longmen Shan, some 850
-km northeast of the easternmost Himalaya. Therefore, contraction across
-the Longmen Shan cannot easily be interpreted to directly reflect
-stresses due to the Indo-Asian collision alone, unless some additional
-mechanism of redirecting crustal stresses is incorporated.
-
-\subsection{Slip on the Beichuan fault vs.~optimally-oriented
-faults}\label{slip-on-the-beichuan-fault-vs.optimally-oriented-faults}
-
-Our highest likelihood estimates of $\mu$ are in the range of 0.2--0.3.
-These values are slightly lower than $\mu \approx 0.4$ inferred in
-laboratory experiments on samples recoverd from the WFSD-1 drill hole
-into the Beichuan fault \citep{kuo2014}. However, it should be noted that
-$\mu \approx 0.4$ is relatively high likelihood in our posteriors
-(Figure \ref{fig:joint_posteriors}). These values of friction 
-are lower than typical values derived from laboratory experiments on
-intact rock \citep[e.g.,][]{byerlee1978}, suggesting that slip occurred on
-preexisting faults because they are weaker than optimally-oriented new
-faults.
-
-The obliquity of slip on the Wenchuan earthquake faults also suggests
-that these faults may not be optimally oriented for slip given the total
-stress state in the region of these faults. However, the Longmen Shan
-fault zone dates back to the Indosinian orogeny, locally late Triassic
-(226-206 Ma) \citep{yong2003} and has had multiple episodes of
-reactivation since \citep[e.g.,][]{burchfiel1995, wang2012}, accumulating
-tens of kilometers of shortening \citep[e.g.,][]{hubbard2010}. Such a mature
-fault system may be expected to have low coefficients of friction due to
-processes such as gouge development \citep[e.g.,][]{kuo2014}, and therefore
-may slip in non-optimal orientations, with high $\phi$ values also
-potentially contributing to this \citep[e.g.,][]{sibson1985}.
-
-Our posterior estimates for $T$, $\phi$ and $\mu$ let us quantitatively
-evaluate to what extent slip on the Wenchuan faults is more favorable
-than on optimally oriented faults with more typical friction
-coefficients. We use the same failure conditions as in determing the
-posteriors above, and assume optimally oriented faults exist
-(i.e., we do not consider the generation of new faults in relatively 
-intact rock).
-Additionally, we evaluate the relative contributions of $T$, $\phi$ and
-$\mu$ on potential fault weakening and reactivation. To explore these
-relationships, we perform some preliminary analysis on a single fault
-model (from \citet{zhang2011}) using a subset of 1000
-samples drawn randomly from the joint posteriors. Given the similarity
-of the fault models and of the posteriors for each model, we do not
-expect that an analysis of all results on all fault models will yield
-different conclusions.
-
-First, we establish a metric with which to evaluate the favorability of
-slip on a given fault plane, which we call the Coulomb failure ratio, or
-CFR:
-
-\begin{equation}
-\mathrm{CFR} = \tau / \mu (1 - \phi) \sigma_n \; .
-\label{eqn:cfr}
-\end{equation}
-
-CFR indicates whether a fault should fail under a given stress state:
-CFR \textgreater{} 1 indicates failure, while a CFR \textless{} 1
-indicates fault stability. We then calculate the CFR on each point in
-the model of the Beichuan fault (594 points describe the fault model of
-\citet{zhang2011}) based on the full stress field $S$ at
-each point on the fault, for each of the 1000 samples of $T$, $\phi$ and
-$\mu$ drawn randomly from the posteriors. We call this CFR$_f$. Then,
-using the same $S$ and $\phi$, we calculate the CFR on an optimally
-oriented fault with $\mu=0.6$ and no cohesion, which we call CFR$_o$.
-Note this $\mu$ is typical for crustal rocks with fault normal stresses
-above 200MPa, but lower than $\mu = 0.85$ for smaller $\sigma_n$
-\citep{byerlee1978}, but may be appropriate for an immature crustal
-fault.
-
-Figure \ref{fig:cfr_ratios} shows
-$\log (\mathrm{CFR}_o / \mathrm{CFR}_f)$ plotted against $\mu$, $\phi$
-and $T^\prime_{xx}$ on the Beichuan fault for all samples. Though
-considerable scatter exists, it is clear that in most instances, slip on
-the Beichuan fault is preferred over slip on an optimal fault. The
-exceptions are at high values of $\mu$, $\phi$, or $T^\prime$, where
-slip on an optimal plane is preferred. Because $T$, $\phi$ and $\mu$ can
-all affect fault reactivation \citep[e.g.,][]{sibson1985}, we compare the
-relative contributions of each with a simple mutliple linear regression,
-using $T^\prime_{xx}$ normalized to [0,1) (the same range $\phi$ and
-$\mu$) as a proxy for $T$ ($T^\prime_{xx}$ is essentially
-$T^\prime_{\mathrm{max}}$ in most of the posteriors). The results are
-shown in Table \ref{table:cfr_regress}. It is clear that $\mu$ is most
-strongly correlated with $\mathrm{CFR}_o/\mathrm{CFR}_f$, followed
-closely by $T^\prime_{xx}$ and then $\phi$; nonetheless, all
-significantly affect the relative ease of faulting on the Beichuan fault
-versus optimal faults.
-
 \begin{figure}%[tb]
@@
 \hline
-Intercept & -0.2566 & 0.0020 & -128.8 & \textless{}0.0001 \\ 
-$\mu$ & 1.1168 & 0.0086 & 130.6 & \textless{}0.0001 \\ 
-$T^\prime_{xx}$ & 0.9138 & 0.0048 & 191.2 & \textless{}0.0001 \\ 
-$\phi$ & 0.5762 & 0.0055 & 105.7 & \textless{}0.0001 \\ 
+Intercept & -1.8211 & 0.002 & -937.5 & \textless{}0.0005 \\ 
+$T^\prime_{xx}$ & 1.4082 & 0.005 & 302.4 & \textless{}0.0005 \\ 
+$\mu$ & 1.1301 & 0.008 & 156.2 & \textless{}0.0005 \\ 
+$\phi$ & 1.1804 & 0.005 & 222.2 & \textless{}0.0005 \\ 
 \hline
@@
 
-\subsection{The role of topography in orogenic development and strain
-localization}\label{the-role-of-topography-in-orogenic-development-and-strain-localization}
-
-Because the rise of broad elevated regions creates significant stresses
-in the crust \citep[e.g.,][]{jeffreys1924}, these stresses have the
-potential to influence orogenesis. The result may be to change the deformation
-state in the interior of the orogen \citep[e.g.,][]{dewey1988, molnar1988}, the
-convergence rates of the plates surrounding the orogen \citep[e.g.][]
-{meade2008}, or the location and style of deformation at the orogen's
-margins \citep[e.g.,][]{beaumont2001, decelles2009}. These studies show that
-as an orogen rises, horizontal contraction via crustal thickening will
-occur at the thin margins of the orogen, and if elevations reach a
-threshold or tectonic compression decreases, extension will take place
-in the orogen's high interior, as is well described in Tibet. Therefore,
-Tibet is commonly suggested to be undergoing some manner of
-gravitational collapse \citep[e.g.,][]{england1989}; where the plateau abuts
-rigid cratons, the deformational patterns resemble `fixed boundary'
-collapse \citep{rey2001}, with concentrated crustal thickening at the
-orogen's margins \citep[e.g.,][]{cook2008}.
-
-Though much work has been done exploring the feedback mechanisms between
-topography and deformation; however, topography is typically greatly smoothed,
-and topographic stresses are folded into gravitational potential energy
-estimates of the entire lithosphere, and the whole lithospheric column
-is often considered viscous \citep{copleymckenzie2007}. Therefore many
-models seeking to predict deformation from gravitational forces yield
-continuous deformation and do not consider how the gravitational
-stresses resolve on heterogeneities embedded in the crust. However, the
-presence of structures such as weak faults \citep{bird1994} or
-low-viscosity channels or shear zones \citep[e.g.,][]{clark2005} can localize
-deformation \citep[e.g.,][]{bird1994, fleschbendick2012}.
-
-Our results that topographic stresses on the Wenchuan
-earthquake faults are largely loaded in the opposite sense to the
-coseismic slip direction indicates that gravitational collapse is probably
-not the driver of reverse faulting in the Longmen Shan. However, as
-noted previously, thrust flats in the middle crust present in some
-coseismic slip models, particularly the Qi model, are loaded in a thrust
-and dextral sense. Given the very low dips of the thrust flats
-($0^{\circ}$--$5^{\circ}$), they receive very little loading from
-horizontal tectonic stress, so any slip on them may be in response to
-topographic loading and thus to gravitational collapse of the high
-Longmen Shan and its hinterland. $M_{H{\mathrm{\max}}$ also changes
-orientation from fault rangefront normal to closer to more oblique
-near the front of the range (Figure \ref{fig:lms_topo_stresses_rot}),
-suggesting that any transfer of mass from the highlands to the lowlands
-due to topographic stresses should terminate near the rangefront.
-
-However, because $M_{zz}$ is greater than $M_{H{\mathrm{max}}$
-underneath the higher topography, topographic stresses can only lead to
-reverse faulting on the Wenchuan faults if subhorizontal shear zones or
-channels are present, and are very weak so shear can occur at 
-low $\tau$ relative to $\sigma_n$. This conclusion has been reached in
-studies of generalized orogens \citep[e.g.,][]{fleschbendick2012} and
-gravitational collapse of volcanic edifices \citep[e.g.,][]{byrne2013},
-indicating that is a general requirement of gravitational spreading. Our
-calculations of $\mu$ and $\phi$ are based on $T$, which is biased
-towards the high slip patches in the upper crust (Equation
-\ref{eqn:rake_misfit}), yield low $\mu$ and low to moderate $\phi$,
-which is probably insufficient for slip at high pressures at depth.
-Similarly, a low viscosity horizontal
-channel that underlies eastern Tibet, likely much deeper than the
-seismogenic zone we consider here, would be able to flow in response to
-$\tau$ regardless of $\sigma_n$ and may be able to facilitate
-gravitational spreading of the orogen \citep[e.g.,][]{clark2005, cook2008,
-fleschbendick2012}.
-
-These arguments are all quite speculative, and we wish only to describe the
-conditions under which gravitational collapse can occur given the
-observations of deformation and the topographic stress field. It is not
-at all clear that a suitably large decollement exists at the base of the
-Beichuan fault (it is only present in one of the slip models
-considered). Nonetheless, this topic has implications not only for
-orogenic development, but for the recurrence interval on the Wenchuan
-earthquake faults. If topographic stress contributes in some fashion to
-the observed displacements on the fault, then those stresses were likely
-barely diminished by coseismic stress change, and earthquake recurrence
-of on the Wenchuan faults may be governed by different processes than
-elastic rebound due to tectonic strain accumulation.
-
-\section{Conclusions}\label{conclusions}
-
-We have calculated shear and normal stresses due to topographic loading
-on the Wenchuan earthquake faults, and used those stresses to constrain
-tectonic stresses, fault friction and pore fluid pressure. Topographic
-stresses on the main Wenchuan faults are large, with $\tau^M$ on the
-faults up to \textbar{}20\textbar{} MPa, and $\sigma^M_n$ up to 80 MPa.
-$\sigma^M_n$ reaches up to 100 MPa on mid-crustal thrust flats present
-in some coseismic slip models. The direction of $\tau_M$ is generally
-opposed to coseismic slip inferred during the 2008 Wenchuan earthquake,
-indicating that weight of the topography resists coseismic slip. High
-values of $\sigma^M_n$ increase the frictional resistance to slip,
-potentially limiting slip magnitude in locations such as below the
-Pengguan massif.
-
-Assuming that the Beichuan faults were at a Mohr-Coulomb fault criterion
-immediately prior to the Wenchuan earthquake, we estimate the tectonic
-stresses required for the faults to fail. We use a Bayesian estimation,
-resulting in posteriors representing likelihood of tectonic stress,
-static friction, and a pore pressure parameter. The posteriors
-indicate that the maximum tectonic stress is oriented
-\textasciitilde{}E-W and has a likely minimum of 13.25 MPa km$^{-1}$.
-The minimum tectonic stress is oriented \textasciitilde{}N-S and is
-fairly low, with highest likelihood one half of the E-W tectonic stress.
-The highest likelihood coefficient of static friction on the fault is
-estimated at about 0.2--0.3, although values up to 0.5--0.6 are
-permissable. Fluid pressures are likely 0--0.5 of the total pressure.
-
-Slip occurred on these faults instead of more favorably-oriented faults
-elsewhere in the region, due to the inferred low coefficient of friction
-and moderate fluid pressures.
-
-\appendix
-
-\section{Green's functions for point-source
-loads}\label{greens-functions-for-point-source-loads}
-For completeness, we reproduce the Boussinesq \citep[e.g.,][]{jeffreys1970}
-and Cerruti \citep[e.g.,][]{love1927} solutions here.
-Note that in these solutions, $\lambda$ and $\mu$ are re-defined, and
-are the first and second Lame's parameters, respectively, instead of
-rake and fault friction as in the body of the manuscript.
-
-\subsection{Boussinesq's solutions for vertical point-source
-loads}\label{boussinesqs-solutions-for-vertical-point-source-loads}
-
-\begin{equation}
-\begin{split}
-G_{xx}^B & = \frac{ F_v }{ 2\pi } \left[ \frac{ 3x^2 }{ r^5 } \right.
-+ \frac{\mu (y^2 + z^2)}{(\lambda + \mu) r^3 (z + r)}
-- \frac{\mu z}{(\lambda + \mu) r^3} \\
-&\qquad \quad \left. - \frac{\mu x^2}{ (\lambda + \mu) r^2 (z + r)^2 }\right ]
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-G_{yy}^B & = \frac{F_v}{2\pi } \left [ \frac{3y^2}{r^5} \right.
-+ \frac{\mu (x^{2} + z^{2})}{(\lambda + \mu) r^{3}(z + r)}
-- \frac{\mu z}{(\lambda + \mu) r ^{3}} \\
-& \qquad \quad \left. - \frac{\mu y^{2}}{(\lambda + \mu ) r^2 (z +r)^2} \right]
-\end{split}
-\end{equation}
-
-\begin{equation}
-G_{xy}^{B} = \frac{F _{v}}{2\pi} \left[ \frac{3xyz}{r^{5}}
-- \frac{\mu x y (z + 2r)}{(\lambda + \mu) r^{3} (z + r)^{2}} \right]
-\end{equation}
-
-\begin{equation}
-G_{zz}^{B} = 3 F _{v} z^{3} / 2 \pi r^{5}
-\end{equation}
-
-\begin{equation}
-G_{xz}^{B} = 3 F _{v} xz^{2} / 2 \pi r^{5}
-\end{equation}
-
-\begin{equation}
-G_{yz}^{B} = 3 F _{v} yz^{2} / 2 \pi r^{5}
-\end{equation}
-
-\subsection{Cerruti's solutions for horizontal point-source
-loads}\label{cerrutis-solutions-for-horizontal-point-source-loads}
-
-\begin{equation}
-\begin{split}
-G_{xx}^{C_x} &= \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3x^2}{r^2} \right.
-- \frac{\mu}{(\lambda + \mu)(z+r)^2} 
-\\
-&\qquad \quad  \left. (r^2 - y^2 - \frac{2ry^2}{r+z}) \right]
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-G_{yy}^{C_x} & = \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3y^2}{r^2} \right.
-- \frac{\mu}{(\lambda + \mu)(z+r)^2} \\
-& \qquad \quad \left. (3r^2 - x^2 - \frac{2rx^2}{r+z}) \right]
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-G_{xy}^{C_x} & = \frac{ F_{h,x} x }{2 \pi r^3} \left[ \frac{ 3x^2}{r^2} \right.
-- \frac{\mu}{(\lambda + \mu)(z+r)^2} 
-\\
-&\qquad \quad \left. \cdot (r^2 - x^2 - \frac{2rx^2}{r+z}) \right]
-\end{split}
-\end{equation}
-
-\begin{equation}
-    G^{C_x}_{zz} = \frac{ 3 F_{h,x} x z^2 }{2 \pi r^5}
-\end{equation}
-
-\begin{equation}
-    G^{C_x}_{xz} = \frac{ 3 F_{h,x} z x^2 }{2 \pi r^5}
-\end{equation}
-
-\begin{equation}
-    G^{C_x}_{yz} = \frac{ 3 F_{h,x} x y z }{2 \pi r^5}
-\end{equation}
-
-\begin{acknowledgements}
-Joe Kington wrote the code to plot Figure \ref{fig:joint_posteriors}c.
-\end{acknowledgements}
-%\section{References}\label{references}
-
-%\bibliography{wench_jgr_bib}{}
-%\bibliographystyle{agufull08}
-
-\begin{thebibliography}{89}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
-
-\bibitem[{\textit{Aagaard et~al.}(2001)\textit{Aagaard, Heaton, and
-  Hall}}]{aagaard2001}
-Aagaard, B.~T., T.~H. Heaton, and J.~F. Hall (2001), Dynamic earthquake
-  ruptures in the presence of lithostatic normal stresses: {I}mplications for
-  friction models and heat production, \textit{Bulletin of the Seismological
-  Society of America}, \textit{91}(6), 1765--1796.
-
-\bibitem[{\textit{Allmann and Shearer}(2009)}]{allmann2009}
-Allmann, B.~P., and P.~M. Shearer (2009), Global variations of stress drop for
-  moderate to large earthquakes, \textit{Journal of Geophysical Research: Solid
-  Earth (1978--2012)}, \textit{114}(B1).
-
-\bibitem[{\textit{Angelier}(1994)}]{angelier1994}
-Angelier, J. (1994), Fault slip analysis and paleostress reconstruction,
-  \textit{Continental deformation}, \textit{4}.
-
-\bibitem[{\textit{Armijo et~al.}(1986)\textit{Armijo, Tapponnier, Mercier, and
-  Han}}]{armijo1986}
-Armijo, R., P.~Tapponnier, J.~Mercier, and T.-L. Han (1986), Quaternary
-  extension in southern {T}ibet: {F}ield observations and tectonic
-  implications, \textit{Journal of Geophysical Research: Solid Earth
-  (1978--2012)}, \textit{91}(B14), 13,803--13,872.
-
-\bibitem[{\textit{Beaumont et~al.}(2001)\textit{Beaumont, Jamieson, Nguyen, and
-  Lee}}]{beaumont2001}
-Beaumont, C., R.~A. Jamieson, M.~Nguyen, and B.~Lee (2001), Himalayan tectonics
-  explained by extrusion of a low-viscosity crustal channel coupled to focused
-  surface denudation, \textit{Nature}, \textit{414}(6865), 738--742.
-
-\bibitem[{\textit{Bird and Kong}(1994)}]{bird1994}
-Bird, P., and X.~Kong (1994), Computer simulations of {C}alifornia tectonics
-  confirm very low strength of major faults, \textit{Geological Society of
-  America Bulletin}, \textit{106}(2), 159--174.
-
-\bibitem[{\textit{Bird and Piper}(1980)}]{birdpiper80}
-Bird, P., and K.~Piper (1980), Plane-stress finite-element models of tectonic
-  flow in southern {C}alifornia, \textit{Physics of the earth and planetary
-  interiors}, \textit{21}(2), 158--175.
-
-\bibitem[{\textit{Bollinger et~al.}(2004)\textit{Bollinger, Avouac, Cattin, and
-  Pandey}}]{bollinger2004}
-Bollinger, L., J.~Avouac, R.~Cattin, and M.~Pandey (2004), Stress buildup in
-  the {H}imalaya, \textit{Journal of Geophysical Research: Solid Earth},
-  \textit{109}(B11).
-
-\bibitem[{\textit{Bott}(1959)}]{bott1959}
-Bott, M. H.~P. (1959), The mechanics of oblique slip faulting,
-  \textit{Geological Magazine}, \textit{96}(02), 109--117.
-
-\bibitem[{\textit{Burchfiel et~al.}(1995)\textit{Burchfiel, Zhiliang, Yupinc,
-  and Royden}}]{burchfiel1995}
-Burchfiel, B., C.~Zhiliang, L.~Yupinc, and L.~Royden (1995), Tectonics of the
-  {L}ongmen {S}han and adjacent regions, central {}china, \textit{International
-  Geology Review}, \textit{37}(8), 661--735.
-
-\bibitem[{\textit{Burchfiel et~al.}(2008)\textit{Burchfiel, Royden, van~der
-  Hilst, Hager, Chen, King, Li, L{\"u}, Yao, and Kirby}}]{burchfiel2008}
-Burchfiel, B., L.~Royden, R.~van~der Hilst, B.~Hager, Z.~Chen, R.~King, C.~Li,
-  J.~L{\"u}, H.~Yao, and E.~Kirby (2008), A geological and geophysical context
-  for the {W}enchuan earthquake of 12 {M}ay 2008, {S}ichuan, {P}eople's
-  {R}epublic of {C}hina, \textit{GSA Today}, \textit{18}(7), 4--11.
-
-\bibitem[{\textit{Byerlee}(1978)}]{byerlee1978}
-Byerlee, J. (1978), Friction of rocks, \textit{Pure and applied Geophysics},
-  \textit{116}(4-5), 615--626.
-
-\bibitem[{\textit{Byrne et~al.}(2013)\textit{Byrne, Holohan, Kervyn, de~Vries,
-  Troll, and Murray}}]{byrne2013}
-Byrne, P., E.~Holohan, M.~Kervyn, B.~v.~W. de~Vries, V.~R. Troll, and J.~Murray
-  (2013), A sagging-spreading continuum of large volcano structure,
-  \textit{Geology}, \textit{41}(3), 339--342.
-
-\bibitem[{\textit{Cattin et~al.}(1997)\textit{Cattin, Lyon-Caen, and
-  Ch{\'e}ry}}]{cattin1997}
-Cattin, R., H.~Lyon-Caen, and J.~Ch{\'e}ry (1997), Quantification of interplate
-  coupling in subduction zones and forearc topography, \textit{Geophysical
-  Research Letters}, \textit{24}(13), 1563--1566.
-
-\bibitem[{\textit{Chen et~al.}(2010)\textit{Chen, Niu, Liu, Huang,
-  Tkal{\v{c}}i{\'c}, Sun, and Chan}}]{chen2010receiverfn}
-Chen, Y., F. Niu, R. Liu, Z. Huang, H. Tkal{\v{c}}i{\'c}, L. Sun, and
-  W. Chan (2010), Crustal structure beneath {C}hina from receiver function
-  analysis, \textit{Journal of Geophysical Research: Solid Earth}, 
-  \textit{115}(B03307), 1--22, doi:10.1029/2009JB006386.
-
-\bibitem[{\textit{Clark and Royden}(2000)}]{clarkroyden2000}
-Clark, M.~K., and L.~H. Royden (2000), Topographic ooze: {B}uilding the eastern
-  margin of {T}ibet by lower crustal flow, \textit{Geology}, \textit{28}(8),
-  703--706.
-
-\bibitem[{\textit{Clark et~al.}(2005)\textit{Clark, Bush, and
-  Royden}}]{clark2005}
-Clark, M.~K., J.~W. Bush, and L.~H. Royden (2005), Dynamic topography produced
-  by lower crustal flow against rheological strength heterogeneities bordering
-  the {T}ibetan {P}lateau, \textit{Geophysical Journal International},
-  \textit{162}(2), 575--590.
-
-\bibitem[{\textit{Coblentz and Richardson}(1996)}]{coblentz1996}
-Coblentz, D.~D., and R.~M. Richardson (1996), Analysis of the {S}outh
-  {A}merican intraplate stress field, \textit{Journal of Geophysical Research:
-  Solid Earth}, \textit{101}(B4), 8643--8657.
-
-\bibitem[{\textit{Cook and Royden}(2008)}]{cook2008}
-Cook, K.~L., and L.~H. Royden (2008), The role of crustal strength variations
-  in shaping orogenic plateaus, with application to {T}ibet, \textit{Journal of
-  Geophysical Research: Solid Earth (1978--2012)}, \textit{113}(B8).
-
-\bibitem[{\textit{Copley and McKenzie}(2007)}]{copleymckenzie2007}
-Copley, A., and D.~McKenzie (2007), Models of crustal flow in the
-  {I}ndia--{A}sia collision zone, \textit{Geophysical Journal International},
-  \textit{169}(2), 683--698.
-
-\bibitem[{\textit{Copley et~al.}(2009)\textit{Copley, Boait, Hollingsworth,
-  Jackson, and McKenzie}}]{copley2009}
-Copley, A., F.~Boait, J.~Hollingsworth, J.~Jackson, and D.~McKenzie (2009),
-  Subparallel thrust and normal faulting in {A}lbania and the roles of
-  gravitational potential energy and rheology contrasts in mountain belts,
-  \textit{Journal of Geophysical Research: Solid Earth}, \textit{114}(B5).
-
-\bibitem[{\textit{Cui et~al.}(2014)\textit{Cui, Lin, Wang, Gao, Huang, Wang,
-  Sun, Li, Zhou, Qian et~al.}}]{cui2014}
-Cui, J., W.~Lin, L.~Wang, L.~Gao, Y.~Huang, W.~Wang, D.~Sun, Z.~Li, C.~Zhou,
-  H.~Qian, et~al. (2014), Determination of three-dimensional in situ stresses
-  by anelastic strain recovery in {W}enchuan {E}arthquake {F}ault {S}cientific
-  {D}rilling {P}roject {H}ole-1 ({W}{F}{S}{D}-1), \textit{Tectonophysics},
-  \textit{619–620}, 123 -- 132.
-
-\bibitem[{\textit{Dahlen}(1990)}]{dahlen1990}
-Dahlen, F. (1990), Critical taper model of fold-and-thrust belts and
-  accretionary wedges, \textit{Annual Review of Earth and Planetary Sciences},
-  \textit{18}, 55.
-
-\bibitem[{\textit{Dalmayrac and Molnar}(1981)}]{dalmayrac1981}
-Dalmayrac, B., and P.~Molnar (1981), Parallel thrust and normal faulting in
-  {P}eru and constraints on the state of stress, \textit{Earth and Planetary
-  Science Letters}, \textit{55}(3), 473--481.
-
-\bibitem[{\textit{Day}(1982)}]{day1982}
-Day, S.~M. (1982), Three-dimensional simulation of spontaneous rupture: {T}he
-  effect of nonuniform prestress, \textit{Bulletin of the Seismological Society
-  of America}, \textit{72}(6A), 1881--1902.
-
-\bibitem[{\textit{DeCelles et~al.}(2009)\textit{DeCelles, Ducea, Kapp, and
-  Zandt}}]{decelles2009}
-DeCelles, P.~G., M.~N. Ducea, P.~Kapp, and G.~Zandt (2009), Cyclicity in
-  cordilleran orogenic systems, \textit{Nature Geoscience}, \textit{2}(4),
-  251--257.
-
-\bibitem[{\textit{Dewey}(1988)}]{dewey1988}
-Dewey, J.~F. (1988), Extensional collapse of orogens, \textit{Tectonics},
-  \textit{7}(6), 1123--1139.
-
-\bibitem[{\textit{England and Houseman}(1989)}]{england1989}
-England, P., and G.~Houseman (1989), Extension during continental convergence,
-  with application to the tibetan plateau, \textit{Journal of Geophysical
-  Research: Solid Earth (1978--2012)}, \textit{94}(B12), 17,561--17,579.
-
-\bibitem[{\textit{Farr et~al.}(2007)\textit{Farr, Rosen, Caro, Crippen, Duren,
-  Hensley, Kobrick, Paller, Rodriguez, Roth et~al.}}]{farr2007srtm}
-Farr, T.~G., P.~A. Rosen, E.~Caro, R.~Crippen, R.~Duren, S.~Hensley,
-  M.~Kobrick, M.~Paller, E.~Rodriguez, L.~Roth, et~al. (2007), The shuttle
-  radar topography mission, \textit{Reviews of Geophysics}, \textit{45}(2).
-
-\bibitem[{\textit{Feng et~al.}(2010)\textit{Feng, Hetland, Ding, Li, and
-  Zhang}}]{feng2010}
-Feng, G., E.~A. Hetland, X.~Ding, Z.~Li, and L.~Zhang (2010), Coseismic fault
-  slip of the 2008 {M}w 7.9 {W}enchuan earthquake estimated from {I}n{S}{A}{R}
-  and {G}{P}{S} measurements, \textit{Geophysical Research Letters},
-  \textit{37}(1).
-
-\bibitem[{\textit{Fialko et~al.}(2005)\textit{Fialko, Riveral, and Kanamori}}]
-    {fialko2005}
-Fialko, Y., L.~Rivera, and H.~Kanamori (2005), Estimate of differential stress
-  in the upper crust from variations in topography and strike along the San
-  Andreas fault, \textit{Geophysical Journal International}, \textit{160},
-  527--532.
-
-\bibitem[{\textit{Fielding et~al.}(2013)\textit{Fielding, Sladen, Li, Avouac,
-  B{\"u}rgmann, and Ryder}}]{fielding2013}
-Fielding, E.~J., A.~Sladen, Z.~Li, J.-P. Avouac, R.~B{\"u}rgmann, and I.~Ryder
-  (2013), Kinematic fault slip evolution source models of the 2008 m7.9
-  wenchuan earthquake in china from sar interferometry, gps and teleseismic
-  analysis and implications for longmen shan tectonics, \textit{Geophysical
-  Journal International}, \textit{194}(2), 1138--1166.
-
-\bibitem[{\textit{Flesch and Bendick}(2012)}]{fleschbendick2012}
-Flesch, L., and R.~Bendick (2012), The relationship between surface kinematics
-  and deformation of the whole lithosphere, \textit{Geology}, \textit{40}(8),
-  711--714.
-
-\bibitem[{\textit{Flesch and Kreemer}(2010)}]{flesch2010gpe}
-Flesch, L.~M., and C.~Kreemer (2010), Gravitational potential energy and
-  regional stress and strain rate fields for continental plateaus: {E}xamples
-  from the central {A}ndes and {C}olorado {P}lateau, \textit{Tectonophysics},
-  \textit{482}(1), 182--192.
-
-\bibitem[{\textit{Heidbach et~al.}(2009)\textit{Heidbach, Tingay, and
-  Barth}}]{heidbach2009}
-Heidbach, O., M.~Tingay, and A.~Barth (2009), \textit{The World Stress Map
-  based on the database release 2008, equatorial scale 1:46,000,000},
-  Commission for the Geological Map of the World, Paris.
-
-\bibitem[{\textit{Hubbard et~al.}(2010)\textit{Hubbard, Shaw, and
-  Klinger}}]{hubbard2010}
-Hubbard, J., J.~H. Shaw, and Y.~Klinger (2010), Structural setting of the 2008
-  {M}w 7.9 {W}enchuan, {C}hina, earthquake, \textit{Bulletin of the
-  Seismological Society of America}, \textit{100}(5B), 2713--2735.
-
-\bibitem[{\textit{Jarvis et~al.}(2008)\textit{Jarvis, Reuter, Nelson, and
-  Guevara}}]{jarvis2008srtm}
-Jarvis, A., H.~Reuter, A.~Nelson, and E.~Guevara (2008), Hole-filled srtm for
-  the globe version 3, available from the cgiar-csi srtm 90m database,
-  \textit{Accessed: March}, \textit{12}, 2012.
-
-\bibitem[{\textit{Jeffreys}(1924)}]{jeffreys1924}
-Jeffreys, H. (1924), \textit{The Earth: its origin, history and physical
-  constitution}, Cambridge University Press.
-
-\bibitem[{\textit{Jeffreys}(1970)}]{jeffreys1970}
-Jeffreys, H. (1970), \textit{The Earth}, Cambridge University Press.
-
-\bibitem[{\textit{Kanamori and Anderson}(1975)}]{kanamori1975}
-Kanamori, H., and D.~L. Anderson (1975), Theoretical basis of some empirical
-  relations in seismology, \textit{Bulletin of the Seismological Society of
-  America}, \textit{65}(5), 1073--1095.
-
-\bibitem[{\textit{Kuo et~al.}(2014)\textit{Kuo, Li, Smith, Di~Toro, Suppe,
-  Song, Nielsen, Sheu, and Si}}]{kuo2014}
-Kuo, L.-W., H.~Li, S.~A. Smith, G.~Di~Toro, J.~Suppe, S.-R. Song, S.~Nielsen,
-  H.-S. Sheu, and J.~Si (2014), Gouge graphitization and dynamic fault
-  weakening during the 2008 {M}w 7.9 {W}enchuan earthquake, \textit{Geology},
-  \textit{42}(1), 47--50.
-
-\bibitem[{\textit{Lamb}(2006)}]{lamb2006}
-Lamb, S. (2006), Shear stresses on megathrusts: {I}mplications for mountain
-  building behind subduction zones, \textit{Journal of Geophysical Research:
-  Solid Earth}, \textit{111}(B7).
-
-\bibitem[{\textit{Liang et~al.}(2013)\textit{Liang, Gan, Shen, Xiao, Liu, Chen,
-  Ding, and Zhou}}]{liang2013}
-Liang, S., W.~Gan, C.~Shen, G.~Xiao, J.~Liu, W.~Chen, X.~Ding, and D.~Zhou
-  (2013), Three-dimensional velocity field of present-day crustal motion of the
-  {T}ibetan {P}lateau derived from {G}{P}{S} measurements, \textit{Journal of
-  Geophysical Research: Solid Earth}, \textit{118}(10), 5722--5732.
-
-\bibitem[{\textit{Lin et~al.}(2009)\textit{Lin, Ren, Jia, and Wu}}]{lin2009}
-Lin, A., Z.~Ren, D.~Jia, and X.~Wu (2009), Co-seismic thrusting rupture and
-slip distribution produced by the 2008 {M}_w 7.9 {W}enchuan
-earthquake, {C}hina, \textit{Tectonophysics}, \textit{471}(3), 203--215.
-
-\bibitem[{\textit{Lisle}(2013)}]{lisle2013}
-Lisle, R.~J. (2013), A critical look at the {W}allace-{B}ott hypothesis in
-  fault-slip analysis, \textit{Bulletin de la Societe Geologique de France},
-  \textit{184}(4-5), 299--306.
-
-\bibitem[{\textit{Liu and Zoback}(1992)}]{liuzoback1992}
-Liu, L., and M.~D. Zoback (1992), The effect of topography on the state of
-  stress in the crust: {A}pplication to the site of the {C}ajon {P}ass
-  {S}cientific {D}rilling {P}roject, \textit{Journal of Geophysical Research:
-  Solid Earth (1978--2012)}, \textit{97}(B4), 5095--5108.
-
-\bibitem[{\textit{Liu and Yang}(2003)}]{liuyang2003}
-Liu, M., and Y.~Yang (2003), Extensional collapse of the {T}ibetan {P}lateau:
-  {R}esults of three-dimensional finite element modeling, \textit{Journal of
-  Geophysical Research: Solid Earth (1978--2012)}, \textit{108}(B8).
-
-\bibitem[{\textit{Liu-Zeng et~al.}(2009)\textit{Liu-Zeng, Zhang, Wen,
-  Tapponnier, Sun, Xing, Hu, Xu, Zeng, Ding et~al.}}]{liu2009}
-Liu-Zeng, J., Z.~Zhang, L.~Wen, P.~Tapponnier, J.~Sun, X.~Xing, G.~Hu, Q.~Xu,
-  L.~Zeng, L.~Ding, et~al. (2009), Co-seismic ruptures of the 12 {M}ay 2008, 
-  {M}_w 8.0 {W}enchuan earthquake, {s}ichuan: {E}ast--{w}est crustal
-  shortening on oblique, parallel thrusts along the eastern edge of {T}ibet,
-  \textit{Earth and Planetary Science Letters}, \textit{286}(3), 355--370.
-
-\bibitem[{\textit{Love}(1927)}]{love1927}
-Love, E.~A.~H. (1927), \textit{The mathematical theory of elasticity.},
-  Cambridge University Press.
-
-\bibitem[{\textit{Loveless and Meade}(2011)}]{loveless2011}
-Loveless, J., and B.~Meade (2011), Partitioning of localized and diffuse
-  deformation in the {T}ibetan {P}lateau from joint inversions of geologic and
-  geodetic observations, \textit{Earth and Planetary Science Letters},
-  \textit{303}(1), 11--24.
-
-\bibitem[{\textit{Luttrell et~al.}(2007)\textit{Luttrell, Sandwell,
-  Smith-Konter, Bills, and Bock}}]{luttrell2007}
-Luttrell, K., D.~Sandwell, B.~Smith-Konter, B.~Bills, and Y.~Bock (2007),
-  Modulation of the earthquake cycle at the southern {S}an {A}ndreas fault by
-  lake loading, \textit{Journal of Geophysical Research: Solid Earth
-  (1978--2012)}, \textit{112}(B8).
-
-\bibitem[{\textit{Luttrell et~al.}(2011)\textit{Luttrell, Tong, Sandwell,
-  Brooks, and Bevis}}]{luttrell2011}
-Luttrell, K.~M., X.~Tong, D.~T. Sandwell, B.~A. Brooks, and M.~G. Bevis (2011),
-  Estimates of stress drop and crustal tectonic stress from the 27 {F}ebruary
-  2010 {M}aule, {C}hile, earthquake: Implications for fault strength,
-  \textit{Journal of Geophysical Research: Solid Earth}, \textit{116}(B11).
-
-\bibitem[{\textit{Martel}(2006)}]{martel2006}
-Martel, S.~J. (2006), Effect of topographic curvature on near-surface stresses
-  and application to sheeting joints, \textit{Geophysical Research Letters},
-  \textit{33}(1).
-
-\bibitem[{\textit{McKenzie}(1969)}]{mckenzie1969}
-McKenzie, D.~P. (1969), The relation between fault plane solutions for
-  earthquakes and the directions of the principal stresses, \textit{Bulletin of
-  the Seismological Society of America}, \textit{59}(2), 591--601.
-
-\bibitem[{\textit{McKinney}(2010)}]{mckinney2010}
-McKinney, W. (2010), Data structures for statistical computing in python, in
-  \textit{Proceedings of the 9th Python in Science Conference}, edited by
-  S.~van~der Walt and J.~Millman, pp. 51 -- 56, SciPy.
-
-\bibitem[{\textit{McTigue and Mei}(1981)}]{mctiguemei1981}
-McTigue, D.~F., and C.~C. Mei (1981), Gravity-induced stresses near topography
-  of small slope, \textit{Journal of Geophysical Research: Solid Earth},
-  \textit{86}(B10), 9268--9278.
-
-\bibitem[{\textit{Meade and Conrad}(2008)}]{meade2008}
-Meade, B.~J., and C.~P. Conrad (2008), Andean growth and the deceleration of
-  {S}outh {A}merican subduction: {T}ime evolution of a coupled
-  orogen-subduction system, \textit{Earth and Planetary Science Letters},
-  \textit{275}(1), 93--101.
-
-\bibitem[{\textit{Medina~Luna and Hetland}(2013)}]{medinaluna2013}
-Medina~Luna, L., and E.~A. Hetland (2013), Regional stresses inferred from
-  coseismic slip models of the 2008 {M}w 7.9 {W}enchuan, {C}hina, earthquake,
-  \textit{Tectonophysics}, \textit{584}, 43--53.
-
-\bibitem[{\textit{Meissner and Strehlau}(1982)}]{meissner1982}
-Meissner, R., and J.~Strehlau (1982), Limits of stresses in continental crusts
-  and their relation to the depth-frequency distribution of shallow
-  earthquakes, \textit{Tectonics}, \textit{1}(1), 73--89.
-
-\bibitem[{\textit{Michael}(1987)}]{michael1987}
-Michael, A.~J. (1987), Use of focal mechanisms to determine stress: a control
-  study, \textit{Journal of Geophysical Research: Solid Earth (1978--2012)},
-  \textit{92}(B1), 357--368.
-
-\bibitem[{\textit{Miller and Dunne}(1996)}]{millerdunne1996}
-Miller, D.~J., and T.~Dunne (1996), Topographic perturbations of regional
-  stresses and consequent bedrock fracturing, \textit{Journal of Geophysical
-  Research: Solid Earth (1978--2012)}, \textit{101}(B11), 25,523--25,536.
-
-\bibitem[{\textit{Molnar and Lyon-Caen}(1988)}]{molnar1988}
-Molnar, P., and H.~Lyon-Caen (1988), Some simple physical aspects of the
-  support, structure, and evolution of mountain belts, \textit{Spec. Pap. Geol.
-  Soc. Am}, \textit{218}, 179--207.
-
-\bibitem[{\textit{Mosegaard and Tarantola}(1995)}]{mosegaard1995}
-Mosegaard, K., and A.~Tarantola (1995), Monte {C}arlo sampling of solutions to
-  inverse problems, \textit{Journal of Geophysical Research: Solid Earth
-  (1978--2012)}, \textit{100}(B7), 12,431--12,447.
-
-\bibitem[{\textit{Nakamura et~al.}(2010)\textit{Nakamura, Tsuboi, Kaneda, and
-  Yamanaka}}]{nakamura2010}
-Nakamura, T., S.~Tsuboi, Y.~Kaneda, and Y.~Yamanaka (2010), Rupture process of
-  the 2008 {W}enchuan, {C}hina earthquake inferred from teleseismic waveform
-  inversion and forward modeling of broadband seismic waves,
-  \textit{Tectonophysics}, \textit{491}(1), 72--84.
-
-\bibitem[{\textit{Oglesby and Day}(2002)}]{oglesbyday2002}
-Oglesby, D.~D., and S.~M. Day (2002), Stochastic fault stress: {I}mplications
-  for fault dynamics and ground motion, \textit{Bulletin of the Seismological
-  Society of America}, \textit{92}(8), 3006--3021.
-
-\bibitem[{\textit{Oliphant}(2007)}]{oliphant2007numpy}
-Oliphant, T.~E. (2007), Python for scientific computing, \textit{Computing in
-  Science \& Engineering}, \textit{9}(3), 10--20.
-
-\bibitem[{\textit{Olsen et~al.}(1997)\textit{Olsen, Madariaga, and
-  Archuleta}}]{olsen1997}
-Olsen, K., R.~Madariaga, and R.~Archuleta (1997), Three-dimensional dynamic
-  simulation of the 1992 {L}anders earthquake, \textit{Science},
-  \textit{278}(5339), 834--838.
-
-\bibitem[{\textit{P\'erez and Granger}(2007)}]{perez2007ipython}
-P\'erez, F., and B.~E. Granger (2007), {IP}ython: a {S}ystem for {I}nteractive
-  {S}cientific {C}omputing, \textit{{C}omput. {S}ci. {E}ng.}, \textit{9}(3),
-  21--29.
-
-\bibitem[{\textit{Qi et~al.}(2011)\textit{Qi, Xuejun, Qigui, Freymueller,
-  Shaomin, Caijun, Yonglin, Xinzhao, Kai, and Gang}}]{qi2011}
-Qi, W., Q.~Xuejun, L.~Qigui, J.~Freymueller, Y.~Shaomin, X.~Caijun, Y.~Yonglin,
-  Y.~Xinzhao, T.~Kai, and C.~Gang (2011), Rupture of deep faults in the 2008
-  {W}enchuan earthquake and uplift of the {L}ongmen {S}han, \textit{Nature
-  Geoscience}, \textit{4}(9), 634--640.
-
-\bibitem[{\textit{Reches}(1987)}]{reches1987}
-Reches, Z. (1987), Determination of the tectonic stress tensor from slip along
-  faults that obey the coulomb yield condition, \textit{Tectonics},
-  \textit{6}(6), 849--861.
-
-\bibitem[{\textit{Rey et~al.}(2001)\textit{Rey, Vanderhaeghe, and
-  Teyssier}}]{rey2001}
-Rey, P., O.~Vanderhaeghe, and C.~Teyssier (2001), Gravitational collapse of the
-  continental crust: definition, regimes and modes, \textit{Tectonophysics},
-  \textit{342}(3), 435--449.
-
-\bibitem[{\textit{Richardson and Coblentz}(1994)}]{richardson1994}
-Richardson, R.~M., and D.~D. Coblentz (1994), Stress modeling in the {A}ndes:
-  Constraints on the {S}outh {A}merican intraplate stress magnitudes,
-  \textit{Journal of Geophysical Research: Solid Earth}, \textit{99}(B11),
-  22,015--22,025.
-
-\bibitem[{\textit{Savage and Swolfs}(1986)}]{savageswolfs1986}
-Savage, W.~Z., and H.~S. Swolfs (1986), Tectonic and gravitational stress in
-  long symmetric ridges and valleys, \textit{Journal of Geophysical Research:
-  Solid Earth}, \textit{91}(B3), 3677--3685.
-
-\bibitem[{\textit{Scholz}(2002)}]{scholz2002}
-Scholz, C.~H. (2002), \textit{The mechanics of earthquakes and faulting},
-  Cambridge university press.
-
-\bibitem[{\textit{Seabold and Perktold}(2010)}]{seabold2010}
-Seabold, S., and J.~Perktold (2010), Statsmodels: econometric and statistical
-  modeling with python, in \textit{Proceedings of the 9th Python in Science
-  Conference}, pp. 57--61, SciPy.
-
-\bibitem[{\textit{Shen et~al.}(2009)\textit{Shen, Sun, Zhang, Wan, Wang,
-  B{\"u}rgmann, Zeng, Gan, Liao, and Wang}}]{shen2009}
-Shen, Z.-K., J.~Sun, P.~Zhang, Y.~Wan, M.~Wang, R.~B{\"u}rgmann, Y.~Zeng,
-  W.~Gan, H.~Liao, and Q.~Wang (2009), Slip maxima at fault junctions and
-  rupturing of barriers during the 2008 {W}enchuan earthquake, \textit{Nature
-  Geoscience}, \textit{2}(10), 718--724.
-
-\bibitem[{\textit{Sibson}(1985)}]{sibson1985}
-Sibson, R.~H. (1985), A note on fault reactivation, \textit{Journal of
-  Structural Geology}, \textit{7}(6), 751--754.
-
-\bibitem[{\textit{Styron et~al.}(2010)\textit{Styron, Taylor, and
-  Okoronkwo}}]{styron2010}
-Styron, R., M.~Taylor, and K.~Okoronkwo (2010), Database of active structures
-  from the {I}ndo-{A}sian collision, \textit{Eos, Transactions American
-  Geophysical Union}, \textit{91}(20), 181--182.
-
-\bibitem[{\textit{Tarantola}(2005)}]{tarantola2005}
-Tarantola, A. (2005), \textit{Inverse problem theory and methods for model
-  parameter estimation}, SIAM.
-
-\bibitem[{\textit{Taylor et~al.}(2003)\textit{Taylor, Yin, Ryerson, Kapp, and
-  Ding}}]{taylor2003}
-Taylor, M., A.~Yin, F.~J. Ryerson, P.~Kapp, and L.~Ding (2003), Conjugate
-  strike-slip faulting along the {B}angong-{N}ujiang suture zone accommodates
-  coeval east-west extension and north-south shortening in the interior of the
-  {T}ibetan {P}lateau, \textit{Tectonics}, \textit{22}(4).
-
-\bibitem[{\textit{Tong et~al.}(2010)\textit{Tong, Sandwell, and
-  Fialko}}]{tong2010}
-Tong, X., D.~T. Sandwell, and Y.~Fialko (2010), Coseismic slip model of the
-  2008 {W}enchuan earthquake derived from joint inversion of interferometric
-  synthetic aperture radar, {G}{P}{S}, and field data, \textit{Journal of
-  Geophysical Research: Solid Earth}, \textit{115}(B4).
-
-\bibitem[{\textit{Townend and Zoback}(2000)}]{townend2000}
-Townend, J., and M.~D. Zoback (2000), How faulting keeps the crust strong,
-  \textit{Geology}, \textit{28}(5), 399--402.
-
-\bibitem[{\textit{Wallace}(1951)}]{wallace1951}
-Wallace, R.~E. (1951), Geometry of shearing stress and relation to faulting,
-  \textit{The Journal of Geology}, pp. 118--130.
-
-\bibitem[{\textit{Wang et~al.}(2012)\textit{Wang, Kirby, Furlong, Van~Soest,
-  Xu, Shi, Kamp, and Hodges}}]{wang2012}
-Wang, E., E.~Kirby, K.~P. Furlong, M.~Van~Soest, G.~Xu, X.~Shi, P.~J. Kamp, and
-  K.~Hodges (2012), Two-phase growth of high topography in eastern {T}ibet
-  during the {C}enozoic, \textit{Nature Geoscience}, \textit{5}(9), 640--645.
-
-\bibitem[{\textit{Xiao et~al.}(1991)\textit{Xiao, Dahlen, and
-  Suppe}}]{xiao1991}
-Xiao, H.-B., F.~Dahlen, and J.~Suppe (1991), Mechanics of extensional wedges,
-  \textit{Journal of Geophysical Research: Solid Earth}, \textit{96}(B6),
-  10,301--10,318.
-
-\bibitem[{\textit{Xu et~al.}(2009)\textit{Xu, Wen, Yu, Chen, Klinger, Hubbard,
-  and Shaw}}]{xu2009}
-Xu, X., X.~Wen, G.~Yu, G.~Chen, Y.~Klinger, J.~Hubbard, and J.~Shaw (2009),
-  Coseismic reverse-and oblique-slip surface faulting generated by the 2008 mw
-  7.9 wenchuan earthquake, china, \textit{Geology}, \textit{37}(6), 515--518.
-
-\bibitem[{\textit{Yong et~al.}(2003)\textit{Yong, Allen, Densmore, and
-  Qiang}}]{yong2003}
-Yong, L., P.~A. Allen, A.~L. Densmore, and X.~Qiang (2003), Evolution of the
-  {L}ongmen {S}han foreland basin (western {S}ichuan, {C}hina) during the
-  {L}ate {T}riassic {I}ndosinian orogeny, \textit{Basin Research},
-  \textit{15}(1), 117--138.
-
-\bibitem[{\textit{Zhang et~al.}(2011)\textit{Zhang, Qu, Shan, Song, Zhang,
-  Wang, Hu, and Wang}}]{zhang2011}
-Zhang, G., C.~Qu, X.~Shan, X.~Song, G.~Zhang, C.~Wang, J.-C. Hu, and R.~Wang
-  (2011), Slip distribution of the 2008 {W}enchuan {M}s 7.9 earthquake by joint
-  inversion from {G}{P}{S} and {I}n{S}{A}{R} measurements: a resolution test
-  study, \textit{Geophysical Journal International}, \textit{186}(1), 207--220.
-
-\bibitem[{\textit{Zhang et~al.}(2010)\textit{Zhang, Wen, Shen, and
-  Chen}}]{zhang2010}
-Zhang, P.-Z., X.-z. Wen, Z.-K. Shen, and J.-h. Chen (2010), Oblique,
-  high-angle, listric-reverse faulting and associated development of strain:
-  The {W}enchuan earthquake of {M}ay 12, 2008, {S}ichuan, {C}hina,
-  \textit{Annual Review of Earth and Planetary Sciences}, \textit{38},
-  353--382.
-
-\end{thebibliography}
-
-\end{article}
-
 \end{document}

### 2015-02-13 revisions/rebuttal.md
eric's changes

@@
 >It is also apparent in Figure 8b that regardless of the magnitude of
-$T^\_{\mathrm{max}$, $p_{J}(T^\prime_{\mathrm{min}} | D)$ is
+$T^\_{\mathrm{max}}$, $p_{J}(T^\prime_{\mathrm{min}} | D)$ is
 consistently 0--0.6 $p_{J}(T^\prime_{\mathrm{max}} | D)$, with a mode

### 2015-02-13 revisions/rebuttal2.md
eric's changes

@@
 
-We believe that most or all of our readers will know the difference between
-tractions and stresses.  However, it is conventional (if slightly improper)
-to refer to tractions as stresses, as in 'shear stress' and 'normal stress',
-instead of 'shear traction' and 'normal traction'. Whether we are referring
-to tensors, vectors, or scalars should be apparent and cause no confusion.
+We believe that most or all of our readers will know the difference
+between tractions and stresses.  Stresses on faults are commonly
+referred to as 'shear stress' and 'normal stress', instead of 'shear
+traction' and 'normal traction'. One of us (EAH) has used both
+terminologies in past papers. We also note that the influential paper
+by Jim Rice ("Spatio-temporal complexities of slip on a fault", 1993,
+JGR) uses "shear stress" for the stress acting on the fault, a
+terminology that has become common in the "Rice school" of earthquake
+dynamics.  Whether we are referring to tensors, vectors, or scalars
+should be apparent and cause no confusion.
 
@@
 
-The term 'posteriors' has been well established in the literature for decades
-and is unlikely to cause offense.
+The term 'posteriors' has been well established in the literature for
+decades and is unlikely to cause offense. Furthermore, the posteriors
+we present in this paper are not function or distributions, but samples
+of an unknown posterior PDF. We believe that our description of the
+sampling methodology in section 4.2 makes this clear. We also note
+that one of us was quite confused as what the the unintended reaction
+from the term "posterior" might be.
 
@@
 
-We have made some minor modifications to figures.
+We have made some minor modifications to figures to address this issue.
 
@@
 
-We clearly state that our results are predicated upon the Wallace-Bott
-assumption that fault slip occurs in the direction of maximum shear stress.
+We agree that this is important to state and did not realize we had
+not in this manuscript. As stated, our model is  predicated
+upon the Wallace-Bott assumption that fault slip occurs in the
+direction of maximum shear stress, implicitly in that assumption is
+that coseismic slip rake reflects accumulated shear stress on the
+fault prior to earthquake nucleation, and that dynamic stresses do not
+result in a significant re-orientation of coseismic slip. To make this
+more clear, we have slightly modified the relevant sentence in the
+abstract to:
+
+"We estimate the tectonic stress needed to overcome topographic and
+lithostatic stresses by assuming that the direction of maximum shear
+accumulated on the faults is roughly collinear with the inferred
+coseismic slip."
+
+And in section 4.2, where we state the Wallace-Bott assumption, we
+added "prior to the initiation of the earthquake" to make this clear.
 
@@
 We have looked over these references, which are largely about volcanic
-deformation on other planets, and have not found a good place to cite them,
-without making a separate paragraph for discussions of topographic stresses in
-volcanoes. This is far from our topic (the interaction of topographic stresses
-and tectonics), and we did not read these papers during the work presented in
-this paper, so we don't feel it appropriate to cite them.
+deformation on other planets, and have not found a good place to cite
+them, without making a separate paragraph for discussions of
+topographic stresses in volcanoes. This is far from our topic (the
+interaction of topographic stresses and tectonics), and we did not
+rely on these papers while developing the work presented in this
+paper, so we don't feel it necessary to cite them.
 
@@
 
-This reference has been eliminated, with the note that it is unfortunate that
-we put up these sorts of walls in scholarship. The history of ideas shouldn't
-be segregated so, and we did not refer to it 'as one might a peer-reviewed
-scientific article'.
+This reference has been eliminated, although we note that we feel it
+ is unfortunate that these sorts of walls exist in scholarship. The
+ history of ideas shouldn't be segregated so. We felt that the
+ quotation was appropriate for the topic of this manuscript, and we
+ did refer back to it later in the text. We are not sure that if we
+ had chosen an opening quotation from a non-religious text, say James
+ Joyce's Finnegans Wake, whether there would have been such a visceral
+ reaction.  Finally, we did not refer to it 'as one might a
+ peer-reviewed scientific article'. That all being said, we have
+ removed the quotation.
 
@@
 We change the sentence introducing the methods to "They show that the
-topographic stress tensor field *beneath (but not within) topography* can be
-determined by a convolution of topographic loading functions with Green's
-functions describing the stresses in an elastic halfspace due to a point load
-at the surface." (changes in italics). Additionally, in Section 2.2 on the
-numerical implementation of the methods, we mention that our calculations only
-apply to depths below 851 m.
+topographic stress tensor field *beneath (but not within) topography*
+can be determined by a convolution of topographic loading functions
+with Green's functions describing the stresses in an elastic halfspace
+due to a point load at the surface." (changes in
+italics). Additionally, in Section 2.2 on the numerical implementation
+of the methods, we mention that our calculations only apply to depths
+below 851 m below sea level, which we take as our reference surface of
+the half-space. We also now specifically state that z=0 corresponds to
+mean sea level in our description of these calculations in Section
+2.1.
 
@@
 readers need to know this information, they can look at the .csv files on
-GitHub or consult the sources for the slip models.
+GitHub or consult the sources for these published slip models.
 
@@
 
-This varies in the different fault models considered. If the reader wants to
-know these specifics, he or she is welcome to read the papers that have
-produced the fault models used.
+This varies in the different fault models considered. All of the
+coseismic slip models we consider in this paper were developed by
+other researchers (with the exception of Feng et al. (2010), of while
+one of use is a co-author) and have been fully described in
+peer-reviewed published papers. If the reader wants to know these
+specifics, he or she is welcome to read the papers that have produced
+the fault models we use.
 
@@
 
-I do not think I have ever seen $\tau_n$ to refer to normal stress/traction;
-we use $\tau$ for shear stress (tractions), which is quite standard.  We also
-use $\sigma_n$ as normal stress (traction) which is standard. We use *M* to
-denote the topographic stress tensor or tensor field, and it is clearly defined
-as such. Capital letters are extremely commonly used to denote arrays,
-matrices, tensors and tensor fields; this is not only a structural geology
-convention but spans the science, mathematics and engineering disciplines.
+We use $\tau$ for shear stress (tractions), which is quite standard
+and we define our usage in the paper.  We also use $\sigma_n$ as
+normal stress (traction) which is also standard and defined in the
+paper. We use *M* to denote the topographic stress tensor or tensor
+field, and it is clearly defined as such. Capital letters are
+commonly, although not exclusively, used to denote tensors and tensor
+fields, in a variety of sources, spanning structural geology, fluid
+dynamics, mathematics, and engineering disciplines.
 
@@
 
-The colors shown are not simple 'background colors' but correspond to the
-magnitudes of the stress component in question. Where they are uniform,
-stresses are uniform. Labeling the contours would add unneccesary clutter to
-the figures, making them harder to read. Nonetheless, the relevant comparisons
-(for example, between high normal stress and low slip) are still quite evident
-to us, both on the screen and in print, and other viewers have not had problems
-with this. Maybe the reviewer has a monitor or printer calibration issue.
+The colors shown are not simple 'background colors' but correspond to
+the magnitudes of the stress component in question, this is stated in
+the figure caption and the associated color bars in the figure. Where
+they are uniform, stresses are uniform. Labeling the contours would
+add unnecessary	 clutter to the figures, making them harder to
+read. The relevant comparisons (for example, between high normal
+stress and low slip) are still quite evident to us, both on the screen
+and in print, and other viewers have not had problems with this.
 
@@
 
-We believe that we have explained the mathematics involved in our calculations
-(Bayes' rule, the misfit equation, and the likelihood equation, which are
-equations 10-12 in the original submission and equations 11-13 in the revised
-submission). These are uncomplicated expressions. Unfortunately, though
-Bayesian statistics are intuitive in practice and a formalized version of our
-thought processes in many instances (revising ideas or beliefs based on
-evidence), there is some sort of barrier for many people in grasping it
-initially. For the first author, this was due to jargon, so we have tried to
-define and explain each of the terms. We have also cited two excellent texts
-[Tarantola, 2005; Mosegaard and Tarantola, 1995] that explain Bayesian
-statistics in more detail.
+We believe that we have explained the mathematics involved in our
+calculations (Bayes' rule, the misfit equation, and the likelihood
+equation, which are equations 10-12 in the original submission and
+equations 11-13 in the revised submission). These are uncomplicated
+expressions, and we believe that they fully describe our estimation
+strategy within a Bayesian sampling methodology.  We have tried to
+define and explain each of the terms, but we do not feel that a more
+basic and general introduction to Bayesian methods is appropriate in
+this paper. We cited two excellent texts [Tarantola, 2005; Mosegaard
+and Tarantola, 1995] that explain Bayesian estimation in more detail.
 
@@
 
-This is true; we have added "and assuming that rock strength (cohesion) is
-negligible" to the sentence describing the Mohr-Coulomb failure envelope.
+This is true; we have added "and assuming that cohesion is negligible"
+to the sentence describing the Mohr-Coulomb failure envelope.
+***NOTE: I will remove "rock strength" from the main text, since you are talking about a failure on a pre-existing fault, it is misleading, let's just leave as "cohesion".***
 
@@
 
-The use of the word 'posterior' in statistics is well-established, and surely
-won't incite any more unintended reactions or schoolboy jokes than 'thrusting'.
-We advise readers who blush at the term 'posterior' to avoid television and
-the internet at all costs.
+The use of the word 'posterior' in statistics is well-established and
+is fairly commonly used in geophysical research papers that utilize
+Bayesian estimation or inferrence. Additionally, as described in
+section 4.2, our sampling estimation strategy results in samples of
+some unknown posterior probability density function, this is the
+solution to the inverse problem in a Bayesian sense, and so refering
+to these collections of posterior samples as "functions" would be
+misleading.
 
@@
 
-The direction of slip is given by the coseismic slip models, which are the net
-slip from the earthquake; this is described in Section 4.2.  A consideration
-of rake variability during the rupture would be excellent but unfortunately
-we do not have the data to do this.
+The direction of slip is given by the coseismic slip models, which are
+the net slip from the earthquake; this is described in Section 4.2.
+We think the reviewer is addressing whether under the Wallace-Bott
+assumption the final coseismic slip rake is parallel to the direction
+of maximum shear stress accumulated on the fault prior to slip
+initiation, and that dynamic stresses do not affect the coseismic slip
+rake. We address this point in comment 8 above.
 
@@
 
-We don't view the value of the minimum tectonic stress to be as significant
-as the value of the maximum stress, and there is simple statement that can
-summarize the importance of this value (i.e. it does not directly affect much).
-We do however discuss the results more in the discussion.
+We don't view the value of the minimum tectonic stress to be as
+significant as the value of the maximum stress, and there is no simple
+statement that can summarize the importance of this value (i.e. it
+does not directly affect much).  We do however discuss the results
+more in the discussion.
 
@@
 
-The highest values of the $T'_{\max}$ marginal histograms at around 0.6 are
-evident. This 'peak' (not our wording) is not particularly sharp, but it is
-clearly there; there is a steep decrease in probability to the left of the
-mode and a more gradual decrease to the right of the mode. As stated in the
-response letter to the previous revision, the probability at 2.5 (the upper
-boundary of $T'$ is 60% of the mode at 0.6.  We have amended this sentence
-to read '0.6-0.8' so it is clear that we do not mean a narrow range of high
-likelihoods around a single value.
+The highest values of the $T'_{\max}$ marginal histograms at around
+0.6 are evident. This 'peak' (not our wording) is not particularly
+sharp, but it is clearly there; there is a steep decrease in
+probability to the left of the mode and a more gradual decrease to the
+right of the mode. As stated in the response letter to the previous
+revision, the likelihood at 2.5 (the upper boundary of $T'$ is 60% of
+the mode at 0.6.  We have amended this sentence to read '0.6-0.8' so
+it is clear that we do not mean a narrow range of high likelihoods
+around a single value.
 
@@
 
-We added a reference to the figure (mentioned in the last response).  We would
-rather readers look at the distribution itself than a number we throw out.
+We added a reference to the figure (mentioned in the last response).
+We would prefer that readers look at the distribution itself, as this
+is the quantitative solution of the inverse problem in a Bayesian
+sense, rather than a number we throw out.
 
@@
 
-We believe most readers will not view slip on the Beichuan fault as impossible,
-because it is believed by all researchers we know of to have happened in the
-Wenchuan event.
-
-The reviewer may be missing a key point, which is that the 'optimally-oriented
-fault' has a canonical value of friction of 0.6 (lines 466-469 of the
-submitted manuscript). In contrast, our model results indicate that the faults
-that are thought to have ruptured in this event have a much lower coefficient
-of static friction.  As is shown in Table 2, the coefficient of friction on the
-fault has a strong influence on the favorability of slip on the Wenchuan
-faults vs. optimally oriented faults of the stress parameters studied.
-This is stated repeatedly throughout this section, so we do not believe that
-we are unclear about this point.
+As described in the manuscript, the 'optimally-oriented fault' has a
+canonical value of friction of 0.6 (lines 466-469 of the submitted
+manuscript). In contrast, our model results indicate that the faults
+that are thought to have ruptured in this event have a much lower
+coefficient of static friction.  As is shown in Table 2, the
+coefficient of friction on the fault has a strong influence on the
+favorability of slip on the Wenchuan faults vs. optimally oriented
+faults of the stress parameters studied.  This is stated repeatedly
+throughout this section, so we do not believe that we are unclear
+about this point.
 
@@
 
-Nonetheless, we have reviewed our code and an error in how the optimal
-slip plane is found; essentially, we the algorithm worked assuming that dip
-slip faulting would be favored (where the optimal plane strikes perpendicular
-to the trend of $\sigma_1$) but produces an incorrect orientation otherwise.
-We have fixed the bug, which has changed these results slightly (the bug only
-affected these calculations; that particular function was not used in the rest
-of the study). The new Figure 10 shows that more of the data plot in the
-favorability region for optimal faulting than previously; however, most of the
-data (with low friction values) still plot in the favorability region for the
-Wenchuan faults.
+Nonetheless, we have reviewed our code and found a slight error in how
+the optimal slip plane was calculated; essentially, the algorithm
+assumed that dip slip faulting would be favored (where the optimal
+plane strikes perpendicular to the trend of $\sigma_1$) but produces
+an incorrect orientation otherwise.  We have fixed the bug, which has
+changed these results slightly (the bug only affected these
+calculations; that particular function was not used in the rest of the
+study). The new Figure 10 shows that more of the data plot in the
+favorability region for optimal faulting than previously; however,
+most of the data (with low friction values) still plot in the
+favorability region for the Wenchuan faults.
 
@@
 
-The equations have been amended (a *z* was missing in the numerator of the
-first term in the brackets). The mistake is not found in the code.  This fixes
-the unit problem mentioned.  Although we reproduce the equations from Liu and
-Zoback [1992], we reproduce them here for reference, and because we actually
-only reproduce the right-hand sides of the equations from that reference, but
-we use different terms on the left-hand sides.  (These equations are
-essentially definitions, so there is no algebra or other operators on the
-left-hand sides).  The elastic coefficients
+The equations have been amended (a *z* was missing in the numerator of
+the first term in the brackets). The mistake is not found in the code.
+This fixes the unit problem mentioned.  We reproduce the equations
+from Liu and Zoback [1992] here for completeness. Since we define
+these equations using different notation, we beleive that if we just
+referred to Liu and Zoback [1992] there might be some confusion for
+readers trying to reproduce our results. We thank the reviewer for
+catching the typo in equations (A1) and (A2).
 
@@
 (6). An explanation of this would be helpful.
+
+Liu and Zoback [1992] devote some attention to the horizontal loading
+function (primarily, see the "Effect of Topography on Tectonic Stress"
+Section in their paper). While we do state in the beginnign of the
+section that we use methods derived by Liu and Zoback [1992], we
+realized that we should also cite their paper where we give the
+horizontal loading function, and we have edited the sentence equation
+(5) to state that these loading functions are derived by Liu and
+Zoback [1992].
+
+Yes, Mxx, Txx, Mxy, and Txy act horizontally, but they represent
+normal tractions in horizontal directions on vertical planes, and not
+tractions on the free surface. Mxx, etc, aboce the reference level,
+here taken to be mean sea level, only excite shear traction on the
+free surface (warranting their inclusion into the horizontal loading
+function) if topography varies. We think of this as due to the fact
+that you are changing the total size of the vertical planes that
+tractions resulting from Mxx, etc, are acting upon. This change in
+horizontal normal tration results in a shear stress on the reference
+surface and stresses are propagated into the half-space via the
+Cerruti solutions.
 

### 2015-02-13 revisions/styron_hetland_wenchuan_topo_revised.tex
eric's changes

@@
 \begin{abstract}
-Though it is widely recognized that large mountain ranges produce significant
-stresses in the Earth's crust, these stresses are not commonly quantified.
-Nonetheless, near large mountains topography may affect fault activity by
-changing the stress balance on the faults. In this work, we calculate the
-stress field from topography in the Longmen Shan (Sichuan, China) and resolve
-those stresses on several models of the faults that ruptured in the 2008
-M$_\mathrm{w}$ 7.9 Wenchuan earthquake. We find that the topography results in
-shear stresses up to 20 MPa and normal stresses up to 80 MPa on the faults,
-with significant variability across the faults. Topographic stresses generally
-load the fault in a normal and left-lateral shear sense, opposite to the
-inferred coseismic slip sense, and thus inhibit the coseismic slip. We estimate
-the accumulated tectonic stress needed to overcome topographic and lithostatic
-stresses so that the direction of maximum shear on the faults is roughly
-collinear with the inferred coseismic slip. We further estimate the static
-friction and pore fluid pressure assuming that the fault was, on average, at
-Mohr-Coulomb failure at the time of the Wenchuan earthquake. We use a Bayesian
-inversion strategy, yielding posterior probability distributions for the
-estimated parameters. We find most likely estimates of maximum tectonic
-compressive stress near 0.6 $\rho g z$ and oriented $\sim$E-W, and minimum
-tectonic stress near 0.2 $\rho g z$. Static friction on the fault is near 0.2,
-and pore fluid pressure is between 0 and 0.4 of the total lithostatic
+Though it is widely recognized that large mountain ranges produce
+significant stresses in the Earth's crust, these stresses are not
+commonly quantified.  Nonetheless, near large mountains topography may
+affect fault activity by changing the stress balance on the faults. In
+this work, we calculate the stress field from topography in the
+Longmen Shan (Sichuan, China) and resolve those stresses on several
+models of the faults that ruptured in the 2008 M$_\mathrm{w}$ 7.9
+Wenchuan earthquake. We find that the topography results in shear
+stresses up to 20 MPa and normal stresses up to 80 MPa on the faults,
+with significant variability across the faults. Topographic stresses
+generally load the fault in a normal and left-lateral shear sense,
+opposite to the inferred coseismic slip sense, and thus inhibit the
+coseismic slip. We estimate the tectonic stress needed to overcome
+topographic and lithostatic stresses by assuming that the direction of
+maximum shear accumulated on the faults is roughly collinear with the
+inferred coseismic slip.  We further estimate the static friction and
+pore fluid pressure assuming that the fault was, on average, at
+Mohr-Coulomb failure at the time of the Wenchuan earthquake. We use a
+Bayesian inversion strategy, yielding posterior probability
+distributions for the estimated parameters. We find most likely
+estimates of maximum tectonic compressive stress near 0.6 $\rho g z$
+and oriented $\sim$E-W, and minimum tectonic stress near 0.2 $\rho g
+z$. Static friction on the fault is near 0.2, and pore fluid pressure
+is between 0 and 0.4 of the total lithostatic
 pressure.  \end{abstract}
@@
 particularly if the region is tectonically active.  For example, shear and
-normal stresses on a fault from topographic loading may push a particular fault
+normal stresses on a fault due to topographic loading may push a particular fault
 closer to or farther from failure, or reorient the
@@
 
-In this work, we investigate the effects of topographic stresses on the faults
-that ruptured in the 2008 \emph{M}7.9 Wenchuan, China earthquake.  This
-earthquake is an ideal candidate for this study because it occurred at the base
-of the Longmen Shan, one of the largest and steepest escarpments on Earth
-(Figure~\ref{fig:lms_map}), and has a well-studied coseismic slip distribution
-characterized by significant along-strike variations in coseismic slip.
-Additionally, because the earthquake occurred after $\sim$2000 years
-of seismic quiescence, postseismic stresses in the lithosphere are likely to be
-negligible, suggesting that the stress state on the faults at the time of
-rupture can be approximated as the sum of topographic, lithostatic, and
-accumulated tectonic stresses. We then use the Mohr-Coulomb failure criterion
-to bracket the tectonic stress field, pore fluid pressure, and static friction
-of the fault, so that these parameters are consistent with the toopographic
-stresses and coseismic slip on the Wenchuan earthquake faults.
+In this work, we investigate the effects of topographic stresses on
+the faults that ruptured in the 2008 \emph{M}7.9 Wenchuan, China
+earthquake.  This earthquake is an ideal candidate for this study
+because it occurred at the base of the Longmen Shan, one of the
+largest and steepest escarpments on Earth (Figure~\ref{fig:lms_map}),
+and has a well-studied coseismic slip distribution characterized by
+significant along-strike variations in coseismic slip.  Additionally,
+because the earthquake occurred after $\sim$2000 years of seismic
+quiescence, postseismic stresses in the lithosphere are likely to be
+negligible, suggesting that the stress state on the faults at the time
+of rupture can be approximated as the sum of topographic, lithostatic,
+and accumulated tectonic stresses. We then use assumptions that
+coseismic slip is collinear with the direction of total accumulated
+shear stress on the fault and that the fault is on average at a
+Mohr-Coulomb failure criterion in order to bracket the tectonic stress
+field, pore fluid pressure, and static friction of the fault, so that
+these parameters are consistent with the toopographic stresses and
+coseismic slip on the Wenchuan earthquake faults.
 
@@
 
-Topographic stresses are only one component of the total stress field in the
-crust \citep{molnar1988}. Coseismic slip in the Wenchaun earthquake is due to
-the total accumulated stress on the faults, which also includes components from
-lithostatic and tectonic stresses. (In the present study, we do not consider
-stresses due to flexure \citep[e.g.,][] {luttrell2007} or buoyancy
-\citep[e.g.,][]{luttrell2011}.) By quantifying both the topographic and
-lithostatic stresses, we can use coseismic slip models to solve for the
-tectonic stress assuming that (1) the total shear stress on the fault is in the
-direction of coseismic slip \citep[e.g.,] []{angelier1994} and (2) the fault is
-at Mohr-Coulomb failure everywhere that it slipped. If topographic stresses are
-significant and produce shear in the direction of fault slip, then for given
-values of static friction and pore fluid pressure, we can calculate the amount
-of tectonic stress that can be added to the ambient stress field before the
-faults should rupture; given limited acceptable ranges for friction and fluid
-pressure, we are essentially able to place maximum constraints on tectonic
-stress. Alternately, if topographic stresses work against coseismic slip, for
-given friction and fluid pressures we can estimate the minimum magnitudes of
-tectonic stresses necessary to overcome shear and frictional resistance to
-slip. In a scenario with complex faulting and topography, it may be possible to
-put bounds on both minimum and maximum magnitudes, in addition to the 
-directions of tectonic
-stresses. To account for the non-uniqueness of the solution, we use a Bayesian
-Monte-Carlo methodology to estimate posterior probability density functions
-(PDFs) of tectonic stresses, static fault friction, and pore fluid pressure.
+Topographic stresses are only one component of the total stress field
+in the crust \citep{molnar1988}. Coseismic slip in the Wenchaun
+earthquake is due to the total accumulated stress on the faults, which
+also includes components from lithostatic and tectonic stresses. (In
+the present study, we do not consider stresses due to flexure
+\citep[e.g.,][] {luttrell2007} or buoyancy
+\citep[e.g.,][]{luttrell2011}.) By quantifying both the topographic
+and lithostatic stresses, we can use coseismic slip models to solve
+for the tectonic stress assuming that (1) coseismic slip is in the
+direction of the the total shear stress on the fault \citep[e.g.,]
+[]{angelier1994} and (2) the fault is at Mohr-Coulomb failure
+everywhere that it slipped.  Assumption (1) is the common
+`Wallace-Bott' assumption (named after \citet{wallace1951} and
+\citet{bott1959}), and note that it ignores dynamic stresses during
+the earthquake process, and only consider that fault slip is collinear
+with the accumulated stress on the fault prior to rupture. While this
+assumption should be subject to further testing, it is standard in all
+studies that infer stress from earthquake data
+\citep[e.g.,][]{mckenzie1969, angelier1994, michael1987, reches1987,
+  luttrell2011, medinaluna2013}.  If topographic stresses are
+significant and produce shear in the direction of fault slip, then for
+given values of static friction and pore fluid pressure, we can
+calculate the amount of tectonic stress that can be added to the
+ambient stress field before the faults should rupture; given limited
+acceptable ranges for friction and fluid pressure, we are essentially
+able to place maximum constraints on tectonic stress. Alternately, if
+topographic stresses work against coseismic slip, for given friction
+and fluid pressures we can estimate the minimum magnitudes of tectonic
+stresses necessary to overcome shear and frictional resistance to
+slip. In a scenario with complex faulting and topography, it may be
+possible to put bounds on both minimum and maximum magnitudes, in
+addition to the directions of tectonic stresses. To account for the
+non-uniqueness of the solution, we use a sampling-based Bayesian
+Monte-Carlo methodology to estimate posterior probability density
+functions (PDFs) of tectonic stresses, static fault friction, and pore
+fluid pressure.
 
@@
 \end{equation}
-where $G(x,y,z)$ is a set of Green's functions for the six stress tensor
-elements, and $F(x,y)$ is a topographic loading function, described
-below. We assume compressive stresses are positive, and $x>0$ is east,
-$y>$ is north, and $z>0$ is depth. \citet{liuzoback1992}
-show that $M(x,y,z)$ can be decomposed into two components as
+where $G(x,y,z)$ is a set of Green's functions for the six stress
+tensor elements, and $F(x,y)$ is a topographic loading function,
+described below. We assume compressive stresses are positive, and
+$x>0$ is east, $y>$ is north, $z=0$ is mean sea level, and $z>0$ is
+depth. \citet{liuzoback1992} show that $M(x,y,z)$ can be decomposed
+into two components as
 
@@
 where $G_i^C(x,y,z)$ are the Cerruti solutions for a horizontal point
-source load in the $i$ direction on the halfspace surface (see Appendix
-A1). The horizontal loading functions are given by
+source load in the $i$ direction on the halfspace surface (see
+Appendix A1). The horizontal loading functions derived by
+\citet{liuzoback1992} are given by
 
@@
 fault geometry but use different regularization in the inference of the slip
-distribution). By using a suite of models in our calculations, we can infer
+distribution, and we consider both their `rough' and `smooth' models here). By using a suite of models in our calculations, we can infer
 that results which are persistent in most or all of the models are more robust,
@@
 more confidently linked to specifics of the model geometry or inferred slip
-distribution in those models.
+distribution in that model.
 
@@
 tensor to fault normal stress, $\sigma_n^M$, down-dip shear stress,
-$\tau_d^M$ and strike-slip shear stress, $\tau_s^M$ at each point in
-describing the fault geometry. In this work, the superscripts on stress
-symbols denote the origin of the stresses.
+$\tau_d^M$, and strike-slip shear stress, $\tau_s^M$, at each point in
+describing the fault geometry (the superscripts on stress
+symbols denote the origin of the stresses).
 
@@
 $\sigma_p$ is the pore fluid pressure \citep[e.g.,][]{sibson1985}, and
-assuming that rock strength (cohesion) is negligible. We
+assuming that cohesion is negligible. We
 describe the pore fluid pressure using a scalar, $0 \leq \phi \leq 1$,
@@
 consistent with published coseismic slip models of the Wenchuan
-earthquake using a Bayesian estimation, resulting in posterior probability 
+earthquake using a Bayesian estimation, resulting in samples of posterior probability 
 density functions of the model parameters. 
@@
 stresses using Bayesian methods, and making the common `Wallace-Bott'
-assumption (named after \citet{wallace1951} and \citet{bott1959}) 
-that slip on the fault occurs in the general direction
-of the maximum resolved shear stress on the fault \citep[e.g.,][]
-{mckenzie1969, angelier1994}. We estimate the tectonic stresses in
-light of the topographic stresses and slip distributions through the
-relation
+assumption (named after \citet{wallace1951} and \citet{bott1959}) that
+slip on the fault occurs in the general direction of the maximum
+resolved shear stress on the fault prior to the initiation of the
+earthquake \citep[e.g.,][] {mckenzie1969, angelier1994}. We estimate
+the tectonic stresses in light of the topographic stresses and slip
+distributions through the relation
 
@@
 constrain pore fluid pressure and friction. (6) We use a Bayesian
-estimation, resulting in a PDF of tectonic stress, as well as $\mu$ and
+estimation, resulting in samples of a PDF of tectonic stress, as well as $\mu$ and
 $\phi$, rather than just solving for the minimum tectonic stress
@@
 trace, suggesting that much of the right-lateral component of shear
-stress was released during the earthquake. These results are similar to
+stress was released during the earthquake. Our results are also similar to
 the orientations of total stress obtained by 
@@
 about $T^\prime_{\mathrm{max}} = 0.5$ (Figure~\ref{fig:joint_posteriors}). 
-These results indicates that $T_{\mathrm{max}}$ of at
+These results indicate that $T_{\mathrm{max}}$ of at
 least $\sim$13.25 MPa km$^{-1}$ is necessary to overcome
@@
 similar to the inferences of strain accumulation rates inferred from
-squishy block modeling by \citet{loveless2011}, who
+`squishy' block modeling by \citet{loveless2011}, who
 concluded that the rate of slip deficit accumulation in the thrust and
@@
 and is in the $\sigma_1$--$\sigma_3$ plane \citep[e.g.,][]{sibson1985}.
-Note this $\mu$ is typical for crustal rocks with fault normal stresses
+Note that $\mu=0.6$ is typical for crustal rocks with fault normal stresses
 above 200 MPa, but lower than $\mu = 0.85$ for smaller $\sigma_n$
@@
 to moderate values of $\phi$, probably yield stress conditions insufficient
-for slip at high pressures found at depths > 10 km.
+for slip at high pressures found at depths $>$ 10 km.
 Similarly, a low viscosity horizontal
@@
 
-Assuming that the Beichuan faults were at a Mohr-Coulomb fault criterion
-immediately prior to the Wenchuan earthquake, we estimate the tectonic
-stresses required for the faults to fail. We use a Bayesian estimation,
-resulting in posteriors representing likelihood of tectonic stress,
-static friction, and a pore pressure parameter. The posteriors
-indicate that the maximum tectonic stress is oriented
-$\sim$E-W and has a likely minimum of about 10 MPa per kilometer of depth
-(i.e. $T^\prime_{\mathrm{max}}$ of at least 0.4).
-The minimum tectonic stress is oriented $\sim$N-S and is
-fairly low, with the most likely values lower than 10-12 MPa per kilometer of
-depth ($T^\prime_{\mathrm{min}} <$ 0.5).
-The highest likelihood coefficient of static friction on the fault is
-estimated at about 0.2--0.3, although values up to 0.5--0.6 are
-permissible. Fluid pressures are likely 0--0.5 of the total pressure.
-Slip occurred on these faults instead of more favorably-oriented faults
-elsewhere in the region, due to the inferred low coefficient of friction
-and moderate fluid pressures.
+Assuming that the Beichuan faults were at a Mohr-Coulomb fault
+criterion immediately prior to the Wenchuan earthquake, we estimate
+the tectonic stresses required for the faults to fail. We use a
+Bayesian estimation, resulting in samples of posterior probability
+distributions representing likelihood of tectonic stress, static
+friction, and a pore pressure parameter. The posteriors indicate that
+the maximum tectonic stress is oriented $\sim$E-W and has a likely
+minimum of about 10 MPa per kilometer of depth
+(i.e. $T^\prime_{\mathrm{max}}$ of at least 0.4).  The minimum
+tectonic stress is oriented $\sim$N-S and is fairly low, with the most
+likely values lower than 10-12 MPa per kilometer of depth
+($T^\prime_{\mathrm{min}} <$ 0.5).  The highest likelihood coefficient
+of static friction on the fault is estimated at about 0.2--0.3,
+although values up to 0.5--0.6 are permissible. Fluid pressures are
+likely 0--0.5 of the total pressure.  Slip occurred on these faults
+instead of more favorably-oriented faults elsewhere in the region, due
+to the inferred low coefficient of friction and moderate fluid
+pressures.
 
@@
 \section{Green's functions for point-source
-loads}\label{greens-functions-for-point-source-loads}
-For completeness, we reproduce the Boussinesq \citep[e.g.,][]{jeffreys1970}
-and Cerruti \citep[e.g.,][]{love1927} solutions here.
-Note that in these solutions, $\lambda$ and $\mu$ are re-defined, and
-are the first and second Lame's parameters, respectively, instead of
-rake and fault friction as in the body of the manuscript.
+loads}\label{greens-functions-for-point-source-loads} For
+completeness, we reproduce the Boussinesq
+\citep[e.g.,][]{jeffreys1970} and Cerruti \citep[e.g.,][]{love1927}
+solutions here.  Note that in these solutions, $\lambda$ and $\mu$ are
+the first and second Lame's parameters, respectively, instead of rake
+and fault friction as in the body of the manuscript.
 

### 2015-02-13 revisions/rebuttal2.md
pushing eric's changes

@@
 
-We have changed the relevant sentence to 'We further estimate the static
+We have changed the relevant sentence to "We further estimate the static
 friction and pore fluid pressure assuming that the fault was*, on average,* at
-Mohr-Coulomb failure at the time of the Wenchuan earthquake.
+Mohr-Coulomb failure at the time of the Wenchuan earthquake."
 
@@
 "It is clear that both $T^\prime_{xx}$ and $\mu$ are strongly correlated with
-$\mathrm{CFR}_o/\mathrm{CFR}_f$, and $\phi$ is to a lesser degree; nonetheless,
+$\mathrm{CFR}_o / \mathrm{CFR}_f$, and $\phi$ is to a lesser degree; nonetheless,
 all significantly affect the relative ease of faulting on the Wenchuan faults
 versus optimal faults. In particular, lower values for any of them favor slip
-on the Wenchuan faults. The lowest $\mathrm{CFR}_o/\mathrm{CFR}_f$ value
-($\sim$0.057, $\sim$ -2.9 in log space) corresponds to the lowest value of
-$\mu$ ($\sim$0.038), and is approximately the ratio of $\mu$ in the model to
+on the Wenchuan faults. The lowest $\mathrm{CFR}_o / \mathrm{CFR}_f$ value
+($\sim$ 0.057, $\sim$ -2.9 in log space ) corresponds to the lowest value of
+$\mu$ ($\sim$ 0.038), and is approximately the ratio of $\mu$ in the model to
 0.6."
@@
 function (primarily, see the "Effect of Topography on Tectonic Stress"
-Section in their paper). While we do state in the beginnign of the
+Section in their paper). While we do state in the beginning of the
 section that we use methods derived by Liu and Zoback [1992], we
@@
 normal tractions in horizontal directions on vertical planes, and not
-tractions on the free surface. Mxx, etc, aboce the reference level,
+tractions on the free surface. Mxx, etc, above the reference level,
 here taken to be mean sea level, only excite shear traction on the

### 2015-02-13 revisions/rebuttal2.md
submitted w/ second revisions

@@
+This document contains the comments from the reviewer, as well as our changes to them.
+
 ## General criticisms
@@
 quotation was appropriate for the topic of this manuscript, and we
- did refer back to it later in the text. We are not sure that if we
- had chosen an opening quotation from a non-religious text, say James
- Joyce's Finnegans Wake, whether there would have been such a visceral
- reaction.  Finally, we did not refer to it 'as one might a
- peer-reviewed scientific article'. That all being said, we have
+did refer back to it later in the text. Finally, we did not refer to it 'as
+one might a peer-reviewed scientific article'. That all being said, we have
 removed the quotation.
@@
 to the sentence describing the Mohr-Coulomb failure envelope.
-***NOTE: I will remove "rock strength" from the main text, since you are talking about a failure on a pre-existing fault, it is misleading, let's just leave as "cohesion".***
 
@@
 is fairly commonly used in geophysical research papers that utilize
-Bayesian estimation or inferrence. Additionally, as described in
+Bayesian estimation or inference. Additionally, as described in
 section 4.2, our sampling estimation strategy results in samples of
 some unknown posterior probability density function, this is the
-solution to the inverse problem in a Bayesian sense, and so refering
+solution to the inverse problem in a Bayesian sense, and so referring
 to these collections of posterior samples as "functions" would be

### 2015-03-18 revisions/styron_hetland_wenchuan_topo_revised.tex

@@
 
-%\begin{figure*}[ht!] \includegraphics[width=40pc]{../figures/lms_map.pdf}
-%    \caption{Map of eastern Tibet and the Sichuan basin, showing active
-%        structures from \citet{styron2010}. Faults that ruptured in the 2008
-%        Wenchuan earthquake are shown in pink. GPS velocities are relative to
-%        the mean velocity of sites within Sichuan basin, with 1$\sigma$
-%        uncertainty, from the dataset of Liang et al.  \citet{liang2013}.
-%        Beachball is from the Global CMT focal mechanism solution for the 2008
-%    Wenchuan earthquake. BF = Beichuan fault. PF = Pengguan fault. P = Pengguan
-%massif. Grey box shows the extent of Figures 3 and 9.} \label{fig:lms_map}
-%\end{figure*}
-
 Surface ruptures during the Wenchuan earthquake are highly variable and show
@@
 topographic stresses.
-
-%\begin{figure*}[t!] \centering
-%    \includegraphics[width=40pc]{../figures/topo_stress_possibilities.pdf}
-%    \caption{Scenarios for topographic effects on rangefront thrust faulting.
-%        (a) Topographic stresses promote thrust faulting. (b) Topographic
-%        stresses inhibit thrust faulting.} \label{fig:topo_fault_scenarios}
-%    \end{figure*}
 
@@
 
-%\begin{table}
-%\centering
-%\begin{tabular}{l c l}
-%\hline
-%Parameter & Value & Unit \\
-%\hline
-%horizontal spacing & 851 & m \\ 
-%vertical spacing & 1000 & m \\ 
-%minimum depth & 851 & m (below sea level) \\ 
-%maximum depth & 35851 & m (below sea level) \\ 
-%density ($\rho$) & 2700 & kg m$^{-3}$ \\ 
-%g & 9.81 & m s$^{-2}$ \\ 
-%Green's function radius & 9e5 & m \\ 
-%Poisson ratio & 0.25 & - \\ 
-%\hline
-%\end{tabular}
-%\caption{Parameters for numerical calculations of topographic stresses.}
-%\label{table:convo_params}
-%\end{table}
-
 \subsection{Topographic fault stress
@@
 
-%\begin{figure}[t]
-%\centering
-%\includegraphics[width=20pc]{../figures/lms_topo_stresses_rot.pdf}
-%\caption{Horizontal topographic stresses in the Longmen Shan region at 5
-%km depth: black and red lines signify most and least compressive
-%horizontal stresses, respectively. Other symbols are the same as in Figure
-%1. Stresses shown are down-sampled from the
-%discretization used in the calculations by a factor of six.}
-%\label{fig:lms_topo_stresses_rot}
-%\end{figure}
-
 Because the compressive stresses are near equal, $M$ contains a large
@@
 of coseismic slip, which is reverse-dextral.
-
-%\begin{figure*}[htbp]
-%\centering
-%\includegraphics[width=40pc]{../figures/fault_stress_3d.pdf}
-%\caption{Southwest-looking views of topographic stresses and coseismic
-%slip on selected slip models. All views share same lateral extent, but
-%the perspectives for B and C are more inclined than A. (A) $\sigma_n^M$
-%(colors), slip magnitude (contours, 1 m interval) and hanging-wall
-%topography on the \citet{feng2010} model of the Beichuan
-%fault. Note the suppression of fault slip where normal stress is
-%highest, such as below the Pengguan massif (P). Fault and topography
-%share the same scale, with no vertical exaggeration. (B) $\tau_d^M$ (colors)
-%and dip slip (contours, 1 m interval) \citet{qi2011}
-%`rough' slip model of the Beichuan fault. (C) $\tau_s^M$ (colors) and strike
-%slip (contours, 1
-%m interval) on the \citet{qi2011} `rough' slip model if the Beichuan fault.}
-%\label{fig:fault_stress_3d}
-%\end{figure*}
 
@@
 forthcoming manuscript.
-
-%\begin{figure}%[htbp]
-%\centering
-%\includegraphics[width=20pc]{../figures/feng_slip_sig_n_scatter.pdf}
-%\caption{Coseismic slip magnitude and $\sigma^M_n$ on the four segments
-%of the \citet{feng2010} coseismic slip model. Trendlines are
-%L1 regressions and do not include points with no slip. ``NE B'' =
-%Northeastern Beichuan fault. ``C B'' = central Beichuan fault. ``SW B'' =
-%Southwestern Beichuan fault. ``P'' = Pengguan fault.}
-%\label{fig:feng_slip_sig_n_scatter}
-%\end{figure}
 
@@
 typically 0--0.4 of $T^\prime_{\mathrm{max}}$, but rarely higher.
-
-%\begin{figure*}%[htbp]
-%\centering
-%\includegraphics[width=40pc]{../figures/T_scatters.pdf}
-%\caption{Samples of
-%$p(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$ associated with
-%each of the coseismic slip models we consider, along with marginals of
-%$T^\prime_{\mathrm{max}}$ and $T^\prime_{\mathrm{min}}$. Inset rose
-%diagrams are histograms of azimuth of
-%$T^\prime_{max}$.}
-%\label{fig:T_scatters}
-%\end{figure*}
 
@@
 
-%\begin{figure*}[htbp]
-%\centering
-%\includegraphics[width=40pc]{../figures/mu_phi_fms.pdf}
-%\caption{Samples of $p(\mu,\phi | D)$ for each coseismic slip model.
-%Colors indicate magnitude of $T^\prime_{\mathrm{max}}$. Contour lines
-%indicate relative density (i.e., likelihood) of posteriors (darker lines
-%signify higher densities).}
-%\label{fig:mu_phi_scatters}
-%\end{figure*}
-
 \subsection{Joint posteriors}\label{joint-posteriors}
@@
 more robust.
-
-%\begin{figure*}[t]
-%\centering
-%\includegraphics[width=40pc]{../figures/joint_pdfs.pdf}
-%\caption{(a) Samples of
-%$p_{\mathrm{J}}(T^\prime_{\mathrm{max}},T^\prime_{\mathrm{min}} | D)$,
-%along with marginals of $T^\prime_{\mathrm{max}}$ and
-%$T^\prime_{\mathrm{min}}$. (b) Samples of
-%$p_{\mathrm{J}}(T^\prime_{\mathrm{\max}}, T^\prime_{\mathrm{\min}}/T^\prime_{\mathrm{max}} | D)$,
-%along with marginals of $T^\prime_{\mathrm{max}}$ and
-%$T^\prime_{\mathrm{min}} / T^\prime_{\mathrm{max}}$. (c) Histogram of
-%azimuths of $T^\prime_{\mathrm{\max}}$. (d) Samples of
-%$p_{\mathrm{J}}(\mu, \phi | D)$, with marginals, where color of the samples
-%indicate magnitudes of $T^\prime_{\mathrm{\max}}$ and contour
-%lines indicate relative density (i.e., likelihood) of posteriors (darker
-%lines signify higher densities).}
-%\label{fig:joint_posteriors}
-%\end{figure*}
 
@@
 
-%\begin{figure}[t]
-%\centering
-%\includegraphics[width=20pc]{../figures/lms_map_stresses_rot.pdf}
-%\caption{Topographic and tectonic horizontal stresses (taken from the
-%most likely estimates of $p(T|D)$ in the Wenchuan rupture region (black
-%and red crosses) with horizontal maximum stress orientation data taken
-%from before the 2008 Wenchuan event from the World Stress Map
-%\citep{heidbach2009} (purple arrows), and horizontal maximum stress
-%orientation data from after the earthquake at the WFSD-1 drill hole
-%\citep{cui2014} (blue arrows). Other symbols are as in Figure
-%\ref{fig:lms_map}. Stresses shown are downsampled from our computational 
-%grid resolution by a factor of nine.}
-%\label{fig:lms_stress_map}
-%\end{figure}
-
 The magnitudes of $T_{\mathrm{max}}$ and $T_{\mathrm{min}}$ are
@@
 similar to the inferences of strain accumulation rates inferred from
-`squishy' block modeling by \citet{loveless2011}, who
+squishy block modeling by \citet{loveless2011}, who
 concluded that the rate of slip deficit accumulation in the thrust and
@@
 
-%\begin{figure}%[tb]
-%\centering
-%\includegraphics{../figures/cfr_plots.pdf}
-%\caption{Comparison of Coulomb failure ratio (CFR) on the Beichuan fault
-%from the \citet{zhang2011} coseismic slip model to CFR on an
-%optimally-oriented fault with $\mu = 0.6$, versus estimated $\mu$ on the
-%Beichuan fault. Values less than 0 (1 in linear space) indicate that
-%slip is favored on the Beichuan fault, even if it is not optimally
-%oriented. Values are calculated for each point in the slip model for
-%1000 randomly-drawn samples from $p(T,\mu,\phi |D)$.}
-%\label{fig:cfr_ratios}
-%\end{figure}
-%
-%\begin{table}[htb]
-%\centering
-%\begin{tabular}{l c c c c}
-%\hline
-%Parameter & Coef. & Std. Err. & t statistic & $P>|t|$ \\
-%\hline
-%Intercept & -0.2566 & 0.0020 & -128.8 & \textless{}0.0001 \\ 
-%$\mu$ & 1.1168 & 0.0086 & 130.6 & \textless{}0.0001 \\ 
-%$T^\prime_{xx}$ & 0.9138 & 0.0048 & 191.2 & \textless{}0.0001 \\ 
-%$\phi$ & 0.5762 & 0.0055 & 105.7 & \textless{}0.0001 \\ 
-%\hline
-%\end{tabular}
-%\caption{Sensitivity of CFR ratios to relevant stress state parameters:
-%Results of multivariate linear regression of
-%$\mathrm{CFR}_o/ \mathrm{CFR}_f)$ against $\mu$, $\phi$ and
-%$T^\prime_{xx}$.}
-%\label{table:cfr_regress}
-%\end{table}
-
 \subsection{The role of topography in orogenic development and strain
@@
 \url{https://github.com/wenchuan_topo_stress}. We thank the editor, Paul
-Tregoning, and Steve Marshak and an anonymous reviewer for reviews. Funding for
+Tregoning, and two anonymous reviewers for reviews. Funding for
 this study was provided by the University of Michigan. Joe Kington wrote the
@@
 \end{acknowledgements}
-
-%\section{References}\label{references}
-
-%\bibliography{wench_jgr_bib}{}
-%\bibliographystyle{agufull08}
 
